### 2014-08-04 tex/lib/DiagramAAM.tex
initial

\begin{center}
\begin{tabular*}{0.66\textwidth}{@{\extracolsep{\fill}} c c}
  \begin{tikzpicture}[ampersand replacement=\&]
    \matrix (m) [matrix of math nodes,row sep=3em,column sep=4em,minimum width=2em]
    { \DSS                                 \& \DSS                                 \\
      \node [lightgray] (m-2-1) {\DSSAbs}; \& \node [lightgray] (m-2-2) {\DSSAbs}; \\
    } ;
    \path[-stealth]
    (m-1-1) edge [lightgray,bend right=40] node [lightgray,left] {$\alpha$} 
                                           node [lightgray,right] {\tiny(2.a)} (m-2-1)
    (m-2-1) edge [lightgray,bend right=40] node [lightgray,right] {$\gamma$} (m-1-1)

    (m-1-2) edge [lightgray,bend right=40] node [lightgray,left] {$\alpha$} (m-2-2)
    (m-2-2) edge [lightgray,bend right=40] node [lightgray,right] {$\gamma$} (m-1-2)

    (m-1-1) edge node [above] {$\Dstep$} (m-1-2)
    (m-2-1) edge node [above] {$\DstepAbs$} 
                 node [below] {\tiny(1.b)} (m-2-2)
    ;
    \draw 
      (m-2-1) node [lightgray,left=0.25cm] {\tiny(1.a)} 
    ;
  \end{tikzpicture}
  &
  \begin{tikzpicture}
    \matrix (m) [matrix of math nodes,row sep=3em,column sep=4em,minimum width=2em]
    { \DstepAbs \\
      \Dstep    \\
    } ;
    \path
    (m-1-1) edge node [right] {$\sqsubseteq$} 
                 node [left] {\tiny(2.b)} (m-2-1)
    ;
  \end{tikzpicture}
\end{tabular*}
\end{center}

### 2014-08-04 tex/sections.old/00Abstract.tex
initial

The design and implementation of static analyses is a difficult process.
%
Verifying the correctness of an implementation is often the most painful step.
%
We present Monadic AAM--an extension of the Abstracting Abstract Machines
methodology introduced by Van Horn and Might~\cite{van-horn:2010:aam}--which
captures a large class of both automatically derivable and correct by
construction abstract interpreters.
%
Monadic AAM is part methodology and part toolkit.
%
In the methodology, semantics are designed in a monadic extension of
traditional AAM.
%
Once designed, a large class of known analyses can be automatically recovered
using our language agnostic toolkit.
%
Both the computational and correctness properties of our framework are realized
through a restricted class of monad transformers.
%
Consequently our framework enjoys the benefits of being highly compositional
and demands minimal proof burdens from the user.

### 2014-08-04 tex/sections.old/01Introduction.tex
initial

\subsection{Basic Order Theory}

\noindent [TODO: Partial Orders. Define ``discrete ordering''.]

\noindent [TODO: Products.]

\noindent [TODO: Function Spaces.]

\noindent [TODO: Sets.]

\noindent [TODO: Monotonicity.]

\noindent [TODO: Fixpoints.]

\subsection{Basic Abstract Interpretation Concepts}

\noindent [TODO: Small-step semantics.]

\noindent [TODO: Galois Connections.]

\noindent [TODO: An extremely simple example (borrow this from Cousot).]

\noindent [TODO: A calculational derivation of the previous example.]

\subsection{Small Step Semantics}

\noindent [TODO: Lambda Calculus Reduction Semantics.]

\noindent [TODO: Lambda Calculus Small Step Semantics.]

\noindent [TODO: CPS motivation and examples.]

### 2014-08-04 tex/sections.old/02ByExample.tex
initial

%%%%%%%%%%%%%%%%%%%% Introduction %%%%%%%%%%%%%%%%%%%%

To demonstrate our monadic abstract interpretation framework we guide the
reader through the design and implementation of 0CFA for a CPS language with
small-step semantics.
%
We present the process guided by three approaches to abstract interpretation:
%
\begin{itemizenobreak}
  \item Ad hoc design and ``a posteriori'' verification
  \item Abstracting Abstract Machines a la Van Horn and Might~\cite{van-horn:2010:aam}
  \item Using our monadic framework
\end{itemizenobreak}
%
General computational, correctness and modularity properties of our framework
are deferred to sections~\ref{section:computational},~\ref{section:correctness}
and~\ref{section:modularity}.
%
0CFA is chosen because it is well-studied, challenging, and useful for analysis of
higher order languages.
%
CPS is chosen to simplify control flow, and also for being well-studied in the
context of 0CFA.

%-----%

Our framework, and the techniques presented in this paper, are specialized to
the traditions of abstract interpretation pioneered by Cousot and Cousot
\cite{cousot:1977:ai}, and small-step operational semantics pioneered by
Friedman and Felleisen \cite{felleisen:1986:cek}.
%
Consequently, each semantics presented will take the form of a state space
$\DSS$ and transition function $\Dstep \in \DSS \to \DSS$, and each static
analysis will take the form of an abstract state space $\DSSAbs$ and abstract
transition function $\DstepAbs \in \DSSAbs \to \DSSAbs$.
%
Possibilties of generalizing to other approaches for semantics are discussed in
section~\ref{section:future-work}.
%
We will use Haskell as both an implementation language and semantics
meta-language.

%%%%%%%%%%%%%%%%%%%% Concrete Semantics %%%%%%%%%%%%%%%%%%%%

\subsection{Concrete Semantics}

Before discussing various approaches to the design and implementation of 0CFA
for CPS, we first present the semantics of CPS using Haskell.

%-----%

%%%%%%%%%%%%%%%%%%%% Ad-hoc Design %%%%%%%%%%%%%%%%%%%%

\subsection{Ad-hoc Design and ``A Posteriori'' Verification }

In the ``a posteriori'' approach to abstract interpretation the analysis
designer invents:
%
\begin{itemizenobreak}
\item (1.a) Abstract state space $\DSSAbs$
\item (1.b) Abstract step function $\DstepAbs \in \DSSAbs \to \DSSAbs$
\end{itemizenobreak}
%
and then attempts to establish correctness properties in an ad-hoc fashion.
%
In the abstract interpretation framework, soundness and tightness are
demonstrated through:
%
\begin{itemizenobreak}
\item (2.a) A Galois connection $\DSS \galois{\alpha}{\gamma} \DSSAbs$
\item (2.b) A proof that 
      $\alpha(\Dstep) \sqsubseteq \DstepAbs$
      or
      $\Dstep \sqsubseteq \gamma(\DstepAbs)$
      (they are equivalent)
\end{itemizenobreak}
%
These points are depicted in the following diagram:
%
\input{lib/DiagramAdHoc.tex}

%-----%

In the ad-hoc approach to static analysis, none of these steps are automated,
nor does the designer follow a methodology for how to proceed.
%
However this gives the analysis designer ultimate freedom and is simultaneously
the appeal of the approach, lending it to be used most often in practice.

%-----%

We present what the analysis designer might come up with for (1.a) and
(1.b) in designing 0CFA, and defer the ``a posteriori'' proofs of (2.a) and
(2.b) to prior work \cite{shivers:1991:cfa, nielson:2004:program-analysis,
van-horn:2010:aam, midtgaard:2008:calculational-cfa}.

%-----%

The state space for 0CFA, and all abstract state spaces to follow,
differ only in the \lstinline|Val|, \lstinline|Env| and $\DSS$ types.
%
\begin{donotbreak}
\lstinputlisting{hs/SSAbstract.hs}
\end{donotbreak}
%
A single global environment \lstinline|Env| is used to approximate all
environments in the concrete image of the abstract state space, which
corrosponds to heap widening in the literature\cite{shivers:1991:cfa}.
%
To maintain soundness in the face of a single global environment, we introduce
a \lstinline|Set| of values in the image of the environment, as it may need to
approximate the binding bindings with all possible 
%
The state space $\DSS$ contains a set of \lstinline|Call| states
because the abstract semantics are non-deterministic.
%
As a consequence of the global environment abstraction, closures no longer
need to capture their environment.
%
This cuts the recursion in the state space allowing $\DSS$ to be finite and
therefore the analysis to be computable.

%-----%

The abstract step function $\DstepAbs$ exhibits a few key differences from the
concrete step function $\Dstep$.
%
\lstinputlisting{hs/SSAbstract.hs}
%
Interpretations of literal constants and primitive operations truncate to a
symbolic type tag.
%
Non-determinism is introduces in both the denotation for \lstinline|Atomic| and
operational behavior of \lstinline|Call| when multiple values are returned from
the environment.
%
The global environment is computed as the join of all environments reached
after a single step, soundly abstracting each individual reachable environment.

%-----%

In summary, we claim advantages and deficiencies of the ad-hoc ``a posteriori''
approach are:
\begin{itemizenobreak}
\item + No restrictive framework in which to funnel the desired analysis
\item + Most ``off the shelf'' analyses from the literature use this method
\item - The analysis chosen is an arbitrary point on a spectrum of possible design choices
\item - Non-modular: Altering the analysis requires reconstructing all points (1.a) - (2.b)
\end{itemizenobreak}

%%%%%%%%%%%%%%%%%%%% AAM %%%%%%%%%%%%%%%%%%%%

\subsection{Abstracting Abstract Machines}

In the abstracting abstract machines approach to abstract interpretation, the
analysis designer:
%
\begin{itemizenobreak}
\item (1.a) Is guided to a particular abstract state space $\DSSAbs$ by a systematic methodology
\item (1.b) Invents abstract step function $\DstepAbs \in \DSS \to \DSS$
\end{itemizenobreak}
%
and in making the correctness argument must establish: 
%
\begin{itemizenobreak}
\item (2.a) A Galois connection $\DSS \galois{\alpha}{\gamma} \DSSAbs$ is
      derived alongside (1.a) by the systematic methodology
\item (2.b) The proof that $\alpha(\Dstep) \sqsubseteq \DstepAbs$ in an ad-hoc
      manner
\end{itemizenobreak}
%
These points are depicted in the following diagram, where grayed out names and
arrows indicate getting something for free (literally or methodologically):
%
\input{lib/DiagramAAM.tex}

%-----%

The derivation process for (1.a) and (2.a) in the Abstracting Abstract Machines
framework begins with identifying recursion in the state space \lstinline|SS|,
and cutting it with a level of indirection.
%
\begin{lstlisting}
data Val = LitV Lit 
         | CloV [Name] Call Env
type Env = Map Name Val                  
type SS  = Maybe (Call, Env)
\end{lstlisting}
%
becomes
%
\begin{lstlisting}
data Val   = LitV Lit
           | CloV [Name] Call Env
type Env   = Map Name Integer
type Store = Map Integer Val
type SS    = Maybe (Call, Env, Store)
\end{lstlisting}
%
A semantics for the new state space allocates unique addresses in the global
store when binding in a local environment.
%
This new semantics can be shown to be equivalent to the initial semantics,
allowing us to recover the concrete semantics with our new ``instrumented''
semantics.

%-----%

To arrive at a static analaysis, the designer must continue to cut infinite
portions of the state space.
%
The Abstracting Abstract Machines framework guided us to trade recursion for
indirection through the integers, and now we need only abstract the integers to
something finite to achieve a finite state space.
%
\begin{lstlisting}
data Val   = LitV Lit
           | CloV [Name] Call Env
type Env   = Map Name Integer
type Store = Map Integer Val
type SS    = (Maybe (Call, Env), Store)
\end{lstlisting}
%
becomes
%
\begin{lstlisting}
data Val   = LitV Lit
           | CloV [Name] Call Env
type Env   = Map Name Name
type Store = Map Name Val
type SS    = (Maybe (Call, Env), Store)
\end{lstlisting}
%
We then notice that \lstinline|Env| will always be the identity map, and
eliminate it from closures and the state space.
%
\begin{lstlisting}
data Val   = LitV Lit
           | CloV [Name] Call
type Store = Map Name Val
type SS    = (Maybe Call, Store)
\end{lstlisting}
%
Finally, we notice that during execution of the semantics, we necessarily must
reuse abstract addresses, which are now \lstinline|Name|s.
%
Because we must reuse addresses, we must do something sensible (meaning sound)
when two values are bound to the same address.
%
To solve this we introduce non-determinism into the semantics, lifting abstract
values to a powerset and taking the join of abstract values as they collide in
the store.
%
\begin{lstlisting}
data Val   = LitV Lit
           | CloV [Name] Call
type Store = Map Name (Set Val)
type SS    = (Set Call, Store)
\end{lstlisting}

%-----%

In the ad-hoc design, we called the global environment \lstinline|Env|, where
here we use \lstinline|Store|.
%
This is an ontological artifact of the systematic design process of Abstracting
Abstract Machines, where the store was introduced purely to cut recursion in
the state space.

%-----%

The above derivation of abstract state space is also amenable to derivation of
Galois connection.
%
The first step, which introduces concrete addresses with integers, is
isomorphic to the inital state space (under assumptions of heap
well-formedness, which can trivially be shown is preserved by the semantics).
%
Abstracting concrete addresses (integers) with abstract addresses (names) forms
a galois connection between state spaces.
%
This galois connection is difficult to fully justify (Van Horn and Might sketch
the proof; see \cite{midtgaard:2008:calculational-cfa} for all the details),
but can be done once and for all within the methodolgy, and applied freely by
the analysis designer.

%-----%

The benefits of the Abstracting Abstract Machines framework extend beyond a
methodology for designing a \textit{single} abstract interpreter.
%
There is a pivot point intentionally placed in the design process--between
introducing addresses and abstracting addresses--where the designer is set up
to recover a wide range of possible analyses.
%
The idea is then to implement an interpreter once parameterized by the choice
of abstract address (and abstract time, to recover the full AAM framework),
which is capable of recovering a wide range of analyses.
%
This is benefical to the designer as properties of the analysis change, or if
multiple analyses of varying precision and power are desired.

%-----%

However, certain classs of design decisions in constructing analyses are
\textit{not} captured by the pivot point in AAM.
%
For example: heap widening, abstract garbage collection and shape analysis each
cannot be recovered by a single abstract interpreter parameterized by abstract
address and time.
%
While Van Horn and Might have demonstrated the seamless integration of these
ideas into their framework, they do not come for free, and must still be
integrated in an ad-hoc manner for the analysis designer.

%-----%

In summary, we claim advantages and deficiencies of the Abstracting Abstract
Machines approach are:
\begin{itemizenobreak}
\item + Nearly systematic design of abstract state space $\DSSAbs$
\item + Pivot point in analysis design that enables a single parameterized implementation
\item - Constructing abstract step function $\DstepAbs$ is non-automated
\item - Pivot point in analysis design doesn't capture certain important
        classes of analysis design decisions
\end{itemizenobreak}

%%%%%%%%%%%%%%%%%%%% Calculation %%%%%%%%%%%%%%%%%%%%

\subsection{Calculation}

In the calculational approach to abstract interpretation, the analysis designer:
%
\begin{itemizenobreak}
\item (1.a \& 2.a) Constructs a Galois connection $\DSS \galois{\alpha}{\gamma}
                   \DSSAbs$ as the composition of several smaller well known
                   components, inducing the abstract state space $\DSSAbs$ .
\item (1.b \& 2.b) Performs calculation on the induced abstract semantics
                   $\DstepAbs = \gamma \comp \Dstep \comp \alpha$ to arrive at
                   a computable abstract semantics.
\end{itemizenobreak}
%
These points are depicted in the following diagram, where grayed out names and
arrows indicate getting something for free (literally or methodologically):
%
\input{lib/DiagramCalculational.tex}

%-----%

The components used to construct (1.a \& 2.a) are typicall typically drawn from
a catalogue of known techniques such as independent attribute method, direct
product, reduced product, total function space, relational method, direct
tensor product, reduced tensor product and monotone function space, to name a
few~\cite{nielson:2004:program-analysis}.
%
After constructing a final Galois connection, the abstract semantics are
derived through a process of calculation--a mixture of equational and partial
order reasoning--over the induced abstract step function $\DstepAbs = \gamma
\comp \Dstep \comp \alpha$.

%%%%%%%%%%%%%%%%%%%% Monadic Derivation %%%%%%%%%%%%%%%%%%%%

\subsection{Monadic Derivation}

Monadic abstract interpretation was introduced by Sergey et al. as a method for
simultaneously defining multiple versions of a semantics in one go.
%
The primary contribution of this work is a proof framework to augment this
approach, showing that the derived abstract interpreters are correct by
construction.
%
For now we contrast the burdon of proof for the analysis designer with the a
posteriori verification described in the previous section.

First we translate the semantics of CPS into monadic form as shown in figure
?, being careful to leave the semantics general enough to capture both concrete
and abstract semantics.
%
We note that certain design decisions must be made up front, like what the
abstract value space looks like, but defer the full choice of state space to a
particular instantiation within the monadic framework
%
Independent of other languages, we need only justify that mstep is monotonic,
which is true by simple inspection.

We now recover both the concrete semantics and multiple abstract semantics from
the monadic semantics by:
%
\begin{itemizenobreak}
  \item Designing delta and literal semantic functions
  \item Building an instance of the abstract monadic effect interface
\end{itemizenobreak}

### 2014-08-04 tex/sections.old/03Computational.tex
initial

%%%%%%%%%%%%%%%%%%%% Computational Framework %%%%%%%%%%%%%%%%%%%%

\section{Computational Framework for Monadic AI}
\label{section:computational}
Our approach slices the monads of the existing approach into reusable,
language-agnostic monad transformer stacks.
%
In Section ? we will establish exactly the space of monad transformers which,
given specific properties can be proven about them, will integrate seamlessly
within our framework.  For now we introduce a few familiar monad transformers
for the sake of demonstration, and defer the explanation of \textit{why} they
can be used.

\begin{minipage}{\linewidth}
\begin{lstlisting}
Identity     A := A
StateT   S m A := S $\to$ m (S $\times$ A)
State    S   A := StateT Identity A
Set          A := {..., A, ...}
SetT       m A := m (Set A)
\end{lstlisting}
\end{minipage}
Next, we introduce interfaces for various monadic behaviors and their laws:
\begin{minipage}{\linewidth}
\begin{lstlisting}
Monad (m : Type -> Type) :=
  return : $\forall$ A, A $\to$ m A
  bind : $\forall$ A B, m A $\to$ (A $\to$ m B) $\to$ m B
  left_unit : bind aM ret = aM
  right_unit : bind (ret x) f = f x
  associativity : bind (bind aM f) g = bind aM ($\lambda$ x $\to$ bind (f x) g)
MonadFunctor (t : (Type -> Type) -> Type -> Type):= 
  mapMonad : $\forall$ m, Monad m -> Monad (t m)
MonadPlus m | Monad m :=
  mzero : $\forall$ A, m A 
  _<+>_ : $\forall$ A, m A -> m A -> m A
  unit : aM <+> mzero = aM
  left_zero : bind mzero f = mzero
  right_zero : bind aM ($\lambda$ x $\to$ mzero) = mzero
  associativity : (aM <+> bM) <+> cM = aM <+> (bM <+> cM)
  commutativity : aM <+> bM = bM + aM
MonadPlusFunctor t | MonadFunctor t :=
  mapMonadPlus : $\forall$ S m, MonadPlus m -> MonadPlus (t m)
MonadState (S : Type) m | Monad m :=
  getState : m S
  putState : S $\to$ m unit
MonadStateFunctor t | MonadFunctor t := 
  mapMonadState : $\forall$ S m, MonadState S m -> MonadState S (t m)
\end{lstlisting}
and prove the following lemmas:
\begin{lstlisting}
forall S:
  - MonadFunctor (StateT S)
  - MonadPlusFunctor (StateT S)
  - MonadStateFunctor (StateT S)
forall S and m | Monad m:
  - MonadState S (StateT S m)

MonadFunctor SetT
MonadStateFunctor SetT
$\forall$ m | Monad m:
  - MonadPlus (SetT m)
\end{lstlisting}
\end{minipage}

We define \textbf{mjoin} for transporting values from the Set monad to an
arbitrary MonadPlus:
\begin{lstlisting}
mjoin : forall A m | MonadPlus m, Set A -> m A
mjoin {} = mzero
mjoin {..., x, ...} = ... <+> return x <+> ...
\end{lstlisting}

\subsection{Methodology}
Languages are embedded in the framework using the following methodology:
\begin{itemizenobreak}
  \item 
  Define a \textit{single} language state space, to be used for both concrete
  and abstract interpreters.
  \item 
  Define a \textit{single} \textit{monadic} small-step function, parameterized
  by a monadic effect interface and $\delta$ function.
\end{itemizenobreak}
For example, consider the following state space for a very simple arithmetic
language:
\begin{lstlisting}
Int := { ... the integers ... }
Symbol := { ... x, y, etc ... }
Sign := IsZero | IsPos | IsNeg
AInt := Exact Int | Symbolic Sign
Lang := Halt Atomic 
      | LetInc Symbol Atomic Lang 
      | IfZero Atomic Lang Lang 
Atomic := Num AInt | Var Symbol
Env A B := A -> Option B
\end{lstlisting}
Notice that \textbf{Lang} is a \textit{mixed} concrete-abstract language,
potentially capable of expressing both concrete and abstract semantics.
%
In this example we will call the delta function \textbf{inc}, because it
interprets increment for atomic expressions. Given \textbf{$\langle$m,
inc$\rangle$} where:
\begin{lstlisting}
- Monad m
- MonadPlus m
- HasState (Env Symbol (Set AInt)) m
- inc : AInt $\to$ AInt $\to$ Set Ant
\end{lstlisting}
The semantics for the language \textbf{AInt} can be defined:
\begin{lstlisting}
lookup :: Symbol -> Env Symbol (dom AInt) -> m (dom AInt)
lookup x e = case e x of
  None -> mzero
  Some aD -> return aD

atomic : Atomic -> m (dom AInt)
atomic (Num n) = return (return n)
atomic (Var x) = do
  e <- getEnv
  v <- lookup x e
  return v

step : Lang -> m Lang
step (Halt a) = return (Halt a)
step (LetInc x na l) = do
  n <- mjoin (atomic na)
  r <- delta n
  modifyEnv (insert x r)
  return l
step (IfZero ca tb fb) = do
  c <- mjoin (atomic ca)
  case c of
    Exact 0 -> return tb
    Exact _ -> return fb
    Symbolic IsZero -> return tb
    Symbolic IsNeg -> return fb
    Symbolic IsPos -> return fb
\end{lstlisting}

This single semantic step function is capable of expressing both the concrete
and abstract semantics of our arithmetic language.
%
To relate our monadic function back to small-step abstract state machines we
introduce another interface which all monads in the framework must obey:
%
\begin{lstlisting}
MonadStateSpace m :=
  ss : Type -> Type
  transition : $\forall$ A B, (A -> m B) -> ss A -> ss B
MonadStateSpaceFunctor t :=
  mapMonadStateSpace : $\forall$ m, MonadStateSpace m -> MonadStateSpace (t m)
\end{lstlisting}
%
and we prove the following lemmas:
\begin{lstlisting}
$\forall$ S, MonadStateSpaceFunctor (StateT S) where 
  ss (StateT S m) A := $\text{ss}_m$ (A $\times$ S)

MonadStateSpaceFunctor SetT where
  ss (SetT m) A := $\text{ss}_m$ (Set A)
\end{lstlisting}

[transition] establishes a relationship between the space of actions in the
monad and the (pure) space of transitions on an abstract machine state space.
%
If \textbf{m} is shown to implement MonadStateSpace, then the semantics of our
language can be described as the least fixed point of iterating [transition
step] from e, i.e.:  $\llbracket e \rrbracket = \mu\; x \to e \sqcup
\text{transition}\; x$

To recover the various semantics we are interested in, the monad \textbf{m} can
then be instantiated with following monad transformer stacks:
\begin{lstlisting}
  m1     := EnvStateT (Env Symbol (Set AInt)) Set
  ss m1 A $\approx$ Set (A $\times$ Env Symbol (Set AInt))
  m2     := SetT (EnvState (Env Symbol (Set AInt)))
  ss m2 A $\approx$ Set A $\times$ Env Symbol (Set AInt)
\end{lstlisting}
and inc functions:
\begin{lstlisting}
  inc-concrete : AInt -> Set AInt
  inc-concrete (Exact n) = return (Exact (n + 1))
  inc-concrete (Symbolic IsZero) = return (Symbolic IsPos)
  inc-concrete (Symbolic IsNeg) = return (Symbolic IsNeg) 
                              <+> return (Symbolic IsZero)
  inc-concrete (Symbolic IsPos) = return (Symbolic IsPos)

  inc-abstract : AInt -> AInt -> Set AInt
  inc-abstract (Exact n) | n < -1  = return IsNeg
                         | n == -1 = return IsZero
                         | n > -1  = return IsPos
  inc-abstract (Symbolic IsZero) = return (Symbolic IsPos)
  inc-abstract (Symbolic IsNeg) = return (Symbolic IsNeg) 
                              <+> return (Symbolic IsZero)
  inc-abstract (Symbolic IsPos) = return (Symbolic IsPos)
\end{lstlisting}

When instantiated with \textbf{$\langle$m1, inc-concrete$\rangle$} a
\textit{concrete} interpreter is recovered.  When instantiated with
\textbf{$\langle$m1, inc-abstract$\rangle$} a computable \textit{abstract}
interpreter is recovered.  When instantiated with \textbf{$\langle$m2,
inc-abstract$\rangle$} a computable \textit{abstract} interpreter with
\textit{heap widening} is recovered.

### 2014-08-04 tex/sections.old/04Correctness.tex
initial

\label{section:correctness}

Novel in this work is the justification that a monadic approach to abstract
interpretation can produce sound and complete abstract interpreters by
construction.
%
As is traditional in abstract interpretation, we capture the soundness and
correctness of derived abstract interpreters by establishing a Galois
connections between concrete and abstract state spaces, and by proving an order
relationship between concrete and abstract transition functions.

We remind the reader of the definition of Galois connection:
\begin{lstlisting}
Galois (A B : Type) :=
  $\alpha$ : A $\nearrow$ B
  $\gamma$ : B $\nearrow$ B
  inverses : $\alpha \circ \gamma \sqsubseteq$ id $\sqsubseteq \gamma \circ \alpha$
\end{lstlisting}
where $A \nearrow B$ is written to mean the \textit{monotonic} function space
between A and B, namely [f : A $\nearrow$ B] means [$\exists$ f' : A $\to$ B
$\and$ $\forall$ x y, x $\sqsubseteq$ y $\to$ f x $\sqsubseteq$ f y].

% We further generalize the notion of Galois connection to the Kleisli category:
% \begin{lstlisting}
% KleisliGalois m A B | Monad m :=
%   $\alpha$ : A $\nearrow$ m B
%   $\gamma$ : B $\nearrow$ m A
%   inverses : $\alpha \circ \gamma \sqsubseteq$ id $\sqsubseteq \gamma \circ \alpha$
% \end{lstlisting}
% where id and $\circ$ are interpreted in the Kleisli category.
% 
% [Note why we need the Kleisli generalization.  It's related to wanting to
% distinguish between computable and non-computable functions, and finite and
% infinite sets.]
% 
% Lemma: KleisliGalois m A B $\to$ Galois (A $\nearrow$ m A) (B $\nearrow$ m B).
% %
% (This is the analogue of [Galois A B $\to$ Galois (A $\nearrow$ A) (B $nearrow$
% B)] in the traditional framework.)

Next we introducing the concept of a \textit{functorial Galois connection},
which establishes a Galois connection between functors:
\begin{lstlisting}
FunctorialGalois ($m_1$ $m_2$: Type -> Type) :=
  mapGalois : Galois A B -> Galois ($m_1$ A) ($m_2$ B)
GaloisTransformer (t : (Type -> Type) -> Type -> Type) :=
  liftGalois : FunctorialGalois $m_1$ $m_2$ -> FunctorialGalois (t $m_1$) (t $m_2$)
\end{lstlisting}
and we say (m : Type $\to$ Type) is a GaloisFunctor if it is proper in the
FunctorialGalois relation:
\begin{lstlisting}
GaloisFunctor m := FunctorialGalois m m
\end{lstlisting}

We prove the following lemmas:
\begin{lstlisting}
GaloisTransformer (StateT S)
GaloisTransformer SetT
GaloisFunctor Set
\end{lstlisting}

For a language $L$ and two monads $m_1$ and $m_2$, the concrete and abstract
state spaces in such a setting are $ss_{m_1} L$ and $ss_{m_2} L$, and the step
functions are (transition $step_{m_1}$) and (transition $step_{m_2}$). To
establish the relation (transition $step_{m_1}$) $\sqsubseteq$ (transition
$step_{m_2})$ we first enrich the (MonadStateSpace m) predicate to establish
an isomorphism between m and $ss_m$:
\begin{lstlisting}
MonadStateSpace m :=
  ss : Type -> Type
  transition : $\forall$ A B, Iso (A $\nearrow$ m B) (ss A $\nearrow$ ss B)
\end{lstlisting}
where
\begin{lstlisting}
Iso A B :=
  to : A $\to$ B
  from : B $\to$ A
  inverses : to $\circ$ from = id = from $\circ$ to
\end{lstlisting}

Given this stronger connection between monadic actions and their induced
state space transitions, we can now prove:
\begin{lstlisting}
FunctorialGalois $m_1$ $m_2$
MonadStateSpace $m_1$
MonadStateSpace $m_2$
step : $\forall$ m, L $\nearrow$ m L
----------------------------
Galois ($ss_{m_1}$ L) ($ss_{m_2}$ L)
$transition_{m_1}$ step : $ss_{m_1}$ L $\to$ $ss_{m_1}$ L
$transition_{m_2}$ step : $ss_{m_2}$ L $\to$ $ss_{m_2}$ L
$transition_{m_1}$ step $\sqsubseteq$ $transition_{m_2}$ step
\end{lstlisting}
which establishs the Galois connection between both the concrete and abstract
state spaces and interpreters interpreters induced by $m_1$ and $m_2$.
%
[... fancy diagram that sketches the proof ...]

Going back to our example, the only thing left to prove is that our step
function is monotonic.
%
Because \textbf{step} is completely generic to an underlying monad, we must
enrich \textbf{all} interfaces that we introduced in Section ? to carry
monotonic functions.
%
Generalizing these interfaces, proving that the suppliers of the interfaces
(the monad transformers) meet the monotonicity requirements, and proving that
the client of the interfaces (the step function) are all monotonic is a
systematic exercise in applying the functionality of the \_$\sqsubseteq$\_
relation.
%
We automate the proofs of such theorems using the Coq proof assistant (see
Appendix).
%
However we note that such a property can be shown to hold by construction for
all morphisms in an embedded logic (as can be done for equality) and therefore
it should come as no surprise that if all arrows are changed from $\to$ to
$\nearrow$ then everything ``just works out''.

Once it is established that our step function is monotonic, the proof burden
remains to establish [inc-concrete $\sqsubseteq$ inc-abstract]. [...]

We give a proof that heap widening is sound and less precise than heap cloning
by proving [FunctorialGalois (StateT S Set) (SetT (State S))]. [...]

Now we have an end-to-end proof that:
\begin{lstlisting}
$\text{transition}\; \text{step}_{\langle\text{m}_1, \text{inc-concrete}\rangle}$
$\sqsubseteq$
$\text{transition}\; \text{step}_{\langle\text{m}_1, \text{inc-abstract}\rangle}$
$\sqsubseteq$
$\text{transition}\; \text{step}_{\langle\text{m}_2, \text{inc-abstract}\rangle}$
\end{lstlisting}
or
\begin{lstlisting}
$\langle$concrete semantics$\rangle$
$\sqsubseteq$
$\langle$abstract semantics$\rangle$
$\sqsubseteq$
$\langle$abstract semantics w/heap widening$\rangle$
\end{lstlisting}
%
Proving this uses the facts [FunctorialGalois $m_1$ $m_2$, MonadStateSpace
$m_1$, MonadStateSpace $m_2$]--facts we get for free from the framework through
composition of monad transformers--and [inc-concrete $\sqsubseteq$
inc-abstract]--a proof obligation of the user of the framework.

%%%%%%%%%%%%%%%%%%%% Expressivity and Modularity %%%%%%%%%%%%%%%%%%%%

### 2014-08-04 tex/sections.old/05Modularity.tex
initial

\label{section:modularity}

The Abstracting Abstract Interpreters (AAI) framework developed by Might and
Van Horne provides an axis on which to develop a spectrum of analyses of
varying precision and intensional behavior.
%
The key benefit of their approach is that it generalizes to arbitrary languages
and their state spaces.

This framework provides a separate orthogonal axis on which extensions to
abstract interpretation can be explained, and is similarly modular and language
agnostic.
%
In particular, the techniques of AAI and Monadic Abstract Interpretation can be
combined arbitrarily.

%%%%%%%%%%%%%%%%%%%% Related Work %%%%%%%%%%%%%%%%%%%%

### 2014-08-04 tex/sections/03AAMByExample/00CPS/main.tex
initial

The state space for CPS syntactically separates \h|Call| expressions from \h|Atom|
expressions.
%
\haskell{sections/03AAMByExample/00CPS/00Language.hs}

%--%

\h|Call| expressions are separated from \h|Atom| expressions because they may
not terminate, and therefore cannot be given a striaghtforward denotation.
%
Each syntactic form in \h|Call| is designed to be recursive in only one position,
eliminating the need for a call stack when we design the operational semantics.

%--%

\h|Atom| expressions are exactly those for which we can compute a denotation.
%
The \h|Lam| syntactic form may look troubling because it is mutually recursive
with \h|Call|, but the denotation of \h|Lam xs c| will be a closure which delays
the evaluation of \p|c| until it is applied in a \h|Call| rule.

%--%

Possibly non-standard additions to our CPS language are conditional branching
(\h|IfC|); two literal types, integer and boolean; and primitive operations
\h|Add1|, \h|Sub1| and \h|IsNonNeg|.
%
We include these to bring CPS closer to a language one might use in practice,
and to examine the design space of their abstract semantics.

### 2014-08-04 tex/sections/03AAMByExample/02AbstractStateSpace/00Cut.tex
initial

In the AAM methodology, we first identify recursion in the state space and
replace it with indirection through addresses and a store.
%
\begin{figure}[H]
\haskell{sections/03AAMByExample/02AbstractStateSpace/00Cut.hs}
\caption*{Cutting Recursion}
\end{figure} 
\noindent
%
Abstract time is introduced into the state space as a constantly increasing
value from which to allocate fresh addresses.

%--%

As a result of this change in the state space, the semantics will need to
change accordingly.
%
Variable lookup must fetch an address from the environment, and then fetch the
corresponding value bound to that address from the store.
%
Variable binding must allocate globally fresh addresses from the current time,
binding the formal parameter to the address in the environment and the address
to the value in the store.

### 2014-08-04 tex/sections/03AAMByExample/02AbstractStateSpace/01AAM.tex
initial

Now that recursion is eliminated from the state space we take aim at unbounded
parts of the state space, with the addresses and time that we just introduced
being the first to go.
%
Rather than make a single choice of abstract address and time, we leave this
choice as a parameter.
%
All that is required is:
\begin{itemize}
\item a time \h|tzero| to begin execution
\item a function \h|tick| which moves time forward
\item a function \h|alloc| which allocates an address for binding a formal
      parameter at the current time
\end{itemize}
%
We encode this interface using a Haskell type class \h|AAM| with associated
types \h|Addr| and \h|Time|.
%
\begin{figure}[H]
\haskell{sections/03AAMByExample/02AbstractStateSpace/01AAM.hs}
\caption*{Address and Time Abstraction}
\end{figure}
\noindent
%
An implementation of \h|AAM aam0| for a particular type \p|aam0| must pick two
types, \h|Addr aam0| and \h|Time aam0|, and then implement \h|tzero|, \h|tick|,
\h|alloc|, for those types.
%
The type |aam0| its self will only used as a type level token for selecting
associated types at the type level, and will have a single inhabitant for
selecting type class functions at the value level.
%
(If Haskell had a proper module system this would all be better expressed as a
module interface--containing a type and functions over that type--rather than
an associated type, type class and proxy singleton type.)

### 2014-08-04 tex/sections/03AAMByExample/03AbstractSemantics/main.tex
initial

Now that we have designed an abstract state space--albeit highly
parameterized--we must turn to implementing its semantics.
%
The goal in this entire exercise is to develop the abstract state space and
semantics in such a way that concrete and abstract interpreters can be derived
from a single implementation.

%--%

Following the structure of the concrete semantics, we adapt each function to
the new abstract state space.
%
\h|op| is now entirely implemented by the designer of the analysis as part of
the \h|Delta| interface.
%
\h|atom| must follow another level of indirection in variable lookups, and must
account for nondeterminism.
%
\haskell{sections/03AAMByExample/03AbstractSemantics/00Atom.hs}
%
\h|call| must use the eliminators provided by \h|Delta d| to guide control
flow.
%
\haskell{sections/03AAMByExample/03AbstractSemantics/01Call.hs}
%
\h|bindMany| must join stores when binding to addresses, in case the address is
already in use.
%
\haskell{sections/03AAMByExample/03AbstractSemantics/02BindMany.hs}
%
Finally, \h|step| must advance time as it advances each \h|Call| state, and
\h|exec| is modified to a collecting semantics.
%
\haskell{sections/03AAMByExample/03AbstractSemantics/03Step.hs}
%
The collecting semantics keeps track of all previously seen states.
%
% TODO: reference kleene fixpoint theorem.  also motivate this idea
%       better/earlier
%
Keeping track of previous states turns exec into a monotonic function, which
means its iteration will terminate if the instantiated state space is finite.

### 2014-08-04 tex/sections/03AAMByExample/07Optimizations/02Composition.tex
initial

Heap widening and abstract garbage collection were introduced independently;
what happens when they compose?
%
The result is semantically what you want but exhibits bad behavior
operationally.
%
The naive composition of the optimizations results in each store being garbage
collected independently of each other, which is expensive.
%
What is worse, \textit{they are all the exact same store}, making the extra
effort entirely redundant.

%--%

The solution to this is to either rewrite the semantics to use a widened store
throughout, or to rewrite abstract garbage collection specifically for the
widened state space.
%
Granted, none of these options are terribly difficult in this setting, but
rewriting semantics is a time-consuming and bug-ridden process in general.
%
A more systematic abstraction is clearly needed for these optimizations: one
where optimizations compose seamlessly with both one another and the abstract
semantics.

### 2014-08-04 tex/sections/04MonadicAAM/00MonadicStyle/main.tex
initial

Monadic AAM begins with recognizing that the abstract semantics from
section~\ref{section:AAMByExample:AbstractSemantics} look suspiciously monadic.
%
In this section we will rewrite the exact same function but in monadic style.
%
After the exercise, the monad will be generalized in
section~\ref{section:MonadicAAM:Generalizing} and this generalization will be
exploited in section~\ref{section:MonadicAAM:Optimizations}.

\subsubsection{Nondeterminism}
\input{sections/04MonadicAAM/00MonadicStyle/00Nondeterminism/Z.tex}

\subsubsection{State}
\input{sections/04MonadicAAM/00MonadicStyle/01State/Z.tex}

### 2014-08-04 tex/sections/04MonadicAAM/04Optimizations/00Widen/Z.tex
initial

Widening can be recovered by simply propagating the state transformer
containing the heap to the very bottom of the stack.
%
The state space
%
\haskell{sections/04MonadicAAM/04Optimizations/00Widen/00BeforeSS.hs}
%
induced by the abstract monad
%
\haskell{sections/04MonadicAAM/04Optimizations/00Widen/01BeforeM.hs}
%
must turn into
%
\haskell{sections/04MonadicAAM/04Optimizations/00Widen/02AfterSS.hs}
%
the desired state space for heap widening.
%
We show that this state space is \textit{induced purely by a reordering of the
monads} in \h|A_M|, lifting \h|Nondet| up one level and \h|StateT (S d aam)| to
the bottom.
%
\haskell{sections/04MonadicAAM/04Optimizations/00Widen/03AfterM.hs}
%
This permutation is only valid if the resulting monad adheres to the
\h|Semantics| interface, which we justify next.

%--%

To show that \h|A_Widen_M| implements \h|Semantics| we need to invent a
transformer variant of \h|Nondet|, which we call \h|NondetT|.
%
Rather than a pure list of values, \h|NondetT| carries a list of values inside
a monadic action.
%
\haskell{sections/04MonadicAAM/04Optimizations/00Widen/04NondetT.hs}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% Aside: List Monad Transformer %%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\paragraph{Aside on List Monad Transformers}
{\it
  It is folklore that the naive implementation of \h|ListT m a| $\equiv$ 
  \h|m [a]| is not a monad.
  %
  Variants of \h|ListT m a| have been proposed which are more akin to monadic
  streams of \p|a| values in the monad \p|m|.
  %
  We are able to interpret \h|m [a]| as both a monad and monad plus given extra
  conditions on the underlying monad, and is to our knowledge the first instance
  where \h|m [a]| is given a valid monadic instance.
}

%%%%%%%%%%%%%%%
%% End Aside %%
%%%%%%%%%%%%%%%

The \h|Monad| instance for \h|NondetT| requires the underlying monad to be a
\h|JoinLattice| \textit{functor}.
%
\haskell{sections/04MonadicAAM/04Optimizations/00Widen/05NondetTMonad.hs}
%
\h|NondetT| also propagates the \h|MonadPlus| interface
%
\haskell{sections/04MonadicAAM/04Optimizations/00Widen/06NondetTMonadPlus.hs}
%
the \h|MonadState| interfaces
%
\haskell{sections/04MonadicAAM/04Optimizations/00Widen/07NondetTMonadState.hs}
%
and implements \h|MonadStep|.
%
\haskell{sections/04MonadicAAM/04Optimizations/00Widen/08NondetTMonadStep.hs}

%--%

Given the permuted state space, heap widening is recovered merely by swapping
\h|A_M| with \h|A_Widen_M|.
%
\haskell{sections/04MonadicAAM/04Optimizations/00Widen/09Exec.hs}

### 2014-08-04 src/FP/Core.hs

@@ -1,0 +1,526 @@
+module FP.Core 
+
+-- Exports {{{
+
+  ( module Prelude
+  , module FP.Core
+  , module GHC.Exts
+  , module Data.Set
+  , module Data.Map
+  ) where
+
+-- }}}
+
+-- Imports {{{
+
+import qualified Prelude
+import Prelude 
+  ( Eq(..), Ord(..)
+  , (.), ($), id, const, flip, curry, uncurry
+  , Bool(..), (||), (&&), not, otherwise
+  , Char, Int, Integer, Double, Rational
+  , Maybe(..)
+  , error, undefined
+  , IO
+  )
+import Data.ByteString.Char8 (ByteString)
+import GHC.Exts (type Constraint)
+import qualified Data.ByteString.Char8 as BS
+import Data.Set (Set)
+import qualified Data.Set as Set
+import qualified Data.Map as Map
+import Data.Map (Map)
+
+-- }}}
+
+-------------
+-- Classes --
+-------------
+
+-- Conversion {{{
+
+class ToInteger a where
+  toInteger :: a -> Integer
+class FromInteger a where
+  fromInteger :: Integer -> a
+
+class ToRational a where
+  toRational :: a -> Rational
+class FromRational a where
+  fromRational :: Rational -> a
+
+class ToDouble a where
+  toDouble :: a -> Double
+class FromDouble a where
+  fromDouble :: Double -> a
+
+class FromChars a where
+  fromChars :: Chars -> a
+class ToChars a where
+  toChars :: a -> Chars
+-- for Overlaoded Strings extension
+fromString :: Chars -> String
+fromString = fromChars
+
+class ToString t where
+  toString :: t -> String
+
+-- }}}
+
+-- Arithmetic {{{ --
+
+class Peano a where
+  pzero :: a
+  psuc :: a -> a
+class Additive a where
+  zero :: a
+  (+) :: a -> a -> a
+class (Additive a) => Subtractive a where
+  (-) :: a -> a -> a
+class (Additive a) => Multiplicative a where
+  one :: a
+  (*) :: a -> a -> a
+class (Multiplicative a) => Divisible a where
+  (/) :: a -> a -> a
[…]

### 2014-08-04 src/MAAM.hs

@@ -1,0 +1,9 @@
+module MAAM
+  ( module MAAM.Classes
+  , module MAAM.Common
+  , module MAAM.M
+  ) where
+
+import MAAM.Classes
+import MAAM.Common
+import MAAM.M

### 2014-08-04 src/MAAM/Classes/AAM.hs

@@ -1,0 +1,12 @@
+module MAAM.Classes.AAM where
+
+import FP
+import MAAM.Common
+
+class AAM addr time μ | μ -> addr, μ -> time where
+  tzero :: P μ -> P 𝓁 -> time 𝓁
+  tick :: P μ -> P 𝓁 -> History 𝓁 -> time 𝓁 -> time 𝓁
+  alloc :: P μ -> P 𝓁 -> Name -> time 𝓁 -> addr
+
+-- data T μ 𝓁 = T μ 𝓁
+-- type instance Cell (T μ 𝓁) = Time μ 𝓁

### 2014-08-04 src/MAAM/M.hs

@@ -1,0 +1,10 @@
+module MAAM.M
+  ( module MAAM.M.Syntax
+  , module MAAM.M.Delta
+  , module MAAM.M.Monadic
+  ) where
+
+import MAAM.M.Syntax
+import MAAM.M.Delta
+import MAAM.M.Monadic
+

### 2014-08-04 src/MAAM/M/Delta.hs

@@ -1,0 +1,27 @@
+module MAAM.M.Delta where
+
+import MAAM.M.Syntax
+import MAAM.Common
+import MAAM.Classes
+import FP
+
+class Delta (val :: * -> *) δ | δ -> val where
+  lit :: P δ -> Lit -> val μ
+  clo :: P δ -> [Name] -> Call -> Env addr time μ -> val μ
+  op :: P δ -> Op -> val μ -> Set (val μ)
+  elimBool :: P δ -> val μ -> Set Bool
+  elimClo :: (AAM addr time μ) => P δ -> val μ -> Set ([Name], Call, Env addr time μ)
+
+elimBoolM :: (Delta val δ, MonadZero m, MonadPlus m) => P δ -> val μ -> m Bool
+elimBoolM = msum .: elimBool
+
+elimCloM :: (AAM addr time μ, Delta val δ, MonadZero m, MonadPlus m, Ord addr) => P δ -> val μ -> m ([Name], Call, Env addr time μ)
+elimCloM = msum .: elimClo
+
+type Env addr time μ = Map Name addr
+-- data E μ = E μ
+-- type instance Cell (E μ) = Env μ
+
+type Store val δ addr time μ = Map addr (Set (val μ))
+-- data S δ μ = S δ μ
+-- type instance Cell (S δ μ) = Store δ μ

### 2014-08-04 src/MAAM/M/Monadic.hs

@@ -1,0 +1,157 @@
+module MAAM.M.Monadic where
+
+import FP
+import MAAM.M.Delta
+import MAAM.M.Syntax
+import MAAM.Classes
+import MAAM.Common
+
+data CPS
+type instance History CPS = Call
+cps :: P CPS
+cps = P
+
+type MonadCPS val δ addr time μ m = 
+  ( MonadZero m
+  , MonadPlus m
+  , MonadState (Env addr time μ) m
+  , MonadState (Store val δ addr time μ) m
+  , MonadState (time CPS) m
+  , MonadMaybeE m
+  , MonadStep m
+  , Ord addr
+  , Ord (val μ)
+  )
+
+type Analysis val δ addr time μ m =
+  ( Delta val δ
+  , AAM addr time μ
+  , MonadCPS val δ addr time μ m
+  )
+
+time :: (AAM addr time μ) => P μ -> P 𝓁 -> P (time 𝓁)
+time P P = P
+
+env :: (AAM addr time μ) => P μ -> P (Env addr time μ)
+env P = P
+
+store :: (AAM addr time μ, Delta val δ) => P δ -> P μ -> P (Store val δ addr time μ)
+store P P = P
+
+new :: (Analysis val δ addr time μ m) => P δ -> P μ -> Name -> m addr
+new _ μ x = do
+  t <- getP $ time μ cps
+  return $ alloc μ cps x t
+
+var :: forall val δ addr time μ m. (Analysis val δ addr time μ m) => P δ -> P μ -> Name -> m (Set (val μ))
+var δ μ x = do
+    e <- getP $ env μ
+    s <- getP $ store δ μ
+    useMaybe $ do
+      index s *$ (index e $ x)
+
+bind :: forall val δ addr time μ m. (Analysis val δ addr time μ m) => P δ -> P μ -> Name -> Set (val μ) -> m ()
+bind δ μ x vD = do
+  l <- new δ μ x
+  modifyP (env μ) $ pinsert x l
+  modifyP (store δ μ) $ pinsertWith (\/) l vD
+
+atomic :: (Analysis val δ addr time μ m) => P δ -> P μ -> Atom -> m (Set (val μ))
+atomic δ _   (LitA l) = return $ ssingleton $ lit δ l
+atomic δ μ (Var x) = var δ μ x
+atomic δ μ (Prim o a) = do
+  vD <- atomic δ μ a
+  return $ vD >>~ op δ o
[…]

### 2014-08-04 src/MAAM/M/Syntax.hs

@@ -1,0 +1,25 @@
+module MAAM.M.Syntax where
+
+import FP
+import MAAM.Common
+
+data Lit = I Integer | B Bool
+  deriving (Eq, Ord)
+instance PartialOrder Lit where pcompare = discreteOrder
+data Op = Add1 | Sub1 | IsNonNeg
+  deriving (Eq, Ord)
+instance PartialOrder Op where pcompare = discreteOrder
+data Atom =
+    LitA Lit
+  | Var Name
+  | Prim Op Atom
+  | Lam [Name] Call
+  deriving (Eq, Ord)
+instance PartialOrder Atom where pcompare = discreteOrder
+data Call =
+    If Atom Call Call
+  | App Atom [Atom]
+  | Halt Atom
+  deriving (Eq, Ord)
+instance PartialOrder Call where pcompare = discreteOrder
+

### 2014-08-04 src/Main.hs

@@ -1,0 +1,7 @@
+module Main where
+
+import FP
+import MAAM ()
+
+main :: IO ()
+main = print "<>"

### 2014-08-04 tex/sections/03AAMByExample/00CPS/00Language.hs

@@ -1,0 +1,5 @@
+type Name = String
+data Lit = I Integer | B Bool
+data Op = Add1 | Sub1 | IsNonNeg
+data Atom = LitA Lit | Var Name | Prim Op Atom | Lam [Name] Call
+data Call = If Atom Call Call | App Atom [Atom] | Halt Atom

### 2014-08-04 tex/sections/03AAMByExample/01Concrete/00SS.hs

@@ -1,0 +1,3 @@
+type Env = Map Name Val
+data Val = LitV Lit | Clo [Name] Call Env
+type StateSpace = Maybe (Call, Env)

### 2014-08-04 tex/sections/03AAMByExample/01Concrete/01Op.hs

@@ -1,0 +1,6 @@
+op :: Op -> Val -> Maybe Val
+op Add1 (LitV (I n)) = Just $ LitV $ I $ n+1
+op Sub1 (LitV (I n)) = Just $ LitV $ I $ n-1
+op IsNonNeg (LitV (I n)) | n >= 0 = Just $ LitV $ B True
+                         | otherwise = Just $ LitV $ B False
+op _ _ = Nothing

### 2014-08-04 tex/sections/03AAMByExample/01Concrete/02Atom.hs

@@ -1,0 +1,7 @@
+atom :: Env -> Atom -> Maybe Val
+atom _ (LitA l) = Just $ LitV l
+atom e (Var x) = mapLookup x e
+atom e (Prim o a) = case atom e a of
+  Nothing -> Nothing
+  Just v -> op o v
+atom e (Lam xs c) = Just $ Clo xs c e

### 2014-08-04 tex/sections/03AAMByExample/01Concrete/03Call.hs

@@ -1,0 +1,13 @@
+call :: Env -> Call -> Maybe (Call, Env)
+call e (If a tc fc) = case atom e a of
+  Just (LitV (B True)) -> Just (tc, e)
+  Just (LitV (B False)) -> Just (fc, e)
+  _ -> Nothing
+call e (App fa xas) =
+  case atom e fa of
+    Just (Clo xs c e') ->
+      case bindMany xs xas e' of
+        Nothing -> Nothing
+        Just e'' -> Just (c, e'')
+    _ -> Nothing
+call e (Halt a) = Just (Halt a, e)

### 2014-08-04 tex/sections/03AAMByExample/01Concrete/04BindMany.hs

@@ -1,0 +1,7 @@
+bindMany :: [Name] -> [Atom] -> Env -> Maybe Env
+bindMany [] [] e = Just e
+bindMany (x:xs) (xa:xas) e = 
+  case (atom xa e, bindMany xs xas e) of
+    (Just xv, Just e') -> Just (mapInsert x xv e')
+    _ -> Nothing
+bindMany _ _ _ = Nothing

### 2014-08-04 tex/sections/03AAMByExample/02AbstractStateSpace/02NonDet.hs

@@ -1,0 +1,2 @@
+type Store aam = Map (Addr aam) (Set (Val aam))
+type StateSpace aam = Set (Call, Env aam, Store aam, Time aam)

### 2014-08-04 tex/sections/03AAMByExample/02AbstractStateSpace/03Delta.hs

@@ -1,0 +1,16 @@
+class AAM aam where
+  type Addr (aam :: *) :: *
+  type Time (aam :: *) :: *
+  tzero :: aam -> Time aam
+  tick :: aam -> Call -> Time aam -> Time aam
+  alloc :: aam -> Name -> Time aam -> Addr aam
+class Delta d where
+  type Val (d :: *) :: * -> *
+  lit :: d -> Lit -> Val d aam
+  clo :: d -> [Name] -> Call -> Env aam -> Val d aam
+  op :: d -> Op -> Val d aam -> Maybe (Val d aam)
+  elimBool :: d -> Val d aam -> Set Bool
+  elimClo :: d -> Val d aam -> Set ([Name], Call, Env aam)
+type Env aam = Map Name (Addr aam)
+type Store d aam = Map (Addr aam) (Set (Val d aam))
+type StateSpace d aam = Set (Call, Env aam, Store d aam, Time aam)

### 2014-08-04 tex/sections/03AAMByExample/03AbstractSemantics/02BindMany.hs

@@ -1,0 +1,11 @@
+bindMany :: (AAM aam) => aam 
+         -> Time aam -> [Name] -> [Atom] 
+         -> Env aam -> Store d aam 
+         -> Maybe (Env aam, Store d aam)
+bindMany _   _ [] [] e s = Just (e, s)
+bindMany aam t (x:xs) (xa:xas) e s = case bindMany aam t xs xas e s of
+  Nothing -> Nothing
+  Just (e', s') ->
+    let l = alloc aam x t
+    in Just (mapInsert x l e', mapInsertWith (\/) l xv s')
+bindMany _ _ _ _ _ _ = Nothing

### 2014-08-04 tex/sections/03AAMByExample/04RecoveringConcrete/01Delta.hs

@@ -1,0 +1,16 @@
+data C_Delta = C_Delta
+data ConcreteVal aam = LitV Lit | Clo [Name] Call (Env aam)
+instance Delta C_Delta where
+  type Val C_Delta = ConcreteVal
+  lit C_Delta l = Lit l
+  clo C_Delta xs c e = Clo xs c e
+  op C_Delta Add1 (LitV (I i)) = Just $ LitV $ I $ i+1
+  op C_Delta Sub1 (LitV (I i)) = Just $ LitV $ I $ i-1
+  op C_Delta IsNonNeg (LitV (I i)) 
+    | i >= 0 = Just $ LitV $ B $ True
+    | otherwise = Just $ LitV $ B $ False
+  op C_Delta _ _ = Nothing
+  elimBool C_Delta (LitV (B b)) = setSingleton b
+  elimBool C_Delta _ = setEmpty
+  elimClo C_Delta (Clo xs c e) = setSingleton (xs, c, e)
+  elimClo C_Delta _ = setEmpty

### 2014-08-04 tex/sections/03AAMByExample/04RecoveringConcrete/02Exec.hs

@@ -1,0 +1,2 @@
+execConcrete :: Call -> StateSpace C_Delta C_AAM
+execConcrete = exec C_Delta C_AAM

### 2014-08-04 tex/sections/03AAMByExample/05Recovering0CFA/00AAM.hs

@@ -1,0 +1,7 @@
+data ZCFA_AAM = ZCFA_AAM
+instance AAM ZCFA_AAM where
+  type Time ZCFA_AAM = ()
+  type Addr ZCFA_AAM = Name
+  tzero ZCFA_AAM = ()
+  tick ZCFA_AAM _ () = ()
+  alloc ZCFA_AAM x () = x

### 2014-08-04 tex/sections/03AAMByExample/05Recovering0CFA/02Exec.hs

@@ -1,0 +1,2 @@
+execSymZCFA :: Call -> StateSpace Sym_Delta ZCFA_AAM
+execSymZCFA = exec Sym_Delta ZCFA_AAM

### 2014-08-04 tex/sections/03AAMByExample/06RecoveringKCFA/00AAM.hs

@@ -1,0 +1,8 @@
+data kCFA_AAM = kCFA_AAM Int
+instance AAM kCFA_AAM where
+  type Time kCFA_AAM = [Call]
+  type Addr kCFA_AAM = (Name, [Call])
+  tzero (kCFA_AAM k) = []
+  tick (kCFA_AAM k) c t = take k (c:t)
+  alloc (kCFA_AAM k) x t = (x, t)
+

### 2014-08-04 tex/sections/04MonadicAAM/00MonadicStyle/00Nondeterminism/01Nondet.hs

@@ -1,0 +1,5 @@
+type Nondet = []
+instance Monad Nondet where
+  return x = [x]
+  [] >>= k = []
+  (x:xs) >>= k = k x ++ (xs >>= k)

### 2014-08-04 tex/sections/04MonadicAAM/00MonadicStyle/00Nondeterminism/02NondetMonadPlus.hs

@@ -1,0 +1,3 @@
+instance MonadPlus [] where
+  mzero = []
+  mplus = (++)

### 2014-08-04 tex/sections/04MonadicAAM/00MonadicStyle/01State/00MonadState.hs

@@ -1,0 +1,3 @@
+class MonadState s m where
+  get :: m s
+  put :: s -> m ()

### 2014-08-04 tex/sections/04MonadicAAM/00MonadicStyle/01State/01State.hs

@@ -1,0 +1,6 @@
+newtype StateT s m a = StateT { runStateT ::  s -> m (a,s) }
+instance (Monad m) => Monad (StateT m) where
+  return x = StateT $ \ s -> return (x, s)
+  xM >>= k = StateT $ \ s -> do
+    (x,s') <- runStateT xM s
+    runStateT (k x) s'

### 2014-08-04 tex/sections/04MonadicAAM/00MonadicStyle/01State/02StateMonadState.hs

@@ -1,0 +1,3 @@
+instance (Monad m) => MonadState s (StateT s m) where
+  get = StateT $ \ s -> return (s,s)
+  put s = StateT $ \ _ -> return ((),s)

### 2014-08-04 tex/sections/04MonadicAAM/00MonadicStyle/01State/04MonadStateTagged.hs

@@ -1,0 +1,5 @@
+class MonadStateTagged s m where
+  type Cell s :: *
+  get :: s -> m (Cell s)
+  put :: s -> Cell s -> m ()
+

### 2014-08-04 tex/sections/04MonadicAAM/00MonadicStyle/01State/05StateTagged.hs

@@ -1,0 +1,1 @@
+newtype StateT s m a = StateT { runStateT ::  Cell s -> m (a,Cell s) }

### 2014-08-04 tex/sections/04MonadicAAM/00MonadicStyle/01State/06MonadStateInterior.hs

@@ -1,0 +1,8 @@
+instance (MonadState s m) => MonadState s (StateT s' m) where
+  get = StateT $ \ s -> do
+    s' <- get
+    return (s',s)
+  put s' = StateT $ \ s -> do
+    put s'
+    return ((),s)
+

### 2014-08-04 tex/sections/04MonadicAAM/00MonadicStyle/01State/07M.hs

@@ -1,0 +1,7 @@
+data E aam = E aam
+type instance Cell (E aam) = Env aam
+data S d aam = S d aam
+type instance Cell (S d aam) = Store d aam
+data T aam = T aam
+type instance Cell (T aam) = Time aam
+type M d aam = StateT (E aam) (StateT (S d aam) (StateT (T aam) []))

### 2014-08-04 tex/sections/04MonadicAAM/00MonadicStyle/01State/09Call.hs

@@ -1,0 +1,10 @@
+call :: (Delta d, AAM aam) => d -> aam -> Call -> M d aam Call
+call d aam (If a tc fc) = do
+  v <- atom a
+  b <- setMSum (elimBool d v)
+  return $ if b then tc else fc
+call d aam (App fa xas) =
+  v <- atom d aam fa
+  (xs,c,e') <- setMSum $ elimClo d v
+  mFromMaybe $ bindMany aam xs xas e'
+call _ _ (Halt a) = return $ Halt a

### 2014-08-04 tex/sections/04MonadicAAM/00MonadicStyle/01State/10StateSpace.hs

@@ -1,0 +1,1 @@
+type StateSpace d aam = Set (Call, Env aam, Store d aam, Time aam)

### 2014-08-04 tex/sections/04MonadicAAM/00MonadicStyle/01State/11Step.hs

@@ -1,0 +1,9 @@
+step :: (Delta d, AAM aam) => d -> aam 
+     -> StateSpace d aam -> StatSpace d aam
+step d aam ss = eachInSet ss $ \ (c,e,s,t) -> setFromList $ do
+  (((c',e'),s'),t') <-
+    flip runStateT e $
+    flip runStateT s $
+    flip runStateT t $
+    call d aam c
+  return (c',e',s',tick aam c' t)

### 2014-08-04 tex/sections/04MonadicAAM/01Generalizing/00Semantics.hs

@@ -1,0 +1,9 @@
+type Semantics d aam m =
+  ( Delta d
+  , AAM aam
+  , Monad m
+  , MonadPlus m
+  , MonadState (E aam) m
+  , MonadState (S d aam) m
+  , MonadState (T aam) m
+  )

### 2014-08-04 tex/sections/04MonadicAAM/01Generalizing/01Atom.hs

@@ -1,0 +1,1 @@
+atom :: (Semantics d aam m) => d -> aam -> Atom -> m (Val d aam)

### 2014-08-04 tex/sections/04MonadicAAM/01Generalizing/02Call.hs

@@ -1,0 +1,1 @@
+call :: (Semantics d aam m) => d -> aam -> Call -> m Call

### 2014-08-04 tex/sections/04MonadicAAM/01Generalizing/03MonadStep.hs

@@ -1,0 +1,3 @@
+class (Pointed (SS m)) => MonadStep m where
+  type SS m :: * -> *
+  step :: (a -> m b) -> SS m a -> SS m b

### 2014-08-04 tex/sections/04MonadicAAM/01Generalizing/04Pointed.hs

@@ -1,0 +1,2 @@
+class Pointed t where
+  point :: a -> t a

### 2014-08-04 tex/sections/04MonadicAAM/01Generalizing/05NondetMonadStep.hs

@@ -1,0 +1,5 @@
+instance Pointed [] where
+  point = return
+instance MonadStep NonDet where
+  type SS NonDet = []
+  step f xs = xs >>= f

### 2014-08-04 tex/sections/04MonadicAAM/01Generalizing/07Exec.hs

@@ -1,0 +1,2 @@
+exec :: (Semantics d aam m) => d -> aam -> Call -> SS m Call
+exec d aam c0 = iter (collect (step $ call d aam)) $ point c0

### 2014-08-04 tex/sections/04MonadicAAM/02RecoveringConcrete/00Point.hs

@@ -1,0 +1,1 @@
+data Point a = Top | Bot | Point a

### 2014-08-04 tex/sections/04MonadicAAM/02RecoveringConcrete/01PointMonad.hs

@@ -1,0 +1,5 @@
+instance Monad Point where
+  return = Point
+  Top >>= k = Top
+  Bot >>= k = Bot
+  Point a >>= k = k a

### 2014-08-04 tex/sections/04MonadicAAM/02RecoveringConcrete/02PointMonadPlus.hs

@@ -1,0 +1,7 @@
+instance MonadPlus Point where
+  mzero = Bot
+  mplus Bot yM = yM
+  mplus xM Bot = xM
+  mplus Top yM = Top
+  mplus xM Top = Top
+  mplus (Point x) (Point y) = Top

### 2014-08-04 tex/sections/04MonadicAAM/02RecoveringConcrete/03PointMonadStep.hs

@@ -1,0 +1,5 @@
+instance Pointed Point where
+  point = Point
+instance MonadStep Point where
+  type SS Point = Point
+  step f xs = xs >>= f

### 2014-08-04 tex/sections/04MonadicAAM/02RecoveringConcrete/04ConcreteMonad.hs

@@ -1,0 +1,1 @@
+type C_M d aam = StateT (E aam) (StateT (S d aam) (StateT (T aam) Point))

### 2014-08-04 tex/sections/04MonadicAAM/03RecoveringKCFA/00AbstractMonad.hs

@@ -1,0 +1,2 @@
+type A_M d aam = StateT (E aam) (StateT (S d aam) (StateT (T aam) Nondet))
+

### 2014-08-04 tex/sections/04MonadicAAM/04Optimizations/00Widen/00BeforeSS.hs

@@ -1,0 +1,1 @@
+[(Call, Env aam, Store d aam, Time aam)]

### 2014-08-04 tex/sections/04MonadicAAM/04Optimizations/00Widen/01BeforeM.hs

@@ -1,0 +1,1 @@
+type A_M d aam = StateT (E aam) (StateT (S d aam) (StateT (T aam) Nondet))

### 2014-08-04 tex/sections/04MonadicAAM/04Optimizations/00Widen/02AfterSS.hs

@@ -1,0 +1,2 @@
+[((Call, Env aam, Time aam), Store d aam)]
+

### 2014-08-04 tex/sections/04MonadicAAM/04Optimizations/00Widen/03AfterM.hs

@@ -1,0 +1,1 @@
+type A_Widen_M d aam = StateT (E aam) (StateT (T aam) (NondetT (State (S d aam))))

### 2014-08-04 tex/sections/04MonadicAAM/04Optimizations/00Widen/04NondetT.hs

@@ -1,0 +1,1 @@
+newtype NondetT m a = NondetT { runNondetT :: m [a] }

### 2014-08-04 tex/sections/04MonadicAAM/04Optimizations/00Widen/07NondetTMonadState.hs

@@ -1,0 +1,7 @@
+instance (MonadState s m) => MonadState s (NondetT m) where
+  get = NondetT $ do
+    s <- get
+    return [s]
+  put s = NondetT $ do
+    put s
+    return [()]

### 2014-08-04 tex/sections/04MonadicAAM/04Optimizations/00Widen/08NondetTMonadStep.hs

@@ -1,0 +1,8 @@
+instance (Pointed m) => Pointed (NondetT m) where
+  point x = NondetT $ point [x]
+instance (MonadStep m) => MonadStep (NondetT m) where
+  type SS (NondetT m) = NondetT m
+  step (f :: a -> NondetT m b) xM = 
+    NondetT $ 
+    step (\ xs -> runNondetT $ NondetT (return xs) >>= f) $ 
+    runNondetT xM

### 2014-08-04 tex/sections/04MonadicAAM/04Optimizations/01AbstractGC/00Free.hs

@@ -1,0 +1,1 @@
+free :: Call -> Env aam -> Set (Addr aam)

### 2014-08-04 tex/sections/04MonadicAAM/04Optimizations/01AbstractGC/01Reachable.hs

@@ -1,0 +1,1 @@
+reachable :: Store d aam -> Set (Addr aam) -> Set (Addr aam)

### 2014-09-17 writeup/sections/03AAMByExample/04RecoveringConcrete/main.tex
writing final version of things...

The final delta abstraction resulted in a highly factorized state space,
amenable to a wide range of analyses.
%
Before manipulating these axes of parameterization, we first show how to
recover a concrete interpreter.

%--%

Our semantics has two axes of parameterization: \h|AAM| and \h|Delta|.
%
For \h|AAM| we choose integers for abstract time and integers paired with
variable names for abstract addresses.
%
\haskell{sections/03AAMByExample/04RecoveringConcrete/00AAM.hs}

%--%

For \h|Delta| we choose the value type and primitive operations from
section~\ref{section:AAMByExample:Concrete}.
%
\haskell{sections/03AAMByExample/04RecoveringConcrete/01Delta.hs}

%--%

Running the abstract semantics from
section~\ref{section:AAMByExample:AbstractSemantics} with these two
parameterizations--\h|C_AAM| and \h|C_Delta|--recovers the concrete
(collecting) semantics.
%
\haskell{sections/03AAMByExample/04RecoveringConcrete/02Exec.hs}

### 2014-09-17 writeup/sections/03AAMByExample/07Optimizations/01AbstractGC.tex
writing final version of things...

Abstract garbage collection has become standard in structural abstract
interpretation and a staple in the AAM methodology.
%
The idea is analogous to garbage collection for memory safe languages.
%
Even though we must accept that abstract addresses can be reused during
execution, it would be nice if we could completely replace the contents of an
abstract location if we knew no future state would need the old value.
%
This concept--of future states depending on values in the store--can be
approximated by the \textit{reachability} of those values in the current state.
%
Abstract garbage collection is the process where unreachable allocations in the
store are removed so that future allocations don't require joins in the store.

%--%

To perform abstract garbage collection we must examine all addresses in the
store and eliminate those which cannot be reached from a \textit{root} in the
state space.
%
The roots $\mathcal{T}(c,e)$ in the state space are the locations bound to free
variables in $c$:
%
\begin{equation*}
\mathcal{T}(c,e) = \{e(v) : v \in free(c)\}
\end{equation*}
%
The reachable addresses $\mathcal{R}(c,e,s,t)$ are those which are reachable
from the roots, which is the least solution to the equation:
%
\begin{equation*}
\mathcal{R}(c,e,s,t) = \mathcal{T}(c,e) 
                  \cup \{ \ell' 
                        : \ell \in \mathcal{R}(c,e,s,t)
                        , \ell \leadsto_{\langle e,s \rangle} \ell'
                       \}
\end{equation*}
where $\leadsto_{\langle e,s \rangle}$ is the single-step points-to relation
in environment $e$ and store $s$.
%
The set of addresses to eliminate is the complement of $\mathcal{R}(c,e,s,t)$

### 2014-09-17 writeup/sections/04MonadicAAM/00MonadicStyle/01State/Z.tex
writing final version of things...

To add state to our monad we apply a state monad transformer on top of the
current nondeterminism monad.
%
Monad transformers are compositional building blocks for building larger
monads.
%
At the interface level, monad transformers consume monads and produce monads;
they are mappings from monads to monads--monad morphisms.
%
Certain monad transformers are known to compose well with others.
%
In our case, nondeterminism and state are known to compose well (in this order
only), meaning the resulting monad will implement both state and nondeterminism
actions.

%--%

As we did with the nondeterminism monad and its interface \h|MonadPlus|, we
introduce an interface for state called \h|MonadState|.
%
\h|MonadState s m| means the monad \p|m| has access to a single cell of state
\p|s|.
%
\haskell{sections/04MonadicAAM/00MonadicStyle/01State/MonadState.hs}

%--%

The state monad transformer contains a function from state values to a monadic
actions of state-result pairs.
%
\haskell{sections/04MonadicAAM/00MonadicStyle/01State/01State.hs}
%
and implements the \h|MonadState| interface
%
\haskell{sections/04MonadicAAM/00MonadicStyle/01State/02StateMonadState.hs}

%--%

The benefit of monad transformers, and the reason why we bother to declare
interfaces for individual monads, is that monad transformer stacks have the
ability to expose the interfaces of inner monads to the fully composed outer
monad.
%
For example, \h|Nondet| has been shown to implement \h|MonadPlus|, but it is
also the case that the transformer stack \h|StateT s NonDet| \textit{also}
implements \h|MonadPlus|.
%
This lifting, or inheritance, of monadic interface can be expressed as a type
class instance rule.
%
\haskell{sections/04MonadicAAM/00MonadicStyle/01State/03StateMonadPlus.hs}

%--%

Our state space has \textit{three} state components, but the \h|MonadState|
interface only encodes one cell of state.
%
One solution is to combine each component of the state into a tuple, but this
would limit our ability to abstract each component in different ways.
%
A better solution is to allow multiple states in one stack, with a way of
selecting which state is desired.
%
The method for selecting the desired state can take many forms:
1) separate type classes could be defined for each state, 2) type singleton
tags (like the singletons used to fake modules in
section~\ref{section:AAMByExample}) could be defined to select the state, or 3)
we could require that each state is of unique type and allow type inference to
infer the desired state.
%
We take the approach of 2)--introducing type tags to select the desired state.
%
This approach adds a little more mechanism and type-trickery, but is much less
work than the 1) and avoids other newtype ``hacks'' required to use 3).

%--%

We redefine the \h|MonadState| class to use the \p|s| class parameter as a tag,
rather than the type of the state cell its self.
%
The actual type of the state is an associated type of the tag.
%
\haskell{sections/04MonadicAAM/00MonadicStyle/01State/04MonadStateTagged.hs}
%
The state type changes accordingly to store \h|Cell s|.
%
\haskell{sections/04MonadicAAM/00MonadicStyle/01State/05StateTagged.hs}
%
(The \h|Monad|, \h|MonadState| and \h|MonadPlus| instance are analagous.)

%--%

Now that we have a tagged state monad, we allow lifting of interior state
interfaces to the outside of a transformer stack.
%
\haskell{sections/04MonadicAAM/00MonadicStyle/01State/06MonadStateInterior.hs}

%--%

We now rewrite the semantics to use a four-layer monad transformer stack, which
we abbreviate to \h|M|.
%
The tags for the environment, store and time components of the monad are called
\h|E|, \h|S| and \h|T| and defined as type family instances.
%
\haskell{sections/04MonadicAAM/00MonadicStyle/01State/07M.hs}
%
The \h|atom| and \h|call| functions are then simplified to monadic actions.
%
\haskell{sections/04MonadicAAM/00MonadicStyle/01State/08Atom.hs}
%
\haskell{sections/04MonadicAAM/00MonadicStyle/01State/09Call.hs}

%--%

We are not finished until we also transform \h|step| to work with our new
monadic type.
%
The purpose of \h|step| is to translate the nondetermistic function \h|call|
into a state space transition.
%
We take this approach to heart and maintain a separation between the monadic
type and the type of its state space transitions.
%
Because the state space is separate from the monad, we reuse the same state
space from before.
%
\haskell{sections/04MonadicAAM/00MonadicStyle/01State/10StateSpace.hs}
%
The \h|step| function now converts between the monadic action 
\h|Call -> M Call| and state space transitions \h|SateSpace -> StateSpace|.
%
\haskell{sections/04MonadicAAM/00MonadicStyle/01State/11Step.hs}

### 2014-09-17 writeup/sections/04MonadicAAM/04Optimizations/01AbstractGC/Z.tex
writing final version of things...

With the semantics in monadic style, abstract garbage collection can be
implemented once--generic to the underlying monad--integrating seamlessly into
both heap-cloning and heap-widening semantics.

%--%

To implement garbage collection we write a monad-generic function that crawls
the expression, environment and store, and replaces the store with a copy
containing only reachable addresses.
%
This implementation captures the generic garbage collection algorithm, and
scales to both heap-cloning and heap-widening stacks.
%
In fact, it's the exact algorithm used to garbage collect concrete semantics,
and behaves as concrete garbage collection when instantiated to concrete
semantics parameters.

%--$

Detecting free variables is a simple recursive function on the \h|Call|
expression in the state space.
%
\haskell{sections/04MonadicAAM/04Optimizations/01AbstractGC/00Free.hs}

%--%

Collecting the set of all reachable addresses demands a fixpoint computation,
iterating from the free variables as roots.
%
\haskell{sections/04MonadicAAM/04Optimizations/01AbstractGC/01Reachable.hs}

%--%

Abstract garbage collection replaces the existing heap with the abstract
garbage collected heap.
%
\haskell{sections/04MonadicAAM/04Optimizations/01AbstractGC/02GC.hs}

### 2014-09-17 writeup/sections/05Correctness/00AbstractSemantics/main.tex
writing final version of things...

In order to obtain Galois connections for our abstract interpreters, it must
first be established that \h|step| is monotonic.
%
Because \h|step| is highly parameterized, we must place monotonicity conditions
on these parameters in order to arrive at a final monotonicity argument for
\h|step|.

%--%

We remind the reader that the parameters to the \h|step| function are:
%
\begin{itemize}
\item The \h|Delta d| parameter, which specifies the value type and semantics of
      primitive operations.
\item The \h|AAM aam| parameter, which specifies abstract time and address types
      and the semantics of allocation.
\item The \h|Monad m| parameter, which specifies the computational properties
      of the abstraction, like nondeterminism or heap widening.
\end{itemize}
%
To justify that \h|step| is monotonic, we require \textit{all} base types in
the parameters (like \h|Val d|, \h|Time aam| and \h|Addr aam|) to come equipped
with partial orders, and \textit{all} functions provided by the the parameters
to be monotonic.
%
We must also clarify that we are using the discrete order for all base types
(\h|Bool|, \h|Name|, etc.), and the standard subset inclusion ordering for
\h|Set| and \h|Point| types.

%--%

\newtheorem{atom-call-step-monotonic}{Lemma}
\begin{atom-call-step-monotonic}
  \h|atom|, \h|call| and \h|step| are monotonic.
\end{atom-call-step-monotonic}
\begin{proof}
  Follows trivially from the assumed monotonicity of \h|Delta|, \h|AAM| and
  \h|Monad| operations.
\end{proof}

### 2014-09-18 writeup/darais-quals.tex
prepping new draft writeup, hoping to give to greg tomorrow morning.

\documentclass{article}

\input{setup.tex}

\title{Abstract Control in Static Analysis}
\author{David Darais}
\date{\today}

\begin{document}
\maketitle

\begin{abstract} % {{{
\end{abstract} % }}}

\tableofcontents

% Introduction {{{
\section{Introduction}
\label{section:Introduction}

Static analyses have two components: 
%
\begin{itemize}
\item A \emph{computational artifact} which computes a set of facts about the execution of a program.
\item A \emph{proof of correctness} about the computational artifact.
\end{itemize}

We are motivated by static analysis techniques which are:
%
\begin{itemize}
\item Compartmental: Various aspects of the analysis have been separated from each other and are understood in isolation.
\item Modular: Design choices in one aspect of the analysis do not restrict design choices of another.  This property is important for both computational and correctness components of an analysis.
\item Language Agnostic: A given analysis technique can be seamlessly transfered from one semantics to another.
\item Correct: The proof of correctness should exist, and there should be a framework for establishing it.
\item Derived: Either the computational artifact or the correctness proof should be obtained "for free" from the other.
\end{itemize}

Our framework compartmentalizes:
%
\begin{itemize}
\item Abstract domain a la Cousot.
\item Abstract time and address a la Van Horn and Might.
\item Intensional optimizations a la Van Horn and Might. 
\item Object-sensitivity a la Smaragdakis.
\item Analysis control properties.
\end{itemize}

Our framework is:
%
\begin{itemize}
\item Compartmental: Each of the above concerns are separated and independent.
\item Modular: Design choices in one axis is completely independent of others
\item Language Agnostic: All axis other than abstract domain are fully language agnostic.
\item Correct: A correctness framework is described and each axis is proven correct in isolation.
\item Derived: The correctness of an analysis using our framework comes for free.
\end{itemize}

Analysis control properties are the primary topic of this work.
%
We merely note that the other axis in the design space are implemented in our framework using existing techniques.

Contributions:
%
\begin{itemize}
\item Computational artifacts and correctness proofs for the axis of abstract control, which captures the choice of flow and path sensitivity.
\item Computational artifacts and correctness proofs for intensional optimizations (gargabe collection and mcfa) independent of abstract control.
\item Language independent proofs of refinement for flow-sensitivity and path-sensitivity choices using the monadic abstraction.
\end{itemize}

% }}}

% Background {{{
\section{Background}
\label{section:Background}

% Notation {{{
\subsection{Notation}
\label{section:Background:Notation}

We use traditional notation for function definitions and applications rather than lambda calculus notation.
%
For example, we will define [f(w, x) := g(y, z)] rather than [f w x := g y z].

We use lambda notation for anonymous functions.
%
For example, defining the identify function could be notated [id := λ(x) → x] rather than the named definition [id(x) := x].

% }}}

% Monads {{{
\subsection{Monads}
\label{section:Background:Monads}

A basic familiarity with the concept of monads will be useful for understanding the rest of the paper.
%
Some find it useful to review the concept of functor before learning about monads. 
%
We adopt this approach for our brief explanation of monads.

A type [𝒯 : Set → Set] is called a functor if one can:
%
\begin{itemize}
\item define [map : ∀ (A : Set), (A → B) → 𝒯(A) → 𝒯(B)]
\item prove [map id = id], where id is the identity function [λ x → x]
\item prove [map (g ∘ f) = map g ∘ map f]
\end{itemize}

Example: Lists are functors, where map is defined:
%
    map(f)(xs) := case xs 
      Nil → Nil 
      Cons(x,xs') → Cons(f(x),map(f)(xs'))

Example: [map(isEven)([1, 2, 3, 4]) = [False, True, False, True]]

Likewise a type [𝒯 : Set → Set] is called a monad if one can:
%
\begin{itemize}
\item define [extend : ∀ (A : Set), (A → 𝒯(B)) → 𝒯(A) → 𝒯(B))]
\item prove unit and associativity laws (not mentioned here).
\end{itemize}

Example: Lists are monads, where extend is defined:
%
    extend(f)(xs) := case xs
      Nil → Nil
      Cons(x,xs') → f(x) ++ extend(f)(xs')
%
[extend] for lists can be understood as a map followed by a concat or flattening operation.

Example: [extend(λ x → [x + 1, x - 1])([10, 100]) = [9, 11, 99, 101]

[extend] for lists can also be understood as concat followed by map:
%
      extend(f)([10, 100]) 
    = concat(map(f)([10, 100])) 
    = concat([[9, 11], [99, 101]]) = [9, 11, 99, 101]
%
For all monads, it is equivalent to define return, map and join, where [join :
∀ (A : Set), 𝒯(𝒯(A)) → 𝒯(A)].
%
As we have seen, join is just concat for the list monad.

% }}}

% Small Step Semantics {{{
\subsection{Small-Step Semantics}
\label{section:Background:SmallStepSemantics}

This work follows the small-step operational approach to language semantics pioneered by ? and popularized by ?.
%
In small-step operational semantics, the meaning of an expression ℯ in a language is given by the transitive closure of a step relation [ℯ ↦ ℯ'].
%
This yields a state-machine approach to semantics.
%
Exploring the meaning of the expression means exploring all possible transitions reachable from ℯ through ↦.

% }}}

% Abstract Interpretation {{{
\subsection{Abstract Interpretation}
\label{section:Background:AbstractInterpretation}

Abstract Iterpretation (AI) is a formal framework for program analysis pioneered by Cousot et Cousot.
%
In the setting of abstract interpretation, a program analysis is an alternate semantics over an abstract domain.

both a discipline for constructing program analyses
and a framework for reasoning about their correctness pioneered by Cousot and
Cousot. Van Horn and Might have extended this tradition in their work on
Abstracting Abstract Machines, showing how to systematically construct analysis
for small step operational semantics.
%
Relating an analysis to a semantics requires a bridge.  In general, such a
relationship will be stated as:
%
∀ (c c': C) (a a': A), c R a ∧ c ↦ c' ∧ a ↦ a' ⇒  c' R a'
%
In words, this is saying that for a concrete program c and a valid abstraction
of it a, if the program can take a step to c' and the abstraction can take a
step to a', then a' is a valid abstraction of c'.
%
Once this theorem is established, the transitive closure of it follows
trivially and justifies performing an analysis by:
%
1: abstracting a program c to it its abstraction a
2: running a using the transitive closure of a ↦* a'
3: inspecting the final a' to infer behavior of running c
%
The question now is how to set up R--how to relate one language and semantics
to another. For this, abstract interpretation turns to order theory.  We say c
and a are related if c is in the image of projecting a to 𝒫(c).  Using this
projection, commonly called γ, we restate the theorem:
%
∀ (c c': C) (a a': A), c ∈ γ(a) ∧ c ↦ c' ∧ a ↦ a' ⇒  c' ∈ γ(a')
%
However, the technique generalizes to arbitrary partial orders, and is done so
profitably in the literature.  Stating the theorem with partial orders gives:
%
∀ (c c': C) (a a': A), c ⊑ γ(a) ∧ c ↦ c' ∧ a ↦ a' ⇒  c' ⊑ γ(a')
%
The partial order for concrete domains is almost always the inclusion order on
𝒫(C), while the partial order for abstract domains is almost always a finite
element lattice.
%
A complete treatment of abstract interpretation is centered around constructing
a _galois connection_ between a concrete and abstract semantics. Establishing a
galois connection between C and A establishes both sides of the relationship
between C and A, whereas until now we've only had one direction.
Having both direction allows us to directly project concrete objects to
abstract objects, and provides a setting to discuss if γ is a "best"
abstraction function relating A to C.
%
A galois connection between two posets C and A is notated [C α⇄ γ A] and
consists of:
%
* [α : C → A] where α is monotonic
* [γ : A → C] where γ is monotonic
* [γ ∘ α] is expansive: ∀ (x : C), x ⊑ γ(α(x))
* [α ∘ γ] is contractive: ∀ (y : A), α(γ(y)) ⊑ y
%
The last two points can be succinctly stated as [α ∘ γ ⊑ id ⊑ γ ∘ α], using the
logical monotonicity relation [f ⊑ g ⇔  (x ⊑ y ⇒  f(x) ⊑ g(y))] for the
function space.
%
Equivalent to the above four rules is:
%
* x ⊑ γ(y) ⇔  α(x) ⊑ y
%
Example: given a galois connection (C α⇄ γ A), there exists a galois connection
(C → C) α⇄ γ (A → A) where:
%
α(f : C → C) = γ ∘ f ∘ α
γ(g : A → A) = α ∘ g ∘ γ
%
Using the constructed galois connection on the function space, we can more
succinctly state the correctness of an analysis as:
%
_↦_ ⊑ _↦ₐ_
%
Example: the language (𝒫(ℤ),+,*) forms a galois connection with the language
(𝒫({EVEN, ODD}),∧,∨) where:
%
α(zs : 𝒫(ℤ)) = ⋃ { { EVEN | ∃ z ∈ zs ∧ Even(z)}, { ODD | ∃ z ∈ zs ∧ Odd(z) } }
γ(ts : 𝒫({EVEN, ODD})) = ⋃ { { z ∈ ℤ | EVEN ∈ ts ∧ Even(z) }, { z ∈ ℤ | ODD ∈ ts ∧ Odd(z) } }
%
α and γ must have the expansive and contractive properties in order
for this to be considered a galois connection.  The expansive property
corresponds to the soundness of the abstraction.  Soundness ensures that if
you abstract and then project back to the concrete, you must cover the points
you started with.  The contractive property corresponds to the tightness of
the abstraction.  Tightness ensures that the abstraction is the most precise
abstraction possible.  For example, picking [α(zs) = {EVEN, ODD}] would be
sound, because projecting back to the concrete gives ℤ which tautologically
covers all integers, but this choice of α is clearly not tight.  Our choice
of α and γ in this example are indeed both sound and tight.
%
In this example, the concrete language is (ℤ,+,*) and the abstract language is
({EVEN,ODD},∧,∨)
%
Using the lifting of galois connections to the function space, we can/should
also justify _+_ ⊑ γ(_∧_) and _*_ ⊑ γ(_∨_)
%
% }}}

% Monadic AAM {{{

\section{Monadic AAM}
\label{section:MonadicAAM}

Monadic AI (Sergey et al. PLDI 2013) showed that computational modularity for
abstract interpreters can be achieved through a monadic abstraction.  We build
directly on this work at the level of intuition--monads serve as our
pivot-point for generalizing abstract control. However, in our pursuit of an
axis-independent correctness framework, our approach deviates greatly from this
work in the specific monadic interfaces used.

We use CPS to simplify presentation, although our implementation supports a
larger scheme-like language.

First we demonstrate a simple, fully instantiated instance of our framework
_without_ monads. This example is the instantiation of:

* A simple abstract domain which tracks control flow, booleans and integer
  constants.
* 0-cfa
* No abstract garbage collection or copied closurse (see mcfa)
* Fully flow- and path-sensitive, which is in some sense the most natural
  structural abstraction of the concrete state space (see AAM).

x,y,k : Var  ::= ..variables..
b,i,l : Lit  ::= ℤ ⋃ 𝔹
a     : Atom ::= x | l | op a | λ x . c | λ x k . c
op    : Op   ::= add1 | sub1 | gez
c     : Call ::= if(a){c}{c} | a(a) | a(a,a) | halt(a)

v  : ^Val^ ::= INT | l | <λ x . c> | <λ x k . c>

σ  : ^Store^ := Var ⇀  𝒫 (^Val^)

𝒜 [_](_) : (^Store^ × Atom) ⇀  𝒫 (^Val^)
𝒜 [σ](x)                      := σ(x)
𝒜 [σ](l)                      := { l }
𝒜 [σ](add1 a) | 𝒜 [ρ,σ] ⊑ INT := { INT }
𝒜 [σ](sub1 a) | 𝒜 [ρ,σ] ⊑ INT := { INT }
𝒜 [σ](gez a)  | 𝒜 [ρ,σ] ⊑ INT := { T , F }
𝒜 [σ](λ x k . c)              := { <λ x k . c> }

(this is often stated as a relation in the general setting of small-step
relational semantics.)

𝒞 <_,_> : (Call × ^Store^) ⇀  𝒫 (Call × ^Store^)
𝒞 <if(a){c₁}{c₂},σ> := { <c,σ> |  c  ∈ ⋃ [ { c1 } if T ∈ 𝒜 [σ](a), { c2 } if F ∈ 𝒜 [σ](a) ] }
𝒞 <a₁(a₂,a₃),σ>     := { <c,σ'> 
                       | <λ x k . c>  ∈ 𝒜 [σ](a₁)
                       |          v₂  ∈ 𝒜 [σ](a₂)
                       |          v₃  ∈ 𝒜 [σ](a₃)
                       |          σ' := σ ⊔ [x ↦ v₂] ⊔ [k ↦ v₃]
                       }

The complete analysis of a program c is defined as:

μ(Σ) . {<c,⊥>} ⊔ Σ ⊔ 𝒞*(Σ)

where

𝒞* : 𝒫 (Call × ^Store^) → 𝒫 (Call × ^Store)
𝒞*(Σ) = ⋃ [ 𝒞 <c,σ> | <c,σ> ∈ Σ ]

The monadic conversion of the above analysis is:

ℳ  : Set → Set
ℳ  a := StateT ^Store^ List a
ℳ  a := ^Store^ → 𝒫 (a × ^Store^)

𝒜 ₘ : Atom → ℳ (^Val^)
𝒜 ₘ(x) = do
  σ ← get-σ
  return σ(x)
𝒜 ₘ(l) = return {l}
𝒜 ₘ(add1 a) = do
  vD ← 𝒜 ₘ (a)
  return {INT | ∃ i ⊑ INT ∈ vD}
𝒜 ₘ(sub1 a) = do
  vD ← 𝒜 ₘ (a)
  return {INT | ∃ i ⊑ INT ∈ vD}
𝒜 ₘ(gez a) = do
  vD ← 𝒜 ₘ (a)
  return {T, F | ∃ i ⊑ INT ∈ vD}
𝒜 ₘ(λ x k . c) = return <λ x k . c>

𝒞ₘ : Call → ℳ (Call)
𝒞ₘ(if(a){c₁}{c₂}) := do
  vD ← 𝒜 ₘ (a)
  v ← lift𝒫 (vD)
  b ← coerce𝔹 (v)
  return { c₁ if b | c₂ if ¬b }
𝒞ₘ(a₁(a₂,a₃)) := do
  vD ← 𝒜 ₘ (a₁)
  v ← lift𝒫 (vD)
  <λ x k . c> <- coerceλ₂ v
  vD₂ ← 𝒜 ₘ (a₂)
  vD₃ ← 𝒜 ₘ (a₃)
  v₂ ← lift𝒫 vD₂
  v₃ ← lift𝒫 vD₃
  modify-σ (λ σ . σ ⊔ [x ↦ v₂] ⊔ [k ↦ v₃])
  return c

The monadic abstraction provides a nice way to simplify the implementation of
the analysis.  In particular, cases which do not modify the store or only
return one result--both of which are instances of having "no effect"--need not
mentioned all members of the state machine.  (Credit Sergey et. al.)

The complete analysis is defined as before where

𝒞ₘ* : 𝒫 (Call × ^Store^) → 𝒫 (Call × ^Store^)
𝒞ₘ*(Σ) = ⋃ [ (𝒞ₘ c)(σ) | <c,σ> ∈ Σ ]

The exact abstract interpreter written directly is recovered through the
definition of ℳ .  Our insight is twofold:

* ℳ  can be _axiomatized_, such that different definitions for ℳ  give rise to
  different control abstractions for the analysis.
* The axiomatized monadic abstraction provides for a modular proof framework
  for the correctness of a range of analyses.

% }}}

% Abstract Control {{{

\section{Abstract Control}
\label{section:AbstractControl}

The current instantiation of ℳ  yields a flow-sensitive path-sensitive
analysis.  Consider a reordering of the List and State monad transformers:

ℳ  : Set → Set
ℳ  a := SetT (State ^Store^) a
ℳ  a := ^Store^ → 𝒫 (a) × ^Store^

[Note: The definition of SetT we use is non-traditional.  However, our
definition for SetT is actually a monad, whereas traditional definitions are
not.  Our definition of SetT and proofs of monad laws are in the
correctness/proofs section.]

As before, we must convert between monadic actions and abstract state space
transitions to achieve an analysis:

𝒞ₘ* : 𝒫 (Call) × ^Store^ → 𝒫 (Call) × ^Store^
𝒞ₘ*(c𝒫,σ) = <{c' | c' ∈ π₁(𝒞ₘ c)(σ) | c ∈ c𝒫 }, ⋃ [ π₂(𝒞ₘ c)(σ) | c ∈ c𝒫 ]>

This instantiation of ℳ  and 𝒞ₘ* yields a flow-insensitive analysis.

For the same monad ℳ , we can change the definition of 𝒞ₘ* to achieve a
flow-sensitive path-insensitive analysis:

𝒞ₘ* : 𝒫 (Call × ^Store^) → 𝒫 (Call × ^Store^)
𝒞ₘ*(Σ) = { <c',π₂(𝒞 ₘ c)(σ)> | c' ∈ π₁(𝒞ₘ c)(σ) | c ∈ c𝒫 }

In the correctness framework, the flow-sensitive path-sensitive analysis and
flow-insensitive analyses are justified through isomorphisms between monadic
actions and abstract state space transitions.  The flow-sensitive
path-insensitive variant is recovered by weakening this isomorphism to a galois
connection.

% }}}

% Intensional Optimizations {{{

\section{Intensional Optimizations}
\label{section:IntensionalOptimizations}

We have now factored the abstract control properties of static analysis behind
a common interface.  Now we show how to implement two intentional
optimizations, abstract garbage collection and mcfa, in a completely general
setting.

The common interface for each variation in abstract control is a functor ℳ  ∈
Set → Set such that ℳ :

* Is a monad
  * return : ∀ (A : Set), A → ℳ  A
  * extend : ∀ (A β : Set), (A → ℳ  β) → (ℳ  A → ℳ  β)
* Is a state monad algebra on ^Store^
  * get : ℳ  ^Store^
  * put : ^Store^ → ℳ  1
* Is a monad plus
  * _+ₘ_ : ∀ (A : Set), ℳ  A → ℳ  A → ℳ  A

We abreviate the conjunction of the above properties holding on some ℳ  as
(Analysis ℳ ).

Abstract garbage collection can now be written once, generic to this interface:

gc : Call → ℳ  1
gc(c) = do
  σ ← get-σ
  touched₀ ← touched-𝒞(c)
  let touched := collect(touched-var)(touched₀)
  modify-σ (λ σ . ⋃ [ {x ↦ v} | σ(x) = v ∧ x ∈ touched ])

The mcfa optimization of copying rather than linking closures also enjoys a
fully generic implementation:

clo-copy : ([Var] × Call) → ℳ  Lam
clo-copy(xs,c) = do
  let ys := free-vars(c)
  σ ← get-σ
  vs ← map* σ(_) ys
  ρ ← { y ↦* v | (y,v) ∈ zip(ys, vs) }
  return <xs,c,ρ>

Using existing techniques, these optimizations must be proved correct for each
instantiation of abstract control.  Our generalization over abstract control
allows us to implement each of these optimizations once, and more importantly,
the proof about these optimizations also transfer to their instantiations for
free.

% }}}

% Correctness Framework {{{

\section{Correctness}
\label{section:Correctness}

The key advantage to our formulation of monadic abstract interpretation--both
over non-monadic approaches like AAM, and the monadic approach described by
Sergey et al. (PLDI 2013)--is that the proofs of correctness for the
constructed analyses can be derived automatically alongside the computational
artifact.

Our monadic abstraction is a proper abstraction in that it mediates between two
parties--the framework and the analysis designer--and gives guarantees to both
sides conditioned on assumptions being met.

We show that language-independent monads exist for instantiating a semantics to
flow-sensitive and path-sensitive analyses.

In the core analysis framework we prove:

* SetT and StateT are monad transformers and compose in both directions.
* Monadic actions (A → ℳ  β) in SetT and StateT form galois connections with
  abstract state machine transitions (𝒮𝒮ₘ A → 𝒮𝒮ₘ β).
* One composition of SetT and StateT is in (⊑) relation with the other.
* For (SetT ∘ StateT), one choice of state machine 𝒮𝒮ₘ is in (⊑) relation with
  the other.

The analysis designer must then prove:

* The semantic step function 𝒞ₘ (which calls 𝒜 ₘ) is monotonic, including the
  choice of 𝓂 .
* The semantics as written, including intensional optimizations, are correct.

Given proofs from both sides of the interface we can conclude:

flow-insensitive 0-cfa 
 ⊑ 
flow-sensitive path-insensitive 0-cfa 
 ⊑ 
flow-sensitive 0-cfa
 ⊑ 
concrete cfa

This proof technique is maximally modular, and greatly reduces the proof burden
on the analysis designer as new languages and analyses are developed.

When adding the rest of the analysis tuning knobs like abstract domain, context
sensitivity, intensional optimizations and object sensitivity, the monotonicity
of 𝒞ₘ must be re-established.  However, like for abstract control, galois
connections can be built between these choices independent of their
instantiation in 𝒞ₘ. Our implementation supports each of these axis in full.

% }}}

% Proofs and Definitions {{{

\section{Proofs and Definitions}
\label{section:Proofs+Definitions}

% Setup {{{

\subsection{Setup}
\label{section:Proofs+Definitions:Setup}

PartialOrder(A : Set) :=
  operators
    _⊑_ : A → A → Prop
  laws
    reflexivity : x ⊑ x
    antisymmetry : x ⊑ y ⇒  y ⊑ x ⇒  x = y
    transitivity : x ⊑ y ⇒  y ⊑ z ⇒  x ⊑ z

Monotonic(𝑓 : A → β) := 
  require
    PartialOrder(A)
    PartialOrder(β)
    ∀ (x y : A), x ⊑ y ⇒  f(x) ⊑ f(y)

(C : Set) α⇄ γ (A : Set) :=
  require
    PartialOrder(C)
    PartialOrder(A)
  operators
    α : C → A
    γ : A → C
  laws
    Monotonic(α)
    Monotonic(γ)
    α ∘ γ ⊑ id ⊑ γ ∘ α

JoinSemilattice(A : Set) :=
  require
    PartialOrder(A)
  operators
    ⊥ : A
    _⊔_ : A → A → A
  laws
    left-unit : ⊥ ⊔ x = x
    right-unit : x ⊔ ⊥ = x
    associativity : x ⊔ (y ⊔ z) = (x ⊔ y) ⊔ z
    commutativity : x ⊔ y = y ⊔ x

Monad(𝓂  : Set → Set) :=
  operators
    return : ∀ (A : Set) → A → 𝓂 (A)
    extend : ∀ (A β : Set) → (A → 𝓂 (β)) → (𝓂 (A) → 𝓂 (β))
  laws
    left-unit : extend(return) = id
    right-unit : extend(k)(return(A)) = k(A)
    associativity : extend(k₂)(extend(k₁)(aM)) = extend(λ x → extend(k₂)(k₁(x)))(A)

Transformer(𝒯 : (Set → Set) → (Set → Set)) :=
  require
    ∀ (𝓂  : Set), Monad(𝓂 ) ⇒  Monad(𝒯(𝓂 ))
  operators
    lift : ∀ (𝓂  : Set), 𝓂 (A) → 𝒯(𝓂 )(A)
  laws
    unit : lift(return) = return

MonadState(𝓈 : Set)(𝓂  : Set → Set) :=
  operators
    get : 𝓂  𝓈
    put : 𝓈 → 𝓂  1

MonadPlus(𝓂  : Set → Set) := 
  require
    ∀ (A : Set), JoinSemilattice(𝓂 (A))
  laws
    left-zero : extend(k)(⊥) = ⊥
    right-zero : extend(const(⊥))(aM) = ⊥
    distributivity : extend(k)(aM₁ ⊔ aM₂) = extend(k)(aM₁) ⊔ extend(k)(aM₂)

Functorial(ℭ : Set → Prop)(𝓂  : Set → Set) := 
  require
    ∀ (A : Set), ℭ (A) → ℭ (𝓂 (A))

Pointed(𝓂  : Set → Set) :=
  unit : ∀ (A : Set) → A → 𝓂 (A)

MonadSmallStep(𝓂  : Set → Set)(𝒮𝒮 : Set → Set) := 
  require
    ∀ (A β : Set), (A → 𝓂 (β)) α⇄ γ (𝒮𝒮(A) → 𝒮𝒮(β))

% }}}

% SetT {{{

\subsection{SetT}
\label{section:Proofs+Definitions:SetT}

SetT : (Set → Set) → (Set → Set)
SetT(𝓂 )(A) := 𝓂 (𝒫 (A))

-----------------
Transformer(SetT)

lift : ∀ (𝓂  : Set → Set) (A : Set), 𝓂 (A) → SetT(𝓂 )(A)
lift := map (λ x → { x })

Monad(𝓂 ) ∧ Functorial(JoinSemilattice)(𝓂 )
-------------------------------------------
Monad(SetT(𝓂 ))

return : ∀ (A : Set), A → SetT 𝓂  A
return := lift ∘ returnₘ

extend : ∀ (A β : Set), (A → SetT(𝓂 )(β)) → (SetT(𝓂 )(A) → SetT(𝓂 )(β))
extend(k) := joinₘ ∘ mapₘ (joins ∘ mapₚ k)

Monad(𝓂 ) ∧ Functorial(JoinSemilattice)(𝓂 )
-------------------------------------------
MonadPlus(SetT(𝓂 ))

m0 : ∀ (A : Set), SetT 𝓂  A
m0 = ⊥ ₘ

_<+>_ : ∀ (A : Set) → SetT 𝓂  A → SetT 𝓂  A → SetT 𝓂  A
_<+>_ := _⊔_

Monad(𝓂 ) ∧ MonadState(𝓈)(𝓂 )
-----------------------------
MonadState(𝓈)(𝓂 )

get : SetT 𝓂  𝓈
get = lift getₘ

put : 𝓈 → SetT 𝓂  1
put = lift ∘ putₘ

MonadSmallStep(𝓂 )(𝒮𝒮) ∧ Functorial(JoinSemilattice)(𝓂 ) ∧ Functor(𝒮𝒮)
-------------------------------
MonadSmallStep(SetT 𝓂 )(𝒮𝒮 ∘ 𝒫)

α : ∀ (A β : Set), (A → SetT(𝓂 )(β)) → (𝒮𝒮(𝒫 (A)) → 𝒮𝒮(𝒫 (β)))
α(f) = αₘ(joinsₘ ∘ mapₚ(f))

γ : ∀ (A β : Set), (𝒮𝒮(𝒫 (A)) → 𝒮𝒮(𝒫 (β))) → (A → SetT(𝓂 )(β))
γ(f) = γₘ(f ∘ mapₛₛ(unitₚ))

MonadSmallStep(𝓂 )(𝒮𝒮) ∧ (𝒫 ∘ 𝒮𝒮) α⇄ γ (𝒮𝒮 ∘ 𝒫 )
-------------------------------------
MonadSmallStep(SetT 𝓂 )(𝒫 ∘ 𝒮)

α : ∀ (A β : Set), (A → SetT(𝓂 )(β)) → (𝒫 (𝒮𝒮(A) → 𝒫 (𝒮𝒮(β))))
α(f) = extendₚ(α ∘ αₘ(f))

γ : ∀ (A β : Set), (𝒫 (𝒮𝒮(A) → 𝒫 (𝒮𝒮(β)))) → (A → SetT(𝓂 )(β))
γ(f) = extendₚ(γ ∘ γₘ(f))

% }}}

% StateT {{{

\subsection{StateT}
\label{section:Proofs+Definitions:StateT}

StateT : Set → (Set → Set) → (Set → Set)
StateT(𝓈)(𝓂 )(A) := 𝓈 → 𝓂  (A × 𝓈)

Transformer(StateT(𝓈))
----------------------

lift : ∀ (𝓂  : Set → Set) (A : Set), 𝓂  a → StateT(𝓈)(𝓂 )(A)
lift aM := λ 𝓈 → mapₘ (,𝓈) aM

Monad(𝓂 )
--------------------
Monad(StateT(𝓈)(𝓂 ))

return : ∀ (A : Set), A → StateT(𝓈)(𝓂 )(A)
return := lift ∘ returnₘ

extend : ∀ (A β : Set), (A → StateT(𝓈)(𝓂 )(β)) → (StateT(𝓈)(𝓂 )(A) → StateT(𝓈)(𝓂 )(β))
extend(k)(aM) := λ 𝓈 → doₘ
  (a,𝓈') ← aM(𝓈)
  k(a)(𝓈')

JoinSemilattice(𝓈) ∧ Monad(𝓂 )
------------------------------------------
Functorial(JoinSemilattice)(StateT(𝓈)(𝓂 ))

bot : StateT(𝓈)(𝓂 )(A)
bot = λ _ → (⊥ ₛ, ⊥ ₐ)

_⊔_ : StateT(𝓈)(𝓂 )(A) → StateT(𝓈)(𝓂 )(A) → StateT(𝓈)(𝓂 )(A)
aM₁ ⊔ aM₂ := λ 𝓈 → doₘ
  (a₁,𝓈₁) ← aM₁
  (a₂,𝓈₂) ← aM₂
  return (a₁ ⊔ a₂, 𝓈₁ ⊔ 𝓈₂)

MonadPlus(𝓂 )
------------------------
MonadPlus(StateT(𝓈)(𝓂 ))

⊥ : ∀ (A : Set), StateT(𝓈)(𝓂 )(A)
⊥ := lift ⊥ ₘ

_⊔_ : ∀ (A : Set), StateT(𝓈)(𝓂 )(A)
aM₁ ⊔ aM₂ := λ 𝓈 → aM₁(𝓈) <+> aM₂(𝓈)

Monad(𝓂 )
----------------------------
MonadState(𝓈)(StateT(𝓈)(𝓂 ))

get : StateT 𝓈 𝓂  𝓈
get = λ 𝓈 → return (𝓈,𝓈)

put : 𝓈 → StateT 𝓈 𝓂  1
put 𝓈 = λ _ → return (∙,𝓈)

MonadSmallStep(𝓂 )(𝒮𝒮)
--------------------------------------
MonadSmallStep(StateT(𝓈)(𝓂 ))(𝒮𝒮(_ × 𝓈))

α : ∀ (A B : Set), (A → StateT(𝓈)(𝓂 )(B)) → (𝒮𝒮(A × 𝓈) → 𝒮𝒮(B × 𝓈))
α(f) = αₘ (λ (a,𝓈) → f(a)(𝓈))

γ : ∀ (A B : Set), (𝒮𝒮(A × 𝓈) → 𝒮𝒮(B × 𝓈)) → (A → StateT(𝓈)(𝓂 )(B))
γ(f) = λ a 𝓈 → γₘ (f)(a,s)

% }}}

% Propositions {{{

\section{Propositions}
\lable{section:Proofs+Definitions:Propositions}

* SetT and StateT are Galois Functors.
* 𝒜 ₘ and 𝒞ₘ are monotonic (including in 𝓂 ).

% }}}

% }}}

\bibliography{davdar}{}
\bibliographystyle{plain}

\end{document}

### 2014-09-18 writeup/old/sections/03AAMByExample/01Concrete/main.tex
prepping new draft writeup, hoping to give to greg tomorrow morning.

To give \h|Call| concrete semantics we first design a state space.
%
The state space introduces an environment to track variable bindings, and a
value type containing closures (which mutually close over environments).
%
\haskell{sections/03AAMByExample/01Concrete/00SS.hs}
%
The state space is partial (uses the \h|Maybe| type) and \h|Nothing| is the meaning
of failed computations.
%
Computations fail when an expression is ill-typed--for example if a literal
flows to function application position--or if a function application is of the
wrong arity.

%--%

The semantics for \h|Op| are given denotationally, and are straightforward.
%
\haskell{sections/03AAMByExample/01Concrete/01Op.hs}

%--%

The semantics for \h|Atom| are given denotationaly.
%
\haskell{sections/03AAMByExample/01Concrete/02Atom.hs}
%
Literals evaluate to themselves.
%
Variables evaluate to a value retrieved from the environment.
%
Lambdas evaluate immediately to closures which capture their environment.

%--%

The semantics for \h|Call| are given \textit{operationally} as a small step
function.
%
\haskell{sections/03AAMByExample/01Concrete/03Call.hs}
%
Conditional statements branch on boolean values.
%
Applications step to a function's body and closure environment with argument
values bound to formal parameters.
%
Termination is signaled with the \h|Halt| command.
%
Helper function \h|bindMany xs xas e| evaluates the function arguments \p|xas| and
binds them to the formal parameters \p|xs| in the environment \p|e|.
%
\haskell{sections/03AAMByExample/01Concrete/04BindMany.hs}
%
\h|bindMany xs xas e| fails if evaluating any argument in \p|xas| fails or if
there is an arity mismatch.

%--%

The full semantics of \h|Call| are given by the transitive closure of \h|step|.
%
\haskell{sections/03AAMByExample/01Concrete/05Step.hs}

### 2014-09-18 writeup/old/sections/03AAMByExample/02AbstractStateSpace/02NonDet.tex
prepping new draft writeup, hoping to give to greg tomorrow morning.

Following the AAM methodology, we realize that in a world with finite adresses
but potentially infinite executions, there is a chance we will need to reuse an
existing address.
%
This causes us to introduce a powerset both around the state space and into the
domain of the heap, as a single address may now point to multiple values, and
variable reference may return multiple values.
%
\begin{figure}[H]
\haskell{sections/03AAMByExample/02AbstractStateSpace/02NonDet.hs}
\caption*{Introducing Nondeterminism}
\end{figure}
\noindent
%
As a result of this abstraction, the semantics must consider multiple control
paths where before there was only one.

### 2014-09-18 writeup/old/sections/03AAMByExample/02AbstractStateSpace/03Delta.tex
prepping new draft writeup, hoping to give to greg tomorrow morning.

Now we have a \textit{nondterministic} semantics paramterized by abstract
addresses--but we are not yet finished.
%
The state space is still infinite due to the unboundedness of integer literal
values.
%
(Even if we took the semantics of finite machine integers, $2^{32}$ is still much
too large a state space to be exploring in a practical analysis.)
%
To curb this we introduce yet another axis of parameterization with the intent
of recovering multiple analyses from the choice later.
%
In the Haskell implementation we represent this axis with the type class
\h|Delta| and associated type \h|Val|.
%
\begin{figure}[H]
\haskell{sections/03AAMByExample/02AbstractStateSpace/03Delta.hs}
\caption*{(Final) Delta Abstraction}
\end{figure}
\noindent
%
Different analyses may want to make different choices in their value
representations, so we don't commit to a particular \h|Val| type, rather we
require that \textit{some} \h|Val| type exist with a particular interface.
%
All that matters is that we have \textit{introduction} forms for literals and
closures, a \textit{denotation function} for primitive operations, and
\textit{elimination} forms for booleans and closures.
%
Eliminators are only required for booleans and closures because those are the
only values scrutenized in the control flow of the semantics.

### 2014-09-18 writeup/old/sections/03AAMByExample/02AbstractStateSpace/main.tex
prepping new draft writeup, hoping to give to greg tomorrow morning.

To arrive at a computable analysis we must abstract the infinite concrete state
space with a finite abstract state space.
%
As we sketch the abstract state space in the AAM methodology we will hint at
corrosponding changes that need to take place in the semantics.
%
The full abstract semantics for the final abstract state space will be given in
section~\ref{section:AAMByExample:AbstractSemantics}.

%--%

\subsubsection{Cutting Recursion}
\input{sections/03AAMByExample/02AbstractStateSpace/00Cut.tex}

\subsubsection{Address and Time Abstraction}
\input{sections/03AAMByExample/02AbstractStateSpace/01AAM.tex}

\subsubsection{Introducing Nondeterminism}
\input{sections/03AAMByExample/02AbstractStateSpace/02NonDet.tex}

\subsubsection{Delta Abstraction}
\input{sections/03AAMByExample/02AbstractStateSpace/03Delta.tex}

### 2014-09-18 writeup/old/sections/03AAMByExample/05Recovering0CFA/main.tex
prepping new draft writeup, hoping to give to greg tomorrow morning.

To recover 0CFA we use variables themselves as abstract addresses and ignore
the time component of the state space.
%
\haskell{sections/03AAMByExample/05Recovering0CFA/00AAM.hs}

%--%

Although the \h|ZCFA_AAM| parameter captures the behavior of 0CFA, we do not
yet have a computable analysis--we need a finite instantiation for the
\h|Delta| parameter.
%
We pick a simple \h|Delta| instance for now, but note that it is not inherent
to 0CFA, it merely serves as \textit{some} abstraction for literals and
primitive operations.
%
\haskell{sections/03AAMByExample/05Recovering0CFA/01Delta.hs}
%
We call this \h|Sym_Delta| because it treats literals purely symbolically,
essentially carrying around only their type information.

%--%

Running the abstract semantics from
section~\ref{section:AAMByExample:AbstractSemantics} with these two
parameterizations--\h|ZCFA_AAM| and \h|Sym_Delta|--recovers a 0CFA analysis.
%
\haskell{sections/03AAMByExample/05Recovering0CFA/02Exec.hs}

### 2014-09-18 writeup/old/sections/03AAMByExample/07Optimizations/00Widen.tex
prepping new draft writeup, hoping to give to greg tomorrow morning.

The state space we arrived at from our methodological abstraction of the
concrete state space in section ~\ref{section:AAMByExample:AbstractStateSpace}
was 
%
\begin{center}
\h|StateSpace d aam = Set (Call, Env aam, Store d aam, Time aam)|
\end{center}
%
This state space will lead to a fully flow sensitive analysis, and consequently
an exponential runtime.
%
The fix is well known in the literature as \textit{heap widening}, where the
desired state space is 
%
\begin{center}
\h|StateSpace d aam = (Set (Call, Env aam, Time aam), Store d aam)|.
\end{center}
%
When widening the heap, each store in the set of states is joined together to
form a single universally approximating store.
%
This results in a less precise but vastly more efficient abstraction.

%--%

Introducing heap widening into the analysis can be accomplished in two ways:
rewriting all the semantics for the new state space, or by introducing a
post-facto widening operation that operates on the original state space.
%
We demonstrate the latter approach, which can be understood as a Galois
connection between the inefficient and efficient state spaces.
%
\begin{center}
\h|Set (Call, Env, Store, Time)| 
\galois{\alpha}{\gamma} 
\h|(Set (Call, Env, Time), Store)|
\end{center}
%
We show only the $\alpha$ portion of the Galois connection and call it
\h|widen|.
%
\haskell{sections/03AAMByExample/07Optimizations/00Widen.hs}

### 2014-09-18 writeup/old/sections/04MonadicAAM/00MonadicStyle/00Nondeterminism/Z.tex
prepping new draft writeup, hoping to give to greg tomorrow morning.

Before introducing the nondeterminism monad, we first describe its interface,
\h|MonadPlus|.
%
\h|MonadPlus| abstractly encodes the notion of nondeterminism.
%
\haskell{sections/04MonadicAAM/00MonadicStyle/00Nondeterminism/00MonadPlus.hs}
%
\h|mzero| encodes failure and \h|mplus| encodes nondeterministic choice.

%--%

The nondeterminism monad is built with the list datatype
%
\haskell{sections/04MonadicAAM/00MonadicStyle/00Nondeterminism/01Nondet.hs}
%
and implements the \h|MonadPlus| interface:
%
\haskell{sections/04MonadicAAM/00MonadicStyle/00Nondeterminism/02NondetMonadPlus.hs}
%
For example, \h|[1,3] >>= \ x -> [x, x+1]| and \h|[1,3] `mplus` [2,4]| both
evaluate to \h|[1,2,3,4]|.

%--%

The \h|atom| denotation function from
section~\ref{section:03AAMByExample:AbstractSemantics} can be refactored into
the nondeterminism monad by replacing \h|setSingleton| with \h|return|,
\h|eachInSet| with \h|(>>=)|, and \h|setEmpty| with \h|mzero|.
%
\haskell{sections/04MonadicAAM/00MonadicStyle/00Nondeterminism/03Atom.hs}

%--%

\h|call| can be factored similarly.
%
\haskell{sections/04MonadicAAM/00MonadicStyle/00Nondeterminism/04Call.hs}

### 2014-09-18 writeup/old/sections/04MonadicAAM/02RecoveringConcrete/main.tex
prepping new draft writeup, hoping to give to greg tomorrow morning.

To recover concrete semantics we can reuse the concrete parameters \h|C_Delta|
and \h|C_AAM|, and need only design a concrete monad to instantiate the
semantics.

%--%

When implementing concrete semantics in section
\ref{section:AAMByExample:RecoveringConcrete} we were forced to reuse the
\h|Set| type even though the semantics were fully deterministic.
%
Because we have abstracted the semantics over a monad, we are now able to
choose an appropriately precise domain for execution.
%
A more appropriate domain would be the \h|Point| type, where the meaning of
\h|Point a| is the extension of \p|a| with explicit \h|Top| and \h|Bot| values.
%
\haskell{sections/04MonadicAAM/02RecoveringConcrete/00Point.hs}

%--%

Point can be shown to be a monad:
%
\haskell{sections/04MonadicAAM/02RecoveringConcrete/01PointMonad.hs}
%
and even a monad plus:
%
\haskell{sections/04MonadicAAM/02RecoveringConcrete/02PointMonadPlus.hs}
%
and finally a monad step:
%
\haskell{sections/04MonadicAAM/02RecoveringConcrete/03PointMonadStep.hs}

%--%

Our final monad stack for recovering concrete semantics will be
%
\haskell{sections/04MonadicAAM/02RecoveringConcrete/04ConcreteMonad.hs}
%
and execution of the concrete semantics can be recovered by
%
\haskell{sections/04MonadicAAM/02RecoveringConcrete/05Exec.hs}

### 2014-09-18 writeup/old/sections/04MonadicAAM/03RecoveringKCFA/main.tex
prepping new draft writeup, hoping to give to greg tomorrow morning.

We skip the recovery of 0CFA in the monadic semantics and head straight to
kCFA.
%
The only thing that changes from the monadic point of view is the arbiter of
nondeterminism at the bottom of the transformer stack.
%
Rather than \h|Point| we use the \h|Nondet| monad as originally introduced in
section~\ref{section:MonadicAAM:MonadicStyle}.

%--%

kCFA can be recovered with monad stack
%
\haskell{sections/04MonadicAAM/03RecoveringKCFA/00AbstractMonad.hs}
%
and the \h|Delta| and \h|AAM| components described in
section~\ref{section:AAMByExample:RecoveringKCFA}
%
\haskell{sections/04MonadicAAM/03RecoveringKCFA/01Exec.hs}

### 2014-09-19 writeup/darais-quals.tex
very close to complete first draft

@@
 \documentclass{article}
+
+% Boilerplate {{{
 
@@
 
-\begin{abstract} % {{{
-\end{abstract} % }}}
+% }}}
+
+% Abstract {{{
+\begin{abstract}
+\end{abstract}
+% }}}
 
@@
 
-Static analyses have two components: 
+% A static analysis has two components: 
+% \begin{itemize}
+% \item A \emph{computation} which computes some set of facts about the execution of a program.
+% \item A \emph{proof of correctness} about the computation.
+% \end{itemize}
 % 
-\begin{itemize}
-\item A \emph{computational artifact} which computes a set of facts about the execution of a program.
-\item A \emph{proof of correctness} about the computational artifact.
-\end{itemize}
-
-We are motivated by static analysis techniques which are:
+% We are motivated by static analysis techniques which are:
+% \begin{itemize}
+% \item Compartmental: Various aspects of the analysis have been separated from each other and are understood in isolation.
+% \item Modular: Design choices in one aspect of the analysis do not restrict design choices of another.  
+%       This property is important for both computational and correctness components of an analysis.
+% \item Language Agnostic: A given analysis technique can be seamlessly transfered from one semantics to another.
+% \item Correct: The proof of correctness should exist, and there should be a framework for establishing it.
+% \item Derived: Either the computational artifact or the correctness proof should be obtained "for free" from the other.
+% \end{itemize}
 % 
-\begin{itemize}
-\item Compartmental: Various aspects of the analysis have been separated from each other and are understood in isolation.
-\item Modular: Design choices in one aspect of the analysis do not restrict design choices of another.  This property is important for both computational and correctness components of an analysis.
-\item Language Agnostic: A given analysis technique can be seamlessly transfered from one semantics to another.
-\item Correct: The proof of correctness should exist, and there should be a framework for establishing it.
-\item Derived: Either the computational artifact or the correctness proof should be obtained "for free" from the other.
-\end{itemize}
-
-Our framework compartmentalizes:
+% Our framework compartmentalizes:
+% \begin{itemize}
+% \item Abstract domain a la Cousot.
+% \item Abstract time and address a la Van Horn and Might.
+% \item Intensional optimizations a la Van Horn and Might. 
+% \item Object-sensitivity a la Smaragdakis.
+% \item Analysis control properties.
+% \end{itemize}
 % 
-\begin{itemize}
-\item Abstract domain a la Cousot.
-\item Abstract time and address a la Van Horn and Might.
-\item Intensional optimizations a la Van Horn and Might. 
-\item Object-sensitivity a la Smaragdakis.
-\item Analysis control properties.
-\end{itemize}
-
-Our framework is:
+% Our framework is:
+% \begin{itemize}
+% \item Compartmental: Each of the above concerns are separated and independent.
+% \item Modular: Design choices in one axis is completely independent of others
+% \item Language Agnostic: All axis other than abstract domain are fully language agnostic.
+% \item Correct: A correctness framework is described and each axis is proven correct in isolation.
+% \item Derived: The correctness of an analysis using our framework comes for free.
+% \end{itemize}
 % 
-\begin{itemize}
-\item Compartmental: Each of the above concerns are separated and independent.
-\item Modular: Design choices in one axis is completely independent of others
-\item Language Agnostic: All axis other than abstract domain are fully language agnostic.
-\item Correct: A correctness framework is described and each axis is proven correct in isolation.
-\item Derived: The correctness of an analysis using our framework comes for free.
-\end{itemize}
-
-Analysis control properties are the primary topic of this work.
+% Analysis control properties are the primary topic of this work.
+% We merely note that the other axis in the design space are implemented in our framework using existing techniques.
 % 
-We merely note that the other axis in the design space are implemented in our framework using existing techniques.
-
-Contributions:
-%
-\begin{itemize}
-\item Computational artifacts and correctness proofs for the axis of abstract control, which captures the choice of flow and path sensitivity.
-\item Computational artifacts and correctness proofs for intensional optimizations (gargabe collection and mcfa) independent of abstract control.
-\item Language independent proofs of refinement for flow-sensitivity and path-sensitivity choices using the monadic abstraction.
-\end{itemize}
+% Contributions:
+% \begin{itemize}
+% \item Computational artifacts and correctness proofs for the axis of abstract control, which captures the choice of flow and path sensitivity.
+% \item Computational artifacts and correctness proofs for intensional optimizations (gargabe collection and mcfa) independent of abstract control.
+% \item Language independent proofs of refinement for flow-sensitivity and path-sensitivity choices using the monadic abstraction.
+% \end{itemize}
 
@@
 We use traditional notation for function definitions and applications rather than lambda calculus notation.
-%
-For example, we will define [f(w, x) := g(y, z)] rather than [f w x := g y z].
+For example, we will write $f(w, x) ≔ g(y, z)$ rather than $f\; w\; x ≔ g\; y\; z$.
 
 We use lambda notation for anonymous functions.
-%
-For example, defining the identify function could be notated [id := λ(x) → x] rather than the named definition [id(x) := x].
+For example, defining the identify function could be notated $id ≔ λ(x) → x$ rather than the named definition $id(x) ≔ x$.
 
@@
 A basic familiarity with the concept of monads will be useful for understanding the rest of the paper.
-%
 Some find it useful to review the concept of functor before learning about monads. 
-%
 We adopt this approach for our brief explanation of monads.
 
-A type [𝒯 : Set → Set] is called a functor if one can:
-%
+A type $𝒯 : Set → Set$ is called a \emph{functor} if one can:
 \begin{itemize}
-\item define [map : ∀ (A : Set), (A → B) → 𝒯(A) → 𝒯(B)]
-\item prove [map id = id], where id is the identity function [λ x → x]
-\item prove [map (g ∘ f) = map g ∘ map f]
+\item Define $map : ∀ (A : Set), (A → B) → (𝒯(A) → 𝒯(B))$
+\item Prove $map(id) = id$, where $id$ is the identity function $λ(x) → x$
+\item Prove $map (g ∘ f) = map(g) ∘ map(f)$
 \end{itemize}
 
-Example: Lists are functors, where map is defined:
-%
-    map(f)(xs) := case xs 
-      Nil → Nil 
-      Cons(x,xs') → Cons(f(x),map(f)(xs'))
+\paragraph{Example:} 
+Lists are functors, where $map$ is defined:
+\begin{align*}
+    map(f)(xs) ≔ &\case(xs):           \\
+          Nil    &→ Nil                \\
+          x ∷ xs &→ f(x) ∷ map(f)(xs') \\
+\end{align*}
 
-Example: [map(isEven)([1, 2, 3, 4]) = [False, True, False, True]]
+\paragraph{Example:} $map(isEven)([1, 2, 3, 4]) = [False, True, False, True]$
 
-Likewise a type [𝒯 : Set → Set] is called a monad if one can:
-%
+Likewise a type $𝒯 : Set → Set$ is called a \emph{monad} if one can:
 \begin{itemize}
-\item define [extend : ∀ (A : Set), (A → 𝒯(B)) → 𝒯(A) → 𝒯(B))]
-\item prove unit and associativity laws (not mentioned here).
+\item Define $extend : ∀ (A : Set), (A → 𝒯(B)) → (𝒯(A) → 𝒯(B))$
+\item Prove unit and associativity laws (not mentioned here).
 \end{itemize}
+The only difference from the definition of functor is in the the first argument $(A → 𝒯(B))$.
+For monads, this argument is allowed to be a \emph{monadic} function, rather than a pure function $(A → B)$.
 
-Example: Lists are monads, where extend is defined:
-%
-    extend(f)(xs) := case xs
-      Nil → Nil
-      Cons(x,xs') → f(x) ++ extend(f)(xs')
-%
-[extend] for lists can be understood as a map followed by a concat or flattening operation.
+\paragraph{Example:} Lists are monads, where $extend$ is defined:
+\begin{align*}
+    extend(f)(xs) ≔ &\case(xs):              \\
+                Nil &→ Nil                   \\
+            x ∷ xs' &→ f(x) ⧺ extend(f)(xs') \\
+\end{align*}
+$extend$ for lists can be understood as $map$ followed by the flattening operation $concat$.
 
-Example: [extend(λ x → [x + 1, x - 1])([10, 100]) = [9, 11, 99, 101]
+\paragraph{Example:} 
+$extend(λ(x) → [x + 1, x - 1])([10, 100]) = [9, 11, 99, 101]$
 
-[extend] for lists can also be understood as concat followed by map:
-%
-      extend(f)([10, 100]) 
-    = concat(map(f)([10, 100])) 
-    = concat([[9, 11], [99, 101]]) = [9, 11, 99, 101]
-%
-For all monads, it is equivalent to define return, map and join, where [join :
-∀ (A : Set), 𝒯(𝒯(A)) → 𝒯(A)].
-%
-As we have seen, join is just concat for the list monad.
+\paragraph{Example:} 
+$extend$ for lists can also be understood as $concat$ followed by $map$:
+\begin{align*}
+extend(f)([10, 100]) &= concat(map(f)([10, 100]))    \\
+                     &= concat([[9, 11], [99, 101]]) \\
+                     &= [9, 11, 99, 101]             \\
+\end{align*}
+For all monads, it is equivalent to define $return$, $map$ and $join$, where $join : ∀ (A : Set), 𝒯(𝒯(A)) → 𝒯(A)$.
+As we have just seen, $join$ for the list monad is just $concat$.
+
+% }}}
+
+% Monad Transformers {{{
+\subsection{Monad Transformers}
+\label{section:Background:MonadTransformers}
+
+Monad transformers are functions between monads.
+Where a monad $𝒯$ will have type $Set → Set$, a monad transformer will have type $(Set → Set) → (Set → Set)$.
+Perhaps a more appropriate type is $Monad → Monad$.
+
+Monad transformers are used to add an effect on top of another monad.
+The three monads used in this work are the state monad, powerset monad, and the identity monad.
+The state monad is notated $\SM(𝓈)$ (carrying a single cell of type $𝓈$), the powerset monad is notated $\PM$, and the identity is notated $ID$.
+The state and powerset monad transformer equivalents are notated $\ST(𝓈)(𝓂 )$ and $\PT(𝓂 )$ where $𝓂 $ is the underlying monad.
+The state monad has $get$ and $put$ effects, whereas the powerset monad has a $nondeterminism$ effect.
+The transformer versions of both of these monads allow you to combine effects piecewise.
+
+\paragraph{Example:}
+$\ST(ℤ)(\PT(ID))$ is a monad which has $get$ and $put$ effects for a single integer state, in addition to nondeterminism effects.
+
+The definitions, details and proofs of all monads used in our work are given in section \ref{section:Proofs+Definitions}.
 
@@
 
-This work follows the small-step operational approach to language semantics pioneered by ? and popularized by ?.
-%
-In small-step operational semantics, the meaning of an expression ℯ in a language is given by the transitive closure of a step relation [ℯ ↦ ℯ'].
-%
+This work follows the small-step operational approach to language semantics pioneered independently by Plotkin and Felleisen.
+In small-step operational semantics, the meaning of an expression $e$ in a language is given by the transitive closure of some step relation $e ↦ e'$.
 This yields a state-machine approach to semantics.
-%
-Exploring the meaning of the expression means exploring all possible transitions reachable from ℯ through ↦.
+Exploring the meaning of $e$ reduces to exploring all possible transitions reachable from $e$ through $↦$.
+This contrasts to previously developed deontational approaches, where the meaning of $e$ would be given by a denotation function $⟦ e ⟧$.
+This denotation function can quickly demand lots of mathematical machinery, and the small-step operational method side-steps this issue.
+
+The small-step operation framework supports arbitrary step relations $e ↦ e'$.  
+However for our purposes $↦$ will be always be a function-like form of \emph{decidable} relation, i.e. a multifunction.
 
@@
 
-Abstract Iterpretation (AI) is a formal framework for program analysis pioneered by Cousot et Cousot.
-%
-In the setting of abstract interpretation, a program analysis is an alternate semantics over an abstract domain.
+Abstract Iterpretation (AI) is a formal framework for program analysis pioneered by Cousot and Cousot.
+In the setting of abstract interpretation, a program analysis is just an alternate semantics over an abstract domain.
+To distinguish the original semantics and the alternate semantics, we call the original semantics the \emph{concrete} semantics, and the alternate semantics the \emph{abstract} semantics\footnote{
+  This terminology quickly becomes confusing because we will eventually construct an abstract machine for both semantics, yielding a concrete abstract machine and an abstract abstract machine. 
+  This pun is exploited mercilessly in the work of Might and Van Horn as they "abstract" abstract abstract machines, yielding an abstract abstract abstract machine.
+}.
 
-both a discipline for constructing program analyses
-and a framework for reasoning about their correctness pioneered by Cousot and
-Cousot. Van Horn and Might have extended this tradition in their work on
-Abstracting Abstract Machines, showing how to systematically construct analysis
-for small step operational semantics.
-%
-Relating an analysis to a semantics requires a bridge.  In general, such a
-relationship will be stated as:
-%
-∀ (c c': C) (a a': A), c R a ∧ c ↦ c' ∧ a ↦ a' ⇒  c' R a'
-%
-In words, this is saying that for a concrete program c and a valid abstraction
-of it a, if the program can take a step to c' and the abstraction can take a
-step to a', then a' is a valid abstraction of c'.
-%
-Once this theorem is established, the transitive closure of it follows
-trivially and justifies performing an analysis by:
-%
-1: abstracting a program c to it its abstraction a
-2: running a using the transitive closure of a ↦* a'
-3: inspecting the final a' to infer behavior of running c
-%
-The question now is how to set up R--how to relate one language and semantics
-to another. For this, abstract interpretation turns to order theory.  We say c
-and a are related if c is in the image of projecting a to 𝒫(c).  Using this
-projection, commonly called γ, we restate the theorem:
-%
-∀ (c c': C) (a a': A), c ∈ γ(a) ∧ c ↦ c' ∧ a ↦ a' ⇒  c' ∈ γ(a')
-%
-However, the technique generalizes to arbitrary partial orders, and is done so
-profitably in the literature.  Stating the theorem with partial orders gives:
-%
-∀ (c c': C) (a a': A), c ⊑ γ(a) ∧ c ↦ c' ∧ a ↦ a' ⇒  c' ⊑ γ(a')
-%
-The partial order for concrete domains is almost always the inclusion order on
-𝒫(C), while the partial order for abstract domains is almost always a finite
-element lattice.
-%
-A complete treatment of abstract interpretation is centered around constructing
-a _galois connection_ between a concrete and abstract semantics. Establishing a
-galois connection between C and A establishes both sides of the relationship
-between C and A, whereas until now we've only had one direction.
-Having both direction allows us to directly project concrete objects to
-abstract objects, and provides a setting to discuss if γ is a "best"
-abstraction function relating A to C.
-%
-A galois connection between two posets C and A is notated [C α⇄ γ A] and
-consists of:
-%
-* [α : C → A] where α is monotonic
-* [γ : A → C] where γ is monotonic
-* [γ ∘ α] is expansive: ∀ (x : C), x ⊑ γ(α(x))
-* [α ∘ γ] is contractive: ∀ (y : A), α(γ(y)) ⊑ y
-%
-The last two points can be succinctly stated as [α ∘ γ ⊑ id ⊑ γ ∘ α], using the
-logical monotonicity relation [f ⊑ g ⇔  (x ⊑ y ⇒  f(x) ⊑ g(y))] for the
-function space.
-%
-Equivalent to the above four rules is:
-%
-* x ⊑ γ(y) ⇔  α(x) ⊑ y
-%
-Example: given a galois connection (C α⇄ γ A), there exists a galois connection
-(C → C) α⇄ γ (A → A) where:
-%
-α(f : C → C) = γ ∘ f ∘ α
-γ(g : A → A) = α ∘ g ∘ γ
-%
-Using the constructed galois connection on the function space, we can more
-succinctly state the correctness of an analysis as:
-%
-_↦_ ⊑ _↦ₐ_
-%
-Example: the language (𝒫(ℤ),+,*) forms a galois connection with the language
-(𝒫({EVEN, ODD}),∧,∨) where:
-%
-α(zs : 𝒫(ℤ)) = ⋃ { { EVEN | ∃ z ∈ zs ∧ Even(z)}, { ODD | ∃ z ∈ zs ∧ Odd(z) } }
-γ(ts : 𝒫({EVEN, ODD})) = ⋃ { { z ∈ ℤ | EVEN ∈ ts ∧ Even(z) }, { z ∈ ℤ | ODD ∈ ts ∧ Odd(z) } }
-%
-α and γ must have the expansive and contractive properties in order
-for this to be considered a galois connection.  The expansive property
-corresponds to the soundness of the abstraction.  Soundness ensures that if
-you abstract and then project back to the concrete, you must cover the points
-you started with.  The contractive property corresponds to the tightness of
-the abstraction.  Tightness ensures that the abstraction is the most precise
-abstraction possible.  For example, picking [α(zs) = {EVEN, ODD}] would be
-sound, because projecting back to the concrete gives ℤ which tautologically
-covers all integers, but this choice of α is clearly not tight.  Our choice
-of α and γ in this example are indeed both sound and tight.
-%
-In this example, the concrete language is (ℤ,+,*) and the abstract language is
-({EVEN,ODD},∧,∨)
-%
-Using the lifting of galois connections to the function space, we can/should
-also justify _+_ ⊑ γ(_∧_) and _*_ ⊑ γ(_∨_)
-%
+Consider a concrete language $\C$ and a program in that language $(\Ce : \C)$.
+In the AI framework, an analysis for $\Ce$ is:
+\begin{itemize}
+\item 
+  An abstract language $\A$.
+\item 
+  A relationship between $\C$ and $\A$ that defines when $(\Ce : \C)$ and $(\Ae : \A)$ are related.
+  This often takes the form of projecting $\A$ to $\PM(\C)$ and using the subset relation.
+\item 
+  An abstract version of $\Ce$ that lives in $\A$: $(\Ae : \A)$.
+\item 
+  An abstract version of the $\Csteps$ relation: $(\Asteps : \A × \A → Prop)$
+\item 
+  A way to explore every state reachable by $\Ae$ under $\Asteps$.
+  This often requires $\A$ to be finite.
+\end{itemize}
+By convention, we put hats ($\widehat{}$) on things to name their abstract cousins.
+The content of the approach is shown in the following picture:
+\input{diagrams/AI.tikz}
+This picture takes place at some point in the analysis process.
+$\Ce$ is the current program (possibly the result of running the base program for a little while).
+$\Ae$ is a valid abstraction of $\Ce$, where this validity is expressed some relation $R$ holding between $\Ce$ and $\Ae$.
+$\Ce'$ is some next state of execution of $\Ce$, and likewise for $\Ae$/$\Ae'$.
+In order to be a correct analysis, it must be guaranteed that $\Ce'$ and $\Ae'$ will be related.
+The logical structure of the picture is given by:
+\begin{equation*}
+∀ (\Ce,\Ce' : \C ) (\Ae,\Ae' : \A), (\Ce R \Ae) ∧ (\Ce \Csteps \Ce') ∧ (\Ae \Asteps \Ae') ⇒  (\Ce' R \Ae')
+\end{equation*}
+
+% }}}
+
+% Galois Connections {{{
+\subsection{Galois Connections}
+\label{section:Background:GaloisConnections}
+
+The AI setting can be elegently simplified and enriched through the use of \emph{galois connections}.
+Galois connections serve as a unifying framework for establishing the "relationship between $\C $ and $\A$" mentioned in the previous section.
+
+A galois connection between two posets (sets with a partial order) $\C$ and $\A$ is notated $\C\galois{α}{γ}\A$ and contains:
+\begin{itemize}
+\item $(α : \C → \A)$ where $α$ is monotonic
+\item $(γ : \A → \C )$ where $γ$ is monotonic
+\item $(γ ∘ α)$ is expansive: $∀ (x : \C ), x ⊑ γ(α(x))$
+\item $(α ∘ γ)$ is contractive: $∀ (y : \A), α(γ(y)) ⊑ y$
+\end{itemize}
+The last two properties can be succinctly stated as $(α ∘ γ ⊑ id ⊑ γ ∘ α)$\footnote{
+  This uses the logical monotonicity relation $f ⊑ g ⇔  (x ⊑ y ⇒  f(x) ⊑ g(y))$ for the function space
+}.
+Equivalent to all four properties is the property $x ⊑ γ(y) ⇔  α(x) ⊑ y$.
+
+\paragraph{Example:}
+Given a galois connection $\C\galois{α}{γ}\A$, there exists a galois connection $\C → \C  \galois{α}{γ} \A → \A$ where:
+\begin{align*}
+α(f : \C → \C) &≔ γ ∘ f ∘ α \\
+γ(g : \A → \A) &≔ α ∘ g ∘ γ \\
+\end{align*}
+
+\paragraph{Example:} 
+The language $(\PM(ℤ),+,*)$ forms a galois connection with the language $(\PM({EVEN, ODD}),∧,∨)$ where:
+\begin{align*}
+α(zs : \PM(ℤ))           &≔ ⋃ { { EVEN | ∃ z ∈ zs ∧ Even(z)}, { ODD | ∃ z ∈ zs ∧ Odd(z) } }      \\
+γ(ts : \PM({EVEN, ODD})) &≔ ⋃ { { z ∈ ℤ | EVEN ∈ ts ∧ Even(z) }, { z ∈ ℤ | ODD ∈ ts ∧ Odd(z) } } \\
+\end{align*}
+
+Galois connections simplify the AI framework by using $x ⊑ γ(y)$ or (equivalently) $α(x) ⊑ y$ as the relation $(x R y)$.
+Put differently, galois connections are a natural and general way of placing partial orders \emph{on sets themselves} (note that x and y live in different sets).
+One can also think of galois connections as something like an isomorphism, but with a weaker round-trip property.
+(An isomorphism would require $α ∘ γ = id = γ ∘ α$.)
+
+Using galois connections, the AI framework introduced in the previous section can be re-stated. In the AI framework, an analysis for $\Ce$ is:
+\begin{itemize}
+\item 
+  An abstract language $\A$.
+\item 
+  A galois connection $\C\galois{α}{γ}\A$.
+\item 
+  An abstract version of the $\Csteps$ relation: $(\Asteps : \A × \A → Prop)$
+\item 
+  A way to explore every state reachable by $\Ae$ under $\Asteps$.
+  This often requires $\A$ to be finite.
+\end{itemize}
+The content of the approach is shown in the following picture:
+\input{diagrams/AIGC.tikz}
+This picture takes place at some point in the analysis process.
+$\Ce$ is the current program (possibly the result of running the base program for a little while).
+$\Ae$ is a valid abstraction of $\Ce$ by construction as $α(\Ce)$.
+$\Ce'$ is some next state of execution of $\Ce$, and likewise for $\Ae$/$\Ae'$.
+In order to be a correct analysis, it must be guaranteed that $\Ce' ⊑ γ(\Ae')$.
+The logical structure of the picture is given by:
+\begin{equation*}
+∀ (\Ce,\Ce' : \C ) (\Ae,\Ae' : \A), (\Ce ⊑ γ(\Ae)) ∧ (\Ce \Csteps \Ce') ∧ (\Ae \Asteps \Ae') ⇒  (\Ce' ⊑ γ(\Ae'))
+\end{equation*}
+
+The logical structure of the previous figure is the \emph{soundness} property of the analysis.
+This property can be simplified further by the statement $\Csteps ⊑ γ(\Asteps)$.
+This uses the definition of galois connections for function spaces shown in a previous example.
+What is more, this more compact statement of soundness also captures another property, the \emph{tightness} of the analysis.
+Unfolded, this property states:
+\begin{equation*}
+∀ (\Ce,\Ce' : \C ) (\Ae,\Ae' : \A), (\Ce' ⊑ γ(\Ae')) ∧ (\Ce \Csteps \Ce') ∧ (\Ae \Asteps \Ae') ⇒  (\Ce ⊑ γ(\Ae))
+\end{equation*}
+Tightness tells us that our abstraction is provably \emph{the best abstraction possible}.
+The beauty of galois connections is how they capture both soundness and tightness of an analysis with one concise statement: $\Csteps ⊑ γ(\Asteps)$.
+
+% }}}
+
+% CPS {{{
+\subsection{Continuation Passing Style}
+\label{section:Background:ContinuationPassingStyle}
+
+We use CPS-IF as an example language to demonstrate the benefits of our framework.
+CPS is a syntactically restricted version of the lambda calculus.
+CPS-IF is an extension of CPS with conditionals.
+We choose CPS because the state machine for the semantics of CPS requires no call stack.
+We add conditionals to make things slightly more interesting.
+
+We define the CPS-IF language as follows:
+\begin{align*}
+x,y,k &: Var  &⩴ ..variables..                          \\
+b,i,l &: Lit  &⩴ ℤ ⋃ 𝔹                                  \\
+f,g   &: Lam  &⩴ λ(x) → c | λ(x,k) → c                  \\
+a     &: Atom &⩴ x | l | f | op(a)                      \\
+op    &: Op   &⩴ add1 | sub1 | gez                      \\
+c     &: Call &⩴ if(a){c}{c} | a(a) | a(a,a) | halt(a)  \\
+\end{align*}
+
+CPS-IF supports integer and boolean literals, addition and subtraction by 1, and testing if an integer is greater than or equal to zero.
+The essence of CPS is in the syntactic restriction on application forms $a(a)$ and $a(a,a)$.  
+Because $a$ cannot be a nested call expression, evaluating a $Call$ requires no evaluation stack.
+
+All lambda calculus terms can be translated to CPS terms through CPS-conversion.
+However, because lambda calculus terms are much easier to read, we will use lambda calculus terms for examples.
+We display the CPS-converted version of examples where appropriate.
+
+% }}}
+
+% Control Flow Analysis {{{
+\subsection{Control Flow Analysis}
+\label{section:Background:ControlFlowAnalysis}
+
+Control flow analysis is a class of analysis which is particularly important for higher-order languages.
+In non-higher-order languages, it is useful to distinguish control-flow--which functions can be called--from data-flow--which values flow where.
+Traditionally one performs a control-flow analysis first, to find out which functions are called, and a data-flow analysis second, using the results of the control flow analysis.
+
+In higher-order languages, data-flow and control-flow become the same problem.
+Before you can tell which functions will be called, you need to know how values flow, because functions are themselves values.
+
+We note that the distinction between "higher-order" languages and "non-higher-order" languages is a red herring.
+Functions can be passed as values in C using function pointers, and object-oriented languages enjoy a similar circularity between control- and data- flow from method dispatch.
+You can read these statements as saying "all languages are higher order, and thus need control flow analysis".
+Or you can read them as saying "all languages are higher order, and we've done just fine without control flow analysis so far".
+The distinction is in the code that you end up analysing.
+
+We build on the tradition of control flow analysis pioneered by Shivers and its recent refinements developed by Might and Van Horn.
+In this tradition, the result of a control flow analysis is closer to what one would think of as a data-flow analysis.
+0CFA, the most basic control flow analysis, computes (conservatively) the set of values which might flow to a particular variable.
+kCFA is a suite of context sensitive extensions to 0CFA which analyses a given function separately for each possible calling context.
+We use 0CFA as the example analysis in this paper.
+However our framework, and our implementation of it, scale to the full range of context sensitive control flow analyses.
+
+% }}}
+
 % }}}
@@
 
-Monadic AI (Sergey et al. PLDI 2013) showed that computational modularity for
-abstract interpreters can be achieved through a monadic abstraction.  We build
-directly on this work at the level of intuition--monads serve as our
-pivot-point for generalizing abstract control. However, in our pursuit of an
-axis-independent correctness framework, our approach deviates greatly from this
-work in the specific monadic interfaces used.
+Monadic AI (Sergey et al. PLDI 2013) showed that computational modularity for abstract interpreters can be achieved through a monadic abstraction.
+We build directly on this work at the level of intuition--monads serve as our pivot-point for generalizing abstract control.
+However, in our pursuit of an axis-independent correctness framework, our approach deviates greatly from this work in the specific interfaces used.
 
-We use CPS to simplify presentation, although our implementation supports a
-larger scheme-like language.
+First we demonstrate a simple 0CFA analysis on CPS-IF which doesn't use our framework.
+The language is the same as introduced in section \ref{section:Background:ContinuationPassingStyle}, and repeated here for convenience:
+\begin{align*}
+x,y,k &: Var  &≔ ..variables..                          \\
+b,i,l &: Lit  &≔ ℤ ⋃ 𝔹                                  \\
+f,g   &: Lam  &⩴ λ(x) → c | λ(x,k) → c                  \\
+a     &: Atom &⩴ x | l | f | op(a)                      \\
+op    &: Op   &⩴ add1 | sub1 | gez                      \\
+c     &: Call &⩴ if(a){c}{c} | a(a) | a(a,a) | halt(a)  \\
+\end{align*}
+The abstract semantics of 0CFA will track literals (including lambdas) that appear in the program text.
+For integers that do not appear in the program text, the analysis uses a single token $INT$ which conservatively approximates any possible integer.
+Because there are a finite many literals in the program text, there are a finite number of possible values.
+Formally, the state space for the abstract machine of the abstract semantics is defined as:
+\begin{align*}
+v &: \AVal   ⩴ INT | l | λ(x) → c | λ(x,k) → c \\
+σ &: \AStore ≔ Var → \PM(\AVal)                 \\
+Σ &≔ \PM(Call × \AStore)                        \\
+\end{align*}
 
-First we demonstrate a simple, fully instantiated instance of our framework
-_without_ monads. This example is the instantiation of:
+The abstract semantics for $Σ$ comes in two parts.  
+First we define a \emph{denotation function} $(𝒜 : \AStore × Atom → \PM(\AVal))$ for $Atom$ expressions.
+Second we define a \emph{step relation} (as a function) $(𝒞 : Call × \AStore → \PM(Call × \AStore))$ for $Call$ expressions.
+\begin{align*}
+𝒜                            &: \AStore × Atom → \PM(\AVal)  \\
+𝒜 (σ,x)                      &≔ σ(x)                        \\
+𝒜 (σ,l)                      &≔ { l }                       \\
+𝒜 (σ,add1 a) | 𝒜 [ρ,σ] ⊑ INT &≔ { INT }                     \\
+𝒜 (σ,sub1 a) | 𝒜 [ρ,σ] ⊑ INT &≔ { INT }                     \\
+𝒜 (σ,gez a)  | 𝒜 [ρ,σ] ⊑ INT &≔ { T , F }                   \\
+𝒜 (σ,λ(x)(k) → c)            &≔ { λ(x)(k) → c }             \\
+\end{align*}
 
-* A simple abstract domain which tracks control flow, booleans and integer
-  constants.
-* 0-cfa
-* No abstract garbage collection or copied closurse (see mcfa)
-* Fully flow- and path-sensitive, which is in some sense the most natural
-  structural abstraction of the concrete state space (see AAM).
+% \begin{align*}
+% 𝒞                   &: &Call × \AStore → \PM(Call × \AStore)                                     \\
+% 𝒞 <if(a){c₁}{c₂},σ> &≔ &{ <c,σ>                                                                 \\
+%                     &  &|             c &∈ ⋃ { { c1 } if T ∈ 𝒜 (σ,a), { c2 } if F ∈ 𝒜 (σ,a) }   \\
+%                     &  &}                                                                       \\
+% 𝒞 <a₁(a₂,a₃),σ>     &≔ &{ <c,σ'>                                                                \\
+%                     &  &| <λ(x)(k) → c> &∈ 𝒜 (σ,a₁)                                             \\
+%                     &  &|            v₂ &∈ 𝒜 (σ,a₂)                                             \\
+%                     &  &|            v₃ &∈ 𝒜 (σ,a₃)                                             \\
+%                     &  &|            σ' &≔ σ ⊔ [x ↦ v₂] ⊔ [k ↦ v₃]                              \\
+%                     &  &}                                                                       \\
+% \end{align*}
 
-x,y,k : Var  ::= ..variables..
-b,i,l : Lit  ::= ℤ ⋃ 𝔹
-a     : Atom ::= x | l | op a | λ x . c | λ x k . c
-op    : Op   ::= add1 | sub1 | gez
-c     : Call ::= if(a){c}{c} | a(a) | a(a,a) | halt(a)
+The complete analysis of a program c is defined as the least fixed point of a
+\emph{collection semantics} for the relation $𝒞$:
+\begin{align*}
+μ(Σ) → {<c,⊥>} ⊔ 𝒞*(Σ)
+\end{align*}
+where
+\begin{align*}
+𝒞*    &: \PM(Call × \AStore) → \PM(Call × \AStore) \\
+𝒞*(Σ) &≔ ⋃ { 𝒞 <c,σ> | <c,σ> ∈ Σ }               \\
+\end{align*}
+(The collecting semantics tracks all states that the program could be in rather than just the final states.)
 
-v  : ^Val^ ::= INT | l | <λ x . c> | <λ x k . c>
+The first insight in monadic static analysis is to abbreviate the definitions of $𝒜 $ and $𝒞$ using a monad.
+At this point, the monad "trick" is nothing more than simplifying the definition. 
+This simplification is similar to how a functional programmer would use a state monad in place of explicit state passing.
 
-σ  : ^Store^ := Var ⇀  𝒫 (^Val^)
+We use a powerset monad and state monad transformer to write the analysis in monadic style.
+The full definitions of the monads used throughout this paper are defered to section \ref{section:Proofs+Definitions}.
+Because monad transformers are just fancy names for simple types, we write the simple types underneath definitions which use monad transformers.
+The monadic conversion of the above analysis is:
+\begin{align*}
+ℳ     &: Set → Set                  \\
+ℳ (a) &≔ \ST(\AStore)(\PM)(a)       \\
+ℳ (a) &≔ \AStore → \PM(a × \AStore) \\
+\end{align*}
 
-𝒜 [_](_) : (^Store^ × Atom) ⇀  𝒫 (^Val^)
-𝒜 [σ](x)                      := σ(x)
-𝒜 [σ](l)                      := { l }
-𝒜 [σ](add1 a) | 𝒜 [ρ,σ] ⊑ INT := { INT }
-𝒜 [σ](sub1 a) | 𝒜 [ρ,σ] ⊑ INT := { INT }
-𝒜 [σ](gez a)  | 𝒜 [ρ,σ] ⊑ INT := { T , F }
-𝒜 [σ](λ x k . c)              := { <λ x k . c> }
+\begin{align*}
+𝒜                &: Atom → ℳ (\PM(\AVal))                    \\
+𝒜 (x)            &≔ do                                       \\
+                 &σ ← get-σ                                  \\
+                 &return(σ(x))                               \\
+𝒜 (l)            &≔ return({l})                              \\
+𝒜 ₘ(add1(a))     &≔ do                                       \\
+                 &v ← 𝒜 (a)                                  \\
+                 &return({INT | ∃ i ⊑ INT ∈ v})              \\
+𝒜 ₘ(sub1(a))     &≔ do                                       \\
+                 &v ← 𝒜 (a)                                  \\
+                 &return({INT | ∃ i ⊑ INT ∈ v})              \\
+𝒜 ₘ(gez(a))      &≔ do                                       \\
+                 &v ← 𝒜 (a)                                  \\
+                 &return(⋃ { {T, F} | ∃ i ⊑ INT ∈ v})        \\
+𝒜 ₘ(λ(x)(k) → c) &≔ return({λ(x)(k) → c})                    \\
+                                                             \\
+𝒞                 &: Call → ℳ (Call)                         \\
+𝒞 (if(a){c₁}{c₂}) &≔ do                                      \\
+                  &vP ← 𝒜 (a)                                \\
+                  &v ← lift\PM(vP)                           \\
+                  &b ← coerce𝔹(v)                            \\
+                  &return { if b then c₁ else c₂ }           \\
+𝒞ₘ(a₁(a₂,a₃))     &≔ do                                      \\
+                  &vP ← 𝒜 (a₁)                               \\
+                  &v ← lift\PM(vP)                           \\
+                  &<λ(x)(k) → c> ← coerceλ₂(v)               \\
+                  &vP₂ ← 𝒜 (a₂)                              \\
+                  &vP₃ ← 𝒜 (a₃)                              \\
+                  &v₂ ← lift\PM(vP₂)                         \\
+                  &v₃ ← lift\PM(vP₃)                         \\
+                  &modify-σ (λ(σ) → σ ⊔ [x ↦ v₂] ⊔ [k ↦ v₃]) \\
+                  &return(c)                                 \\
+\end{align*}
 
-(this is often stated as a relation in the general setting of small-step
-relational semantics.)
+The monadic abstraction provides a nice way to simplify the implementation of the analysis.  
+In particular, cases which do not modify the store or only return one result.
+Both of these cases are instances of having "no effect", and they need not mentioned all members of the state machine.  
+(This was observed in Sergey et. al. PLDI 2013.)
 
-𝒞 <_,_> : (Call × ^Store^) ⇀  𝒫 (Call × ^Store^)
-𝒞 <if(a){c₁}{c₂},σ> := { <c,σ> |  c  ∈ ⋃ [ { c1 } if T ∈ 𝒜 [σ](a), { c2 } if F ∈ 𝒜 [σ](a) ] }
-𝒞 <a₁(a₂,a₃),σ>     := { <c,σ'> 
-                       | <λ x k . c>  ∈ 𝒜 [σ](a₁)
-                       |          v₂  ∈ 𝒜 [σ](a₂)
-                       |          v₃  ∈ 𝒜 [σ](a₃)
-                       |          σ' := σ ⊔ [x ↦ v₂] ⊔ [k ↦ v₃]
-                       }
+As before, we must complete the analysis by building an abstract machine transition function:
+\begin{align*}
+𝒞*    &: \PM(Call × \AStore) → \PM(Call × \AStore) \\
+𝒞*(Σ) &≔ ⋃ { (𝒞ₘ c)(σ) | <c,σ> ∈ Σ }               \\
+\end{align*}
 
-The complete analysis of a program c is defined as:
-
-μ(Σ) . {<c,⊥>} ⊔ Σ ⊔ 𝒞*(Σ)
-
-where
-
-𝒞* : 𝒫 (Call × ^Store^) → 𝒫 (Call × ^Store)
-𝒞*(Σ) = ⋃ [ 𝒞 <c,σ> | <c,σ> ∈ Σ ]
-
-The monadic conversion of the above analysis is:
-
-ℳ  : Set → Set
-ℳ  a := StateT ^Store^ List a
-ℳ  a := ^Store^ → 𝒫 (a × ^Store^)
-
-𝒜 ₘ : Atom → ℳ (^Val^)
-𝒜 ₘ(x) = do
-  σ ← get-σ
-  return σ(x)
-𝒜 ₘ(l) = return {l}
-𝒜 ₘ(add1 a) = do
-  vD ← 𝒜 ₘ (a)
-  return {INT | ∃ i ⊑ INT ∈ vD}
-𝒜 ₘ(sub1 a) = do
-  vD ← 𝒜 ₘ (a)
-  return {INT | ∃ i ⊑ INT ∈ vD}
-𝒜 ₘ(gez a) = do
-  vD ← 𝒜 ₘ (a)
-  return {T, F | ∃ i ⊑ INT ∈ vD}
-𝒜 ₘ(λ x k . c) = return <λ x k . c>
-
-𝒞ₘ : Call → ℳ (Call)
-𝒞ₘ(if(a){c₁}{c₂}) := do
-  vD ← 𝒜 ₘ (a)
-  v ← lift𝒫 (vD)
-  b ← coerce𝔹 (v)
-  return { c₁ if b | c₂ if ¬b }
-𝒞ₘ(a₁(a₂,a₃)) := do
-  vD ← 𝒜 ₘ (a₁)
-  v ← lift𝒫 (vD)
-  <λ x k . c> <- coerceλ₂ v
-  vD₂ ← 𝒜 ₘ (a₂)
-  vD₃ ← 𝒜 ₘ (a₃)
-  v₂ ← lift𝒫 vD₂
-  v₃ ← lift𝒫 vD₃
-  modify-σ (λ σ . σ ⊔ [x ↦ v₂] ⊔ [k ↦ v₃])
-  return c
-
-The monadic abstraction provides a nice way to simplify the implementation of
-the analysis.  In particular, cases which do not modify the store or only
-return one result--both of which are instances of having "no effect"--need not
-mentioned all members of the state machine.  (Credit Sergey et. al.)
-
-The complete analysis is defined as before where
-
-𝒞ₘ* : 𝒫 (Call × ^Store^) → 𝒫 (Call × ^Store^)
-𝒞ₘ*(Σ) = ⋃ [ (𝒞ₘ c)(σ) | <c,σ> ∈ Σ ]
-
-The exact abstract interpreter written directly is recovered through the
-definition of ℳ .  Our insight is twofold:
-
-* ℳ  can be _axiomatized_, such that different definitions for ℳ  give rise to
-  different control abstractions for the analysis.
-* The axiomatized monadic abstraction provides for a modular proof framework
-  for the correctness of a range of analyses.
+The monad we used recovers exactly the analysis we wrote before.
+On top of the convenience of writing things monadically, our insight and contribution is twofold:
+\begin{itemize}
+\item ℳ  can be \emph{axiomatized}, such that different definitions for ℳ  give rise to different control abstractions for the analysis.
+\item The monadic abstraction provides for a modular proof framework for establishing the correctness of the analysis.
+\end{itemize}
 
@@
 % Abstract Control {{{
-
 \section{Abstract Control}
@@
 
-The current instantiation of ℳ  yields a flow-sensitive path-sensitive
-analysis.  Consider a reordering of the List and State monad transformers:
+The current instantiation of ℳ  yields a flow-sensitive path-sensitive analysis.  
+Consider a reordering of the powerset\footnote{
+  The definition of $\PT$ we use is non-traditional.  However, our definition
+  for $\PT$ is actually a monad, whereas traditional definitions are not.  Our
+  definition and proofs for $\PT$ are given in section \ref{section:Proofs+Definitions}.
+}
+and state monad transformers:
+\begin{align*}
+ℳ     &: Set → Set                  \\
+ℳ (a) &≔ \PT(\SM(\AStore))(a)       \\
+ℳ (a) &≔ \AStore → \PM(a) × \AStore \\
+\end{align*}
 
-ℳ  : Set → Set
-ℳ  a := SetT (State ^Store^) a
-ℳ  a := ^Store^ → 𝒫 (a) × ^Store^
+As before, we must convert between monadic actions and abstract state space transitions to achieve an analysis:
+\begin{align*}
+𝒞*       &: \PM(Call) × \AStore → \PM(Call) × \AStore                          \\
+𝒞*<cP,σ> &≔ <{c' | c' ∈ π₁(𝒞(c))(σ) | c ∈ c\PM }, ⋃ { π₂(𝒞(c))(σ) | c ∈ cP }>  \\
+\end{align*}
+This instantiation of $ℳ $ and $𝒞*$ yields a flow-insensitive analysis.
 
-[Note: The definition of SetT we use is non-traditional.  However, our
-definition for SetT is actually a monad, whereas traditional definitions are
-not.  Our definition of SetT and proofs of monad laws are in the
-correctness/proofs section.]
+For the same monad $ℳ $, we can change the definition of $𝒞*$ to achieve a flow-sensitive path-insensitive analysis:
+\begin{align*}
+𝒞*    &: \PM(Call × ^Store^) → \PM(Call × ^Store^)           \\
+𝒞*(Σ) &≔ { <c',π₂(𝒞(c))(σ)> | c' ∈ π₁(𝒞(c))(σ) | <c,σ> ∈ Σ } \\
+\end{align*}
 
-As before, we must convert between monadic actions and abstract state space
-transitions to achieve an analysis:
-
-𝒞ₘ* : 𝒫 (Call) × ^Store^ → 𝒫 (Call) × ^Store^
-𝒞ₘ*(c𝒫,σ) = <{c' | c' ∈ π₁(𝒞ₘ c)(σ) | c ∈ c𝒫 }, ⋃ [ π₂(𝒞ₘ c)(σ) | c ∈ c𝒫 ]>
-
-This instantiation of ℳ  and 𝒞ₘ* yields a flow-insensitive analysis.
-
-For the same monad ℳ , we can change the definition of 𝒞ₘ* to achieve a
-flow-sensitive path-insensitive analysis:
-
-𝒞ₘ* : 𝒫 (Call × ^Store^) → 𝒫 (Call × ^Store^)
-𝒞ₘ*(Σ) = { <c',π₂(𝒞 ₘ c)(σ)> | c' ∈ π₁(𝒞ₘ c)(σ) | c ∈ c𝒫 }
-
-In the correctness framework, the flow-sensitive path-sensitive analysis and
-flow-insensitive analyses are justified through isomorphisms between monadic
-actions and abstract state space transitions.  The flow-sensitive
-path-insensitive variant is recovered by weakening this isomorphism to a galois
-connection.
+In the correctness framework, both the flow-sensitive path-sensitive analysis and flow-insensitive analyses are justified through isomorphisms between monadic actions and abstract state space transitions.
+The flow-sensitive path-insensitive variant is recovered by weakening this isomorphism to a galois connection.
 
@@
 
-We have now factored the abstract control properties of static analysis behind
-a common interface.  Now we show how to implement two intentional
-optimizations, abstract garbage collection and mcfa, in a completely general
-setting.
+Up to this point we have factored the abstract control properties of static analysis behind a common interface.  
+Now we show how to implement two intentional optimizations, abstract garbage collection and mcfa, in a completely general setting.
 
-The common interface for each variation in abstract control is a functor ℳ  ∈
-Set → Set such that ℳ :
+The common interface for each variation in abstract control is a monad $ℳ  ∈ Set → Set$ with $get$, $put$, and nondeterminism effects.
+We abreviate the conjunction of the above properties holding on some ℳ  as $Analysis(ℳ )$.
 
-* Is a monad
-  * return : ∀ (A : Set), A → ℳ  A
-  * extend : ∀ (A β : Set), (A → ℳ  β) → (ℳ  A → ℳ  β)
-* Is a state monad algebra on ^Store^
-  * get : ℳ  ^Store^
-  * put : ^Store^ → ℳ  1
-* Is a monad plus
-  * _+ₘ_ : ∀ (A : Set), ℳ  A → ℳ  A → ℳ  A
+\paragraph{Abstract Garbage Collection}
+Abstract garbage collection is a technique in abstract interpretation where unreachable abstract adresses are pruned from the state space.
+This is analagous to "real" garbage collection, where unreachable pointers are reclaimed for space efficiency.
+However, in abstract semantics, adresses are \emph{re-used} to soundly and finitely aproximate an infinite address space.
+Impreceision in control flow analyses arise when an abstract address present in the current store must be used for allocation.
+Abstract garbage collection is a purely precision and performance improving optimization which removes unreachable abstract adresses.
 
-We abreviate the conjunction of the above properties holding on some ℳ  as
-(Analysis ℳ ).
+We will use our generic monadic interfact $Analysis(ℳ )$ for an arbitrary monad $ℳ $.
+Using this interface, abstract garbage collection can be implemented once in a purely generic way.
+\begin{align*}
+gc    &: Call → ℳ (1)                                      \\
+gc(c) &≔ do                                                \\
+      &σ ← get-σ                                           \\
+      &𝓉₀ ← 𝓉-𝒞(c)                                         \\
+      &let 𝓉 ≔ μ(𝓉) → 𝓉 ⊔ 𝓉-Var(𝓉)                         \\
+      &modify-σ (λ(σ) → ⋃ { [x ↦ v] | x ∈ 𝓉 ∧ σ(x) = v } ) \\
+\end{align*}
+This is literally the implementation of \emph{concrete} garbage collection.
+However, it seamlessly becomes abstract garbage collection when instantiated with the appropriate monad.
 
-Abstract garbage collection can now be written once, generic to this interface:
+\paragraph{MCFA}
+MCFA is an optimization that improves the asymptotic complexity of context-sensitive control flow analyses.
+kCFA explodes exponentially for $k > 0$ in functional analyses, causing extremely poor analysis performance in the worst case.
+However, an apparent paradox was discovered as kCFA was proven to be polynomial in the worst case for object-oriented programs.
+MCFA resolved the paradox by identifying the difference in analyses and transfering the polynomial behavior to functional analysis.
+The crux of MCFA is to create packed, copied closures rather than linked closures.
 
-gc : Call → ℳ  1
-gc(c) = do
-  σ ← get-σ
-  touched₀ ← touched-𝒞(c)
-  let touched := collect(touched-var)(touched₀)
-  modify-σ (λ σ . ⋃ [ {x ↦ v} | σ(x) = v ∧ x ∈ touched ])
+So far, our analysis has not used closures for abstract values; 0CFA only tracks lambdas without their closing context.
+Adding closures to abstract values is a straightforward addition to the examples we've shown.
+Also, our implementation implements all of these features in their full glory.
 
-The mcfa optimization of copying rather than linking closures also enjoys a
-fully generic implementation:
+The mcfa optimization of copying rather than linking closures similarly enjoys a fully generic implementation:
+\begin{align*}
+clo-copy       &: [Var] × Call → ℳ (Clo)             \\
+clo-copy(xs,c) &≔ do                                 \\
+               &let ys ≔ free-vars(c)                \\
+               &σ ← get-σ                            \\
+               &vs ← map* σ ys                       \\
+               &ρ ← { y ↦* v | (y,v) ∈ zip(ys, vs) } \\
+               &return <xs,c,ρ>                      \\
+\end{align*}
 
-clo-copy : ([Var] × Call) → ℳ  Lam
-clo-copy(xs,c) = do
-  let ys := free-vars(c)
-  σ ← get-σ
-  vs ← map* σ(_) ys
-  ρ ← { y ↦* v | (y,v) ∈ zip(ys, vs) }
-  return <xs,c,ρ>
-
-Using existing techniques, these optimizations must be proved correct for each
-instantiation of abstract control.  Our generalization over abstract control
-allows us to implement each of these optimizations once, and more importantly,
-the proof about these optimizations also transfer to their instantiations for
-free.
+Using existing techniques, these optimizations would need to be both implemented and proven correct for each instantiation of abstract control.
+Our generalization over abstract control allows us to implement each of these optimizations once.
+More importantly, the proofs about these optimizations need only happen once as well, and the proofs transfer to their instantiations for free.
 
@@
 
-The key advantage to our formulation of monadic abstract interpretation--both
-over non-monadic approaches like AAM, and the monadic approach described by
-Sergey et al. (PLDI 2013)--is that the proofs of correctness for the
-constructed analyses can be derived automatically alongside the computational
-artifact.
+The key advantage to our framework is that the proofs of correctness for constructed analyses are derived automatically.
 
-Our monadic abstraction is a proper abstraction in that it mediates between two
-parties--the framework and the analysis designer--and gives guarantees to both
-sides conditioned on assumptions being met.
+We must now relate monadic actions like 𝒞 back to small step semantics.
+To do this we establish a galois connection between monadic actions in ℳ  and transitions functions for \emph{some} abstract state space $𝒮𝒮$.
+This abstract state space is constructed from the monad transformer stack, however some transformer stacks support multiple abstract state spaces.
+This step is notated $(A → ℳ (B))\galois{α}{\gamma}(𝒮𝒮(A) → 𝒮𝒮(B))$.
+We call a particular monad $ℳ $ which enjoys this property a \emph{small-step monad}.
 
-We show that language-independent monads exist for instantiating a semantics to
-flow-sensitive and path-sensitive analyses.
+In our framework we prove that not only are $\ST$ and $\PT$ monad transformers, they're small-step monad transformers.
+This means that for any stack of interleaving $\ST$ and $\PT$ monad transformers, one can construct the needed galois connection.
 
-In the core analysis framework we prove:
+In our framework we prove that $\ST(𝓈)(\PM) ⊑ \PT(\SM(𝓈))$.
+These monad interleavings correspond to path-sensitivity and path-insensitivity respectively.
 
-* SetT and StateT are monad transformers and compose in both directions.
-* Monadic actions (A → ℳ  β) in SetT and StateT form galois connections with
-  abstract state machine transitions (𝒮𝒮ₘ A → 𝒮𝒮ₘ β).
-* One composition of SetT and StateT is in (⊑) relation with the other.
-* For (SetT ∘ StateT), one choice of state machine 𝒮𝒮ₘ is in (⊑) relation with
-  the other.
+In our framework we prove that for the $\PT(\SM(𝓈))$ monad, two galois connections are possible to the state space $\PM(\_ × 𝓈)$.
+These choices for galois connection correspond to flow-sensitivity and flow-insensitivity respectively.
 
-The analysis designer must then prove:
+Independent of language or application, we can now conclude abstractly (yet precisely):
+\begin{align*}
+flow-insensitive ⊑ flow-sensitive path-insensitive ⊑ flow-sensitive 
+\end{align*}
 
-* The semantic step function 𝒞ₘ (which calls 𝒜 ₘ) is monotonic, including the
-  choice of 𝓂 .
-* The semantics as written, including intensional optimizations, are correct.
+When using our framework, the analysis designer need only prove prove:
+\begin{itemize}
+\item The semantic step function 𝒞 (which calls 𝒜 ) is monotonic, including monotonic in the choice of 𝓂 .
+\item The semantics as written, including intensional optimizations, are correct.
+\end{itemize}
+After supplying these proofs, the analysis designer enjoys:
+\begin{itemize}
+\item An automatically derived analysis for their language with their choice of abstract control.
+\item A proof-by-construction of correctness for the derived analysis.
+\item Seamless extensions to AAM and intensional optimizations (like gc or mcfa).
+\end{itemize}
 
-Given proofs from both sides of the interface we can conclude:
+Our proof technique is maximally modular in that new monads can be composed seamlessly into the framework with disruption.
+This modularity greatly reduces the proof burden on the analysis designer as new languages and analyses are developed.
 
-flow-insensitive 0-cfa 
- ⊑ 
-flow-sensitive path-insensitive 0-cfa 
- ⊑ 
-flow-sensitive 0-cfa
- ⊑ 
-concrete cfa
-
-This proof technique is maximally modular, and greatly reduces the proof burden
-on the analysis designer as new languages and analyses are developed.
-
-When adding the rest of the analysis tuning knobs like abstract domain, context
-sensitivity, intensional optimizations and object sensitivity, the monotonicity
-of 𝒞ₘ must be re-established.  However, like for abstract control, galois
-connections can be built between these choices independent of their
-instantiation in 𝒞ₘ. Our implementation supports each of these axis in full.
+In practice, there will be many more state space components, like abstract time and abstract address in the AAM framework.
+To add these features, the analysis designer need only stack more state space monads together to augment the resulting abstract state machine.
+However, once monotonicity of the $𝒞$ action is established, both concrete and abstract interpreters can be derived for free.
+This is only possible because all proofs have been decomposed to the unit of monad transformer.
 
@@
 
+This section details all structures used in this paper and their proofs.
+
 % Setup {{{
@@
 
-PartialOrder(A : Set) :=
+\begin{verbatim}
+PartialOrder(A : Set) ≔
   operators
-    _⊑_ : A → A → Prop
+    ⊑ : A → A → Prop
   laws
@@
     transitivity : x ⊑ y ⇒  y ⊑ z ⇒  x ⊑ z
+\end{verbatim}
 
-Monotonic(𝑓 : A → β) := 
+\begin{verbatim}
+Monotonic(𝑓 : A → B) ≔ 
   require
     PartialOrder(A)
-    PartialOrder(β)
+    PartialOrder(B)
     ∀ (x y : A), x ⊑ y ⇒  f(x) ⊑ f(y)
+\end{verbatim}
 
-(C : Set) α⇄ γ (A : Set) :=
+\begin{verbatim}
+(C : Set) α⇄ γ (A : Set) ≔
   require
@@
     α ∘ γ ⊑ id ⊑ γ ∘ α
+\end{verbatim}
 
-JoinSemilattice(A : Set) :=
+\begin{verbatim}
+JoinSemilattice(A : Set) ≔
   require
@@
     ⊥ : A
-    _⊔_ : A → A → A
+    ⊔ : A → A → A
   laws
@@
     commutativity : x ⊔ y = y ⊔ x
+\end{verbatim}
 
-Monad(𝓂  : Set → Set) :=
+\begin{verbatim}
+Monad(𝓂  : Set → Set) ≔
   operators
     return : ∀ (A : Set) → A → 𝓂 (A)
-    extend : ∀ (A β : Set) → (A → 𝓂 (β)) → (𝓂 (A) → 𝓂 (β))
+    extend : ∀ (A B : Set) → (A → 𝓂 (B)) → (𝓂 (A) → 𝓂 (B))
   laws
@@
     associativity : extend(k₂)(extend(k₁)(aM)) = extend(λ x → extend(k₂)(k₁(x)))(A)
+\end{verbatim}
 
-Transformer(𝒯 : (Set → Set) → (Set → Set)) :=
+\begin{verbatim}
+Transformer(𝒯 : (Set → Set) → (Set → Set)) ≔
   require
@@
     unit : lift(return) = return
+\end{verbatim}
 
-MonadState(𝓈 : Set)(𝓂  : Set → Set) :=
+\begin{verbatim}
+MonadState(𝓈 : Set)(𝓂  : Set → Set) ≔
   operators
@@
     put : 𝓈 → 𝓂  1
+\end{verbatim}
 
-MonadPlus(𝓂  : Set → Set) := 
+\begin{verbatim}
+MonadPlus(𝓂  : Set → Set) ≔ 
   require
@@
     distributivity : extend(k)(aM₁ ⊔ aM₂) = extend(k)(aM₁) ⊔ extend(k)(aM₂)
+\end{verbatim}
 
-Functorial(ℭ : Set → Prop)(𝓂  : Set → Set) := 
+\begin{verbatim}
+Functorial(ℭ : Set → Prop)(𝓂  : Set → Set) ≔ 
   require
     ∀ (A : Set), ℭ (A) → ℭ (𝓂 (A))
+\end{verbatim}
 
-Pointed(𝓂  : Set → Set) :=
+\begin{verbatim}
+Pointed(𝓂  : Set → Set) ≔
   unit : ∀ (A : Set) → A → 𝓂 (A)
+\end{verbatim}
 
-MonadSmallStep(𝓂  : Set → Set)(𝒮𝒮 : Set → Set) := 
+\begin{verbatim}
+MonadSmallStep(𝓂  : Set → Set)(𝒮𝒮 : Set → Set) ≔ 
   require
-    ∀ (A β : Set), (A → 𝓂 (β)) α⇄ γ (𝒮𝒮(A) → 𝒮𝒮(β))
+    ∀ (A B : Set), (A → 𝓂 (B)) α⇄ γ (𝒮𝒮(A) → 𝒮𝒮(B))
+\end{verbatim}
 
@@
 % SetT {{{
-
-\subsection{SetT}
+\subsection{\ST}
 \label{section:Proofs+Definitions:SetT}
 
-SetT : (Set → Set) → (Set → Set)
-SetT(𝓂 )(A) := 𝓂 (𝒫 (A))
+\begin{verbatim}
+\ST : (Set → Set) → (Set → Set)
+\ST(𝓂 )(A) ≔ 𝓂 (\PM(A))
 
 -----------------
-Transformer(SetT)
+Transformer(\ST)
 
-lift : ∀ (𝓂  : Set → Set) (A : Set), 𝓂 (A) → SetT(𝓂 )(A)
-lift := map (λ x → { x })
+lift : ∀ (𝓂  : Set → Set) (A : Set), 𝓂 (A) → \ST(𝓂 )(A)
+lift ≔ map (λ x → { x })
 
@@
 -------------------------------------------
-Monad(SetT(𝓂 ))
+Monad(\ST(𝓂 ))
 
-return : ∀ (A : Set), A → SetT 𝓂  A
-return := lift ∘ returnₘ
+return : ∀ (A : Set), A → \ST 𝓂  A
+return ≔ lift ∘ returnₘ
 
-extend : ∀ (A β : Set), (A → SetT(𝓂 )(β)) → (SetT(𝓂 )(A) → SetT(𝓂 )(β))
-extend(k) := joinₘ ∘ mapₘ (joins ∘ mapₚ k)
+extend : ∀ (A B : Set), (A → \ST(𝓂 )(B)) → (\ST(𝓂 )(A) → \ST(𝓂 )(B))
+extend(k) ≔ joinₘ ∘ mapₘ (joins ∘ mapₚ k)
 
@@
 -------------------------------------------
-MonadPlus(SetT(𝓂 ))
+MonadPlus(\ST(𝓂 ))
 
-m0 : ∀ (A : Set), SetT 𝓂  A
-m0 = ⊥ ₘ
+m0 : ∀ (A : Set), \ST 𝓂  A
+m0 ≔ ⊥ ₘ
 
-_<+>_ : ∀ (A : Set) → SetT 𝓂  A → SetT 𝓂  A → SetT 𝓂  A
-_<+>_ := _⊔_
+<+> : ∀ (A : Set) → \ST 𝓂  A → \ST 𝓂  A → \ST 𝓂  A
+<+> ≔ ⊔
 
@@
 
-get : SetT 𝓂  𝓈
-get = lift getₘ
+get : \ST 𝓂  𝓈
+get ≔ lift getₘ
 
-put : 𝓈 → SetT 𝓂  1
-put = lift ∘ putₘ
+put : 𝓈 → \ST 𝓂  1
+put ≔ lift ∘ putₘ
 
@@
 -------------------------------
-MonadSmallStep(SetT 𝓂 )(𝒮𝒮 ∘ 𝒫)
+MonadSmallStep(\ST 𝓂 )(𝒮𝒮 ∘ \PM)
 
-α : ∀ (A β : Set), (A → SetT(𝓂 )(β)) → (𝒮𝒮(𝒫 (A)) → 𝒮𝒮(𝒫 (β)))
-α(f) = αₘ(joinsₘ ∘ mapₚ(f))
+α : ∀ (A B : Set), (A → \ST(𝓂 )(B)) → (𝒮𝒮(\PM(A)) → 𝒮𝒮(\PM(B)))
+α(f) ≔ αₘ(joinsₘ ∘ mapₚ(f))
 
-γ : ∀ (A β : Set), (𝒮𝒮(𝒫 (A)) → 𝒮𝒮(𝒫 (β))) → (A → SetT(𝓂 )(β))
-γ(f) = γₘ(f ∘ mapₛₛ(unitₚ))
+γ : ∀ (A B : Set), (𝒮𝒮(\PM(A)) → 𝒮𝒮(\PM(B))) → (A → \ST(𝓂 )(B))
+γ(f) ≔ γₘ(f ∘ mapₛₛ(unitₚ))
 
-MonadSmallStep(𝓂 )(𝒮𝒮) ∧ (𝒫 ∘ 𝒮𝒮) α⇄ γ (𝒮𝒮 ∘ 𝒫 )
+MonadSmallStep(𝓂 )(𝒮𝒮) ∧ (\PM ∘ 𝒮𝒮) α⇄ γ (𝒮𝒮 ∘ \PM)
 -------------------------------------
-MonadSmallStep(SetT 𝓂 )(𝒫 ∘ 𝒮)
+MonadSmallStep(\ST 𝓂 )(\PM ∘ 𝒮)
 
-α : ∀ (A β : Set), (A → SetT(𝓂 )(β)) → (𝒫 (𝒮𝒮(A) → 𝒫 (𝒮𝒮(β))))
-α(f) = extendₚ(α ∘ αₘ(f))
+α : ∀ (A B : Set), (A → \ST(𝓂 )(B)) → (\PM(𝒮𝒮(A) → \PM(𝒮𝒮(B))))
+α(f) ≔ extendₚ(α ∘ αₘ(f))
 
-γ : ∀ (A β : Set), (𝒫 (𝒮𝒮(A) → 𝒫 (𝒮𝒮(β)))) → (A → SetT(𝓂 )(β))
-γ(f) = extendₚ(γ ∘ γₘ(f))
+γ : ∀ (A B : Set), (\PM(𝒮𝒮(A) → \PM(𝒮𝒮(B)))) → (A → \ST(𝓂 )(B))
+γ(f) ≔ extendₚ(γ ∘ γₘ(f))
+\end{verbatim}
 
@@
 
-\subsection{StateT}
+\subsection{$\ST$}
 \label{section:Proofs+Definitions:StateT}
 
-StateT : Set → (Set → Set) → (Set → Set)
-StateT(𝓈)(𝓂 )(A) := 𝓈 → 𝓂  (A × 𝓈)
+\begin{verbatim}
+\ST : Set → (Set → Set) → (Set → Set)
+\ST(𝓈)(𝓂 )(A) ≔ 𝓈 → 𝓂  (A × 𝓈)
 
-Transformer(StateT(𝓈))
+Transformer(\ST(𝓈))
 ----------------------
 
-lift : ∀ (𝓂  : Set → Set) (A : Set), 𝓂  a → StateT(𝓈)(𝓂 )(A)
-lift aM := λ 𝓈 → mapₘ (,𝓈) aM
+lift : ∀ (𝓂  : Set → Set) (A : Set), 𝓂  a → \ST(𝓈)(𝓂 )(A)
+lift aM ≔ λ 𝓈 → mapₘ (,𝓈) aM
 
@@
 --------------------
-Monad(StateT(𝓈)(𝓂 ))
+Monad(\ST(𝓈)(𝓂 ))
 
-return : ∀ (A : Set), A → StateT(𝓈)(𝓂 )(A)
-return := lift ∘ returnₘ
+return : ∀ (A : Set), A → \ST(𝓈)(𝓂 )(A)
+return ≔ lift ∘ returnₘ
 
-extend : ∀ (A β : Set), (A → StateT(𝓈)(𝓂 )(β)) → (StateT(𝓈)(𝓂 )(A) → StateT(𝓈)(𝓂 )(β))
-extend(k)(aM) := λ 𝓈 → doₘ
+extend : ∀ (A B : Set), (A → \ST(𝓈)(𝓂 )(B)) → (\ST(𝓈)(𝓂 )(A) → \ST(𝓈)(𝓂 )(B))
+extend(k)(aM) ≔ λ 𝓈 → doₘ
   (a,𝓈') ← aM(𝓈)
@@
 ------------------------------------------
-Functorial(JoinSemilattice)(StateT(𝓈)(𝓂 ))
+Functorial(JoinSemilattice)(\ST(𝓈)(𝓂 ))
 
-bot : StateT(𝓈)(𝓂 )(A)
-bot = λ _ → (⊥ ₛ, ⊥ ₐ)
+bot : \ST(𝓈)(𝓂 )(A)
+bot ≔ λ(𝓈)→ (⊥ ₛ, ⊥ ₐ)
 
-_⊔_ : StateT(𝓈)(𝓂 )(A) → StateT(𝓈)(𝓂 )(A) → StateT(𝓈)(𝓂 )(A)
-aM₁ ⊔ aM₂ := λ 𝓈 → doₘ
+⊔ : \ST(𝓈)(𝓂 )(A) → \ST(𝓈)(𝓂 )(A) → \ST(𝓈)(𝓂 )(A)
+aM₁ ⊔ aM₂ ≔ λ 𝓈 → doₘ
   (a₁,𝓈₁) ← aM₁
@@
 ------------------------
-MonadPlus(StateT(𝓈)(𝓂 ))
+MonadPlus(\ST(𝓈)(𝓂 ))
 
-⊥ : ∀ (A : Set), StateT(𝓈)(𝓂 )(A)
-⊥ := lift ⊥ ₘ
+⊥ : ∀ (A : Set), \ST(𝓈)(𝓂 )(A)
+⊥ ≔ lift ⊥ ₘ
 
-_⊔_ : ∀ (A : Set), StateT(𝓈)(𝓂 )(A)
-aM₁ ⊔ aM₂ := λ 𝓈 → aM₁(𝓈) <+> aM₂(𝓈)
+⊔ : ∀ (A : Set), \ST(𝓈)(𝓂 )(A)
+aM₁ ⊔ aM₂ ≔ λ 𝓈 → aM₁(𝓈) <+> aM₂(𝓈)
 
@@
 ----------------------------
-MonadState(𝓈)(StateT(𝓈)(𝓂 ))
+Monad\ST(𝓈)(\ST(𝓈)(𝓂 ))
 
-get : StateT 𝓈 𝓂  𝓈
-get = λ 𝓈 → return (𝓈,𝓈)
+get : \ST 𝓈 𝓂  𝓈
+get ≔ λ 𝓈 → return (𝓈,𝓈)
 
-put : 𝓈 → StateT 𝓈 𝓂  1
-put 𝓈 = λ _ → return (∙,𝓈)
+put : 𝓈 → \ST 𝓈 𝓂  1
+put(𝓈) ≔ λ (𝓈') → return (∙,𝓈)
 
@@
 --------------------------------------
-MonadSmallStep(StateT(𝓈)(𝓂 ))(𝒮𝒮(_ × 𝓈))
+MonadSmallStep(\ST(𝓈)(𝓂 ))(𝒮𝒮(\_ × 𝓈))
 
-α : ∀ (A B : Set), (A → StateT(𝓈)(𝓂 )(B)) → (𝒮𝒮(A × 𝓈) → 𝒮𝒮(B × 𝓈))
-α(f) = αₘ (λ (a,𝓈) → f(a)(𝓈))
+α : ∀ (A B : Set), (A → \ST(𝓈)(𝓂 )(B)) → (𝒮𝒮(A × 𝓈) → 𝒮𝒮(B × 𝓈))
+α(f) ≔ αₘ (λ (a,𝓈) → f(a)(𝓈))
 
-γ : ∀ (A B : Set), (𝒮𝒮(A × 𝓈) → 𝒮𝒮(B × 𝓈)) → (A → StateT(𝓈)(𝓂 )(B))
-γ(f) = λ a 𝓈 → γₘ (f)(a,s)
+γ : ∀ (A B : Set), (𝒮𝒮(A × 𝓈) → 𝒮𝒮(B × 𝓈)) → (A → \ST(𝓈)(𝓂 )(B))
+γ(f) ≔ λ(a,𝓈) → γₘ (f)(a,s)
+\end{verbatim}
 
@@
 \section{Propositions}
-\lable{section:Proofs+Definitions:Propositions}
+\label{section:Proofs+Definitions:Propositions}
 
-* SetT and StateT are Galois Functors.
-* 𝒜 ₘ and 𝒞ₘ are monotonic (including in 𝓂 ).
+\begin{itemize}
+\item $\PT$ and $\ST$ are Galois Functors.
+\item $\PT$ and $\ST$ are MonadStep Functors.
+\item 𝒜  and 𝒞 are monotonic (including in 𝓂 ).
+\end{itemize}
 
@@
 
+% Bibliography {{{
 \bibliography{davdar}{}
 \bibliographystyle{plain}
+% }}}
 

### 2014-09-21 writeup/darais-quals.tex
things are getting polished.  next to polish is proofs.

@@
 We use traditional notation for function definitions and applications rather than lambda calculus notation.
-For example, we will write $f(w, x) ≔ g(y, z)$ rather than $f\; w\; x ≔ g\; y\; z$.
+For example, we will write:
+\begin{equation*}
+compose(g, f, x) ≔ g(f(x))
+\end{equation*}
+rather than:
+\begin{equation*}
+compose\; g\; f\; x ≔ g\; (f\; x)
+\end{equation*}
 
 We use lambda notation for anonymous functions.
-For example, defining the identify function could be notated $id ≔ λ(x) → x$ rather than the named definition $id(x) ≔ x$.
+For example, the anonymous composition function is written:
+\begin{equation*}
+λ(g, f, x) → g(f(x))
+\end{equation*}
 
@@
 
-A basic familiarity with the concept of monads will be useful for understanding the rest of the paper.
+A basic familiarity with the concept of monads will be necessary for understanding this paper.
 Some find it useful to review the concept of functor before learning about monads. 
@@
 
-A type $𝒯 : Set → Set$ is called a \emph{functor} if one can:
+A type $(𝒯 : Set → Set)$ is called a \emph{functor} if one can:
 \begin{itemize}
-\item Define $map : ∀ (A : Set), (A → B) → (𝒯(A) → 𝒯(B))$
-\item Prove $map(id) = id$, where $id$ is the identity function $λ(x) → x$
+\item Define $map : ∀ (A, B : Set), (A → B) → (𝒯(A) → 𝒯(B))$
+\item Prove $map(id) = id$, where $id$ is the identity function $(λ(x) → x)$
 \item Prove $map (g ∘ f) = map(g) ∘ map(f)$
@@
           Nil    &→ Nil                \\
-          x ∷ xs &→ f(x) ∷ map(f)(xs') \\
+         x ∷ xs' &→ f(x) ∷ map(f)(xs') \\
 \end{align*}
@@
 
-Likewise a type $𝒯 : Set → Set$ is called a \emph{monad} if one can:
+Likewise a type $(𝒯 : Set → Set)$ is called a \emph{monad} if one can:
 \begin{itemize}
-\item Define $extend : ∀ (A : Set), (A → 𝒯(B)) → (𝒯(A) → 𝒯(B))$
+\item Define $extend : ∀ (A, B : Set), (A → 𝒯(B)) → (𝒯(A) → 𝒯(B))$
 \item Prove unit and associativity laws (not mentioned here).
 \end{itemize}
-The only difference from the definition of functor is in the the first argument $(A → 𝒯(B))$.
-For monads, this argument is allowed to be a \emph{monadic} function, rather than a pure function $(A → B)$.
+The only difference from the definition of functor is in the the first argument.
+For monads, this argument is allowed to be a \emph{monadic} $(A → 𝒯(B))$, rather than a pure function $(A → B)$.
 
@@
 \end{align*}
-$extend$ for lists can be understood as $map$ followed by the flattening operation $concat$.
 
@@
 
+$extend$ for lists can also be understood as $concat$ followed by $map$:
 \paragraph{Example:} 
-$extend$ for lists can also be understood as $concat$ followed by $map$:
 \begin{align*}
@@
 \end{align*}
-For all monads, it is equivalent to define $return$, $map$ and $join$, where $join : ∀ (A : Set), 𝒯(𝒯(A)) → 𝒯(A)$.
+For all monads, it is equivalent to define $return$, $map$ and $join$, where:
+\begin{equation*}
+join : ∀ (A : Set), 𝒯(𝒯(A)) → 𝒯(A)
+\end{equation*}
 As we have just seen, $join$ for the list monad is just $concat$.
@@
 Where a monad $𝒯$ will have type $Set → Set$, a monad transformer will have type $(Set → Set) → (Set → Set)$.
-Perhaps a more appropriate type is $Monad → Monad$.
 
-Monad transformers are used to add an effect on top of another monad.
+Monad transformers are used to extend an existing monad to support another effect.
 The three monads used in this work are the state monad, powerset monad, and the identity monad.
@@
 \paragraph{Example:}
-$\ST(ℤ)(\PT(ID))$ is a monad which has $get$ and $put$ effects for a single integer state, in addition to nondeterminism effects.
+$\ST(ℤ)(\PT(ID))$ is a monad which has $get$ and $put$ effects for a single cell of integer state, in addition to nondeterminism effects.
 
@@
 
-This work follows the small-step operational approach to language semantics pioneered independently by Plotkin and Felleisen.
-In small-step operational semantics, the meaning of an expression $e$ in a language is given by the transitive closure of some step relation $e ↦ e'$.
+This work follows the small-step operational approach to language semantics pioneered by Plotkin and Felleisen.
+In small-step operational semantics, the meaning of an expression $e$ of a language is given by the transitive closure of some step relation $e ↦ e'$.
 This yields a state-machine approach to semantics.
 Exploring the meaning of $e$ reduces to exploring all possible transitions reachable from $e$ through $↦$.
-This contrasts to previously developed deontational approaches, where the meaning of $e$ would be given by a denotation function $⟦ e ⟧$.
-This denotation function can quickly demand lots of mathematical machinery, and the small-step operational method side-steps this issue.
+This contrasts to previously developed deontational approaches, where the meaning of $e$ must be given all at once by a denotation function $⟦ e ⟧$.
+This denotation function can quickly demand lots of mathematical machinery.
+Small-step operational methods side-step the need for such machinery.
 
@@
 In the setting of abstract interpretation, a program analysis is just an alternate semantics over an abstract domain.
-To distinguish the original semantics and the alternate semantics, we call the original semantics the \emph{concrete} semantics, and the alternate semantics the \emph{abstract} semantics\footnote{
-  This terminology quickly becomes confusing because we will eventually construct an abstract machine for both semantics, yielding a concrete abstract machine and an abstract abstract machine. 
-  This pun is exploited mercilessly in the work of Might and Van Horn as they "abstract" abstract abstract machines, yielding an abstract abstract abstract machine.
+To distinguish the original semantics from the alternate semantics we call the original semantics the \emph{concrete} semantics.
+The alternate semantics is called the \emph{abstract} semantics\footnote{
+  This terminology quickly becomes confusing because we will eventually construct an abstract machine for both semantics.
+  This yields a concrete abstract machine and an abstract abstract machine. 
+  This linguistic pun is exploited mercilessly in the work of Might and Van Horn as they “abstract” abstract abstract machines, yielding abstract abstract abstract machines.
 }.
@@
 \item 
-  A relationship between $\C$ and $\A$ that defines when $(\Ce : \C)$ and $(\Ae : \A)$ are related.
+  A relationship between $\C$ and $\A$ that defines when some $(\Ce' : \C)$ and $(\Ae' : \A)$ are related.
   This often takes the form of projecting $\A$ to $\PM(\C)$ and using the subset relation.
 \item 
-  An abstract version of $\Ce$ that lives in $\A$: $(\Ae : \A)$.
+  An abstract version of our concrete program $\Ce$: $(\Ae : \A)$.
 \item 
@@
 \item 
-  A way to explore every state reachable by $\Ae$ under $\Asteps$.
+  A method to explore every state reachable by $\Ae$ under $\Asteps$.
   This often requires $\A$ to be finite.
 \end{itemize}
-By convention, we put hats ($\widehat{}$) on things to name their abstract cousins.
-The content of the approach is shown in the following picture:
+By convention, we put hats ($\widehat{\;\;}$) on things to name their abstract cousins.
+The overall approach is summarized in the following picture:
 \input{diagrams/AI.tikz}
@@
 In order to be a correct analysis, it must be guaranteed that $\Ce'$ and $\Ae'$ will be related.
-The logical structure of the picture is given by:
+The logical structure of the picture, which states the correctness of the analysis, is given by:
 \begin{equation*}
@@
 The AI setting can be elegently simplified and enriched through the use of \emph{galois connections}.
-Galois connections serve as a unifying framework for establishing the "relationship between $\C $ and $\A$" mentioned in the previous section.
+Galois connections serve as a unifying framework for establishing the “relationship between $\C $ and $\A$” mentioned in the previous section.
 
@@
 \item $(γ : \A → \C )$ where $γ$ is monotonic
-\item $(γ ∘ α)$ is expansive: $∀ (x : \C ), x ⊑ γ(α(x))$
-\item $(α ∘ γ)$ is contractive: $∀ (y : \A), α(γ(y)) ⊑ y$
+\item A proof that $(γ ∘ α)$ is expansive: $∀ (x : \C ), x ⊑ γ(α(x))$
+\item A proof that $(α ∘ γ)$ is contractive: $∀ (y : \A), α(γ(y)) ⊑ y$
 \end{itemize}
 The last two properties can be succinctly stated as $(α ∘ γ ⊑ id ⊑ γ ∘ α)$\footnote{
-  This uses the logical monotonicity relation $f ⊑ g ⇔  (x ⊑ y ⇒  f(x) ⊑ g(y))$ for the function space
+  This uses the logical monotonicity relation $f ⊑ g ⇔  (x ⊑ y ⇒  f(x) ⊑ g(y))$ for the function space.
 }.
@@
 \paragraph{Example:}
-Given a galois connection $\C\galois{α}{γ}\A$, there exists a galois connection $\C → \C  \galois{α}{γ} \A → \A$ where:
+Given a galois connection $\C\galois{α}{γ}\A$, there exists a galois connection $(\C → \C)\galois{α'}{γ'}(\A → \A)$ where:
 \begin{align*}
-α(f : \C → \C) &≔ γ ∘ f ∘ α \\
-γ(g : \A → \A) &≔ α ∘ g ∘ γ \\
+α'(f : \C → \C) &≔ α ∘ f ∘ γ \\
+γ'(g : \A → \A) &≔ γ ∘ g ∘ α \\
 \end{align*}
@@
 \paragraph{Example:} 
-The language $(\PM(ℤ),+,*)$ forms a galois connection with the language $(\PM({EVEN, ODD}),∧,∨)$ where:
+The language $(\PM(ℤ),+,*)$ forms a galois connection with the language $(\PM(\{ EVEN, ODD \}),∧,∨)$ where:
 \begin{align*}
-α(zs : \PM(ℤ))           &≔ ⋃ { { EVEN | ∃ z ∈ zs ∧ Even(z)}, { ODD | ∃ z ∈ zs ∧ Odd(z) } }      \\
-γ(ts : \PM({EVEN, ODD})) &≔ ⋃ { { z ∈ ℤ | EVEN ∈ ts ∧ Even(z) }, { z ∈ ℤ | ODD ∈ ts ∧ Odd(z) } } \\
+α(zs : \PM(ℤ))               &≔ ⋃ \{ \{ EVEN  \;|\; ∃ z ∈ zs ∧ Even(z) \},  \{ ODD   \;|\; ∃ z ∈ zs ∧ Odd(z) \} \} \\
+γ(ts : \PM(\{ EVEN, ODD \})) &≔ ⋃ \{ \{ z ∈ ℤ \;|\; EVEN ∈ ts ∧ Even(z) \}, \{ z ∈ ℤ \;|\; ODD ∈ ts ∧ Odd(z) \} \} \\
 \end{align*}
@@
 Galois connections simplify the AI framework by using $x ⊑ γ(y)$ or (equivalently) $α(x) ⊑ y$ as the relation $(x R y)$.
-Put differently, galois connections are a natural and general way of placing partial orders \emph{on sets themselves} (note that x and y live in different sets).
+Galois connections are a natural and general way of placing partial orders \emph{on sets themselves}.
+$x ⊑ γ(y)$ can be seen as a heterogenous extension of $x ⊑ y$ when $(x : A)$ and $(y : B)$ live in different sets. 
+This heterogenous order is given meaning when there exists a galois connection $A\galois{α}{γ}B$.
 One can also think of galois connections as something like an isomorphism, but with a weaker round-trip property.
@@
 \end{itemize}
-The content of the approach is shown in the following picture:
+The overall approach is summarized in the following picture:
 \input{diagrams/AIGC.tikz}
-This picture takes place at some point in the analysis process.
-$\Ce$ is the current program (possibly the result of running the base program for a little while).
-$\Ae$ is a valid abstraction of $\Ce$ by construction as $α(\Ce)$.
-$\Ce'$ is some next state of execution of $\Ce$, and likewise for $\Ae$/$\Ae'$.
-In order to be a correct analysis, it must be guaranteed that $\Ce' ⊑ γ(\Ae')$.
-The logical structure of the picture is given by:
+As before, the logical structure of the picture is given by:
 \begin{equation*}
@@
 \end{equation*}
+The statement of this property can be simplified further as $\Csteps ⊑ γ(\Asteps)$.
+This uses the definition of galois connections for function spaces shown in a previous example.
 
-The logical structure of the previous figure is the \emph{soundness} property of the analysis.
-This property can be simplified further by the statement $\Csteps ⊑ γ(\Asteps)$.
-This uses the definition of galois connections for function spaces shown in a previous example.
-What is more, this more compact statement of soundness also captures another property, the \emph{tightness} of the analysis.
-Unfolded, this property states:
-\begin{equation*}
-∀ (\Ce,\Ce' : \C ) (\Ae,\Ae' : \A), (\Ce' ⊑ γ(\Ae')) ∧ (\Ce \Csteps \Ce') ∧ (\Ae \Asteps \Ae') ⇒  (\Ce ⊑ γ(\Ae))
-\end{equation*}
-Tightness tells us that our abstraction is provably \emph{the best abstraction possible}.
-The beauty of galois connections is how they capture both soundness and tightness of an analysis with one concise statement: $\Csteps ⊑ γ(\Asteps)$.
+Using a galois connection $\C\galois{α}{γ}\A$ and abstract step function $\Asteps ⊑ γ(\Csteps)$, the analysis process becomes more specific:
+\begin{itemize}
+\item Translate $(\Ce : \C)$ to $(α(\Ce) : \A)$
+\item Explore all abstract states $\Ae'$ states reachable from $α(\Ce)$.
+\item All reachable concrete states are summarized by projecting each $γ(\Ae')$ back to $\C$.
+\end{itemize}
+
+The galois connection framework simultaneously guarantees the \emph{soundness} and \emph{tightness} of this method.
+Soundness tells us that if we abstract, take an abstract step, and then concretize, that the result is an approximation of taking a concrete step.
+Soundness tells us that we can trust the results of the analysis.
+\input{diagrams/GCSoundness.tikz}
+Tightness tells us that if we concretize, take a concrete step, and then abstract, then the result must be just as precise as taking an abstract step.
+Tightness tells us that our abstract step function isn't losing precision unnecessarily, and that our abstract step $\Asteps$ is provably as strong as possible.
+\input{diagrams/GCTightness.tikz}
 
@@
 We define the CPS-IF language as follows:
-\begin{align*}
-x,y,k &: Var  &⩴ ..variables..                          \\
-b,i,l &: Lit  &⩴ ℤ ⋃ 𝔹                                  \\
-f,g   &: Lam  &⩴ λ(x) → c | λ(x,k) → c                  \\
-a     &: Atom &⩴ x | l | f | op(a)                      \\
-op    &: Op   &⩴ add1 | sub1 | gez                      \\
-c     &: Call &⩴ if(a){c}{c} | a(a) | a(a,a) | halt(a)  \\
-\end{align*}
+\input{snips/CPS-IF.tex}
 
@@
 Control flow analysis is a class of analysis which is particularly important for higher-order languages.
-In non-higher-order languages, it is useful to distinguish control-flow--which functions can be called--from data-flow--which values flow where.
+In non-higher-order languages, it is useful to distinguish control-flow from data-flow.
+The control flow of the program is a graph of what functions are called from where.
+The data flow of a program is a mapping of which values can flow to which variables.
 Traditionally one performs a control-flow analysis first, to find out which functions are called, and a data-flow analysis second, using the results of the control flow analysis.
 
-In higher-order languages, data-flow and control-flow become the same problem.
+In higher-order languages, data-flow and control-flow are tightly coupled.
 Before you can tell which functions will be called, you need to know how values flow, because functions are themselves values.
 
-We note that the distinction between "higher-order" languages and "non-higher-order" languages is a red herring.
-Functions can be passed as values in C using function pointers, and object-oriented languages enjoy a similar circularity between control- and data- flow from method dispatch.
-You can read these statements as saying "all languages are higher order, and thus need control flow analysis".
-Or you can read them as saying "all languages are higher order, and we've done just fine without control flow analysis so far".
-The distinction is in the code that you end up analysing.
+We note that the distinction between “higher-order” languages and “non-higher-order” languages is a red herring.
+Functions can be passed as values in C using function pointers, and object-oriented languages enjoy a similar circularity between control- and data- flow due to method dispatch.
+You can read these statements as saying “all languages are higher order, and thus all languages need higher order control flow analysis”.
+Or you can read them as saying “all languages are higher order, and we've done just fine without higher oder control flow analysis so far”.
+The real distinction is in the code that you end up analysing.
 
-We build on the tradition of control flow analysis pioneered by Shivers and its recent refinements developed by Might and Van Horn.
-In this tradition, the result of a control flow analysis is closer to what one would think of as a data-flow analysis.
-0CFA, the most basic control flow analysis, computes (conservatively) the set of values which might flow to a particular variable.
-kCFA is a suite of context sensitive extensions to 0CFA which analyses a given function separately for each possible calling context.
+We build on the tradition of control flow analysis (CFA) pioneered by Shivers, as well as recent refinements of CFA developed by Might and Van Horn.
+In this tradition, the result of a control flow analysis is closer to what one would think of as a data-flow analysis, only the process accounts for higher order flow.
+0CFA, the most basic control flow analysis, merely computes the set of values which might flow to a particular variable.
+kCFA is a suite of context sensitive extensions to 0CFA for which functions are analyzed separately for each possible calling context.
 We use 0CFA as the example analysis in this paper.
@@
 
-Monadic AI (Sergey et al. PLDI 2013) showed that computational modularity for abstract interpreters can be achieved through a monadic abstraction.
-We build directly on this work at the level of intuition--monads serve as our pivot-point for generalizing abstract control.
-However, in our pursuit of an axis-independent correctness framework, our approach deviates greatly from this work in the specific interfaces used.
+Monadic abstract interpretation[Sergey et al. PLDI 2013] demonstrated that abstract interpreters can be built modularly using a monadic abstraction.
+We build directly on this work at the level of intuition—monads serve as our pivot-point for achieving modularity.
+However, in our pursuit of a modular correctness framework, our approach deviates greatly from prior work in the specific interfaces used.
 
 First we demonstrate a simple 0CFA analysis on CPS-IF which doesn't use our framework.
-The language is the same as introduced in section \ref{section:Background:ContinuationPassingStyle}, and repeated here for convenience:
+The CPS-IF language is introduced in section \ref{section:Background:ContinuationPassingStyle}.
+We repeat the definition here for convenience.
+\input{snips/CPS-IF.tex}
+
+Our abstract semantics of 0CFA will track literals (including lambdas) that appear in the program text.
+For integers that do not appear in the program text, the analysis uses a single token $INT$ which conservatively approximates any possible integer.
+Formally, the state space $Σ$ for the abstract machine of the abstract semantics is defined as:
 \begin{align*}
-x,y,k &: Var  &≔ ..variables..                          \\
-b,i,l &: Lit  &≔ ℤ ⋃ 𝔹                                  \\
-f,g   &: Lam  &⩴ λ(x) → c | λ(x,k) → c                  \\
-a     &: Atom &⩴ x | l | f | op(a)                      \\
-op    &: Op   &⩴ add1 | sub1 | gez                      \\
-c     &: Call &⩴ if(a){c}{c} | a(a) | a(a,a) | halt(a)  \\
-\end{align*}
-The abstract semantics of 0CFA will track literals (including lambdas) that appear in the program text.
-For integers that do not appear in the program text, the analysis uses a single token $INT$ which conservatively approximates any possible integer.
-Because there are a finite many literals in the program text, there are a finite number of possible values.
-Formally, the state space for the abstract machine of the abstract semantics is defined as:
-\begin{align*}
-v &: \AVal   ⩴ INT | l | λ(x) → c | λ(x,k) → c \\
-σ &: \AStore ≔ Var → \PM(\AVal)                 \\
+v : \AVal   &⩴ \INT \;|\; l \;|\; \lam(x) → c \;|\; \lam(x,k) → c \\
+σ : \AStore &≔ Var → \PM(\AVal)                                   \\
 Σ           &≔ \PM(Call × \AStore)                                \\
 \end{align*}
+Because there are a finite many literals in the program text, we can claim there are a finite number of possible abstract values.
 
 The abstract semantics for $Σ$ comes in two parts.  
-First we define a \emph{denotation function} $(𝒜 : \AStore × Atom → \PM(\AVal))$ for $Atom$ expressions.
-Second we define a \emph{step relation} (as a function) $(𝒞 : Call × \AStore → \PM(Call × \AStore))$ for $Call$ expressions.
+First we define a \emph{denotation function} $𝒜 $ for $Atom$ expressions.
 \begin{align*}
@@
 𝒜 (σ,x)                      &≔ σ(x)                        \\
-𝒜 (σ,l)                      &≔ { l }                       \\
-𝒜 (σ,add1 a) | 𝒜 [ρ,σ] ⊑ INT &≔ { INT }                     \\
-𝒜 (σ,sub1 a) | 𝒜 [ρ,σ] ⊑ INT &≔ { INT }                     \\
-𝒜 (σ,gez a)  | 𝒜 [ρ,σ] ⊑ INT &≔ { T , F }                   \\
-𝒜 (σ,λ(x)(k) → c)            &≔ { λ(x)(k) → c }             \\
+𝒜 (σ,l)                      &≔ \{ l \}                     \\
+𝒜 (σ,\add a) | 𝒜 (σ,a) ⊑ INT &≔ \{ \INT \}                  \\
+𝒜 (σ,\sub a) | 𝒜 (σ,a) ⊑ INT &≔ \{ \INT \}                  \\
+𝒜 (σ,\gez a) | 𝒜 (σ,a) ⊑ INT &≔ \{ \TRUE , \FALSE \}        \\
+𝒜 (σ,\lam(x) → c)            &≔ \{ \lam(x) → c \}           \\
+𝒜 (σ,\lam(x)(k) → c)         &≔ \{ \lam(x)(k) → c \}        \\
 \end{align*}
 
-% \begin{align*}
-% 𝒞                   &: &Call × \AStore → \PM(Call × \AStore)                                     \\
-% 𝒞 <if(a){c₁}{c₂},σ> &≔ &{ <c,σ>                                                                 \\
-%                     &  &|             c &∈ ⋃ { { c1 } if T ∈ 𝒜 (σ,a), { c2 } if F ∈ 𝒜 (σ,a) }   \\
-%                     &  &}                                                                       \\
-% 𝒞 <a₁(a₂,a₃),σ>     &≔ &{ <c,σ'>                                                                \\
-%                     &  &| <λ(x)(k) → c> &∈ 𝒜 (σ,a₁)                                             \\
-%                     &  &|            v₂ &∈ 𝒜 (σ,a₂)                                             \\
-%                     &  &|            v₃ &∈ 𝒜 (σ,a₃)                                             \\
-%                     &  &|            σ' &≔ σ ⊔ [x ↦ v₂] ⊔ [k ↦ v₃]                              \\
-%                     &  &}                                                                       \\
-% \end{align*}
+Second we define a \emph{step relation} (as a function) $𝒞$ for $Call$ expressions.
+\begin{align*}
+𝒞                        : &Call × \AStore → \PM(Call × \AStore)                                              \\
+𝒞(\iif(a)\{c₁\}\{c₂\},σ) ≔ &\{ (c,σ)                                                                          \\
+                           &|\; c ∈ ⋃ \{ \{ c₁ \;|\; \TRUE ∈ 𝒜 (σ,a) \}, \{ c₂ \;|\; \FALSE ∈ 𝒜 (σ,a) \} \} \\
+                           &\}                                                                                \\
+𝒞(a₁(a₂,a₃),σ)           ≔ &\{ (c,σ')                                                                         \\
+                           &|\; (\lam(x)(k) → c) ∈ 𝒜 (σ,a₁)                                                   \\
+                           &|\;             v₂ ∈ 𝒜 (σ,a₂)                                                     \\
+                           &|\;             v₃ ∈ 𝒜 (σ,a₃)                                                     \\
+                           &|\;             σ' ≔ σ ⊔ [x ↦ v₂] ⊔ [k ↦ v₃]                                      \\
+                           &\}                                                                                \\
+𝒞(\halt(a),σ)            ≔ &\{ (\halt(a),σ) \}                                                                \\
+\end{align*}
+(We omit the $(a₁(a₂))$ case. It is directly analogous to $(a₁(a₂,a₃))$).
 
-The complete analysis of a program c is defined as the least fixed point of a
-\emph{collection semantics} for the relation $𝒞$:
+The complete analysis of a program $c$ is defined as the least fixed point of a \emph{collection semantics} for the relation $𝒞$:
 \begin{align*}
-μ(Σ) → {<c,⊥>} ⊔ 𝒞*(Σ)
+\text{analysis} ≔ μ(Σ) → \{(c,⊥)\} ⊔ 𝒞^{⋆}(Σ)
 \end{align*}
@@
 \begin{align*}
-𝒞*    &: \PM(Call × \AStore) → \PM(Call × \AStore) \\
-𝒞*(Σ) &≔ ⋃ { 𝒞 <c,σ> | <c,σ> ∈ Σ }               \\
+𝒞^{⋆}    &: \PM(Call × \AStore) → \PM(Call × \AStore) \\
+𝒞^{⋆}(Σ) &≔ ⋃ \{ 𝒞(c,σ) | (c,σ) ∈ Σ \}                  \\
 \end{align*}
-(The collecting semantics tracks all states that the program could be in rather than just the final states.)
+The collecting semantics tracks all states that the program could be in rather than just the final states.
 
 The first insight in monadic static analysis is to abbreviate the definitions of $𝒜 $ and $𝒞$ using a monad.
-At this point, the monad "trick" is nothing more than simplifying the definition. 
-This simplification is similar to how a functional programmer would use a state monad in place of explicit state passing.
+At this point, the monad “trick” is nothing more than a technique to simplify the definition of $𝒜 $ and $𝒞$.
+This simplification is similar to how a functional programmer would use a state monad in place of state passing style.
 
-We use a powerset monad and state monad transformer to write the analysis in monadic style.
-The full definitions of the monads used throughout this paper are defered to section \ref{section:Proofs+Definitions}.
-Because monad transformers are just fancy names for simple types, we write the simple types underneath definitions which use monad transformers.
-The monadic conversion of the above analysis is:
+We wll use a powerset monad and state monad transformer to write the analysis in monadic style.
+The full definitions of the monads (and their proofs) used throughout this paper are defered to section \ref{section:Proofs+Definitions}.
+Because monad transformers are just fancy names for simple types, we write the equivalent simple type underneath definitions which reference monad transformer.
+The monadic conversion of the above analysis is as follows.
 \begin{align*}
@@
 𝒜                &: Atom → ℳ (\PM(\AVal))                    \\
-𝒜 (x)            &≔ do                                       \\
-                 &σ ← get-σ                                  \\
-                 &return(σ(x))                               \\
-𝒜 (l)            &≔ return({l})                              \\
-𝒜 ₘ(add1(a))     &≔ do                                       \\
+𝒜 (x)            &≔ \ddo                                     \\
+                 &σ ← \getstore                              \\
+                 &\return(σ(x))                               \\
+𝒜 (l)            &≔ \return(\{l\})                              \\
+𝒜 ₘ(add1(a))     &≔ \ddo                                       \\
                  &v ← 𝒜 (a)                                  \\
-                 &return({INT | ∃ i ⊑ INT ∈ v})              \\
-𝒜 ₘ(sub1(a))     &≔ do                                       \\
+                 &\return(\{\INT \;|\; ∃ i ⊑ \INT ∈ v\})              \\
+𝒜 ₘ(sub1(a))     &≔ \ddo                                       \\
                  &v ← 𝒜 (a)                                  \\
-                 &return({INT | ∃ i ⊑ INT ∈ v})              \\
-𝒜 ₘ(gez(a))      &≔ do                                       \\
+                 &\return(\{\INT \;|\; ∃ i ⊑ \INT ∈ v\})              \\
+𝒜 ₘ(gez(a))      &≔ \ddo                                       \\
                  &v ← 𝒜 (a)                                  \\
-                 &return(⋃ { {T, F} | ∃ i ⊑ INT ∈ v})        \\
-𝒜 ₘ(λ(x)(k) → c) &≔ return({λ(x)(k) → c})                    \\
-                                                             \\
+                 &\return(\{\TRUE, \FALSE \;|\; ∃ i ⊑ \INT ∈ v\})      \\
+𝒜 ₘ(\lam(x)(k) → c) &≔ \return(\{\lam(x)(k) → c\})                    \\
+\end{align*}
+
+\begin{align*}
 𝒞                 &: Call → ℳ (Call)                         \\
-𝒞 (if(a){c₁}{c₂}) &≔ do                                      \\
+𝒞 (\iif(a)\{c₁\}\{c₂\}) &≔ \ddo                                      \\
                   &vP ← 𝒜 (a)                                \\
-                  &v ← lift\PM(vP)                           \\
-                  &b ← coerce𝔹(v)                            \\
-                  &return { if b then c₁ else c₂ }           \\
-𝒞ₘ(a₁(a₂,a₃))     &≔ do                                      \\
+                  &v ← \liftpowerset(vP)                           \\
+                  &b ← \coercebool(v)                            \\
+                  &\return(\{ \text{if}\; b\; \text{then}\; c₁\; \text{else}\; c₂ \}) \\
+𝒞ₘ(a₁(a₂,a₃))     &≔ \ddo                                      \\
                   &vP ← 𝒜 (a₁)                               \\
-                  &v ← lift\PM(vP)                           \\
-                  &<λ(x)(k) → c> ← coerceλ₂(v)               \\
+                  &v ← \liftpowerset(vP)                           \\
+                  &(\lam(x)(k) → c) ← \coercefun(v)               \\
                   &vP₂ ← 𝒜 (a₂)                              \\
                   &vP₃ ← 𝒜 (a₃)                              \\
-                  &v₂ ← lift\PM(vP₂)                         \\
-                  &v₃ ← lift\PM(vP₃)                         \\
-                  &modify-σ (λ(σ) → σ ⊔ [x ↦ v₂] ⊔ [k ↦ v₃]) \\
-                  &return(c)                                 \\
+                  &v₂ ← \liftpowerset(vP₂)                         \\
+                  &v₃ ← \liftpowerset(vP₃)                         \\
+                  &\modifystore(λ(σ) → σ ⊔ [x ↦ v₂] ⊔ [k ↦ v₃]) \\
+                  &\return(c)                                    \\
 \end{align*}
@@
 \begin{align*}
-𝒞*    &: \PM(Call × \AStore) → \PM(Call × \AStore) \\
-𝒞*(Σ) &≔ ⋃ { (𝒞ₘ c)(σ) | <c,σ> ∈ Σ }               \\
+𝒞^{⋆}    &: \PM(Call × \AStore) → \PM(Call × \AStore) \\
+𝒞^{⋆}(Σ) &≔ ⋃ \{ 𝒞(c)(σ) \;|\; (c,σ) ∈ Σ \}             \\
 \end{align*}
 
-The monad we used recovers exactly the analysis we wrote before.
+Our current definition of $ℳ $ used recovers exactly the analysis we wrote before.
 On top of the convenience of writing things monadically, our insight and contribution is twofold:
 \begin{itemize}
-\item ℳ  can be \emph{axiomatized}, such that different definitions for ℳ  give rise to different control abstractions for the analysis.
+\item $ℳ $ can be \emph{axiomatized}, such that different definitions for $ℳ $ give rise to different control abstractions for the analysis.
 \item The monadic abstraction provides for a modular proof framework for establishing the correctness of the analysis.
@@
 \begin{align*}
-𝒞*       &: \PM(Call) × \AStore → \PM(Call) × \AStore                          \\
-𝒞*<cP,σ> &≔ <{c' | c' ∈ π₁(𝒞(c))(σ) | c ∈ c\PM }, ⋃ { π₂(𝒞(c))(σ) | c ∈ cP }>  \\
+𝒞^{⋆}       &: \PM(Call) × \AStore → \PM(Call) × \AStore                                         \\
+𝒞^{⋆}(cP,σ) &≔ (\{ c' \;|\; c' ∈ π₁(𝒞(c))(σ) \;|\; c ∈ c\PM \}, ⋃ \{ π₂(𝒞(c))(σ) \;|\; c ∈ cP \}) \\
 \end{align*}
@@
 \begin{align*}
-𝒞*    &: \PM(Call × ^Store^) → \PM(Call × ^Store^)           \\
-𝒞*(Σ) &≔ { <c',π₂(𝒞(c))(σ)> | c' ∈ π₁(𝒞(c))(σ) | <c,σ> ∈ Σ } \\
+𝒞^{⋆}    &: \PM(Call × \AStore) → \PM(Call × \AStore)             \\
+𝒞^{⋆}(Σ) &≔ \{ (c',π₂(𝒞(c))(σ)) \;|\; c' ∈ π₁(𝒞(c))(σ) \;|\; (c,σ) ∈ Σ \} \\
 \end{align*}
@@
 
-The common interface for each variation in abstract control is a monad $ℳ  ∈ Set → Set$ with $get$, $put$, and nondeterminism effects.
-We abreviate the conjunction of the above properties holding on some ℳ  as $Analysis(ℳ )$.
+\paragraph{Abstract Garbage Collection}
+Abstract garbage collection is a technique in abstract interpretation where unreachable abstract addresses are pruned from the state space.
+This is analagous to “real” garbage collection, where unreachable pointers are reclaimed for space efficiency.
+However, in abstract semantics, addresses are \emph{re-used} to soundly and finitely aproximate an infinite address space.
+Impreceision in control flow analyses arise when an abstract address present in the current store must be used for allocation.
+Abstract garbage collection is a purely precision and performance improving optimization which removes unreachable abstract addresses.
 
-\paragraph{Abstract Garbage Collection}
-Abstract garbage collection is a technique in abstract interpretation where unreachable abstract adresses are pruned from the state space.
-This is analagous to "real" garbage collection, where unreachable pointers are reclaimed for space efficiency.
-However, in abstract semantics, adresses are \emph{re-used} to soundly and finitely aproximate an infinite address space.
-Impreceision in control flow analyses arise when an abstract address present in the current store must be used for allocation.
-Abstract garbage collection is a purely precision and performance improving optimization which removes unreachable abstract adresses.
-
-We will use our generic monadic interfact $Analysis(ℳ )$ for an arbitrary monad $ℳ $.
-Using this interface, abstract garbage collection can be implemented once in a purely generic way.
+For a generic implementation of garbage collection we assume an arbitrary $ℳ $ that has $get$, $put$ and nondeterminism effects.
+Using this axiomatized interface for $ℳ $, abstract garbage collection can be implemented once in a purely generic way.
 \begin{align*}
 gc    &: Call → ℳ (1)                                        \\
-gc(c) &≔ do                                                \\
-      &σ ← get-σ                                           \\
-      &𝓉₀ ← 𝓉-𝒞(c)                                         \\
-      &let 𝓉 ≔ μ(𝓉) → 𝓉 ⊔ 𝓉-Var(𝓉)                         \\
-      &modify-σ (λ(σ) → ⋃ { [x ↦ v] | x ∈ 𝓉 ∧ σ(x) = v } ) \\
+gc(c) &≔ \ddo                                                  \\
+      &σ ← \getstore                                             \\
+      &𝓉₀ ← \touchedcall(c)                                           \\
+      &\llet 𝓉 ≔ μ(𝓉) → 𝓉 ⊔ \touchedvar(𝓉)                           \\
+      &\modifystore (λ(σ) → ⋃ \{ [x ↦ v] | x ∈ 𝓉 ∧ σ(x) = v \} ) \\
 \end{align*}
 This is literally the implementation of \emph{concrete} garbage collection.
-However, it seamlessly becomes abstract garbage collection when instantiated with the appropriate monad.
+However, because we have not yet committed to the underlying monad, it seamlessly becomes abstract garbage collection when instantiated with the appropriate monad.
 
@@
 MCFA resolved the paradox by identifying the difference in analyses and transfering the polynomial behavior to functional analysis.
-The crux of MCFA is to create packed, copied closures rather than linked closures.
+The key insight of MCFA is to use packed, copied closures rather than linked closures in the abstract semantics.
 
-So far, our analysis has not used closures for abstract values; 0CFA only tracks lambdas without their closing context.
+So far our analysis has not used closures for abstract values; 0CFA only tracks lambdas without their closing context.
 Adding closures to abstract values is a straightforward addition to the examples we've shown.
-Also, our implementation implements all of these features in their full glory.
+(Our implementation implements all of these features in full.)
 
@@
 \begin{align*}
-clo-copy       &: [Var] × Call → ℳ (Clo)             \\
-clo-copy(xs,c) &≔ do                                 \\
-               &let ys ≔ free-vars(c)                \\
-               &σ ← get-σ                            \\
-               &vs ← map* σ ys                       \\
-               &ρ ← { y ↦* v | (y,v) ∈ zip(ys, vs) } \\
-               &return <xs,c,ρ>                      \\
+  ρ : Env  &≔  Var ⇀ \widehat{Addr}                       \\
+clo : Clo  &⩴ <\lam(x) → c, ρ> \;|\; <\lam(x,k) → c, ρ>   \\
+\end{align*}
+
+\begin{align*}
+clo-copy    &: Lam → ℳ (Clo)                               \\
+clo-copy(l) &≔ \ddo                                        \\
+            &\llet ys ≔ \operatorname{free-vars}(l)        \\
+            &vs ← map^{⋆}(lookup)(ys)                      \\
+            &ρ ← \{ y ↦^{⋆} v \;|\; (y,v) ∈ zip(ys, vs) \} \\
+            &\return (l,ρ)                                 \\
 \end{align*}
@@
 Our generalization over abstract control allows us to implement each of these optimizations once.
-More importantly, the proofs about these optimizations need only happen once as well, and the proofs transfer to their instantiations for free.
+More importantly, the correctness of these optimizations can be established generically as well.
 
@@
 The key advantage to our framework is that the proofs of correctness for constructed analyses are derived automatically.
+To establish the correctness of the monadic approach, we do not merely inline the definitions and resort to reasoning about fully instantiated instances of monadic definitions.
+Instead, we relate monadic actions directly to abstract state space transitions for each monad transformer combination.
+Given these proofs, along with an argument about monotonicity of the monadic semantics, we can establish a proof of galois connection between concrete and abstract interpreters.
 
-We must now relate monadic actions like 𝒞 back to small step semantics.
-To do this we establish a galois connection between monadic actions in ℳ  and transitions functions for \emph{some} abstract state space $𝒮𝒮$.
-This abstract state space is constructed from the monad transformer stack, however some transformer stacks support multiple abstract state spaces.
-This step is notated $(A → ℳ (B))\galois{α}{\gamma}(𝒮𝒮(A) → 𝒮𝒮(B))$.
+To relate back to small step semantics, we establish a galois connection between monadic actions in $ℳ $ and transitions functions for \emph{some} abstract state space $𝒮𝒮$.
+This abstract state space is constructed from the monad transformer stack, although some transformer stacks support multiple abstract state spaces.
+This property is notated $(A → ℳ (B))\galois{α}{\gamma}(𝒮𝒮(A) → 𝒮𝒮(B))$.
 We call a particular monad $ℳ $ which enjoys this property a \emph{small-step monad}.
+Likewise, we call a monad transformer $𝒯$ which enjoys \emph{transports} this property a \emph{small-step monad transformer}.
 
 In our framework we prove that not only are $\ST$ and $\PT$ monad transformers, they're small-step monad transformers.
-This means that for any stack of interleaving $\ST$ and $\PT$ monad transformers, one can construct the needed galois connection.
+This means that for any stack of interleaving $\ST$ and $\PT$ monad transformers, one can construct the necessary galois connection back to small step semantics.
 
-In our framework we prove that $\ST(𝓈)(\PM) ⊑ \PT(\SM(𝓈))$.
+We prove that $\ST(𝓈)(\PM) ⊑ \PT(\SM(𝓈))$.
 These monad interleavings correspond to path-sensitivity and path-insensitivity respectively.
 
-In our framework we prove that for the $\PT(\SM(𝓈))$ monad, two galois connections are possible to the state space $\PM(\_ × 𝓈)$.
+For the $\PT(\SM(𝓈))$ monad, we show two galois connections are possible to the state space $\PM(\_ × 𝓈)$.
 These choices for galois connection correspond to flow-sensitivity and flow-insensitivity respectively.
@@
 
+A given concrete semantics will always use the flow-sensitive path-sensitive monad.
+We can use this bridge to construct a galois connection between concrete and a flow-insensitive abstract semantics.
+
 When using our framework, the analysis designer need only prove prove:
 \begin{itemize}
-\item The semantic step function 𝒞 (which calls 𝒜 ) is monotonic, including monotonic in the choice of 𝓂 .
+\item The semantic step function 𝒞 (which calls 𝒜 ) is monotonic.
 \item The semantics as written, including intensional optimizations, are correct.
@@
 \item A proof-by-construction of correctness for the derived analysis.
-\item Seamless extensions to AAM and intensional optimizations (like gc or mcfa).
+\item 
+  Seamless extensions to AAM and intensional optimizations (like gc or mcfa), which we demonstrate in our implementation.
 \end{itemize}
@@
 However, once monotonicity of the $𝒞$ action is established, both concrete and abstract interpreters can be derived for free.
-This is only possible because all proofs have been decomposed to the unit of monad transformer.
+This is only possible because all proofs have been decomposed to the unit of monad transformer, and these extensions reduce to just adding more transformers to the stack.
 
@@
 % SetT {{{
-\subsection{\ST}
+\subsection{$\PT$}
 \label{section:Proofs+Definitions:SetT}
@@
 lift : ∀ (𝓂  : Set → Set) (A : Set), 𝓂 (A) → \ST(𝓂 )(A)
-lift ≔ map (λ x → { x })
+lift ≔ map (λ x → \{ x \})
 

### 2014-09-21 writeup/darais-quals.tex
pretty proofs for associativity of my monad.  so happy :)

@@
 
-A type $(𝒯 : Set → Set)$ is called a \emph{functor} if one can:
-\begin{itemize}
-\item Define $map : ∀ (A, B : Set), (A → B) → (𝒯(A) → 𝒯(B))$
+A type $(F : Set → Set)$ is called a \emph{functor} if one can:
+\begin{itemize}
+\item Define $map : ∀ (A, B : Set), (A → B) → (F(A) → F(B))$
 \item Prove $map(id) = id$, where $id$ is the identity function $(λ(x) → x)$
@@
 
-Likewise a type $(𝒯 : Set → Set)$ is called a \emph{monad} if one can:
-\begin{itemize}
-\item Define $extend : ∀ (A, B : Set), (A → 𝒯(B)) → (𝒯(A) → 𝒯(B))$
+Likewise a type $(ℳ  : Set → Set)$ is called a \emph{monad} if one can:
+\begin{itemize}
+\item Define $extend : ∀ (A, B : Set), (A → ℳ (B)) → (ℳ (A) → ℳ (B))$
 \item Prove unit and associativity laws (not mentioned here).
@@
 The only difference from the definition of functor is in the the first argument.
-For monads, this argument is allowed to be a \emph{monadic} $(A → 𝒯(B))$, rather than a pure function $(A → B)$.
+For monads, this argument is allowed to be a \emph{monadic} $(A → ℳ (B))$, rather than a pure function $(A → B)$.
 
@@
 \begin{equation*}
-join : ∀ (A : Set), 𝒯(𝒯(A)) → 𝒯(A)
+join : ∀ (A : Set), ℳ (ℳ (A)) → ℳ (A)
 \end{equation*}
@@
 Monad transformers are functions between monads.
-Where a monad $𝒯$ will have type $Set → Set$, a monad transformer will have type $(Set → Set) → (Set → Set)$.
+Where a monad $ℳ $ will have type $Set → Set$, a monad transformer $𝒯$ will have type $(Set → Set) → (Set → Set)$.
 
@@
 The state monad is notated $\SM(𝓈)$ (carrying a single cell of type $𝓈$), the powerset monad is notated $\PM$, and the identity is notated $ID$.
-The state and powerset monad transformer equivalents are notated $\ST(𝓈)(𝓂 )$ and $\PT(𝓂 )$ where $𝓂 $ is the underlying monad.
+The state and powerset monad transformer equivalents are notated $\ST(𝓈)(ℳ )$ and $\PT(ℳ )$ where $ℳ $ is the underlying monad.
 The state monad has $get$ and $put$ effects, whereas the powerset monad has a $nondeterminism$ effect.
@@
 
-The definitions, details and proofs of all monads used in our work are given in section \ref{section:Proofs+Definitions}.
+The definitions, details and proofs of all monads used in our work are given in section \ref{section:Proofs}.
 
@@
 We wll use a powerset monad and state monad transformer to write the analysis in monadic style.
-The full definitions of the monads (and their proofs) used throughout this paper are defered to section \ref{section:Proofs+Definitions}.
+The full definitions of the monads (and their proofs) used throughout this paper are defered to section \ref{section:Proofs}.
 Because monad transformers are just fancy names for simple types, we write the equivalent simple type underneath definitions which reference monad transformer.
@@
   for $\PT$ is actually a monad, whereas traditional definitions are not.  Our
-  definition and proofs for $\PT$ are given in section \ref{section:Proofs+Definitions}.
+  definition and proofs for $\PT$ are given in section \ref{section:Proofs}.
 }
@@
 
-% Proofs and Definitions {{{
-
-\section{Proofs and Definitions}
-\label{section:Proofs+Definitions}
-
-This section details all structures used in this paper and their proofs.
-
-% Setup {{{
-
-\subsection{Setup}
-\label{section:Proofs+Definitions:Setup}
-
-\begin{verbatim}
-PartialOrder(A : Set) ≔
-  operators
-    ⊑ : A → A → Prop
-  laws
-    reflexivity : x ⊑ x
-    antisymmetry : x ⊑ y ⇒  y ⊑ x ⇒  x = y
-    transitivity : x ⊑ y ⇒  y ⊑ z ⇒  x ⊑ z
-\end{verbatim}
-
-\begin{verbatim}
-Monotonic(𝑓 : A → B) ≔ 
-  require
-    PartialOrder(A)
-    PartialOrder(B)
-    ∀ (x y : A), x ⊑ y ⇒  f(x) ⊑ f(y)
-\end{verbatim}
-
-\begin{verbatim}
-(C : Set) α⇄ γ (A : Set) ≔
-  require
-    PartialOrder(C)
-    PartialOrder(A)
-  operators
-    α : C → A
-    γ : A → C
-  laws
-    Monotonic(α)
-    Monotonic(γ)
-    α ∘ γ ⊑ id ⊑ γ ∘ α
-\end{verbatim}
-
-\begin{verbatim}
-JoinSemilattice(A : Set) ≔
-  require
-    PartialOrder(A)
-  operators
-    ⊥ : A
-    ⊔ : A → A → A
-  laws
-    left-unit : ⊥ ⊔ x = x
-    right-unit : x ⊔ ⊥ = x
-    associativity : x ⊔ (y ⊔ z) = (x ⊔ y) ⊔ z
-    commutativity : x ⊔ y = y ⊔ x
-\end{verbatim}
-
-\begin{verbatim}
-Monad(𝓂  : Set → Set) ≔
-  operators
-    return : ∀ (A : Set) → A → 𝓂 (A)
-    extend : ∀ (A B : Set) → (A → 𝓂 (B)) → (𝓂 (A) → 𝓂 (B))
-  laws
-    left-unit : extend(return) = id
-    right-unit : extend(k)(return(A)) = k(A)
-    associativity : extend(k₂)(extend(k₁)(aM)) = extend(λ x → extend(k₂)(k₁(x)))(A)
-\end{verbatim}
-
-\begin{verbatim}
-Transformer(𝒯 : (Set → Set) → (Set → Set)) ≔
-  require
-    ∀ (𝓂  : Set), Monad(𝓂 ) ⇒  Monad(𝒯(𝓂 ))
-  operators
-    lift : ∀ (𝓂  : Set), 𝓂 (A) → 𝒯(𝓂 )(A)
-  laws
-    unit : lift(return) = return
-\end{verbatim}
-
-\begin{verbatim}
-MonadState(𝓈 : Set)(𝓂  : Set → Set) ≔
-  operators
-    get : 𝓂  𝓈
-    put : 𝓈 → 𝓂  1
-\end{verbatim}
-
-\begin{verbatim}
-MonadPlus(𝓂  : Set → Set) ≔ 
-  require
-    ∀ (A : Set), JoinSemilattice(𝓂 (A))
-  laws
-    left-zero : extend(k)(⊥) = ⊥
-    right-zero : extend(const(⊥))(aM) = ⊥
-    distributivity : extend(k)(aM₁ ⊔ aM₂) = extend(k)(aM₁) ⊔ extend(k)(aM₂)
-\end{verbatim}
-
-\begin{verbatim}
-Functorial(ℭ : Set → Prop)(𝓂  : Set → Set) ≔ 
-  require
-    ∀ (A : Set), ℭ (A) → ℭ (𝓂 (A))
-\end{verbatim}
-
-\begin{verbatim}
-Pointed(𝓂  : Set → Set) ≔
-  unit : ∀ (A : Set) → A → 𝓂 (A)
-\end{verbatim}
-
-\begin{verbatim}
-MonadSmallStep(𝓂  : Set → Set)(𝒮𝒮 : Set → Set) ≔ 
-  require
-    ∀ (A B : Set), (A → 𝓂 (B)) α⇄ γ (𝒮𝒮(A) → 𝒮𝒮(B))
-\end{verbatim}
+% Proofs {{{
+
+\section{Proofs}
+\label{section:Proofs}
+
+% Basic Definitions {{{
+\subsection{Basic Definitions}
+\label{section:Proofs:BasicDefinitions}
+
+\begin{definition}
+For a type $(A : Set)$, a \emph{partial order} structure on $A$ contains an operator:
+\begin{align*}
+\_⊑\_ : A → A → Prop
+\end{align*}
+which respect the following properties:
+\begin{align*}
+ \operatorname{reflexivity} &: x ⊑ x                   \\
+\operatorname{antisymmetry} &: x ⊑ y ⇒  y ⊑ x ⇒  x = y \\
+\operatorname{transitivity} &: x ⊑ y ⇒  y ⊑ z ⇒  x ⊑ z \\
+\end{align*}
+\end{definition}
+
+\begin{definition}
+For types $(A, B : Set)$ which both have partial orders, a function $(f : A → B)$ is called \emph{monotonic} if:
+\begin{align*}
+∀ (x, y : A), x ⊑ y ⇒  f(x) ⊑ f(y)
+\end{align*}
+\end{definition}
+
+\begin{lemma}
+For types $(A, B : Set)$ which both have partial orders, there exists a partial order on the \emph{monotonic} function space $(A →^{mon} B : Set)$ with operator:
+\begin{align*}
+f ⊑ g ≔ ∀ (x, y : A), x ⊑ y → f(x) ⊑ g(y)
+\end{align*}
+\begin{proof}
+Reflexivity is justified by monotonicity.  
+Antisymmetry and transitivity are justified through the partial orders on $A$ and $B$.
+\end{proof}
+\end{lemma}
+
+\begin{definition}
+For types $(C, A : Set)$ which both have partial orders, a \emph{galois connection} between $C$ and $A$, written $C\galois{α}{γ}A$, contains two operators:
+\begin{align*}
+α &: C → A \\
+γ &: A → C \\
+\end{align*}
+which respect the following properties:
+\begin{align*}
+                           &α \text{is monotonic} \\
+                           &γ \text{is monotonic} \\
+\operatorname{contractive} &: α ∘ γ ⊑ id          \\
+  \operatorname{expansive} &: id ⊑ γ ∘ α          \\
+\end{align*}
+\end{definition}
+
+\begin{definition}
+For a type $(A : Set)$, we call $A$ a \emph{join semilattice}, written $JoinSemilattice(A)$, if one can define two operators:
+\begin{align*}
+    ⊥ &: A         \\
+\_⊔\_ &: A → A → A \\
+\end{align*}
+which respect the following properties:
+\begin{align*}
+    \operatorname{left-unit} &: ⊥ ⊔ x = x                  \\
+   \operatorname{right-unit} &: x ⊔ bot = x                \\
+\operatorname{associativity} &: x ⊔ (y ⊔ z) = (x ⊔ y) ⊔ z  \\
+\operatorname{commutativity} &: x ⊔ y = y ⊔ x              \\
+\end{align*}
+\end{definition}
+
+\begin{definition}
+For a type $(F : Set → Set)$, we call $F$ a \emph{functor}, written $Functor(F)$, if one can define the operator:
+\begin{align*}
+\operatorname{map} : ∀ (A , B : Set), (A → B) → (F(A) → F(B))
+\end{align*}
+which respects the following properties:
+\begin{align*}
+          unit &: map(id) = id                  \\
+distributivity &: map(g ∘ f) = map(g) ∘ map(f)  \\
+\end{align*}
+\end{definition}
+
+\begin{definition}
+For a type $(ℳ  : Set → Set)$, we call $ℳ $ a \emph{monad}, written $Monad(ℳ )$, if one can define two operators:
+\begin{align*}
+\operatorname{return} &: ∀ (A : Set), A → ℳ (A)                          \\
+\operatorname{extend} &: ∀ (A, B : Set), (A → ℳ (B)) → (ℳ (A) → (ℳ (B))) \\
+\end{align*}
+which respect the following properties:
+\begin{align*}
+    \operatorname{left-unit} &: extend(return) = id                               \\
+   \operatorname{right-unit} &: extend(k) ∘ return = k                            \\
+\operatorname{associativity} &: extend(k₂) ∘ extend(k₁) = extend(extend(k₂) ∘ k₁) \\
+\end{align*}
+\end{definition}
+
+\begin{corrolary}
+All monads are functors.
+\end{corrolary}
+
+\begin{definition}
+For a type $(𝒯 : (Set → Set) → (Set → Set))$, we call $𝒯$ a \emph{monad transformer}, written $Transformer(𝒯)$, if one can define a single operator:
+\begin{align*}
+lift : ∀ (ℳ  : Set → Set) (A : Set), ℳ (A) → 𝒯(ℳ )(A)
+\end{align*}
+and the following property holds:
+\begin{align*}
+∀ (ℳ  : Set → Set), Monad(ℳ ) ⇒  Monad(𝒯(ℳ ))
+\end{align*}
+\end{definition}
+
+\begin{definition}
+For types $(𝓈 : Set)$ and $(ℳ  : Set → Set)$, we call $ℳ $ a \emph{monad state over $𝓈$}, written $MonadState(𝓈)(ℳ )$, if one can define operators:
+\begin{align*}
+\operatorname{get} &: ℳ (𝓈)     \\
+\operatorname{put} &: 𝓈 → ℳ (1) \\
+\end{align*}
+\end{definition}
+
+\begin{definition}
+For a type $(ℳ  : Set → Set)$, we call $ℳ $ a \emph{monad plus}, written $MonadPlus(ℳ )$, if the following property holds:
+\begin{align*}
+∀ (A : Set), JoinSemilattice(ℳ (A))
+\end{align*}
+and the following additional properties hold:
+\begin{align*}
+     \operatorname{left-zero} &: extend(k)(⊥) = ⊥                               \\
+    \operatorname{right-zero} &: extend(const(⊥))(x) = ⊥                        \\
+\operatorname{distributivity} &: extend(k)(x ⊔ y) = extend(k)(x) ⊔ extend(k)(y) \\
+\end{align*}
+\end{definition}
+
+\begin{definition}
+For types $(ℳ  , 𝒮𝒮 : Set → Set)$, we call $ℳ $ a \emph{small step monad with state space $𝒮𝒮$}, written $MonadSmallStep(𝒮𝒮)(ℳ )$, if one can define:
+\begin{align*}
+∀ (A, B : Set), (A → ℳ (B)) \galois{α}{γ} (𝒮𝒮(A) → 𝒮𝒮(B))
+\end{align*}
+\end{definition}
+
+\begin{definition}
+For a type $(F : Set → Set)$ and property $(P : Set → Prop)$, we call $F$ \emph{functorial in $P$}, written $Functorial(P)(F)$, if the following property holds:
+\begin{align*}
+∀ (A : Set), P(A) ⇒ P (F(A))
+\end{align*}
+and all operations in $P$ distribute through monadic operations in $F$.
+\end{definition}
+
+\begin{example}
+For a type $(F : Set → Set)$ to be \emph{functorial in $JoinSemilattice$}, the following additional laws must hold:
+\begin{align*}
+return (x ⊔ y) = return(x) ⊔ return(y)
+\end{align*}
+\end{example}
+
+% }}}
+
+% ID {{{
+\subsection{ID}
+\label{section:Proofs:ID}
+
+\begin{definition}
+The \emph{identity monad}, written $ID$, is defined:
+\begin{align*}
+   ID &: Set → Set \\
+ID(A) &≔ A         \\
+\end{align*}
+\end{definition}
+
+\begin{lemma}
+The identity monad is a monad with operators:
+\begin{align*}
+return &≔ id \\
+extend &≔ id \\
+\end{align*}
+\begin{proof}
+Unit and associativity laws are established trivially by definition.
+\end{proof}
+\end{lemma}
+
+\begin{lemma}
+The identity monad is \emph{functorial in all structures}.
+\begin{proof}
+Holds by definition of $ID$.
+\end{proof}
+\end{lemma}
 
@@
 \subsection{$\PT$}
-\label{section:Proofs+Definitions:SetT}
+\label{section:Proofs:SetT}
+
+\begin{definition}
+The \emph{set monad transformer}, written $\PT$, is defined:
+\begin{align*}
+\PT(ℳ )(A) ≔ ℳ (\PM(A))
+\end{align*}
+\end{definition}
+
+\begin{lemma}
+For a given $(ℳ  : Set → Set)$ where:
+\begin{itemize}
+\item $ℳ $ is a monad.
+\item $ℳ $ is functorial in JoinSemilattice.
+\end{itemize}
+then $\PT(ℳ )$ is a monad plus.
+\begin{proof}
+By applying functoriality of $ℳ $ to the semilattice $\PM$.
+\end{proof}
+\end{lemma}
+
+\begin{lemma}
+For a given $(ℳ  : Set → Set)$ where:
+\begin{itemize}
+\item $ℳ $ is a monad.
+\item $ℳ $ is functorial in JoinSemilattice.
+\end{itemize}
+then $\PT(ℳ )$ is a monad with operators:
+\begin{align*}
+   return &≔ return_{ℳ } ∘ singleton \\
+extend(k) &≔ extend_{ℳ }(joins ∘ map_{\PM}(k))  \\
+\end{align*}
+\begin{proof}
+\begin{align*}
+\operatorname{left-unit} &: extend(return) = id                                                                                        \\
+                                                                                                                                       \\
+extend(return) &= extend_{ℳ }(joins ∘ map_{\PM}(return_{ℳ } ∘ singleton_{\PM}))            \tag{definition of extend and join}         \\
+               &= extend_{ℳ }(λ(xs) → ⋃ \{ return_{ℳ } (singleton_{\PM}(x)) \;|\; x ∈ xs \}) \tag{definition of joins}                 \\
+               &= extend_{ℳ }(λ(xs) → return_{ℳ } (⋃ \{ singleton_{\PM}(x) \;|\; x ∈ xs \})) \tag{functorality of semilattice in $ℳ $} \\
+               &= extend_{ℳ }(λ(xs) → return_{ℳ }(xs))                                       \tag{join identity for $\PM$}             \\
+               &= extend_{ℳ }(return_{ℳ })                                                 \tag{η reduction for λ}                     \\
+               &= id                                                                       \tag{left unit monad law for $ℳ $}          \\
+\end{align*}
+\end{proof}
+\begin{proof}
+\begin{align*}
+\operatorname{right-unit} &: extend(k)(return(x)) = k(x)                                                                                 \\
+                                                                                                                                         \\
+extend(k)(return(x)) &= extend_{ℳ }(joins ∘ map_{\PM}(k))(return_{ℳ }(singleton_{\PM}(x))) \tag{definition of extend and join}           \\
+                     &= joins(map_{\PM}(k)(return_{ℳ }(singleton_{\PM}(x))))               \tag{right unit monad law for $ℳ $}           \\
+                     &= joins(return_{ℳ }(singleton_{\PM}(k(x))))                          \tag{map distribution law for $ℳ $ and $\PM$} \\
+                     &= k(x)                                                               \tag{definition of joins}                     \\
+\end{align*}
+\end{proof}
+\begin{proof}
+\begin{align*}
+\operatorname{associativity} &: extend(k₂) ∘ extend(k₁) = extend(extend(k₂) ∘ k₁)                                                              \\
+                                                                                                                                               \\
+extend(k₂)∘ extend(k₁) &= extend_{ℳ }(joins ∘ map_{\PM}(k₂)) ∘ extend_{ℳ }(joins ∘ map_{\PM}(k₁)) \tag{definition of $extend$}                 \\
+                       &= extend_{ℳ }(extend_{ℳ }(joins ∘ map_{\PM}(k₂)) ∘ joins ∘ map_{\PM}(k₁)) \tag{associativity law for $ℳ $}             \\
+                       &= extend_{ℳ }(extend(k₂) ∘ joins ∘ map_{\PM}(k₁))                         \tag{definition of $extend$}                 \\
+                       &= extend_{ℳ }(λ(xs) → extend(k₂) (⋃ \{ k₁(x) \;|\; x ∈ xs\}))                   \tag{definitions of $joins$ and $map_{\PM}$} \\
+                       &= extend_{ℳ }(λ(xs) → ⋃ \{ extend(k₂)(k₁(x)) \;|\; x ∈ xs\})                    \tag{distributivity of $∪$ for $ℳ $}         \\
+                       &= extend_{ℳ }(joins ∘ map_{\PM}(extend(k₂) ∘ k₁))                         \tag{definition of $map_{\PM}$}              \\
+                       &= extend(extend(k₂) ∘ k₁)                                                 \tag{definition of $extend$}                 \\
+%                        = extend_m(λ(xs) → extend_m(joins ∘ map_P(k2)) (joins { k1(x) | x ∈ xs}))
+%                        = extend_m(λ(xs) → joins { extend_m(joins ∘ map_P(k2))(k1(x)) | x ∈ xs })
+%                        = extend_m(λ(xs) → joins ∘ map_P(extend_m(joins ∘ map_P(k2)) ∘ k1)(xs))
+% 
+%                            = do
+%                             y <- x
+%                             z <- joins (map(k1(y)))
+%                             joins(map(k2(z))
+% 
+% 
+% 
+%                            = do
+%                           {y1..yn} <- x
+%                           ⋃ {do
+%                                {z1..zn} <- k1(y1)
+%                                ∪ {k2(z1)..k2(zn)}
+%                                ..
+%                              do
+%                                {z1..zn} <- k1(yn)
+%                                ∪ {k2(z1)..k2(zn)}}
+% 
+% 
+%                           ⋃
+%                           { x | x ∈ map (\ a -> do
+%                                   z <- k1 a
+%                                   joins (map k2 z)) y }
+%                              
+% 
+%                           = extend_m(extend_m(joins ∘ map_P(extend_m(joins ∘ map_P(k2)) ∘ k1)) ∘ return)
+% 
+% 
+% 
+% 
+%                           = extend(extend(k2) ∘ k1)
+%                           = extend_m(joins ∘ map_P(extend_m(joins ∘ map_P(k2)) ∘ k1))
+%                           = extend_m(λ(xs) → joins { extend_m(joins ∘ map_P(k2)) (k1(x)) | x ∈ xs }
+%                           = extend_m(λ(x) → extend_m(joins ∘ map_P(k2))(joins { k1(x) | x ∈ xs }))
+%                           = extend_m(extend_m(joins ∘ map_P(k2)) ∘ joins ∘ map(k1))
+% 
+\end{align*}
+\end{proof}
+\end{lemma}
 
 \begin{verbatim}
-\ST : (Set → Set) → (Set → Set)
-\ST(𝓂 )(A) ≔ 𝓂 (\PM(A))
-
------------------
-Transformer(\ST)
-
-lift : ∀ (𝓂  : Set → Set) (A : Set), 𝓂 (A) → \ST(𝓂 )(A)
-lift ≔ map (λ x → \{ x \})
-
-Monad(𝓂 ) ∧ Functorial(JoinSemilattice)(𝓂 )
--------------------------------------------
-Monad(\ST(𝓂 ))
-
-return : ∀ (A : Set), A → \ST 𝓂  A
-return ≔ lift ∘ returnₘ
-
-extend : ∀ (A B : Set), (A → \ST(𝓂 )(B)) → (\ST(𝓂 )(A) → \ST(𝓂 )(B))
-extend(k) ≔ joinₘ ∘ mapₘ (joins ∘ mapₚ k)
-
-Monad(𝓂 ) ∧ Functorial(JoinSemilattice)(𝓂 )
--------------------------------------------
-MonadPlus(\ST(𝓂 ))
-
-m0 : ∀ (A : Set), \ST 𝓂  A
-m0 ≔ ⊥ ₘ
-
-<+> : ∀ (A : Set) → \ST 𝓂  A → \ST 𝓂  A → \ST 𝓂  A
-<+> ≔ ⊔
 
@@
 \subsection{$\ST$}
-\label{section:Proofs+Definitions:StateT}
+\label{section:Proofs:StateT}
 
@@
 \section{Propositions}
-\label{section:Proofs+Definitions:Propositions}
+\label{section:Proofs:Propositions}
 

### 2014-09-27 writeup/darais-quals.tex
abstract

@@
 
-\title{Abstract Control in Static Analysis}
+\title{Abstract Control in Static Analysis (Draft)}
 \author{David Darais}
@@
 \begin{abstract}
+
+The field of static analysis offers many techniques for automatically discovering facts about programs.
+For example, a static analysis might be used to verify that a program is memory safe, or that it obeys a privacy policy.
+Techniques based on abstract interpretation allow the analysis designer to vary some properties of an analysis.
+However, features like context, object, and flow sensitivity are often hard-coded into a given analysis framework.
+One option for altering such features is to redesign an analysis from scratch.
+Unfortunately, designing new analyses is a difficult and error-prone process.
+Abstracting Abstract Machines (AAM) is a methodology for building analyses which simplifies this process by adhering to a restricted design philosophy.
+However, AAM lacks reusable tools for building analyses, and not all analysis features fit nicely into the framework.
+
+We propose a new framework which gives the analysis designer compositional building blocks for designing analyses, and which supports one key feature missing in AAM: abstract control.
+We achieve compositionality by refactoring AAM into monad transformers which can be related back to abstract state machines.
+When put together, these monad transformers form analyses which are correct by construction.
+Using these building blocks we are able to uncover a new tuning knob for static analyses which we call abstract control.
+Abstract control allows one to adjust the flow and path sensitivity of an analysis independent of other design choices.
+
 \end{abstract}
@@
 \item 
-  An abstract version of our concrete program $\Ce$: $(\Ae : \A)$.
+  An abstract version of $\Ce$ our concrete program: $(\Ae : \A)$.
 \item 
@@
 $\Ce$ is the current program (possibly the result of running the base program for a little while).
-$\Ae$ is a valid abstraction of $\Ce$, where this validity is expressed some relation $R$ holding between $\Ce$ and $\Ae$.
+$\Ae$ is a valid abstraction of $\Ce$, where this validity is expressed by some relation $R$ holding between $\Ce$ and $\Ae$.
 $\Ce'$ is some next state of execution of $\Ce$, and likewise for $\Ae$/$\Ae'$.
@@
 \end{equation*}
+The \emph{meaning} of the analysis is entirely subject to the relation $R$ which relates the abstract to the concrete.
 
@@
 Equivalent to all four properties is the property $x ⊑ γ(y) ⇔  α(x) ⊑ y$.
+
+The expansive property corresponds to \emph{soundness}, and the contractive property corresponds to \emph{tightness}.
+A sound analysis gives results you can trust.  
+A tight analysis promises to be the “best” analysis possible.
 
@@
 
-Using a galois connection $\C\galois{α}{γ}\A$ and abstract step function $\Asteps ⊑ γ(\Csteps)$, the analysis process becomes more specific:
+Using a galois connection $\C\galois{α}{γ}\A$ and abstract step function $\Asteps ⊑ γ(\Csteps)$, the analysis story becomes simplified even further:
 \begin{itemize}
@@
 The galois connection framework simultaneously guarantees the \emph{soundness} and \emph{tightness} of this method.
-Soundness tells us that if we abstract, take an abstract step, and then concretize, that the result is an approximation of taking a concrete step.
+Soundness tells us that if we abstract, take an abstract step, and then concretize, then the result is an approximation of taking a concrete step.
 Soundness tells us that we can trust the results of the analysis.
@@
 Instead, we relate monadic actions directly to abstract state space transitions for each monad transformer combination.
-Given these proofs, along with an argument about monotonicity of the monadic semantics, we can establish a proof of galois connection between concrete and abstract interpreters.
+Given these proofs, along with an argument about monotonicity of the monadic semantics, we can establish a proof of galois connection for any given interpreter instantiation.
 
@@
 
-Independent of language or application, we can now conclude abstractly (yet precisely):
+Independent of language or application we have proven:
 \begin{align*}
@@
 \begin{itemize}
-\item The semantic step function 𝒞 (which calls 𝒜 ) is monotonic.
+\item The semantic step function $𝒞$ is monotonic.
 \item The semantics as written, including intensional optimizations, are correct.
@@
 \item An automatically derived analysis for their language with their choice of abstract control.
-\item A proof-by-construction of correctness for the derived analysis.
-\item 
-  Seamless extensions to AAM and intensional optimizations (like gc or mcfa), which we demonstrate in our implementation.
+\item A proof-by-construction of galois connection (soundness and tightness) for the derived analysis.
+\item Seamless extensions to AAM and intensional optimizations (like gc or mcfa), which we demonstrate in our implementation.
 \end{itemize}
@@
 \end{align*}
-\begin{proof}
+\begin{proof} of left unit.
 \begin{align*}
@@
 \end{proof}
-\begin{proof}
+\begin{proof} of right unit.
 \begin{align*}
@@
 \end{proof}
-\begin{proof}
+\begin{proof} of associativity.
 \begin{align*}
@@
                        &= extend(extend(k₂) ∘ k₁)                                                 \tag{definition of $extend$}                 \\
-%                        = extend_m(λ(xs) → extend_m(joins ∘ map_P(k2)) (joins { k1(x) | x ∈ xs}))
-%                        = extend_m(λ(xs) → joins { extend_m(joins ∘ map_P(k2))(k1(x)) | x ∈ xs })
-%                        = extend_m(λ(xs) → joins ∘ map_P(extend_m(joins ∘ map_P(k2)) ∘ k1)(xs))
-% 
-%                            = do
-%                             y <- x
-%                             z <- joins (map(k1(y)))
-%                             joins(map(k2(z))
-% 
-% 
-% 
-%                            = do
-%                           {y1..yn} <- x
-%                           ⋃ {do
-%                                {z1..zn} <- k1(y1)
-%                                ∪ {k2(z1)..k2(zn)}
-%                                ..
-%                              do
-%                                {z1..zn} <- k1(yn)
-%                                ∪ {k2(z1)..k2(zn)}}
-% 
-% 
-%                           ⋃
-%                           { x | x ∈ map (\ a -> do
-%                                   z <- k1 a
-%                                   joins (map k2 z)) y }
-%                              
-% 
-%                           = extend_m(extend_m(joins ∘ map_P(extend_m(joins ∘ map_P(k2)) ∘ k1)) ∘ return)
-% 
-% 
-% 
-% 
-%                           = extend(extend(k2) ∘ k1)
-%                           = extend_m(joins ∘ map_P(extend_m(joins ∘ map_P(k2)) ∘ k1))
-%                           = extend_m(λ(xs) → joins { extend_m(joins ∘ map_P(k2)) (k1(x)) | x ∈ xs }
-%                           = extend_m(λ(x) → extend_m(joins ∘ map_P(k2))(joins { k1(x) | x ∈ xs }))
-%                           = extend_m(extend_m(joins ∘ map_P(k2)) ∘ joins ∘ map(k1))
-% 
 \end{align*}
@@
 
+\begin{lemma}
+For a given $(ℳ  : Set → Set)$ and $(𝓈 : Set)$ where:
+\begin{itemize}
+\item $ℳ $ is a monad.
+\item $ℳ $ is a monad state over $𝓈$.
+\item $ℳ $ is functorial in JoinSemilattice.
+\end{itemize}
+then $\PT(ℳ )$ is a monad state over $𝓈$ with operators:
+\begin{align*}
+   get    &≔ map_{ℳ }(singleton)(get_{ℳ })    \\
+   put(𝓈) &≔ map_{ℳ }(singleton)(put_{ℳ }(𝓈)) \\
+\end{align*}
+\end{lemma}
+
+\begin{lemma}
+For a given $(ℳ  : Set → Set)$ and $(𝒮𝒮 : Set → Set)$ where:
+\begin{itemize}
+\item $ℳ $ is a monad.
+\item $ℳ $ is a small-step monad over $𝒮𝒮$.
+\item $ℳ $ is functorial in join semilattice.
+\item $𝒮𝒮$ is a functor.
+\end{itemize}
+then $\PT(ℳ )$ is a small step monad over $(𝒮𝒮 ∘ \PM)$ with galois connection maps:
+\begin{align*}
+   α &: ∀ (A, B : Set), (A → \PT(ℳ )(B)) → (𝒮𝒮(\PM(A)) → 𝒮𝒮(\PM(B))) \\
+α(f) &≔ α_{ℳ }(joins ∘ map_{\PM}(f))                                 \\
+                                                                     \\
+   γ &: ∀ (A, B : Set), (𝒮𝒮(\PM(A)) → 𝒮𝒮(\PM(B))) → (A → \PT(ℳ )(B)) \\
+γ(f) &≔ γ_{ℳ }(f ∘ map_{𝒮𝒮}(return_{\PM}))                           \\
+\end{align*}
+\begin{proof} of contractivity.
+\begin{align*}
+α ∘ γ &⊑ id                                                                   \\
+                                                                              \\
+(α ∘ γ)(f) &= α_{ℳ }(joins ∘ map_{\PM}(γ_{ℳ }(f ∘ map_{𝒮 𝒮 }(return_{\PM})))) \\
+\end{align*}
+TODO.
+(Requires $α(1) = 1$ and $α(g ∘ f) = α(g) ∘ α(f)$ (maybe $⊑$?), and likewise for $γ$.)
+\end{proof}
+\begin{proof} of expansivity.
+\begin{align*}
+\end{align*}
+TODO.
+\end{proof}
+\end{lemma}
+
 \begin{verbatim}
-
-Monad(𝓂 ) ∧ MonadState(𝓈)(𝓂 )
------------------------------
-MonadState(𝓈)(𝓂 )
-
-get : \ST 𝓂  𝓈
-get ≔ lift getₘ
-
-put : 𝓈 → \ST 𝓂  1
-put ≔ lift ∘ putₘ
-
-MonadSmallStep(𝓂 )(𝒮𝒮) ∧ Functorial(JoinSemilattice)(𝓂 ) ∧ Functor(𝒮𝒮)
--------------------------------
-MonadSmallStep(\ST 𝓂 )(𝒮𝒮 ∘ \PM)
-
-α : ∀ (A B : Set), (A → \ST(𝓂 )(B)) → (𝒮𝒮(\PM(A)) → 𝒮𝒮(\PM(B)))
-α(f) ≔ αₘ(joinsₘ ∘ mapₚ(f))
-
-γ : ∀ (A B : Set), (𝒮𝒮(\PM(A)) → 𝒮𝒮(\PM(B))) → (A → \ST(𝓂 )(B))
-γ(f) ≔ γₘ(f ∘ mapₛₛ(unitₚ))
 
@@
 
-% Propositions {{{
-
-\section{Propositions}
-\label{section:Proofs:Propositions}
-
-\begin{itemize}
-\item $\PT$ and $\ST$ are Galois Functors.
-\item $\PT$ and $\ST$ are MonadStep Functors.
-\item 𝒜  and 𝒞 are monotonic (including in 𝓂 ).
-\end{itemize}
-
-% }}}
-
 % }}}

### 2014-09-28 writeup/darais-quals.tex
new intro, back to working on content

@@
 For example, a static analysis might be used to verify that a program is memory safe, or that it obeys a privacy policy.
-Techniques based on abstract interpretation allow the analysis designer to vary some properties of an analysis.
-However, features like context, object, and flow sensitivity are often hard-coded into a given analysis framework.
-One option for altering such features is to redesign an analysis from scratch.
-Unfortunately, designing new analyses is a difficult and error-prone process.
-Abstracting Abstract Machines (AAM) is a methodology for building analyses which simplifies this process by adhering to a restricted design philosophy.
-However, AAM lacks reusable tools for building analyses, and not all analysis features fit nicely into the framework.
-
-We propose a new framework which gives the analysis designer compositional building blocks for designing analyses, and which supports one key feature missing in AAM: abstract control.
-We achieve compositionality by refactoring AAM into monad transformers which can be related back to abstract state machines.
+Techniques based on Abstract Interpretation (AI) allow the analysis designer to vary some properties of an analysis.
+However, features like context, object, and flow sensitivity must be hard-coded into an analysis based on classic AI.
+Abstracting Abstract Machines (AAM) is one design methodology for analysis which exposes more of these features as tuning knobs.
+Unfortunately, AAM lacks reusable tools for building analyses, and not all analysis features fit nicely into the framework.
+
+We propose a new compositional framework for designing analyses, and one which supports a tuning knob missing from AAM: abstract control.
+We achieve compositionality by refactoring AAM into monad transformers, each of which we relate back to abstract state machines.
 When put together, these monad transformers form analyses which are correct by construction.
-Using these building blocks we are able to uncover a new tuning knob for static analyses which we call abstract control.
+Furthermore, the monadic abstraction allows us to recover a tuning knob for abstract control when designing analyses.
 Abstract control allows one to adjust the flow and path sensitivity of an analysis independent of other design choices.
@@
 
-% A static analysis has two components: 
-% \begin{itemize}
-% \item A \emph{computation} which computes some set of facts about the execution of a program.
-% \item A \emph{proof of correctness} about the computation.
-% \end{itemize}
-% 
-% We are motivated by static analysis techniques which are:
-% \begin{itemize}
-% \item Compartmental: Various aspects of the analysis have been separated from each other and are understood in isolation.
-% \item Modular: Design choices in one aspect of the analysis do not restrict design choices of another.  
-%       This property is important for both computational and correctness components of an analysis.
-% \item Language Agnostic: A given analysis technique can be seamlessly transfered from one semantics to another.
-% \item Correct: The proof of correctness should exist, and there should be a framework for establishing it.
-% \item Derived: Either the computational artifact or the correctness proof should be obtained "for free" from the other.
-% \end{itemize}
-% 
-% Our framework compartmentalizes:
-% \begin{itemize}
-% \item Abstract domain a la Cousot.
-% \item Abstract time and address a la Van Horn and Might.
-% \item Intensional optimizations a la Van Horn and Might. 
-% \item Object-sensitivity a la Smaragdakis.
-% \item Analysis control properties.
-% \end{itemize}
-% 
-% Our framework is:
-% \begin{itemize}
-% \item Compartmental: Each of the above concerns are separated and independent.
-% \item Modular: Design choices in one axis is completely independent of others
-% \item Language Agnostic: All axis other than abstract domain are fully language agnostic.
-% \item Correct: A correctness framework is described and each axis is proven correct in isolation.
-% \item Derived: The correctness of an analysis using our framework comes for free.
-% \end{itemize}
-% 
-% Analysis control properties are the primary topic of this work.
-% We merely note that the other axis in the design space are implemented in our framework using existing techniques.
-% 
-% Contributions:
-% \begin{itemize}
-% \item Computational artifacts and correctness proofs for the axis of abstract control, which captures the choice of flow and path sensitivity.
-% \item Computational artifacts and correctness proofs for intensional optimizations (gargabe collection and mcfa) independent of abstract control.
-% \item Language independent proofs of refinement for flow-sensitivity and path-sensitivity choices using the monadic abstraction.
-% \end{itemize}
+Static analysis is useful for many different situations.
+A software developer might use an analysis to find bugs in a program.
+A software merchant might use an analysis to detect malware or other security vulnerabilities in software hosted by their app store.
+A manufacturer might use an analysis to establish high assurance in the reliability of control software for a car or airplane.
+Each of these situations requires a different type of analysis.
+To meet the needs of a given problem, an analysis designer has two options: retrofit an existing analysis or design one from scratch.
+However, retrofitting an existing analysis might be insufficient because the tension between precision and performance is often too high.
+Getting an analysis right might require more modification than an existing analysis was designed to support.
+Because of this, complex analyses often need to be designed from scratch, and analysis designer need tools to help them with this process.
+
+Designing a static analysis is hard.
+Designing \emph{the right} static analysis is even more difficult.
+One possible process for arriving at the right static analysis is:
+\begin{enumerate}
+\item Describe the semantics of your language.
+\item Design an analysis \emph{which might have the right properties}.
+\item Exercise the analysis.
+\item If the analysis is too slow or imprecise: repeat from 2.
+\end{enumerate}
+To get a complex analysis right this process must be repeated multiple times.
+Iterating quickly between the design and exercise of an analysis is the key to getting the analysis you wanted without too much pain.
+
+One methodology which aims to ease the design process for static analysis is the Abstracting Abstract Machines (AAM) framework\cite{AAM}.
+AAM separates two design steps which are traditionally tightly coupled: describing features of interest and abstracting those features.
+Furthermore, AAM shows that the abstraction step is systematic, freeing the analysis designer to spend effort in designing features of interest.
+However, AAM doesn't provide reusable tools for the design process; the analysis designer must describe their state space transition from scratch.
+Additionally, some analysis features like flow and path sensitivity don't fit nicely into the AAM framework.
+
+We propose a new compositional framework for designing analyses, and one which supports a tuning knob missing from AAM: abstract control.
+To achieve this we refactor AAM into monad transformers.
+Using monad transformers, the analysis designer can easily add and remove features from their semantics independently of other features.
+The monadic abstraction also allows us to recover a tuning knob for abstract control.
+Abstract control allows one to adjust the flow and path sensitivity of an analysis independent of other design choices.
+
+To justify the soundness of the approach, we show that the monad transformers we use are also galois connection transformers.
+This means that any combination of our monad transformers will yield a sound approximation of a state machine semantics.
+
+Our contributions in this work are:
+\begin{enumerate}
+\item A monadic formulation of AAM which exposes abstract control as a tuning knob.
+\item A compositional \emph{implementation} framework for building AAM machines using monad transformers.
+\item A compositional \emph{proof} framework for justifying the correctness off AAM machines using galois connection transformers.
+\item A new monad transformer for nondeterminism which satisfies the compositional interface of our framework.
+\end{enumerate}
+
+The rest of the paper is structured as follows:
+Section \section{Background} covers necessary background material in semantics, analysis, and the CPS language which we use for examples.
+Section \section{MAAM} demonstrates a monadic variant of AAM which supports defining compositional interpreters for analysis.
+Section \section{AbstractControl} discusses abstract control and how we exploit the monadic abstraction to allow tuning it.
+Section \section{Optimizations} gives two examples of analysis optimizations and how they adapt seamlessly to variations in abstract control.
+Section \section{Correctness} discusses the correctness framework and what key lemmas are involved.
+Section \section{Proofs} shows proof steps in more detail for the arguments made in section \section{Correctness}.
+Sections \section{FutureWork} and \section{RelatedWork} discuss future and related work, and \section{Conclusion} concludes.
 
@@
 \label{section:Background}
-
-% Notation {{{
-\subsection{Notation}
-\label{section:Background:Notation}
-
-We use traditional notation for function definitions and applications rather than lambda calculus notation.
-For example, we will write:
-\begin{equation*}
-compose(g, f, x) ≔ g(f(x))
-\end{equation*}
-rather than:
-\begin{equation*}
-compose\; g\; f\; x ≔ g\; (f\; x)
-\end{equation*}
-
-We use lambda notation for anonymous functions.
-For example, the anonymous composition function is written:
-\begin{equation*}
-λ(g, f, x) → g(f(x))
-\end{equation*}
-
-% }}}
-
-% Monads {{{
-\subsection{Monads}
-\label{section:Background:Monads}
-
-A basic familiarity with the concept of monads will be necessary for understanding this paper.
-Some find it useful to review the concept of functor before learning about monads. 
-We adopt this approach for our brief explanation of monads.
-
-A type $(F : Set → Set)$ is called a \emph{functor} if one can:
-\begin{itemize}
-\item Define $map : ∀ (A, B : Set), (A → B) → (F(A) → F(B))$
-\item Prove $map(id) = id$, where $id$ is the identity function $(λ(x) → x)$
-\item Prove $map (g ∘ f) = map(g) ∘ map(f)$
-\end{itemize}
-
-\paragraph{Example:} 
-Lists are functors, where $map$ is defined:
-\begin{align*}
-    map(f)(xs) ≔ &\case(xs):           \\
-          Nil    &→ Nil                \\
-         x ∷ xs' &→ f(x) ∷ map(f)(xs') \\
-\end{align*}
-
-\paragraph{Example:} $map(isEven)([1, 2, 3, 4]) = [False, True, False, True]$
-
-Likewise a type $(ℳ  : Set → Set)$ is called a \emph{monad} if one can:
-\begin{itemize}
-\item Define $extend : ∀ (A, B : Set), (A → ℳ (B)) → (ℳ (A) → ℳ (B))$
-\item Prove unit and associativity laws (not mentioned here).
-\end{itemize}
-The only difference from the definition of functor is in the the first argument.
-For monads, this argument is allowed to be a \emph{monadic} $(A → ℳ (B))$, rather than a pure function $(A → B)$.
-
-\paragraph{Example:} Lists are monads, where $extend$ is defined:
-\begin{align*}
-    extend(f)(xs) ≔ &\case(xs):              \\
-                Nil &→ Nil                   \\
-            x ∷ xs' &→ f(x) ⧺ extend(f)(xs') \\
-\end{align*}
-
-\paragraph{Example:} 
-$extend(λ(x) → [x + 1, x - 1])([10, 100]) = [9, 11, 99, 101]$
-
-$extend$ for lists can also be understood as $concat$ followed by $map$:
-\paragraph{Example:} 
-\begin{align*}
-extend(f)([10, 100]) &= concat(map(f)([10, 100]))    \\
-                     &= concat([[9, 11], [99, 101]]) \\
-                     &= [9, 11, 99, 101]             \\
-\end{align*}
-For all monads, it is equivalent to define $return$, $map$ and $join$, where:
-\begin{equation*}
-join : ∀ (A : Set), ℳ (ℳ (A)) → ℳ (A)
-\end{equation*}
-As we have just seen, $join$ for the list monad is just $concat$.
-
-% }}}
-
-% Monad Transformers {{{
-\subsection{Monad Transformers}
-\label{section:Background:MonadTransformers}
-
-Monad transformers are functions between monads.
-Where a monad $ℳ $ will have type $Set → Set$, a monad transformer $𝒯$ will have type $(Set → Set) → (Set → Set)$.
-
-Monad transformers are used to extend an existing monad to support another effect.
-The three monads used in this work are the state monad, powerset monad, and the identity monad.
-The state monad is notated $\SM(𝓈)$ (carrying a single cell of type $𝓈$), the powerset monad is notated $\PM$, and the identity is notated $ID$.
-The state and powerset monad transformer equivalents are notated $\ST(𝓈)(ℳ )$ and $\PT(ℳ )$ where $ℳ $ is the underlying monad.
-The state monad has $get$ and $put$ effects, whereas the powerset monad has a $nondeterminism$ effect.
-The transformer versions of both of these monads allow you to combine effects piecewise.
-
-\paragraph{Example:}
-$\ST(ℤ)(\PT(ID))$ is a monad which has $get$ and $put$ effects for a single cell of integer state, in addition to nondeterminism effects.
-
-The definitions, details and proofs of all monads used in our work are given in section \ref{section:Proofs}.
-
-% }}}
+% * Background
+%   - Small Step Semantics
+%     - CPS Language
+%   - Abstract Interpretation (Galois Connections)
+%     - Control Flow Analysis
+%   - AAM
+%   - Monads and Monad Transformers
 
@@
 However for our purposes $↦$ will be always be a function-like form of \emph{decidable} relation, i.e. a multifunction.
+
+% }}}
+
+% CPS {{{
+\subsection{Continuation Passing Style}
+\label{section:Background:ContinuationPassingStyle}
+
+We use CPS-IF as an example language to demonstrate the benefits of our framework.
+CPS is a syntactically restricted version of the lambda calculus.
+CPS-IF is an extension of CPS with conditionals.
+We choose CPS because the state machine for the semantics of CPS requires no call stack.
+We add conditionals to make things slightly more interesting.
+
+We define the CPS-IF language as follows:
+\input{snips/CPS-IF.tex}
+
+CPS-IF supports integer and boolean literals, addition and subtraction by 1, and testing if an integer is greater than or equal to zero.
+The essence of CPS is in the syntactic restriction on application forms $a(a)$ and $a(a,a)$.  
+Because $a$ cannot be a nested call expression, evaluating a $Call$ requires no evaluation stack.
+
+All lambda calculus terms can be translated to CPS terms through CPS-conversion.
+However, because lambda calculus terms are much easier to read, we will use lambda calculus terms for examples.
+We display the CPS-converted version of examples where appropriate.
 
@@
 
-% CPS {{{
-\subsection{Continuation Passing Style}
-\label{section:Background:ContinuationPassingStyle}
-
-We use CPS-IF as an example language to demonstrate the benefits of our framework.
-CPS is a syntactically restricted version of the lambda calculus.
-CPS-IF is an extension of CPS with conditionals.
-We choose CPS because the state machine for the semantics of CPS requires no call stack.
-We add conditionals to make things slightly more interesting.
-
-We define the CPS-IF language as follows:
-\input{snips/CPS-IF.tex}
-
-CPS-IF supports integer and boolean literals, addition and subtraction by 1, and testing if an integer is greater than or equal to zero.
-The essence of CPS is in the syntactic restriction on application forms $a(a)$ and $a(a,a)$.  
-Because $a$ cannot be a nested call expression, evaluating a $Call$ requires no evaluation stack.
-
-All lambda calculus terms can be translated to CPS terms through CPS-conversion.
-However, because lambda calculus terms are much easier to read, we will use lambda calculus terms for examples.
-We display the CPS-converted version of examples where appropriate.
-
-% }}}
-
 % Control Flow Analysis {{{
@@
 
+% Monads {{{
+\subsection{Monads}
+\label{section:Background:Monads}
+
+A basic familiarity with the concept of monads will be necessary for understanding this paper.
+Some find it useful to review the concept of functor before learning about monads. 
+We adopt this approach for our brief explanation of monads.
+
+A type $(F : Set → Set)$ is called a \emph{functor} if one can:
+\begin{itemize}
+\item Define $map : ∀ (A, B : Set), (A → B) → (F(A) → F(B))$
+\item Prove $map(id) = id$, where $id$ is the identity function $(λ(x) → x)$
+\item Prove $map (g ∘ f) = map(g) ∘ map(f)$
+\end{itemize}
+
+\paragraph{Example:} 
+Lists are functors, where $map$ is defined:
+\begin{align*}
+    map(f)(xs) ≔ &\case(xs):           \\
+          Nil    &→ Nil                \\
+         x ∷ xs' &→ f(x) ∷ map(f)(xs') \\
+\end{align*}
+
+\paragraph{Example:} $map(isEven)([1, 2, 3, 4]) = [False, True, False, True]$
+
+Likewise a type $(ℳ  : Set → Set)$ is called a \emph{monad} if one can:
+\begin{itemize}
+\item Define $extend : ∀ (A, B : Set), (A → ℳ (B)) → (ℳ (A) → ℳ (B))$
+\item Prove unit and associativity laws (not mentioned here).
+\end{itemize}
+The only difference from the definition of functor is in the the first argument.
+For monads, this argument is allowed to be a \emph{monadic} $(A → ℳ (B))$, rather than a pure function $(A → B)$.
+
+\paragraph{Example:} Lists are monads, where $extend$ is defined:
+\begin{align*}
+    extend(f)(xs) ≔ &\case(xs):              \\
+                Nil &→ Nil                   \\
+            x ∷ xs' &→ f(x) ⧺ extend(f)(xs') \\
+\end{align*}
+
+\paragraph{Example:} 
+$extend(λ(x) → [x + 1, x - 1])([10, 100]) = [9, 11, 99, 101]$
+
+$extend$ for lists can also be understood as $concat$ followed by $map$:
+\paragraph{Example:} 
+\begin{align*}
+extend(f)([10, 100]) &= concat(map(f)([10, 100]))    \\
+                     &= concat([[9, 11], [99, 101]]) \\
+                     &= [9, 11, 99, 101]             \\
+\end{align*}
+For all monads, it is equivalent to define $return$, $map$ and $join$, where:
+\begin{equation*}
+join : ∀ (A : Set), ℳ (ℳ (A)) → ℳ (A)
+\end{equation*}
+As we have just seen, $join$ for the list monad is just $concat$.
+
+% }}}
+
+% Monad Transformers {{{
+\subsection{Monad Transformers}
+\label{section:Background:MonadTransformers}
+
+Monad transformers are functions between monads.
+Where a monad $ℳ $ will have type $Set → Set$, a monad transformer $𝒯$ will have type $(Set → Set) → (Set → Set)$.
+
+Monad transformers are used to extend an existing monad to support another effect.
+The three monads used in this work are the state monad, powerset monad, and the identity monad.
+The state monad is notated $\SM(𝓈)$ (carrying a single cell of type $𝓈$), the powerset monad is notated $\PM$, and the identity is notated $ID$.
+The state and powerset monad transformer equivalents are notated $\ST(𝓈)(ℳ )$ and $\PT(ℳ )$ where $ℳ $ is the underlying monad.
+The state monad has $get$ and $put$ effects, whereas the powerset monad has a $nondeterminism$ effect.
+The transformer versions of both of these monads allow you to combine effects piecewise.
+
+\paragraph{Example:}
+$\ST(ℤ)(\PT(ID))$ is a monad which has $get$ and $put$ effects for a single cell of integer state, in addition to nondeterminism effects.
+
+The definitions, details and proofs of all monads used in our work are given in section \ref{section:Proofs}.
+
+% }}}
+
 % }}}
@@
 \label{section:MonadicAAM}
+%* Monadic AAM
+%  - 0CFA Before Monads
+%  - 0CFA After Monads
+%  - Recovering the Previous Analysis
 
@@
 \label{section:AbstractControl}
+%* Abstract Control
+%  - Axiomatize monadic interface
+%  - Build interface from legos
+%  - Swapping the legos gives you PS - FS - FI
 
@@
 % Intensional Optimizations {{{
-
 \section{Intensional Optimizations}
 \label{section:IntensionalOptimizations}
+% * Intensional Optimizations
+%   - Garbage Collection
+%   - MCFA
 
@@
 
-% Correctness Framework {{{
+% Correctness {{{
 
@@
 \label{section:Correctness}
+% * Correctness
+%   - All Legos give correct analyses
+%   - Nondeterminism MT
+%   - Monad Transformers relate to state machine functors.
+%   - Galois connections with functors for α γ maps
 
@@
 
+% Future Work {{{
+\section{Future Work}
+\label{FutureWork}
+% * Future Work:
+%   - Verified compilation:
+%     + CompCert, but whole program analysis, functional languages.
+%     + Requires embedding whole framework in proof assistant.
+%     + Mechanizing galois connections is hard, because of calculation.
+%   - More knobs for analysis:
+%     + Lexical Time
+%     + Object sensitivity for functional languages.
+%   - Denotational variants of MAAM
+%     + Picking small-step semantics has consequences.
+%     + Add MonadFix to effects, show a memoizing fixpoint is an abstraction of Y.
+
+% }}}
+
+% Related Work {{{
+\section{Related Work}
+\label{RelatedWork}
+% * Abstract Control:
+%   - Johnson and Van Horn. 2014. Abstracting Abstract Control.
+%     + Discusses push-down as the missing knob.
+%     + Different knob settings for the same knob; hopefully they can be combined.
+% * Compositional Analysis Design: Calculational AI:
+%   - Cousot. 1999. The calculational design of a generic abstract interpreter.
+%     + Works for anything in principle, but hasn't been demonstrated.
+%   - Midgaard and Jensen. 2008. A Calculational Approach to Control-flow Analysis by Abstract Interpretation.
+%     + Calculation is hard; maybe proof assistants can help with this.
+% * New knobs: object sensitivity:
+%   - Kastrinis and Smaragdakis. 2013. Hybrid Context-Sensitivity for Points-To Analysis.
+% * Using Path Sensitivity: 
+%   - Das et al. 2002. ESP: Path-Sensitive Program Verification in Polynomial Time.
+
+% }}}
+
+% Conclusion {{{
+\section{Conclusion}
+\label{Conclusion}
+
+% }}}
+
 % Proofs {{{
-
 \section{Proofs}

### 2014-09-29 writeup/darais-quals.tex
next major revision is to move kcfa into the story

@@
 
-We propose a new compositional framework for designing analyses, and one which supports a tuning knob missing from AAM: abstract control.
+We propose a new compositional framework for designing analyses which supports a tuning knob missing from AAM: abstract control.
 We achieve compositionality by refactoring AAM into monad transformers, each of which we relate back to abstract state machines.
-When put together, these monad transformers form analyses which are correct by construction.
+When put together, these monad transformers build analyses which are correct by construction.
 Furthermore, the monadic abstraction allows us to recover a tuning knob for abstract control when designing analyses.
@@
 To meet the needs of a given problem, an analysis designer has two options: retrofit an existing analysis or design one from scratch.
-However, retrofitting an existing analysis might be insufficient because the tension between precision and performance is often too high.
+Retrofitting an existing analysis can be insufficient because the tension between precision and performance is often too high.
 Getting an analysis right might require more modification than an existing analysis was designed to support.
@@
 \item Describe the semantics of your language.
-\item Design an analysis \emph{which might have the right properties}.
+\item Design an analysis which you hope has right properties.
 \item Exercise the analysis.
@@
 Furthermore, AAM shows that the abstraction step is systematic, freeing the analysis designer to spend effort in designing features of interest.
-However, AAM doesn't provide reusable tools for the design process; the analysis designer must describe their state space transition from scratch.
+However, AAM doesn't provide reusable tools for the design process; the analysis designer must describe their state machine transition from scratch.
 Additionally, some analysis features like flow and path sensitivity don't fit nicely into the AAM framework.
 
-We propose a new compositional framework for designing analyses, and one which supports a tuning knob missing from AAM: abstract control.
+We propose a new compositional framework for designing analyses which supports a tuning knob missing from AAM: abstract control.
 To achieve this we refactor AAM into monad transformers.
 Using monad transformers, the analysis designer can easily add and remove features from their semantics independently of other features.
-The monadic abstraction also allows us to recover a tuning knob for abstract control.
+The monadic abstraction additionaly exposes a new tuning knob for analyses which we call abstract control.
 Abstract control allows one to adjust the flow and path sensitivity of an analysis independent of other design choices.
@@
 \item A monadic formulation of AAM which exposes abstract control as a tuning knob.
-\item A compositional \emph{implementation} framework for building AAM machines using monad transformers.
-\item A compositional \emph{proof} framework for justifying the correctness off AAM machines using galois connection transformers.
-\item A new monad transformer for nondeterminism which satisfies the compositional interface of our framework.
+\item A compositional analysis \emph{implementation} framework for building abstract state machines using monad transformers.
+\item A compositional analysis \emph{proof} framework for justifying the correctness of abstract state machines using galois connection transformers.
+\item A new monad transformer for nondeterminism which is the key to recovering abstract control as a tuning knob.
 \end{enumerate}
@@
 The rest of the paper is structured as follows:
-Section \section{Background} covers necessary background material in semantics, analysis, and the CPS language which we use for examples.
-Section \section{MAAM} demonstrates a monadic variant of AAM which supports defining compositional interpreters for analysis.
-Section \section{AbstractControl} discusses abstract control and how we exploit the monadic abstraction to allow tuning it.
-Section \section{Optimizations} gives two examples of analysis optimizations and how they adapt seamlessly to variations in abstract control.
-Section \section{Correctness} discusses the correctness framework and what key lemmas are involved.
-Section \section{Proofs} shows proof steps in more detail for the arguments made in section \section{Correctness}.
-Sections \section{FutureWork} and \section{RelatedWork} discuss future and related work, and \section{Conclusion} concludes.
+Section \ref{Background} covers necessary background material in semantics, analysis, and the CPS language which we use for examples.
+Section \ref{MAAM} demonstrates a monadic variant of AAM which supports defining compositional interpreters for analysis.
+Section \ref{AbstractControl} discusses abstract control and how we exploit the monadic abstraction to expose a tuning knob for it.
+Section \ref{Optimizations} gives two examples of analysis optimizations and how they adapt seamlessly to variations in abstract control.
+Section \ref{Correctness} discusses the correctness framework and what key lemmas are involved.
+Section \ref{Proofs} shows proof steps in more detail for the arguments made in section \ref{Correctness}.
+Sections \ref{RelatedWork} and \ref{FutureWork} discuss relate and future work, and section \ref{Conclusion} concludes.
 
@@
 
-Monadic abstract interpretation[Sergey et al. PLDI 2013] demonstrated that abstract interpreters can be built modularly using a monadic abstraction.
-We build directly on this work at the level of intuition—monads serve as our pivot-point for achieving modularity.
-However, in our pursuit of a modular correctness framework, our approach deviates greatly from prior work in the specific interfaces used.
-
-First we demonstrate a simple 0CFA analysis on CPS-IF which doesn't use our framework.
-The CPS-IF language is introduced in section \ref{section:Background:ContinuationPassingStyle}.
+Monadic abstract interpretation[Sergey et al. PLDI 2013] demonstrated that abstract interpreters can be built using a monadic abstraction.
+We build directly on the intuition of this work: monadic interpreters are easier to develop than state machines.
+However, our approach takes this idea further in pursuit of a more compositional framework.
+
+First we demonstrate a simple 0CFA analysis on CPS-IF using classic AAM.
+The CPS-IF language is introduced in section \ref{CPS}.
 We repeat the definition here for convenience.
@@
 
-Our abstract semantics of 0CFA will track literals (including lambdas) that appear in the program text.
+The abstract semantics of 0CFA will track literals (including lambdas) that appear in the program text.
 For integers that do not appear in the program text, the analysis uses a single token $INT$ which conservatively approximates any possible integer.
+We note that many more abstractions of integers are possible, like range analysis or relational integer analysis.
+These decisions do not interfere with AAM or abstract control, so we do not discuss them for simplicity.
+
 Formally, the state space $Σ$ for the abstract machine of the abstract semantics is defined as:
-\begin{align*}
-v : \AVal   &⩴ \INT \;|\; l \;|\; \lam(x) → c \;|\; \lam(x,k) → c \\
-σ : \AStore &≔ Var → \PM(\AVal)                                   \\
-Σ           &≔ \PM(Call × \AStore)                                \\
-\end{align*}
-Because there are a finite many literals in the program text, we can claim there are a finite number of possible abstract values.
+\input{snips/CPS-IF-Sigma.tex}
+Because there are finite many literals in the program text, we can claim that $\AVal$ and $\AStore$ are finite.
 
@@
 First we define a \emph{denotation function} $𝒜 $ for $Atom$ expressions.
-\begin{align*}
-𝒜                            &: \AStore × Atom → \PM(\AVal) \\
-𝒜 (σ,x)                      &≔ σ(x)                        \\
-𝒜 (σ,l)                      &≔ \{ l \}                     \\
-𝒜 (σ,\add a) | 𝒜 (σ,a) ⊑ INT &≔ \{ \INT \}                  \\
-𝒜 (σ,\sub a) | 𝒜 (σ,a) ⊑ INT &≔ \{ \INT \}                  \\
-𝒜 (σ,\gez a) | 𝒜 (σ,a) ⊑ INT &≔ \{ \TRUE , \FALSE \}        \\
-𝒜 (σ,\lam(x) → c)            &≔ \{ \lam(x) → c \}           \\
-𝒜 (σ,\lam(x)(k) → c)         &≔ \{ \lam(x)(k) → c \}        \\
-\end{align*}
+\input{snips/CPS-IF-A.tex}
 
 Second we define a \emph{step relation} (as a function) $𝒞$ for $Call$ expressions.
-\begin{align*}
-𝒞                        : &Call × \AStore → \PM(Call × \AStore)                                              \\
-𝒞(\iif(a)\{c₁\}\{c₂\},σ) ≔ &\{ (c,σ)                                                                          \\
-                           &|\; c ∈ ⋃ \{ \{ c₁ \;|\; \TRUE ∈ 𝒜 (σ,a) \}, \{ c₂ \;|\; \FALSE ∈ 𝒜 (σ,a) \} \} \\
-                           &\}                                                                                \\
-𝒞(a₁(a₂,a₃),σ)           ≔ &\{ (c,σ')                                                                         \\
-                           &|\; (\lam(x)(k) → c) ∈ 𝒜 (σ,a₁)                                                   \\
-                           &|\;             v₂ ∈ 𝒜 (σ,a₂)                                                     \\
-                           &|\;             v₃ ∈ 𝒜 (σ,a₃)                                                     \\
-                           &|\;             σ' ≔ σ ⊔ [x ↦ v₂] ⊔ [k ↦ v₃]                                      \\
-                           &\}                                                                                \\
-𝒞(\halt(a),σ)            ≔ &\{ (\halt(a),σ) \}                                                                \\
-\end{align*}
+\input{snips/CPS-IF-C.tex}
 (We omit the $(a₁(a₂))$ case. It is directly analogous to $(a₁(a₂,a₃))$).
 
-The complete analysis of a program $c$ is defined as the least fixed point of a \emph{collection semantics} for the relation $𝒞$:
-\begin{align*}
-\text{analysis} ≔ μ(Σ) → \{(c,⊥)\} ⊔ 𝒞^{⋆}(Σ)
+The complete analysis of an expression $e$ is defined as the least fixed point of a \emph{collection semantics} for the relation $𝒞$:
+\begin{align*}
+\text{analysis} ≔ μ(Σ) → \{(e,⊥)\} ⊔ 𝒞^{⋆}(Σ)
 \end{align*}
@@
 𝒞^{⋆}    &: \PM(Call × \AStore) → \PM(Call × \AStore) \\
-𝒞^{⋆}(Σ) &≔ ⋃ \{ 𝒞(c,σ) | (c,σ) ∈ Σ \}                  \\
+𝒞^{⋆}(Σ) &≔ ⋃ \{ 𝒞(e,σ) | (e,σ) ∈ Σ \}                  \\
 \end{align*}
@@
 
-The first insight in monadic static analysis is to abbreviate the definitions of $𝒜 $ and $𝒞$ using a monad.
+The first insight in monadic abstract interpretation is to abbreviate the definitions of $𝒜 $ and $𝒞$ monadic state rather than explicit state-passing-style.
 At this point, the monad “trick” is nothing more than a technique to simplify the definition of $𝒜 $ and $𝒞$.
-This simplification is similar to how a functional programmer would use a state monad in place of state passing style.
-
-We wll use a powerset monad and state monad transformer to write the analysis in monadic style.
-The full definitions of the monads (and their proofs) used throughout this paper are defered to section \ref{section:Proofs}.
-Because monad transformers are just fancy names for simple types, we write the equivalent simple type underneath definitions which reference monad transformer.
-The monadic conversion of the above analysis is as follows.
-\begin{align*}
-ℳ     &: Set → Set                  \\
-ℳ (a) &≔ \ST(\AStore)(\PM)(a)       \\
-ℳ (a) &≔ \AStore → \PM(a × \AStore) \\
-\end{align*}
-
-\begin{align*}
-𝒜                &: Atom → ℳ (\PM(\AVal))                    \\
-𝒜 (x)            &≔ \ddo                                     \\
-                 &σ ← \getstore                              \\
-                 &\return(σ(x))                               \\
-𝒜 (l)            &≔ \return(\{l\})                              \\
-𝒜 ₘ(add1(a))     &≔ \ddo                                       \\
-                 &v ← 𝒜 (a)                                  \\
-                 &\return(\{\INT \;|\; ∃ i ⊑ \INT ∈ v\})              \\
-𝒜 ₘ(sub1(a))     &≔ \ddo                                       \\
-                 &v ← 𝒜 (a)                                  \\
-                 &\return(\{\INT \;|\; ∃ i ⊑ \INT ∈ v\})              \\
-𝒜 ₘ(gez(a))      &≔ \ddo                                       \\
-                 &v ← 𝒜 (a)                                  \\
-                 &\return(\{\TRUE, \FALSE \;|\; ∃ i ⊑ \INT ∈ v\})      \\
-𝒜 ₘ(\lam(x)(k) → c) &≔ \return(\{\lam(x)(k) → c\})                    \\
-\end{align*}
-
-\begin{align*}
-𝒞                 &: Call → ℳ (Call)                         \\
-𝒞 (\iif(a)\{c₁\}\{c₂\}) &≔ \ddo                                      \\
-                  &vP ← 𝒜 (a)                                \\
-                  &v ← \liftpowerset(vP)                           \\
-                  &b ← \coercebool(v)                            \\
-                  &\return(\{ \text{if}\; b\; \text{then}\; c₁\; \text{else}\; c₂ \}) \\
-𝒞ₘ(a₁(a₂,a₃))     &≔ \ddo                                      \\
-                  &vP ← 𝒜 (a₁)                               \\
-                  &v ← \liftpowerset(vP)                           \\
-                  &(\lam(x)(k) → c) ← \coercefun(v)               \\
-                  &vP₂ ← 𝒜 (a₂)                              \\
-                  &vP₃ ← 𝒜 (a₃)                              \\
-                  &v₂ ← \liftpowerset(vP₂)                         \\
-                  &v₃ ← \liftpowerset(vP₃)                         \\
-                  &\modifystore(λ(σ) → σ ⊔ [x ↦ v₂] ⊔ [k ↦ v₃]) \\
-                  &\return(c)                                    \\
-\end{align*}
-
-The monadic abstraction provides a nice way to simplify the implementation of the analysis.  
-In particular, cases which do not modify the store or only return one result.
-Both of these cases are instances of having "no effect", and they need not mentioned all members of the state machine.  
-(This was observed in Sergey et. al. PLDI 2013.)
+
+We use a powerset monad underneath a state monad transformer to write the analysis in monadic style.
+\input{snips/CPS-IF-M.tex}
+The full definitions of the monads (and their proofs) used throughout this paper are defered to section \ref{Proofs}.
+Because monad transformers are just fancy names for simple types, we write the equivalent simple type underneath definitions which use monad transformer.
+
+The monadic conversion of $𝒜 $ interacts with the store $σ$ using get and put operations.
+\input{snips/CPS-IF-Am.tex}
+
+The monadic conversion of $𝒞$ doesn't use monadic actions directly.
+\input{snips/CPS-IF-Cm.tex}
+The “effects” in $𝒞_{m}$ happen in its calls to $𝒜_{m}$, $\liftpowerset$, $\coercebool$ and $\coercefun$.
+$𝒜_{m}$ will have state effects and $\liftpowerset$, $\coercebool$ and $\coercefun$ will have nondeterminism effects.
+
+$\liftpowerset$, $\coercebool$ and $\coercefun$ use the nondeterminism monadic effect and are defined as follows:
+\input{snips/CPS-IF-P.tex}
+
+The monadic abstraction leads to a simplified definition of the analysis.
+This simplification becomes much more apparent as a state space grows to have more than two states.
+Metafunctions which interact with only one state component can be written generically to support any state space which has that component.
 
@@
 \begin{align*}
-𝒞^{⋆}    &: \PM(Call × \AStore) → \PM(Call × \AStore) \\
-𝒞^{⋆}(Σ) &≔ ⋃ \{ 𝒞(c)(σ) \;|\; (c,σ) ∈ Σ \}             \\
-\end{align*}
-
-Our current definition of $ℳ $ used recovers exactly the analysis we wrote before.
+𝒞_{m}^{⋆}    &: \PM(Call × \AStore) → \PM(Call × \AStore) \\
+𝒞_{m}^{⋆}(Σ) &≔ ⋃ \{ 𝒞_{m}(e)(σ) \;|\; (e,σ) ∈ Σ \}          \\
+\end{align*}
+
+Our current definition of $ℳ $ recovers exactly the analysis we wrote before.
 On top of the convenience of writing things monadically, our insight and contribution is twofold:
 \begin{itemize}
-\item $ℳ $ can be \emph{axiomatized}, such that different definitions for $ℳ $ give rise to different control abstractions for the analysis.
-\item The monadic abstraction provides for a modular proof framework for establishing the correctness of the analysis.
+\item State spaces $Σ$ can be defined and proven correct compositionally using monad transformers.
+\item $𝒞_{m}$ can be written generic to a monadic effect signature, rather than a specific monad.
+      Multiple different monads can be shown to meet this signature, and swapping the monad allows one to tune the flow and path sensitivity of an analysis.
 \end{itemize}
@@
 
-The current instantiation of ℳ  yields a flow-sensitive path-sensitive analysis.  
-Consider a reordering of the powerset\footnote{
-  The definition of $\PT$ we use is non-traditional.  However, our definition
-  for $\PT$ is actually a monad, whereas traditional definitions are not.  Our
-  definition and proofs for $\PT$ are given in section \ref{section:Proofs}.
-}
-and state monad transformers:
+The current instantiation of ℳ  in our example happens to give a flow-sensitive path-sensitive analysis.
+Consider a reordering of powerset and state monad transformers:
 \begin{align*}
@@
 \end{align*}
-
 As before, we must convert between monadic actions and abstract state space transitions to achieve an analysis:
 \begin{align*}
-𝒞^{⋆}       &: \PM(Call) × \AStore → \PM(Call) × \AStore                                         \\
-𝒞^{⋆}(cP,σ) &≔ (\{ c' \;|\; c' ∈ π₁(𝒞(c))(σ) \;|\; c ∈ c\PM \}, ⋃ \{ π₂(𝒞(c))(σ) \;|\; c ∈ cP \}) \\
-\end{align*}
-This instantiation of $ℳ $ and $𝒞*$ yields a flow-insensitive analysis.
-
-For the same monad $ℳ $, we can change the definition of $𝒞*$ to achieve a flow-sensitive path-insensitive analysis:
-\begin{align*}
-𝒞^{⋆}    &: \PM(Call × \AStore) → \PM(Call × \AStore)             \\
-𝒞^{⋆}(Σ) &≔ \{ (c',π₂(𝒞(c))(σ)) \;|\; c' ∈ π₁(𝒞(c))(σ) \;|\; (c,σ) ∈ Σ \} \\
-\end{align*}
-
-In the correctness framework, both the flow-sensitive path-sensitive analysis and flow-insensitive analyses are justified through isomorphisms between monadic actions and abstract state space transitions.
-The flow-sensitive path-insensitive variant is recovered by weakening this isomorphism to a galois connection.
+𝒞_{m}^{⋆}       &: \PM(Call) × \AStore → \PM(Call) × \AStore                                              \\
+𝒞_{m}^{⋆}(cP,σ) &≔ (\{ e' \;|\; e' ∈ π₁(𝒞_{m}(e))(σ) \;|\; e ∈ e\PM \}, ⋃ \{ π₂(𝒞(e))(σ) \;|\; e ∈ cP \}) \\
+\end{align*}
+This instantiation of $ℳ $ and $𝒞_{m}^{⋆}$ yields a flow-insensitive path-insensitive analysis.
+For the same monad $ℳ $, we can change the definition of $𝒞_{m}^{⋆}$ to achieve a flow-sensitive path-insensitive analysis:
+\begin{align*}
+𝒞_{m}^{⋆}    &: \PM(Call × \AStore) → \PM(Call × \AStore)                             \\
+𝒞_{m}^{⋆}(Σ) &≔ \{ (e',π₂(𝒞_{m}(e))(σ)) \;|\; e' ∈ π₁(𝒞_{m}(e))(σ) \;|\; (e,σ) ∈ Σ \} \\
+\end{align*}
+By altering the monad behind the abstraction, we are able to tune the path sensitivity of the analysis.
+For the path-insensitive monad, we were able to further alter the resulting state space $Σ$ to tune the flow sensitivity of the analysis.
+
+% Talk about monad transformers as building blocks.  How these computations (and their proofs) fall out fully compositional components.
 
@@
 \paragraph{Abstract Garbage Collection}
-Abstract garbage collection is a technique in abstract interpretation where unreachable abstract addresses are pruned from the state space.
+Abstract garbage collection is an optimization technique in abstract interpretation where unreachable abstract addresses are pruned from the state space.
 This is analagous to “real” garbage collection, where unreachable pointers are reclaimed for space efficiency.
-However, in abstract semantics, addresses are \emph{re-used} to soundly and finitely aproximate an infinite address space.
-Impreceision in control flow analyses arise when an abstract address present in the current store must be used for allocation.
-Abstract garbage collection is a purely precision and performance improving optimization which removes unreachable abstract addresses.
+However, in abstract semantics, live addresses are \emph{re-used} as part of the sound and finite approximation of an infinite address space.
+When a live address present in the heap must be used for allocation, a static analysis is forced to lose precision.
+Abstract garbage collection exploits the realization that these live addresses can be collected before they are reallocated, resulting in no loss of precision.
 
 For a generic implementation of garbage collection we assume an arbitrary $ℳ $ that has $get$, $put$ and nondeterminism effects.
-Using this axiomatized interface for $ℳ $, abstract garbage collection can be implemented once in a purely generic way.
-\begin{align*}
-gc    &: Call → ℳ (1)                                        \\
-gc(c) &≔ \ddo                                                  \\
+This allows us to change the flow and path sensitivity of the garbage collection without having to reimplement that analysis.
+\input{snips/GC.tex}
+This is literally the implementation of \emph{concrete} garbage collection, but with control properties abstracted behind a monad.
+When instantiated with path-sensitive, flow-sensitive or flow insensitive monads, we can recover the appropriate version of control for abstract garbage collection.
+
+In the original AAM paper, abstract garbage collection is described using a path-sensitive CESK* machine.
+Later in the paper, the path-sensitive CESK* machine is abstracted to use a single global store, which yields a less precise but more tractable analysis.
+However, the single-store variant of the analysis requires one to repeat the “instrument then abstract” design process for abstract garbage collection.
+By abstracting control and exposing it as an orthogonal tuning knob, we are able to design flow and path sensitive variants of garbage collection at once.
+
+\paragraph{MCFA}
+
+MCFA is an optimization that improves the asymptotic complexity of context-sensitive control flow analyses for functional languages.
+Context sensitive analysis were shown to be exponential for functional analyses, yet polynomial for OO analysis, creating an apparent paradox.
+MCFA resolves this paradox by identifying the difference in context sensitive analyses and porting the polynomial behavior from OO to functional analyses.
+The key insight of MCFA is to use packed, copied closures rather than linked closures in the abstract semantics.
+Like garbage collection, using packed rather than linked closures is a simple optimization a compiler or concrete interpreter might use to increase performance.
+Like \emph{abstract} garbage collection, transporting this simple idea from concrete to abstract interpreters reveals a surprising analysis optimization.
+
+The examples so far have covered 0CFA, a context insensitive analysis.
+Before discussing MCFA, we add context sensitivity to our example analysis.
+
+To incorporate context sensitivity into the analysis we do the following:
+\begin{itemize}
+\item Add another state space component to $Σ$ for environments: $\AEnv$.
+      This environment will track which calling function is responsible for binding a given parameter.
+\item Capture this environment when evaluating $λ$ expressions, just like a concrete interpreter would for concrete environments.
+\item Add another state space component which tracks the program points visited so far by the program: $\ATime$.
+\item When interpreting a command $𝒞_{m}(e)$, move time forward by adding $e$ to the list of program points visited.
+\item When calling a function, record the current time next to the function arguments in the environment.
+\end{itemize}
+These steps have the effect of re-analyzing a function twice if it is called from two separate places.
+
+Changing the state space two have two new components is achieved by adding two more state effects two our monad, one for $\AEnv$ and one for $\ATime$:
+\begin{align*}
+  ρ : \AEnv &≔ Var ⇀ \ATime                              \\
+σ : \AStore &≔ (Var,\ATime) → \PM(\AVal)                 \\
+ τ : \ATime &≔ Call \;\text{(the last program location)} \\
+ f,g : Lam  &⩴ \lam(x) → e \;|\; \lam(x,k) → e           \\
+  clo : Clo &⩴ ⟨f, ρ⟩                                    \\
+  v : \AVal &⩴ l \;|\; clo \;|\; INT                     \\
+\end{align*}
+
+Capturing the environment when evaluating $λ$ expressions is a simple monadic action:
+\begin{align*}
+𝒜_{m}                &: Atom → ℳ (\PM(\AVal))      \\
+                 ... &                             \\
+𝒜_{m}(\lam(x) → e)   &≔ \ddo                       \\
+                     &ρ ← \getenv                  \\
+                     &\return (⟨\lam(x) → e, ρ⟩)   \\
+𝒜_{m}(\lam(x,k) → e) &≔ \ddo                       \\
+                     &ρ ← \getenv                  \\
+                     &\return (⟨\lam(x,k) → e, ρ⟩) \\
+                 ... &                             \\
+\end{align*}
+This is the “linked” implementation of closures, which MCFA will replace with a packed version.
+
+Moving abstract time forward is a simple update of the $\ATime$ state to be the current call expression:
+\begin{align*}
+   𝒞_{m}' &: Call → ℳ (Call) \\
+𝒞_{m}'(e) &≔ \ddo            \\
+          &\puttime(e)       \\
+          &𝒞_{m}(e)          \\
+\end{align*}
+
+Calling a function now uses time as an address for storing values in $\AStore$.
+\begin{align*}
+           𝒞_{m} &: Call → ℳ (Call)                            \\
+             ... &                                             \\
+𝒞_{m}(a₁(a₂,a₃)) &≔ \ddo                                       \\
+                 &vP ← 𝒜_{m}(a₁)                               \\
+                 &v ← \liftpowerset(vP)                        \\
+                 &(⟨\lam(x)(k) → e, ρ⟩) ← \coercefun(v)        \\
+                 &vP₂ ← 𝒜_{m}(a₂)                              \\
+                 &vP₃ ← 𝒜_{m}(a₃)                              \\
                  &σ ← \getstore                                \\
-      &𝓉₀ ← \touchedcall(c)                                           \\
-      &\llet 𝓉 ≔ μ(𝓉) → 𝓉 ⊔ \touchedvar(𝓉)                           \\
-      &\modifystore (λ(σ) → ⋃ \{ [x ↦ v] | x ∈ 𝓉 ∧ σ(x) = v \} ) \\
-\end{align*}
-This is literally the implementation of \emph{concrete} garbage collection.
-However, because we have not yet committed to the underlying monad, it seamlessly becomes abstract garbage collection when instantiated with the appropriate monad.
-
-\paragraph{MCFA}
-MCFA is an optimization that improves the asymptotic complexity of context-sensitive control flow analyses.
-kCFA explodes exponentially for $k > 0$ in functional analyses, causing extremely poor analysis performance in the worst case.
-However, an apparent paradox was discovered as kCFA was proven to be polynomial in the worst case for object-oriented programs.
-MCFA resolved the paradox by identifying the difference in analyses and transfering the polynomial behavior to functional analysis.
-The key insight of MCFA is to use packed, copied closures rather than linked closures in the abstract semantics.
-
-So far our analysis has not used closures for abstract values; 0CFA only tracks lambdas without their closing context.
-Adding closures to abstract values is a straightforward addition to the examples we've shown.
-(Our implementation implements all of these features in full.)
-
-The mcfa optimization of copying rather than linking closures similarly enjoys a fully generic implementation:
-\begin{align*}
-  ρ : Env  &≔  Var ⇀ \widehat{Addr}                       \\
-clo : Clo  &⩴ <\lam(x) → c, ρ> \;|\; <\lam(x,k) → c, ρ>   \\
-\end{align*}
-
-\begin{align*}
-clo-copy    &: Lam → ℳ (Clo)                               \\
-clo-copy(l) &≔ \ddo                                        \\
-            &\llet ys ≔ \operatorname{free-vars}(l)        \\
-            &vs ← map^{⋆}(lookup)(ys)                      \\
-            &ρ ← \{ y ↦^{⋆} v \;|\; (y,v) ∈ zip(ys, vs) \} \\
-            &\return (l,ρ)                                 \\
-\end{align*}
-
-Using existing techniques, these optimizations would need to be both implemented and proven correct for each instantiation of abstract control.
-Our generalization over abstract control allows us to implement each of these optimizations once.
-More importantly, the correctness of these optimizations can be established generically as well.
+                 &τ ← \gettime                                 \\
+                 &\putenv(ρ[x ↦ τ][k ↦ τ])                     \\
+                 &\putstore(σ ⊔ [(x,τ) ↦ vP₂] ⊔ [(k,τ) ↦ vP₃]) \\
+                 &\return(e)                                   \\
+             ... &                                             \\
+\end{align*}
+
+Now that we have a context sensitive analysis, MCFA is implemented by creating a new packed environment, rather than just linking with the current environment:
+\input{snips/MCFA.tex}
+
+Like abstract garbage collection, our implementation of MCFA is tunable from one implementation which has abstracted control through a monad.
+Flow and path sensitive variants of MCFA fall out of a simple rearranging of the monad transformer stack.
 
@@
 
-The key advantage to our framework is that the proofs of correctness for constructed analyses are derived automatically.
-To establish the correctness of the monadic approach, we do not merely inline the definitions and resort to reasoning about fully instantiated instances of monadic definitions.
-Instead, we relate monadic actions directly to abstract state space transitions for each monad transformer combination.
+The key advantage to our framework is that the proofs of correctness for constructed analyses are constructed compositionally and derived automatically.
+The proofs are compositional because they are established at the level of monad transformers, not hard-coded for a full monad stack.
+The proofs are derived automatically through the galois-functorality of the monad transformers we use; they transport galois connections in addition to monadic actions.
 Given these proofs, along with an argument about monotonicity of the monadic semantics, we can establish a proof of galois connection for any given interpreter instantiation.
 
-To relate back to small step semantics, we establish a galois connection between monadic actions in $ℳ $ and transitions functions for \emph{some} abstract state space $𝒮𝒮$.
+To relate back to abstract machine semantics, we establish a galois connection between monadic actions in $ℳ $ and transitions functions for \emph{some} abstract state space $𝒮𝒮$.
 This abstract state space is constructed from the monad transformer stack, although some transformer stacks support multiple abstract state spaces.
@@
 
-For the $\PT(\SM(𝓈))$ monad, we show two galois connections are possible to the state space $\PM(\_ × 𝓈)$.
+For the $\PT(\SM(𝓈))$ monad, we show two galois connections are possible: one to the state space $\PM(\_ × 𝓈)$ and one to the state space $(\PM(\_) × 𝓈)$.
 These choices for galois connection correspond to flow-sensitivity and flow-insensitivity respectively.
@@
 A given concrete semantics will always use the flow-sensitive path-sensitive monad.
-We can use this bridge to construct a galois connection between concrete and a flow-insensitive abstract semantics.
-
-When using our framework, the analysis designer need only prove prove:
-\begin{itemize}
-\item The semantic step function $𝒞$ is monotonic.
+Using this bridge we can construct a galois connection between concrete and a flow-insensitive abstract semantics with no added effort.
+
+This table summarizes these results, showing both the monad $ℳ $ and state space $Σ$ which correspond to each analysis:
+\begin{center}\begin{tabular}{| c | c | c |}
+  \hline Sensitivity      & Path Sensitive                    & Path Insensitive                  \\
+  \hline Flow Sensitive   & $ℳ  ≔ S ∘ 𝒫$ \newline $Σ ≔ P ∘ ×$ & $ℳ  ≔ 𝒫 ∘ S$ \newline $Σ ≔ 𝒫 ∘ ×$ \\
+  \hline Flow Insensitive & $DNE$                             & $ℳ  ≔ × ∘ 𝒫$ \newline $Σ ≔ × ∘ 𝒫$ \\
+  \hline                                                                                          \\
+\end{tabular}\end{center}
+(We write $S and 𝒫$ to abbreviate the state and nondeterminism monad
+transformers, and $× and 𝒫$ to abbreviate the cartesian product and
+nondeterminism state space functors.)
+
+In summary, when using our framework, the analysis designer need only prove prove:
+\begin{itemize}
+\item The semantic step function $𝒞_{m}$ is monotonic.
 \item The semantics as written, including intensional optimizations, are correct.
@@
 \begin{itemize}
-\item An automatically derived analysis for their language with their choice of abstract control.
+\item An automatically derived analysis for their language with abstract control left as a tuning knob.
 \item A proof-by-construction of galois connection (soundness and tightness) for the derived analysis.
-\item Seamless extensions to AAM and intensional optimizations (like gc or mcfa), which we demonstrate in our implementation.
-\end{itemize}
-
-Our proof technique is maximally modular in that new monads can be composed seamlessly into the framework with disruption.
-This modularity greatly reduces the proof burden on the analysis designer as new languages and analyses are developed.
-
-In practice, there will be many more state space components, like abstract time and abstract address in the AAM framework.
-To add these features, the analysis designer need only stack more state space monads together to augment the resulting abstract state machine.
-However, once monotonicity of the $𝒞$ action is established, both concrete and abstract interpreters can be derived for free.
-This is only possible because all proofs have been decomposed to the unit of monad transformer, and these extensions reduce to just adding more transformers to the stack.
-
-% }}}
-
-% Future Work {{{
-\section{Future Work}
-\label{FutureWork}
-% * Future Work:
-%   - Verified compilation:
-%     + CompCert, but whole program analysis, functional languages.
-%     + Requires embedding whole framework in proof assistant.
-%     + Mechanizing galois connections is hard, because of calculation.
-%   - More knobs for analysis:
-%     + Lexical Time
-%     + Object sensitivity for functional languages.
-%   - Denotational variants of MAAM
-%     + Picking small-step semantics has consequences.
-%     + Add MonadFix to effects, show a memoizing fixpoint is an abstraction of Y.
+\item Easy integration of new analysis features, a la. AAM.
+\item Single-definition implementation of intensional optimizations like abstract garbage collection or MCFA.
+\end{itemize}
 
@@
 %   - Das et al. 2002. ESP: Path-Sensitive Program Verification in Polynomial Time.
+\paragraph{Abstract Control}
+Johnson and Van Horn recently explored the concept of abstract control in their paper Abstracting Abstract Control (AAC)\cite{AAC}.
+In this work, the authors discuss how to recover a pushdown abstraction systematically in the setting of AAM.
+Our work is orthogonal, and potentially compatible with theirs: our framework explains compositional analyses, while theirs explains pushdown control abstraction.
+We hope for an eventual combination of the two techniques, and see this as a ripe area for combination of results.
+
+As far as the question “what is abstract control?”, it appears that we paint a similar picture as Johnson and Van Horn in AAC.
+In both works, abstract control has something to do with the fixpoint iteration of the semantic step function, and is noted as lacking in classic AAM.
+Our work gives one possible definition for what abstract control is: it's the control effects used by an interpreter, coupled with the way these effects are related to abstract machines.
+
+\paragraph{Compositional Approaches}
+Another compositional approach to abstract interpretation is described by Cousot and Cousot: \emph{calculational abstract interpretation}\cite{CalculationalAI}.
+Calculational AI shares a similar philosophy to our work: abstract interpreters should be build piecewise from compositional components.
+In calculational AI, analyses are derived from applying a sequence of galois connections.
+This contrasts to the approach of designing an analysis first, and manually justifying the galois connection after.
+After applying a sequence of galois connections, the analysis designer must \emph{calculate} the computable analysis from the one derived by galois connection composition.
+
+The downfall of calculational AI is in the effort required during the calculation step.
+For example, Midgaard and Jensen\cite{Calculational0CFA} describe a calculational derivation of 0CFA, the same analysis described in this work, and it is a difficult process.
+
+While work in calculational AI has seen less success than AAM, we note that the two approaches are actually compatible, and thus tackle different parts of the analysis design process.
+AAM is a methodology for splitting the design process to first instrument a concrete semantics, and then systematically abstract.
+Both of these steps must be justified with galois connections, and one could use calculation for either.
+Our work provides computations and proofs for the second step, the systematic abstraction, but one might happily justify the instrumentation step via calculation.
+
+\paragraph{Tuning Knobs}
+Tuning knobs for static analysis are great when they're general enough for a wide range of applications.
+For example, the work Kastrinis and Smaragdakis\cite{HybridContextObject} shows how to leave both context-sensitivity and object-sensitivity as orthogonal tuning knobs in an OO analysis.
+These knobs have shown to be sufficient for interesting applications like malware detection for Android applications\cite{Apposcopy}.
+(\cite{Apposcopy} directly uses the analysis framework described in\cite{HybridContextObject}.)
+
+\paragraph{Path Sensitivity}
+Path sensitivity in particular was shown to be profitible and tractable by Das et al.\cite{ESP}.
+In this work the authors verify the absence of open/close file handle bugs in the GNU C compiler.
+Path sensitivity was crucial achieve the precision needed for their analysis, and their paper describes to to achieve this precision without too much added cost in complexity.
+These subtle variants of path sensitivity are important for applying path sensitive analysis in practice.
+We do not account for these particular variants of path sensitivity in our framework, although we hope to explore how they fit into our framework in future work.
+% }}}
+
+% Future Work {{{
+\section{Future Work}
+\label{FutureWork}
+% * Future Work:
+%   - Verified compilation:
+%     + CompCert, but whole program analysis, functional languages.
+%     + Requires embedding whole framework in proof assistant.
+%     + Mechanizing galois connections is hard, because of calculation.
+%   - More knobs for analysis:
+%     + Lexical Time
+%     + Object sensitivity for functional languages.
+%   - Denotational variants of MAAM
+%     + Picking small-step semantics has consequences.
+%     + Add MonadFix to effects, show a memoizing fixpoint is an abstraction of Y.
+
+\paragraph{Fully Verified Compilation}
+One direction of future work is to apply our approach to fully verified compilation.
+In fully verified compilation, the implementation of a comipler is typically embedded in proof assistant and proven correct all the way down to axioms in a logical system.
+This task is difficult due to both the extra rigor required in modelling and the proof engineering required in interacting with proof assistants.
+We hope the compositionality of the \emph{correctness} of our approach will apply especially in this setting, because proofs need to be developed and maintained side-by-side with code.
+
+CompCert\cite{CompCert} is perhaps the most well known fully verified compiler to date.
+CompCert compiles C programs down to machine code and performs simple local optimizations.
+A verified compiler for languages with rich features like objects and first-class functions benefit especially from approaches like AAM and our extension of it.
+Whole program analysis has also not been studied yet in the context of fully verified compilation, perhaps because establishing the proof of correctness is too burdensome.
+We hope our technique can be extended to support fully verified compilation of whole program analysis for functional programs in the future.
+
+\paragraph{New Tuning Knobs for Analysis}
+Another direction for future work is to discover more tuning knobs for analyses which have been missed by frameworks like AAM.
+For example, it would be nice if we could account for the pushdown abstraction from Johnson and Van Horn's work on \emph{Abstracting Abstract Control} into our monadic formulation.
+Less related to abstract control, we believe there is work to be done in porting object sensitivity to functional language analysis.
+We have preliminary results suggesting that object sensitivity becomes lexically scoped time in the functional setting.
+Exploring this would uncover yet another useful knob to analysis designers to quickly interate through the analysis design process.
+
+\paragraph{New Semantic Settings}
+Our framework small-step abstract machines semantic framework of AAM.
+The small-step setting is convenient, but it has less-well understood consequences when it serves as the basis for abstract interpretation.
+
+For example, we can imagine a more denotational variant of this work where we relate our interpreters to a denotational space rather than state machines.
+We have preliminary results suggesting that this can be done by allowing our monadic interpreter to be recursive.
+When this is done, the meaning of the program is not given by an external fixpoint of the step function, but rather directly through the denotation of recursion.
+To recover an analysis, one would abstract a precise fixpoint finder like the Y-combinator with an approximating fixpoint finder.
 
@@
 \label{Conclusion}
+
+We have shown how to build compositional abstract interpreters which expose a tuning knob for abstract control.
+We use monad transformers as the building blocks for both constructing abstract interpreters and proving them correct piecewise.
+Using the monad abstraction, we showed how to expose abstract control as a tuning knob to recover flow and path sensitivities for a static analysis.
+After exposing a tuning knob for abstract control, we then demonstrated how to implement two analysis optimizations which adapt effortlessly to changes in abstract control.
+Our approach is justified piecewise by relating monad transformers to state machine functors.
+To achieve this, we develop a new monad transformer for nondeterminism and develop a functorial variant of galois connection.
 

### 2014-09-29 writeup-old/sections/03AAMByExample/07Optimizations/main.tex
draft is starting to look good, but still not perfect

We now introduce two optimizations to AAM--heap widening and abstract garbage
collection--and show how they integrate into the computational framework
described thus far.

%--%

\subsubsection{Heap Widening}
\input{sections/03AAMByExample/07Optimizations/00Widen.tex}

\subsubsection{Abstract Garbage Collection}
\input{sections/03AAMByExample/07Optimizations/01AbstractGC.tex}

\subsubsection{Composing Heap Widening and Abstract GC}
\input{sections/03AAMByExample/07Optimizations/02Composition.tex}

### 2014-09-29 writeup-old/sections/04MonadicAAM/01Generalizing/main.tex
draft is starting to look good, but still not perfect

The previous section demonstrated how to write the abstract semnatics from
section~\ref{section:AAMByExample:AbstractSemantics} in monadic style.
%
In this section we generalize the monad, allowing us to instantiate to
different monad stacks so long as they meet an interface.

%--%

The generalized interface will have parameters \p|d|, \p|aam| and \p|m| such
that:
%
\begin{itemize}
\item \h|Delta d| implements a value space primitive operations
\item \h|AAM aam| implements abstract address and time operations
\item \p|m| is a \h|Monad|, \h|MonadPlus| and \h|MonadState| for environment,
      store and time types.
\end{itemize}
%
This interface is encoded in full by the type class constraint 
\h|Semantics d aam m|.
%
\haskell{sections/04MonadicAAM/01Generalizing/00Semantics.hs}

%--%

We next manipulate the semantics into a form generic to parameters \p|d|,
\p|aam| and \p|m| such that \h|Semantics d aam m| holds.
%
\haskell{sections/04MonadicAAM/01Generalizing/01Atom.hs}
%
\haskell{sections/04MonadicAAM/01Generalizing/02Call.hs}
%
The code is omitted because it is identical to the code for the concrete \h|M|
in section~\ref{section:MonadicAAM:MonadicStyle}

%--%

Now we're in some trouble.
%
How do we implement \h|step| when we don't know what monad we're using?
%
The answer is to require the monad to come with an implementation of \h|step|.
%
Fortunately \h|step| can be implemented at the \textit{monad transformer}
level, so the implementation can be derived for an arbitrary monad transformer
stack.

%--%

We capture the ability for a monad to convert monadic actions to state space
transitions in another type class: \h|MonadStep|.
%
\haskell{sections/04MonadicAAM/01Generalizing/03MonadStep.hs}
%
\h|MonadStep| has associated type \h|SS m| for describing the type of step
functions \h|step :: SS m a -> SS m b|.
%
In our framework, we will always be instantiating step to 
\h|step :: SS m Call -> SS m Call|.
%
However this polymorphism is crucial to building state space transitions for
monad transformers, as they may instantiate the polymorphic type \p|a| to
something else.

%--%

We further require that the associated state space \h|SS m| be \h|Pointed| so
that we can inject pure values into the state space.
%
\haskell{sections/04MonadicAAM/01Generalizing/04Pointed.hs}

%--%

The only monad thus far, nondeterminism, is a \h|MonadStep| with its state
space \h|SS| being list and \h|step| being simply monadic bind.
%
\haskell{sections/04MonadicAAM/01Generalizing/05NondetMonadStep.hs}
%
The only monad transformer thus far, state, is a \h|MonadStep| transformer as
well, taking \h|MonadStep| things to \h|MonadStep| things.
%
\haskell{sections/04MonadicAAM/01Generalizing/06StateTMonadStep.hs}
%
Note that in order for the state space associated with \h|StateT| to be
\h|Pointed| we require \h|Cell s| to be a \h|JoinLattice|, meaning it has a
bottom element.

%--%

Executing the semantics can be achieved by injecting the initial \h|Call| value
into the state space and iterating the collecting function as before.
%
\haskell{sections/04MonadicAAM/01Generalizing/07Exec.hs}

### 2014-09-29 writeup-old/sections/04MonadicAAM/main.tex
draft is starting to look good, but still not perfect

We now present monadic AAM.
%
Monadic AAM is part an extension of the AAM methodology, and part extension of
the AAM toolkit.
%
In methodology, interpreters are first abstracted in the AAM tradition, and
then further factored to be monadic.
%
In toolkit, several monads are provided by our library to recover various
analysis properties in a language agnostic fashion.

\subsection{AAM in Monadic Style}
\label{section:MonadicAAM:00}
\input{sections/04MonadicAAM/00MonadicStyle/main.tex}

\subsection{Generalizing the Monad}
\label{section:MonadicAAM:01}
\input{sections/04MonadicAAM/01Generalizing/main.tex}

\subsection{Recovering a Concrete Interpreter}
\label{section:MonadicAAM:02}
\input{sections/04MonadicAAM/02RecoveringConcrete/main.tex}

\subsection{Recovering kCFA}
\label{section:MonadicAAM:03}
\input{sections/04MonadicAAM/03RecoveringKCFA/main.tex}

\subsection{Optimizations}
\label{section:MonadicAAM:04}
\input{sections/04MonadicAAM/04Optimizations/main.tex}

### 2014-09-29 writeup/darais-quals.tex
draft is starting to look good, but still not perfect

@@
 
-\title{Abstract Control in Static Analysis (Draft)}
+\title{Abstract Control in Static Analysis}
 \author{David Darais}
@@
 Techniques based on Abstract Interpretation (AI) allow the analysis designer to vary some properties of an analysis.
-However, features like context, object, and flow sensitivity must be hard-coded into an analysis based on classic AI.
+However, features like context, object, and flow sensitivity are often hard-coded into an analysis based on classic AI.
 Abstracting Abstract Machines (AAM) is one design methodology for analysis which exposes more of these features as tuning knobs.
@@
 
-We propose a new compositional framework for designing analyses which supports a tuning knob missing from AAM: abstract control.
-We achieve compositionality by refactoring AAM into monad transformers, each of which we relate back to abstract state machines.
+We propose a new compositional framework for designing analyses which support a tuning knob missing from AAM: abstract control.
+We achieve compositionality by refactoring AAM into monad transformers, each of which we relate back to abstract state machines piecewise.
 When put together, these monad transformers build analyses which are correct by construction.
-Furthermore, the monadic abstraction allows us to recover a tuning knob for abstract control when designing analyses.
+Furthermore, the monadic abstraction allows us to recover a new tuning knob which we call abstract control.
 Abstract control allows one to adjust the flow and path sensitivity of an analysis independent of other design choices.
@@
 \section{Introduction}
-\label{section:Introduction}
-
-Static analysis is useful for many different situations.
+\label{Intro}
+
+Static analysis has many applications.
 A software developer might use an analysis to find bugs in a program.
@@
 Designing a static analysis is hard.
-Designing \emph{the right} static analysis is even more difficult.
+Designing \emph{the right} static analysis is even harder.
 One possible process for arriving at the right static analysis is:
@@
 To get a complex analysis right this process must be repeated multiple times.
-Iterating quickly between the design and exercise of an analysis is the key to getting the analysis you wanted without too much pain.
-
-One methodology which aims to ease the design process for static analysis is the Abstracting Abstract Machines (AAM) framework\cite{AAM}.
+Iterating quickly between the design and exercise of an analysis is the key to getting the analysis you wanted without too much unnecessary effort.
+
+One methodology which aims to ease the design process for static analysis is the Abstracting Abstract Machines (AAM) 
+  framework\cite{van-horn:2010:aam}.
 AAM separates two design steps which are traditionally tightly coupled: describing features of interest and abstracting those features.
 Furthermore, AAM shows that the abstraction step is systematic, freeing the analysis designer to spend effort in designing features of interest.
-However, AAM doesn't provide reusable tools for the design process; the analysis designer must describe their state machine transition from scratch.
+However, AAM doesn't provide reusable tools for constructing analyses.
+The analysis designer must describe their state machine transition function from scratch.
 Additionally, some analysis features like flow and path sensitivity don't fit nicely into the AAM framework.
@@
 Using monad transformers, the analysis designer can easily add and remove features from their semantics independently of other features.
-The monadic abstraction additionaly exposes a new tuning knob for analyses which we call abstract control.
+The monadic abstraction additionally exposes a new tuning knob for analyses which we call abstract control.
 Abstract control allows one to adjust the flow and path sensitivity of an analysis independent of other design choices.
 
-To justify the soundness of the approach, we show that the monad transformers we use are also galois connection transformers.
+To justify the soundness of the approach, we show that the monad transformers we use are also Galois connection transformers.
 This means that any combination of our monad transformers will yield a sound approximation of a state machine semantics.
@@
 \item A compositional analysis \emph{implementation} framework for building abstract state machines using monad transformers.
-\item A compositional analysis \emph{proof} framework for justifying the correctness of abstract state machines using galois connection transformers.
-\item A new monad transformer for nondeterminism which is the key to recovering abstract control as a tuning knob.
+\item A compositional analysis \emph{proof} framework for justifying the correctness of abstract state machines using Galois connection transformers.
+\item A new monad transformer for nondeterminism which serves as our gateway to recovering abstract control as a tuning knob.
 \end{enumerate}
@@
 Section \ref{Background} covers necessary background material in semantics, analysis, and the CPS language which we use for examples.
-Section \ref{MAAM} demonstrates a monadic variant of AAM which supports defining compositional interpreters for analysis.
+Section \ref{CompositionalAnalysis} demonstrates a monadic variant of AAM which supports defining compositional interpreters for analysis.
 Section \ref{AbstractControl} discusses abstract control and how we exploit the monadic abstraction to expose a tuning knob for it.
@@
 \section{Background}
-\label{section:Background}
-% * Background
-%   - Small Step Semantics
-%     - CPS Language
-%   - Abstract Interpretation (Galois Connections)
-%     - Control Flow Analysis
-%   - AAM
-%   - Monads and Monad Transformers
+\label{Background}
+
+In this section we discuss background material potentially useful for understanding our compositional analysis framework.
+This section provides merely a sketch these concepts, and does not motivate them directly in the context of our work.
+Some readers may find it useful to skip ahead to section \ref{CompositionalAnalysis} and chase backlinks as they are needed and motivated.
+Or some may prefer to proceed directly with this section, consuming the document in a more bottom-up fashion.
+For these readers, we have done our best to keep this section brief.
 
@@
 \subsection{Small-Step Semantics}
-\label{section:Background:SmallStepSemantics}
-
-This work follows the small-step operational approach to language semantics pioneered by Plotkin and Felleisen.
-In small-step operational semantics, the meaning of an expression $e$ of a language is given by the transitive closure of some step relation $e ↦ e'$.
-This yields a state-machine approach to semantics.
-Exploring the meaning of $e$ reduces to exploring all possible transitions reachable from $e$ through $↦$.
-This contrasts to previously developed deontational approaches, where the meaning of $e$ must be given all at once by a denotation function $⟦ e ⟧$.
-This denotation function can quickly demand lots of mathematical machinery.
-Small-step operational methods side-step the need for such machinery.
-
-The small-step operation framework supports arbitrary step relations $e ↦ e'$.  
-However for our purposes $↦$ will be always be a function-like form of \emph{decidable} relation, i.e. a multifunction.
+\label{SS-Semantics}
+
+We use the small-step operational approach to semantics pioneered by independently by 
+  Plotkin\cite{plotkin:1981:sos} 
+  and Felleisen\cite{felleisen:1992:red-sem}.
+In small-step operational semantics, expressions in a language are given meaning by mapping them to state machines.
+Formally, a small-step semantics for a language $L$ is given by:
+\begin{itemize}
+\item A state machine with configuration states $\C$.
+\item A transition relation $σ \Csteps σ'$ for $(\Ce, \Ce' : \C)$.
+\item An embedding function $η : L → \C$.
+\end{itemize}
+The meaning of an expression $e$ is then given by a final configuration $σ$ reachable by $η(e)$ through $\Csteps$.
+
+Small-step semantics contrast to previously developed denotational approaches.
+A denotational semantics for a language $L$ is given by a denotation function $⟦\_⟧$ mapping terms $e$ to objects in a mathematical space.
+This denotation function can quickly demand non-trivial mathematical machinery for even very simple languages.
+Small-step methods side-step the need for such machinery at the cost of potentially useful mathematical structure.
+Many useful analysis properties can be states as properties of execution states of an abstract machine.
+Small-step semantics therefore serve as a nice middle-ground marrying formality and simplicity.
 
@@
 \subsection{Continuation Passing Style}
-\label{section:Background:ContinuationPassingStyle}
-
-We use CPS-IF as an example language to demonstrate the benefits of our framework.
-CPS is a syntactically restricted version of the lambda calculus.
-CPS-IF is an extension of CPS with conditionals.
-We choose CPS because the state machine for the semantics of CPS requires no call stack.
-We add conditionals to make things slightly more interesting.
+\label{CPS}
+
+We use CPS-IF as an example language to demonstrate our framework.
+CPS-IF is a variant of continuation passing style λ-calculus extended with conditionals and branching.
+We use λ-calculus because it is the simplest language with support for higher-order control flow.
+λ-calculus can serve as the core target language for larger languages which appear in practice.
+We use CPS because the state machine for the semantics of CPS requires one less component and it simplifies our presentation.
+All lambda calculus terms can be translated to CPS terms through CPS-conversion, so there is no loss in language expressivity.
+We add conditionals to help exercise analyses which track branching structure in different ways.
 
@@
 The essence of CPS is in the syntactic restriction on application forms $a(a)$ and $a(a,a)$.  
-Because $a$ cannot be a nested call expression, evaluating a $Call$ requires no evaluation stack.
-
-All lambda calculus terms can be translated to CPS terms through CPS-conversion.
-However, because lambda calculus terms are much easier to read, we will use lambda calculus terms for examples.
-We display the CPS-converted version of examples where appropriate.
+Because $a$ cannot be a nested expression, evaluating an $Exp$ requires no evaluation stack.
+
+The state machine $\C$ for CPS-IF is defined as:
+\input{snips/CPS-IF-CSigma.tex}
+Environments map variables to addresses, and the store maps addresses to values.
+Values contain closures, which are lambda expressions paired with their closure environment.
+
+The small-step semantics for CPS-IF are defined in two parts.
+First we define a \emph{denotation function} $𝒜 $ for $Atom$ expressions.
+\input{snips/CPS-IF-CA.tex}
+
+Second we define a \emph{step relation} (as a function) $ℰ$ for $Exp$ expressions.
+\input{snips/CPS-IF-CE.tex}
 
@@
 \subsection{Abstract Interpretation}
-\label{section:Background:AbstractInterpretation}
-
-Abstract Iterpretation (AI) is a formal framework for program analysis pioneered by Cousot and Cousot.
+\label{AI}
+
+Abstract Iterpretation (AI) is a formal framework for program analysis pioneered by Cousot and Cousot\cite{cousot:1977:ai}.
 In the setting of abstract interpretation, a program analysis is just an alternate semantics over an abstract domain.
@@
 The alternate semantics is called the \emph{abstract} semantics\footnote{
-  This terminology quickly becomes confusing because we will eventually construct an abstract machine for both semantics.
+  This terminology quickly becomes confusing when we construct an abstract machine for both semantics.
   This yields a concrete abstract machine and an abstract abstract machine. 
-  This linguistic pun is exploited mercilessly in the work of Might and Van Horn as they “abstract” abstract abstract machines, yielding abstract abstract abstract machines.
+  This linguistic pun is exploited mercilessly in the work of Might and Van Horn as they “abstract” abstract abstract machines.
+  We simplify the matter by calling them abstract state machines and concrete state machines.
 }.
 
-Consider a concrete language $\C$ and a program in that language $(\Ce : \C)$.
+Consider a concrete state machine $\C$ and a state $(\Ce : \C)$.
 In the AI framework, an analysis for $\Ce$ is:
@@
 \item 
-  An abstract language $\A$.
+  An abstract state machine $\A$.
 \item 
@@
 \item 
-  An abstract version of $\Ce$ our concrete program: $(\Ae : \A)$.
+  An abstract version of $\Ce$: $(\Ae : \A)$.
 \item 
@@
 This picture takes place at some point in the analysis process.
-$\Ce$ is the current program (possibly the result of running the base program for a little while).
+$\Ce$ is the current state (possibly the result of running the original program for a little while).
 $\Ae$ is a valid abstraction of $\Ce$, where this validity is expressed by some relation $R$ holding between $\Ce$ and $\Ae$.
@@
 In order to be a correct analysis, it must be guaranteed that $\Ce'$ and $\Ae'$ will be related.
-The logical structure of the picture, which states the correctness of the analysis, is given by:
+The logical structure of the picture, which states the correctness of the analysis $\Asteps$, is given by:
 \begin{equation*}
@@
 \end{equation*}
-The \emph{meaning} of the analysis is entirely subject to the relation $R$ which relates the abstract to the concrete.
+The meaning of the analysis is entirely subject to the relation $R$ which relates the abstract to the concrete.
 
@@
 \subsection{Galois Connections}
-\label{section:Background:GaloisConnections}
-
-The AI setting can be elegently simplified and enriched through the use of \emph{galois connections}.
+\label{GaloisConnections}
+
+The AI setting can be elegantly simplified and enriched through the use of \emph{Galois connections}.
 Galois connections serve as a unifying framework for establishing the “relationship between $\C $ and $\A$” mentioned in the previous section.
 
-A galois connection between two posets (sets with a partial order) $\C$ and $\A$ is notated $\C\galois{α}{γ}\A$ and contains:
+A Galois connection between two posets (sets with a partial order) $\C$ and $\A$ is notated $\C\galois{α}{γ}\A$ and contains:
 \begin{itemize}
@@
 The last two properties can be succinctly stated as $(α ∘ γ ⊑ id ⊑ γ ∘ α)$\footnote{
-  This uses the logical monotonicity relation $f ⊑ g ⇔  (x ⊑ y ⇒  f(x) ⊑ g(y))$ for the function space.
+  This uses the usual monotonicity relation $f ⊑ g ⇔  (x ⊑ y ⇒  f(x) ⊑ g(y))$ for the function space.
 }.
@@
 
+A Galois connection $\C\galois{α}{γ}\A$ can be read “$\A$ is an abstraction of $\C$” and $x ⊑ y$ can be read “$x$ is more precise than $y$”.
 The expansive property corresponds to \emph{soundness}, and the contractive property corresponds to \emph{tightness}.
@@
 \paragraph{Example:}
-Given a galois connection $\C\galois{α}{γ}\A$, there exists a galois connection $(\C → \C)\galois{α'}{γ'}(\A → \A)$ where:
-\begin{align*}
-α'(f : \C → \C) &≔ α ∘ f ∘ γ \\
-γ'(g : \A → \A) &≔ γ ∘ g ∘ α \\
+Given a Galois connection $\C\galois{α}{γ}\A$, there exists a Galois connection 
+  $(\C \xrightarrow{mon} \C)\galois{α'}{γ'}(\A \xrightarrow{mon} \A)$ where:
+\begin{align*}
+  α'(f : \C \xrightarrow{mon} \C) &≔ α ∘ f ∘ γ \\
+  γ'(g : \A \xrightarrow{mon} \A) &≔ γ ∘ g ∘ α \\
 \end{align*}
@@
 \paragraph{Example:} 
-The language $(\PM(ℤ),+,*)$ forms a galois connection with the language $(\PM(\{ EVEN, ODD \}),∧,∨)$ where:
-\begin{align*}
-α(zs : \PM(ℤ))               &≔ ⋃ \{ \{ EVEN  \;|\; ∃ z ∈ zs ∧ Even(z) \},  \{ ODD   \;|\; ∃ z ∈ zs ∧ Odd(z) \} \} \\
-γ(ts : \PM(\{ EVEN, ODD \})) &≔ ⋃ \{ \{ z ∈ ℤ \;|\; EVEN ∈ ts ∧ Even(z) \}, \{ z ∈ ℤ \;|\; ODD ∈ ts ∧ Odd(z) \} \} \\
+The language $(\PM(ℤ),+,*)$ forms a Galois connection with the language $(\PM(\{ EVEN, ODD \}),∧,∨)$ where:
+\begin{align*}
+α(zs : \PM(ℤ))               &≔ \{ EVEN  \;|\; ∃ z ∈ zs ∧ Even(z) \} ∪ \{ ODD   \;|\; ∃ z ∈ zs ∧ Odd(z) \}  \\
+γ(ts : \PM(\{ EVEN, ODD \})) &≔ \{ z ∈ ℤ \;|\; EVEN ∈ ts ∧ Even(z) \} ∪ \{ z ∈ ℤ \;|\; ODD ∈ ts ∧ Odd(z) \} \\
 \end{align*}
@@
 $x ⊑ γ(y)$ can be seen as a heterogenous extension of $x ⊑ y$ when $(x : A)$ and $(y : B)$ live in different sets. 
-This heterogenous order is given meaning when there exists a galois connection $A\galois{α}{γ}B$.
-One can also think of galois connections as something like an isomorphism, but with a weaker round-trip property.
+One can also think of Galois connections as something like an isomorphism, but with a weaker round-trip property.
 (An isomorphism would require $α ∘ γ = id = γ ∘ α$.)
 
-Using galois connections, the AI framework introduced in the previous section can be re-stated. In the AI framework, an analysis for $\Ce$ is:
+Using Galois connections, the AI framework introduced in the previous section can be re-stated. In the AI framework, an analysis for $\Ce$ is:
 \begin{itemize}
@@
 \item 
-  A galois connection $\C\galois{α}{γ}\A$.
+  A Galois connection $\C\galois{α}{γ}\A$.
 \item 
@@
 \input{diagrams/AIGC.tikz}
-As before, the logical structure of the picture is given by:
+As before, the logical structure of the picture, which states the correctness of the analysis $\Asteps$, is given by:
 \begin{equation*}
@@
 The statement of this property can be simplified further as $\Csteps ⊑ γ(\Asteps)$.
-This uses the definition of galois connections for function spaces shown in a previous example.
-
-Using a galois connection $\C\galois{α}{γ}\A$ and abstract step function $\Asteps ⊑ γ(\Csteps)$, the analysis story becomes simplified even further:
-\begin{itemize}
+This uses the definition of Galois connections for function spaces shown in a previous example.
+
+Using a Galois connection $\C\galois{α}{γ}\A$ and abstract step function $\Asteps ⊑ γ(\Csteps)$, an algorithm for analysis can be stated:
+\begin{enumerate}
 \item Translate $(\Ce : \C)$ to $(α(\Ce) : \A)$
 \item Explore all abstract states $\Ae'$ states reachable from $α(\Ce)$.
-\item All reachable concrete states are summarized by projecting each $γ(\Ae')$ back to $\C$.
-\end{itemize}
-
-The galois connection framework simultaneously guarantees the \emph{soundness} and \emph{tightness} of this method.
-Soundness tells us that if we abstract, take an abstract step, and then concretize, then the result is an approximation of taking a concrete step.
+\end{enumerate}
+All reachable concrete states are then summarized by projecting each $γ(\Ae')$ back to $\C$.
+
+The Galois connection framework simultaneously justifies the \emph{soundness} and \emph{tightness} of this method.
+Soundness tells us that if we (1) abstract, (2) take an abstract step, and then (3) concretize, 
+  then the result is an approximation of (4) taking a concrete step.
 Soundness tells us that we can trust the results of the analysis.
 \input{diagrams/GCSoundness.tikz}
-Tightness tells us that if we concretize, take a concrete step, and then abstract, then the result must be just as precise as taking an abstract step.
-Tightness tells us that our abstract step function isn't losing precision unnecessarily, and that our abstract step $\Asteps$ is provably as strong as possible.
+Tightness tells us that if we (1) concretize, (2) take a concrete step, and then (3) abstract, 
+  then the result must be just as precise as (4) taking an abstract step.
+Tightness tells us that our abstract step function isn't losing precision unnecessarily, 
+  and that our abstract step $\Asteps$ is provably as precise as possible.
 \input{diagrams/GCTightness.tikz}
+The above diagrams compose to state similar properties about $\Csteps*$ and $\Asteps*$.
 
@@
 \subsection{Control Flow Analysis}
-\label{section:Background:ControlFlowAnalysis}
+\label{CFA}
 
 Control flow analysis is a class of analysis which is particularly important for higher-order languages.
-In non-higher-order languages, it is useful to distinguish control-flow from data-flow.
-The control flow of the program is a graph of what functions are called from where.
-The data flow of a program is a mapping of which values can flow to which variables.
-Traditionally one performs a control-flow analysis first, to find out which functions are called, and a data-flow analysis second, using the results of the control flow analysis.
-
+In non-higher-order languages, it can be useful to distinguish control-flow from data-flow.
+The control flow of the program is the graph of what functions are called from where.
+The data flow of a program is the mapping of which values flow to which variables.
+Traditionally one performs a control-flow analysis first to find out which functions are called, 
+  and a data-flow analysis second using the results of the control flow analysis.
 In higher-order languages, data-flow and control-flow are tightly coupled.
-Before you can tell which functions will be called, you need to know how values flow, because functions are themselves values.
+Before you can tell which functions will be called, you need to know how values will flow through the program.
 
 We note that the distinction between “higher-order” languages and “non-higher-order” languages is a red herring.
-Functions can be passed as values in C using function pointers, and object-oriented languages enjoy a similar circularity between control- and data- flow due to method dispatch.
+Functions can be passed as values in C using function pointers.
+Object-oriented languages enjoy a similar circularity between control and data flow from dynamic method dispatch.
 You can read these statements as saying “all languages are higher order, and thus all languages need higher order control flow analysis”.
 Or you can read them as saying “all languages are higher order, and we've done just fine without higher oder control flow analysis so far”.
-The real distinction is in the code that you end up analysing.
-
-We build on the tradition of control flow analysis (CFA) pioneered by Shivers, as well as recent refinements of CFA developed by Might and Van Horn.
-In this tradition, the result of a control flow analysis is closer to what one would think of as a data-flow analysis, only the process accounts for higher order flow.
+The distinction in practice comes from the properties you are looking for and the code that you end up analysing.
+
+We build on the tradition of Control Flow Analysis (CFA) pioneered by Shivers\cite{shivers:1991:cfa}.
+In this tradition, the result of a control flow analysis is closer to what one would think of as a data-flow analysis.
+CFA is nothing more than getting data-flow right in a higher-order setting.
 0CFA, the most basic control flow analysis, merely computes the set of values which might flow to a particular variable.
-kCFA is a suite of context sensitive extensions to 0CFA for which functions are analyzed separately for each possible calling context.
-We use 0CFA as the example analysis in this paper.
-However our framework, and our implementation of it, scale to the full range of context sensitive control flow analyses.
+kCFA is a family of context sensitive extensions to 0CFA which separates analysis results for functions based on call site.
+We use 0CFA and kCFA as the example analyses in this paper.
 
@@
 \subsection{Monads}
-\label{section:Background:Monads}
-
-A basic familiarity with the concept of monads will be necessary for understanding this paper.
+\label{Monads}
+
 Some find it useful to review the concept of functor before learning about monads. 
@@
 
+\paragraph{Functors}
 A type $(F : Set → Set)$ is called a \emph{functor} if one can:
@@
 \item Define $map : ∀ (A, B : Set), (A → B) → (F(A) → F(B))$
-\item Prove $map(id) = id$, where $id$ is the identity function $(λ(x) → x)$
+\item Prove $map(id) = id$
 \item Prove $map (g ∘ f) = map(g) ∘ map(f)$
@@
 
-\paragraph{Example:} 
+\begin{example}
 Lists are functors, where $map$ is defined:
@@
 \end{align*}
-
-\paragraph{Example:} $map(isEven)([1, 2, 3, 4]) = [False, True, False, True]$
-
-Likewise a type $(ℳ  : Set → Set)$ is called a \emph{monad} if one can:
+\end{example}
+
+\begin{example} 
+  $map(isEven)([1, 2, 3, 4]) = [False, True, False, True]$ 
+\end{example}
+
+\paragraph{Monads}
+A type $(ℳ  : Set → Set)$ is called a \emph{monad} if one can:
 \begin{itemize}
 \item Define $extend : ∀ (A, B : Set), (A → ℳ (B)) → (ℳ (A) → ℳ (B))$
-\item Prove unit and associativity laws (not mentioned here).
+\item Prove $extend(return) = id$, called the left-unit law.
+\item Prove $extend(f)(return(x)) = f(x)$, called the right-unit law.
+\item Prove $extend(g)(extend(f)(x)) = extend(λ(y) → extend(g)(f(y)))(x)$ called associativity.
 \end{itemize}
-The only difference from the definition of functor is in the the first argument.
+The only computational difference from the definition of functor is in the first argument of extend.
 For monads, this argument is allowed to be a \emph{monadic} $(A → ℳ (B))$, rather than a pure function $(A → B)$.
 
-\paragraph{Example:} Lists are monads, where $extend$ is defined:
+\begin{example} Lists are monads, where $extend$ is defined:
 \begin{align*}
@@
                 Nil &→ Nil                                                                      \\
-            x ∷ xs' &→ f(x) ⧺ extend(f)(xs') \\
-\end{align*}
-
-\paragraph{Example:} 
-$extend(λ(x) → [x + 1, x - 1])([10, 100]) = [9, 11, 99, 101]$
-
-$extend$ for lists can also be understood as $concat$ followed by $map$:
-\paragraph{Example:} 
-\begin{align*}
-extend(f)([10, 100]) &= concat(map(f)([10, 100]))    \\
-                     &= concat([[9, 11], [99, 101]]) \\
-                     &= [9, 11, 99, 101]             \\
-\end{align*}
-For all monads, it is equivalent to define $return$, $map$ and $join$, where:
-\begin{equation*}
-join : ∀ (A : Set), ℳ (ℳ (A)) → ℳ (A)
-\end{equation*}
-As we have just seen, $join$ for the list monad is just $concat$.
+            x ∷ xs' &→ f(x) ⧺\footnote{⧺ is our notation for list concatenation} extend(f)(xs') \\
+\end{align*}
+\end{example}
+
+\begin{example} 
+Let $adjacent(x) ≔ [x - 1, x + 1]$.  Then $extend(adjacent)([10, 100]) = [9, 11, 99, 101]$
+\end{example}
+
+\paragraph{Monad Notation}
+One of the biggest benefits of monads is perhaps the most superficial: do-notation.
+Do-notation transforms imperative-looking code into uses of $extend$.
+This enables kind of “abstract semicolon”, where the meaning of sequencing statements is give by an arbitrary implementation of $extend$.
+
+Do-notation is simply a translation from:
+\begin{align*}
+  do&                  \\
+    &x ← get           \\
+    &y ← get           \\
+    &put(x + y)        \\
+\end{align*}
+to
+\begin{align*}
+  extend(λ(x) → extend(λ(y) → put(x + y)(get)))(get)
+\end{align*}
+
+Leveraging even more notation, we can write $extend(f)(x)$ infix and backwards as $x ≫= f$, yielding the following equivalent translation:
+\begin{align*}
+  &get ≫= λ(x) →    \\
+  &get ≫= λ(y) →    \\
+  &put(x + y)       \\
+\end{align*}
+which looks closer to the original expression written in do-notation.
+For the rest of the paper we will write exclusively using do-notation and assume this lightweight translation to uses of $extend$.
+
+% }}}
+
+% Monadic Effects {{{
+\subsection{Monadic Effects}
+\label{MonadicEffects}
+
+In addition to $return$ and $extend$, monads typically come with their own flavor of \emph{monadic effect}.
+We use two monadic effects in this paper: state and nondeterminism.
+
+\paragraph{State}
+The state monadic effect models manipulation of a single cell of state with type $𝓈$ using operations $get$ and $put$:
+\begin{align*}
+  get &: ℳ (𝓈)      \\
+  put &: 𝓈 → ℳ (1)  \\
+\end{align*}
+(We write $1$ as the unit type, i.e. the singleton set.)
+$get$ and $put$ come with the following laws:
+\begin{align*}
+               put(𝓈_1) ; put(𝓈_2) &= put (𝓈_2)           \\
+                      put(𝓈) ; get &= return(𝓈)           \\
+                  𝓈 ← get ; put(𝓈) &= return(∙)           \\
+𝓈_1 ← get ; 𝓈_2 ← get ; k(𝓈_1,𝓈_2) &= 𝓈 ← get ; k(𝓈,𝓈)    \\
+\end{align*}
+Multiple cells of state can be modeled by stacking this effect multiple times.
+
+\paragraph{Nondeterminism}
+The nondeterminism monadic effect models branching control flow using a single operation $⟨+⟩$:
+\begin{align*}
+    ⊥ &: ∀ (a : A), ℳ (A)                  \\
+  ⟨+⟩ &: ∀ (a : A), ℳ (A) → ℳ (A) → ℳ (A)  \\
+\end{align*}
+$⟨+⟩$ comes with the following laws:
+\begin{align*}
+         extend(\_)(⊥) &= ⊥                                \\
+  extend(const(⊥))(\_) &= ⊥                                \\
+    extend(f)(x\; ⟨+⟩\; y) &= extend(f)(x)\; ⟨+⟩\; extend(f)(y)    \\
+\end{align*}
+A null branch of a computation is written $⊥ $ and branching with two possible values $v_1 v_2$ is written $return(v_1)\; ⟨+⟩\; return(v_2)$.
+Note that $ℳ $ having a nondeterminism effect is equivalent to $ℳ $ being a join-semilattice for every possible $ℳ (a)$.
 
@@
 \subsection{Monad Transformers}
-\label{section:Background:MonadTransformers}
-
-Monad transformers are functions between monads.
+\label{MonadTransformers}
+
+Monad transformers are building blocks for constructing larger monads.
 Where a monad $ℳ $ will have type $Set → Set$, a monad transformer $𝒯$ will have type $(Set → Set) → (Set → Set)$.
-
 Monad transformers are used to extend an existing monad to support another effect.
-The three monads used in this work are the state monad, powerset monad, and the identity monad.
-The state monad is notated $\SM(𝓈)$ (carrying a single cell of type $𝓈$), the powerset monad is notated $\PM$, and the identity is notated $ID$.
-The state and powerset monad transformer equivalents are notated $\ST(𝓈)(ℳ )$ and $\PT(ℳ )$ where $ℳ $ is the underlying monad.
-The state monad has $get$ and $put$ effects, whereas the powerset monad has a $nondeterminism$ effect.
+
+The three monads used in this work are the state monad, nondeterminism monad, and identity monad.
+The state monad is notated $\SM(𝓈)$ (carrying a single cell of type $𝓈$), nondeterminism $\PM$, and identity $ID$.
+The transformer analogues for state and nondeterminism are notated $\ST(𝓈)(ℳ )$ and $\PT(ℳ )$ where $ℳ $ is the underlying monad.
 The transformer versions of both of these monads allow you to combine effects piecewise.
 
-\paragraph{Example:}
-$\ST(ℤ)(\PT(ID))$ is a monad which has $get$ and $put$ effects for a single cell of integer state, in addition to nondeterminism effects.
-
-The definitions, details and proofs of all monads used in our work are given in section \ref{section:Proofs}.
-
-% }}}
-
-% }}}
-
-% Monadic AAM {{{
-
-\section{Monadic AAM}
-\label{section:MonadicAAM}
-%* Monadic AAM
-%  - 0CFA Before Monads
-%  - 0CFA After Monads
-%  - Recovering the Previous Analysis
-
-Monadic abstract interpretation[Sergey et al. PLDI 2013] demonstrated that abstract interpreters can be built using a monadic abstraction.
-We build directly on the intuition of this work: monadic interpreters are easier to develop than state machines.
-However, our approach takes this idea further in pursuit of a more compositional framework.
-
-First we demonstrate a simple 0CFA analysis on CPS-IF using classic AAM.
-The CPS-IF language is introduced in section \ref{CPS}.
-We repeat the definition here for convenience.
+\begin{example}
+  $\ST(ℤ)(\PT(ID))$ and $\PT(\ST(ℤ)(ID))$ are both monads with state (where $𝓈 = ℤ$) and nondeterminism effects.
+\end{example}
+
+The non-transformer versions of state and nondeterminism monads can be defined directly, or simply as transformers on top of $ID$.
+The nondeterminism monad transformer we use in this work is new.
+The full definitions and proofs of all monad transformers used in this paper are given in section \ref{Proofs}.
+
+% }}}
+
+% }}}
+
+% A Compositional Analysis Framework {{{
+
+\section{A Compositional Analysis Framework}
+\label{CompositionalAnalysis}
+
+A compositional analysis framework must support adding and removing features from the analysis without touching the entire implementation.
+We achieve such compositionality by building abstract interpreters using monad transformers.
+To demonstrating our compositional approach to analysis design, we first show a classic definition of 0CFA.
+Then, we will show how to design such an analysis using our framework.
+After porting 0CFA to our framework, we will show how to grow this analysis to include context sensitivity in a compositional fashion.
+
+We will use the CPS-IF language introduced in section \ref{CPS} as the target language for 0CFA.
+We repeat the definition of CPS-IF syntax here for convenience.
 \input{snips/CPS-IF.tex}
 
-The abstract semantics of 0CFA will track literals (including lambdas) that appear in the program text.
+\subsection{Classic 0CFA}
+\label{AAM-0CFA}
+
+Our abstract semantics for classic 0CFA will track literals (including lambdas) that appear in the program text.
 For integers that do not appear in the program text, the analysis uses a single token $INT$ which conservatively approximates any possible integer.
-We note that many more abstractions of integers are possible, like range analysis or relational integer analysis.
-These decisions do not interfere with AAM or abstract control, so we do not discuss them for simplicity.
-
-Formally, the state space $Σ$ for the abstract machine of the abstract semantics is defined as:
-\input{snips/CPS-IF-Sigma.tex}
+We note that many more abstractions for integers are possible, like integer ranges or symbolic binary relations among variables.
+These decisions do not interfere with AAM or abstract control, so we pick the simplest abstraction to simplify presentation.
+
+The state space $\A$ for 0CFA is defined as:
+\input{snips/CPS-IF-ASigma.tex}
 Because there are finite many literals in the program text, we can claim that $\AVal$ and $\AStore$ are finite.
 
-The abstract semantics for $Σ$ comes in two parts.  
+The abstract semantics for $\A$ comes in two parts.  
 First we define a \emph{denotation function} $𝒜 $ for $Atom$ expressions.
-\input{snips/CPS-IF-A.tex}
-
-Second we define a \emph{step relation} (as a function) $𝒞$ for $Call$ expressions.
-\input{snips/CPS-IF-C.tex}
+\input{snips/CPS-IF-AA.tex}
+
+Second we define a \emph{step relation} (as a function) $ℰ$ for $Exp$ expressions.
+\input{snips/CPS-IF-AE.tex}
 (We omit the $(a₁(a₂))$ case. It is directly analogous to $(a₁(a₂,a₃))$).
 
-The complete analysis of an expression $e$ is defined as the least fixed point of a \emph{collection semantics} for the relation $𝒞$:
-\begin{align*}
-\text{analysis} ≔ μ(Σ) → \{(e,⊥)\} ⊔ 𝒞^{⋆}(Σ)
+The complete analysis of an expression $e$ is defined as the least fixed point of a \emph{collection semantics} for the relation $ℰ$:
+\begin{align*}
+\text{analysis} ≔ μ(\Ae) → \{(e,⊥)\} ⊔ ℰ^{⋆}(\Ae)
 \end{align*}
@@
 \begin{align*}
-𝒞^{⋆}    &: \PM(Call × \AStore) → \PM(Call × \AStore) \\
-𝒞^{⋆}(Σ) &≔ ⋃ \{ 𝒞(e,σ) | (e,σ) ∈ Σ \}                  \\
-\end{align*}
-The collecting semantics tracks all states that the program could be in rather than just the final states.
-
-The first insight in monadic abstract interpretation is to abbreviate the definitions of $𝒜 $ and $𝒞$ monadic state rather than explicit state-passing-style.
-At this point, the monad “trick” is nothing more than a technique to simplify the definition of $𝒜 $ and $𝒞$.
-
-We use a powerset monad underneath a state monad transformer to write the analysis in monadic style.
+ℰ^{⋆}    &: \PM(Exp × \AStore) → \PM(Exp × \AStore) \\
+  ℰ^{⋆}(\Ae) &≔ ⋃_{e, θ ∈ \AE} ℰ(e,θ) \\
+\end{align*}
+A collecting semantics is used to track all states that the program could be in rather than just the final states.
+
+\subsection{Monadic 0CFA}
+\label{MAAM-0CFA}
+
+The first insight in monadic abstract interpretation is to abbreviate the definitions of $𝒜 $ and $ℰ$ using monadic state.
+We use a nondeterminism monad underneath a state monad transformer to write the analysis in monadic style.
+Monad transformers are just fancy names for simple types, 
+  and we write simple types underneath definitions which use monad transformer to aid understanding.
 \input{snips/CPS-IF-M.tex}
-The full definitions of the monads (and their proofs) used throughout this paper are defered to section \ref{Proofs}.
-Because monad transformers are just fancy names for simple types, we write the equivalent simple type underneath definitions which use monad transformer.
-
-The monadic conversion of $𝒜 $ interacts with the store $σ$ using get and put operations.
-\input{snips/CPS-IF-Am.tex}
-
-The monadic conversion of $𝒞$ doesn't use monadic actions directly.
-\input{snips/CPS-IF-Cm.tex}
-The “effects” in $𝒞_{m}$ happen in its calls to $𝒜_{m}$, $\liftpowerset$, $\coercebool$ and $\coercefun$.
-$𝒜_{m}$ will have state effects and $\liftpowerset$, $\coercebool$ and $\coercefun$ will have nondeterminism effects.
+
+The monadic conversion of $𝒜 $ interacts with the store $θ$ using get and put operations.
+\input{snips/CPS-IF-AAm.tex}
+
+The monadic conversion of $ℰ$ uses $\liftpowerset$ to convert from powerset to monadic values.
+$\coercebool$ and $\coercefun$ are used to coerce lambdas out of abstract values, failing (with $⊥$) otherwise.
+The “effects” in $ℰ_{m}$ happen in calls to $𝒜_{m}$, $\liftpowerset$, $\coercebool$ and $\coercefun$.
+\input{snips/CPS-IF-AEm.tex}
 
@@
 
-The monadic abstraction leads to a simplified definition of the analysis.
-This simplification becomes much more apparent as a state space grows to have more than two states.
-Metafunctions which interact with only one state component can be written generically to support any state space which has that component.
+The monadic abstraction leads to a more compositional definition of the analysis.
+Pieces of the analysis which only use one effect need not mention the others.
+This compositionality will become more apparent when we extend this analysis to include context sensitivity.
 
@@
 \begin{align*}
-𝒞_{m}^{⋆}    &: \PM(Call × \AStore) → \PM(Call × \AStore) \\
-𝒞_{m}^{⋆}(Σ) &≔ ⋃ \{ 𝒞_{m}(e)(σ) \;|\; (e,σ) ∈ Σ \}          \\
-\end{align*}
-
-Our current definition of $ℳ $ recovers exactly the analysis we wrote before.
-On top of the convenience of writing things monadically, our insight and contribution is twofold:
+ℰ_{m}^{⋆}    &: \PM(Exp × \AStore) → \PM(Exp × \AStore) \\
+  ℰ_{m}^{⋆}(\Ae) &≔ ⋃_{e, θ ∈ \Ae} ℰ_{m}(e)(θ) \\
+\end{align*}
+This definition of $ℳ $ and $ ℰ_{m}^{⋆}$ recovers exactly the analysis we wrote before.
+
+\subsection{Monadic kCFA}
+\label{MAAM-kCFA}
+
+To incorporate context sensitivity (for $k=1$) into the analysis we do the following:
 \begin{itemize}
-\item State spaces $Σ$ can be defined and proven correct compositionally using monad transformers.
-\item $𝒞_{m}$ can be written generic to a monadic effect signature, rather than a specific monad.
-      Multiple different monads can be shown to meet this signature, and swapping the monad allows one to tune the flow and path sensitivity of an analysis.
+\item Add another state space component to $\A$ for environments: $\AEnv$.
+      This environment will track which calling function is responsible for binding a given parameter.
+\item Capture this environment when evaluating $λ$ expressions, just the concrete semantics does for concrete environments.
+\item Add another state space component which tracks the program points visited so far by the program: $\ATime$.
+\item When interpreting a command $ℰ_{m}(e)$ set the time to $e$.
+\item When calling a function, record the current time next to the function arguments in the environment.
 \end{itemize}
+These steps have the effect of re-analyzing a function twice if it is called from two separate places.
+
+Changing the state space two have two new components is achieved simply by adding two more state effects two our monad, 
+  one for $\AEnv$ and one for $\ATime$.
+The resulting abstract state space looks like this:
+\begin{align*}
+  ρ : \AEnv &≔ Var ⇀ \ATime                                     \\
+θ : \AStore &≔ (Var,\ATime) → \PM(\AVal)                        \\
+ τ : \ATime &≔ Exp \;\text{(the last program location visited)} \\
+ f,g : Lam  &⩴ \lam(x) → e \;|\; \lam(x,k) → e                  \\
+  clo : Clo &⩴ ⟨f, ρ⟩                                           \\
+  v : \AVal &⩴ l \;|\; clo \;|\; INT                            \\
+  \Ae : \A  &≔ \PM(Exp × \ATime × \AEnv × \AStore)              \\
+\end{align*}
+
+Using monad transformers, we can easily add two new state components to our monadic interpreter.
+\begin{align*}
+ℳ     &: Set → Set                                                    \\
+  ℳ (a) &≔ \ST(\AEnv)(\ST(\ATime)(\ST(\AStore)(\PM))(a)               \\
+ℳ (a) &≔ \AEnv × \ATime × \AStore → \PM(a × \AEnv × \ATime × \AStore) \\
+\end{align*}
+
+Capturing the environment when evaluating $λ$ expressions is a simple use of monadic state:
+\begin{align*}
+𝒜_{m}                &: Atom → ℳ (\PM(\AVal)) \\
+                 ... &                        \\
+𝒜_{m}(f)             &≔ \ddo                  \\
+                     &ρ ← \getenv             \\
+                     &\return (\{⟨f, ρ⟩\})    \\
+                 ... &                        \\
+\end{align*}
+
+Moving abstract time forward is a simple update of the $\ATime$ state to be the current expression:
+\begin{align*}
+   ℰ_{m}' &: Exp → ℳ (Exp) \\
+ℰ_{m}'(e) &≔ \ddo          \\
+          &\puttime(e)     \\
+          &ℰ_{m}(e)        \\
+\end{align*}
+
+Calling a function now uses time as an address for storing values in $\AStore$.
+This is analogous to the allocation concrete semantics shown in section \ref{CPS}.
+\begin{align*}
+           ℰ_{m} &: Exp → ℳ (Exp)                              \\
+             ... &                                             \\
+ℰ_{m}(a₁(a₂,a₃)) &≔ \ddo                                       \\
+                 &vP ← 𝒜_{m}(a₁)                               \\
+                 &v ← \liftpowerset(vP)                        \\
+                 &(⟨\lam(x)(k) → e, ρ⟩) ← \coercefun(v)        \\
+                 &vP₂ ← 𝒜_{m}(a₂)                              \\
+                 &vP₃ ← 𝒜_{m}(a₃)                              \\
+                 &θ ← \getstore                                \\
+                 &τ ← \gettime                                 \\
+                 &\putenv(ρ[x ↦ τ][k ↦ τ])                     \\
+                 &\putstore(θ ⊔ [(x,τ) ↦ vP₂] ⊔ [(k,τ) ↦ vP₃]) \\
+                 &\return(e)                                   \\
+             ... &                                             \\
+\end{align*}
+
+Finally, lookup for variables must follow the indirection of addresses through $\AStore$.
+\begin{align*}
+𝒜_{m}                &: Atom → ℳ (\PM(\AVal)) \\
+                 ... &                        \\
+𝒜_{m}(x)             &≔ \ddo                  \\
+                     &ρ ← \getenv             \\
+                     &θ ← \getstore           \\
+                     &\return (θ(x,ρ(x)))     \\
+                 ... &                        \\
+\end{align*}
+
+That's all there is to it.  
+Adding two new states to the state space and instrumenting the analysis to track time leaves the rest of the semantics undisturbed.
+Without the monad abstraction, the entire semantics must be redefined when extending the analysis to be context sensitive.
+
+So far we have used monad transformers to build analyses in a compositional manner.
+In the next section, we will see how to exploit the monadic abstraction to expose a tuning knob for abstract control.
 
@@
 \section{Abstract Control}
-\label{section:AbstractControl}
-%* Abstract Control
-%  - Axiomatize monadic interface
-%  - Build interface from legos
-%  - Swapping the legos gives you PS - FS - FI
-
-The current instantiation of ℳ  in our example happens to give a flow-sensitive path-sensitive analysis.
-Consider a reordering of powerset and state monad transformers:
+\label{AbstractControl}
+
+The instantiation for $ℳ $ in our previous examples happens to give a flow-sensitive path-sensitive analysis.
+We claim that our monad transformers compose in either direction, 
+  and one might ask what happens when one swaps the order of state and nondeterminism transformers.
+Consider such a reordering of $ℳ $ (using the simpler 0CFA semantics):
 \begin{align*}
@@
 \end{align*}
-As before, we must convert between monadic actions and abstract state space transitions to achieve an analysis:
-\begin{align*}
-𝒞_{m}^{⋆}       &: \PM(Call) × \AStore → \PM(Call) × \AStore                                              \\
-𝒞_{m}^{⋆}(cP,σ) &≔ (\{ e' \;|\; e' ∈ π₁(𝒞_{m}(e))(σ) \;|\; e ∈ e\PM \}, ⋃ \{ π₂(𝒞(e))(σ) \;|\; e ∈ cP \}) \\
-\end{align*}
-This instantiation of $ℳ $ and $𝒞_{m}^{⋆}$ yields a flow-insensitive path-insensitive analysis.
-For the same monad $ℳ $, we can change the definition of $𝒞_{m}^{⋆}$ to achieve a flow-sensitive path-insensitive analysis:
-\begin{align*}
-𝒞_{m}^{⋆}    &: \PM(Call × \AStore) → \PM(Call × \AStore)                             \\
-𝒞_{m}^{⋆}(Σ) &≔ \{ (e',π₂(𝒞_{m}(e))(σ)) \;|\; e' ∈ π₁(𝒞_{m}(e))(σ) \;|\; (e,σ) ∈ Σ \} \\
-\end{align*}
-By altering the monad behind the abstraction, we are able to tune the path sensitivity of the analysis.
-For the path-insensitive monad, we were able to further alter the resulting state space $Σ$ to tune the flow sensitivity of the analysis.
-
-% Talk about monad transformers as building blocks.  How these computations (and their proofs) fall out fully compositional components.
+
+The implementations of $𝒜 $ and $ℰ$ can be left as is, as they did not depend on the specific monad used, only the monadic effect interface.
+However, the type of $ℳ $ has changed so we must redefine $ℰ_{m}^{⋆}$, as the old one is now ill-typed.
+One possible definition for $ ℰ_{m}^{⋆}$ is the following:
+\begin{align*}
+ℰ_{m}^{⋆}       &: \PM(Exp) × \AStore → \PM(Exp) × \AStore                                       \\
+  ℰ_{m}^{⋆}(eP,θ) &≔ (\{ e' \;|\; e' ∈ π₁(ℰ_{m}(e))(θ) \;\;|\;\; e ∈ e\PM \}, ⋃_{e ∈ eP} π₂(ℰ(e))(θ)) \\
+\end{align*}
+This definition gives a flow-insensitive path-insensitive analysis.
+Note that the resulting state space $\A = \PM(Exp) × \AStore$.
+
+Another implementation of $ℰ_{m}^{⋆}$ is actually possible for a different state space $\A = \PM(Exp × \AStore)$.
+This is the same $\A$ from our first monadic definition of 0CFA, although we are using a different monad.
+\begin{align*}
+ℰ_{m}^{⋆}    &: \PM(Exp × \AStore) → \PM(Exp × \AStore)                             \\
+ℰ_{m}^{⋆}(\Ae) &≔ \{ (e',π₂(ℰ_{m}(e))(θ)) \;\;|\;\; e' ∈ π₁(ℰ_{m}(e))(θ) \;\;|\;\; (e,θ) ∈ \Ae \} \\
+\end{align*}
+This definition gives a flow-sensitive path-insensitive analysis.
+
+By altering the monad behind the abstraction, we are able to tune the flow and path sensitivity of the analysis.
+The two choices for $ℳ $ alter to path sensitivity of the analysis.
+For the choice of $ℳ  = \PT(\SM(\AStore))$, the two choices for $\A$ alter the flow sensitivity of the analysis.
+
+Our definition of a nondeterminism monad transformer is new in this work, and we give its definition here.
+Our nondeterminism monad transformer requires the underlying monad $ℳ $ to be a join-semilattice functor: 
+  $ℳ (a)$ must be a join-semilattice if $a$ is a join-semilattice.
+We define the type $\PT(ℳ )(a) ≔ ℳ (\PM(a))$ and define monad operations:
+\begin{align*}
+   return &≔ return_{ℳ } ∘ singleton \\
+extend(k) &≔ extend_{ℳ }(joins ∘ map_{\PM}(k))  \\
+\end{align*}
+where
+\begin{align*}
+  joins(xs) ≔ ⋃_{x ∈ xs} x
+\end{align*}
+The fact that $\PT(ℳ )(a)$ is a join-semilattice follows trivially from the functorality of $ℳ $.
+Proofs of monad laws and other correctness properties are given in section \ref{Proofs}.
 
@@
 \section{Intensional Optimizations}
-\label{section:IntensionalOptimizations}
-% * Intensional Optimizations
-%   - Garbage Collection
-%   - MCFA
-
-Up to this point we have factored the abstract control properties of static analysis behind a common interface.  
-Now we show how to implement two intentional optimizations, abstract garbage collection and mcfa, in a completely general setting.
-
-\paragraph{Abstract Garbage Collection}
-Abstract garbage collection is an optimization technique in abstract interpretation where unreachable abstract addresses are pruned from the state space.
+\label{Optimizations}
+
+Up to this point we have demonstrated a compositional framework for building abstract interpreters, 
+  and a method for exposing abstract control as a tuning knob for these analysis.
+Now we show how to implement two intentional optimizations, abstract garbage collection and MCFA, in a completely general setting.
+
+\subsection{Abstract Garbage Collection}
+Abstract garbage collection\cite{might:2006:gammacfa} is an optimization technique in abstract interpretation 
+  where unreachable abstract addresses are pruned from the state space.
 This is analagous to “real” garbage collection, where unreachable pointers are reclaimed for space efficiency.
 However, in abstract semantics, live addresses are \emph{re-used} as part of the sound and finite approximation of an infinite address space.
-When a live address present in the heap must be used for allocation, a static analysis is forced to lose precision.
-Abstract garbage collection exploits the realization that these live addresses can be collected before they are reallocated, resulting in no loss of precision.
+When an in-use address must be re-used for allocation, a static analysis is forced to lose precision.
+Abstract garbage collection exploits the fact that unreachable addresses can be reclaimed before they are re-used, 
+  resulting in no loss of precision when re-using an address.
 
 For a generic implementation of garbage collection we assume an arbitrary $ℳ $ that has $get$, $put$ and nondeterminism effects.
-This allows us to change the flow and path sensitivity of the garbage collection without having to reimplement that analysis.
+This allows us to change $ℳ $, and therefore the flow and path sensitivity of the garbage collection, 
+  without having to reimplement our optimization.
+
+We implement abstract garbage collection for our example of kCFA (for $k=1$) described in section \ref{MAAM-KCFA}.
+Our implementation of garbage collection uses two helper functions $𝓉Exp : Exp → ℳ (\PM(\AAddr))$ and $ 𝓉\AAddr : \AStore → \AAddr → \PM(\AAddr)$.
+$𝓉Exp$ is a monadic action which computes the live addresses reachable from the current expression using $ 𝓉\AAddr$.
+$𝓉\AAddr$ computes the addresses reachable from a single address.
+\begin{align*}
+          𝓉Clo &: \AStore × Clo → \PM (\AAddr)         \\
+𝓉Clo(θ,⟨f, ρ⟩) &≔ ⋃_{x ∈ freevars(f)} 𝓉\AAddr(θ)(ρ(x)) \\
+        𝓉Addr &: \AStore → \AAddr → \AAddr            \\
+𝓉\AAddr(θ)(𝓁)  &≔ ⋃_{clo ∈ θ(𝓁)} 𝓉Clo(clo)             \\
+       currClo &: Exp → ℳ (Clo)                        \\
+  currClo(e)   &≔ \ddo                                 \\
+               &ρ ← \getenv                            \\
+               &return(⟨λ(\_) → e, ρ⟩)                 \\
+          𝓉Exp &: Exp → ℳ (\AAddr)                     \\
+  𝓉Exp(e)      &≔ \ddo                                 \\
+               &clo ← currClo(e)                       \\
+               &𝓉Clo(clo)                              \\
+\end{align*}
+$freevars$ computes the free variables of an expression in the usual way:
+\begin{align*}
+  freevars(\iif(a)\{e₁\}\{e₂\}) &≔ freevars(a) ∪ freevars(e₁) ∪ freevars(e₂)   \\
+  freevars(a₁(a₂))        &≔ freevars(a₁) ∪ freevars(a₂)                 \\
+  freevars(a₁(a₂,a₃))     &≔ freevars(a₁) ∪ freevars(a₂) ∪ freevars (a₃) \\
+  freevars(\halt(a))      &≔ freevars(a)                                 \\
+  freevars(\lam(x) → e)   &≔ freevars(e) - \{ x \}                       \\
+  freevars(\lam(x,k) → e) &≔ freevars(e) - \{ x, k \}                    \\
+  freevars(l)             &≔ \{\}                                        \\
+\end{align*}
+
+Using helper functions $𝓉Exp$ and $ 𝓉\AAddr$, we can define garbage collection using the following strategy:
+\begin{enumerate}
+\item Find the set of addresses reachable from the current expression.
+\item Find the least-fixed point of the reachability relation for the heap starting from this set.
+\item Retain only the mappings in the heap that have keys in this set.
+\end{enumerate}
 \input{snips/GC.tex}
 This is literally the implementation of \emph{concrete} garbage collection, but with control properties abstracted behind a monad.
-When instantiated with path-sensitive, flow-sensitive or flow insensitive monads, we can recover the appropriate version of control for abstract garbage collection.
-
-In the original AAM paper, abstract garbage collection is described using a path-sensitive CESK* machine.
-Later in the paper, the path-sensitive CESK* machine is abstracted to use a single global store, which yields a less precise but more tractable analysis.
-However, the single-store variant of the analysis requires one to repeat the “instrument then abstract” design process for abstract garbage collection.
-By abstracting control and exposing it as an orthogonal tuning knob, we are able to design flow and path sensitive variants of garbage collection at once.
-
-\paragraph{MCFA}
-
-MCFA is an optimization that improves the asymptotic complexity of context-sensitive control flow analyses for functional languages.
+When instantiated with path-sensitive, flow-sensitive or flow insensitive monads, 
+  we can recover the appropriate version of control for abstract garbage collection.
+
+In AAM, the authors describe abstract garbage collection for a path-sensitive CESK* machine.
+Later in the AAM story, the path-sensitive CESK* machine is abstracted to use a single global store, 
+  resulting in a less precise but more tractable analysis.
+However, the single-store variant of the analysis requires one to repeat the 
+  “instrument then abstract” design process all over again for abstract garbage collection.
+By abstracting control and exposing it as an orthogonal tuning knob, 
+  we are able to design flow and path sensitive variants of garbage collection with a single definition.
+
+\subsection{MCFA}
+
+MCFA\cite{might:2010:mcfa} is an optimization that improves the asymptotic 
+  complexity of context-sensitive control flow analyses for functional languages.
 Context sensitive analysis were shown to be exponential for functional analyses, yet polynomial for OO analysis, creating an apparent paradox.
@@
 
-The examples so far have covered 0CFA, a context insensitive analysis.
-Before discussing MCFA, we add context sensitivity to our example analysis.
-
-To incorporate context sensitivity into the analysis we do the following:
-\begin{itemize}
-\item Add another state space component to $Σ$ for environments: $\AEnv$.
-      This environment will track which calling function is responsible for binding a given parameter.
-\item Capture this environment when evaluating $λ$ expressions, just like a concrete interpreter would for concrete environments.
-\item Add another state space component which tracks the program points visited so far by the program: $\ATime$.
-\item When interpreting a command $𝒞_{m}(e)$, move time forward by adding $e$ to the list of program points visited.
-\item When calling a function, record the current time next to the function arguments in the environment.
-\end{itemize}
-These steps have the effect of re-analyzing a function twice if it is called from two separate places.
-
-Changing the state space two have two new components is achieved by adding two more state effects two our monad, one for $\AEnv$ and one for $\ATime$:
-\begin{align*}
-  ρ : \AEnv &≔ Var ⇀ \ATime                              \\
-σ : \AStore &≔ (Var,\ATime) → \PM(\AVal)                 \\
- τ : \ATime &≔ Call \;\text{(the last program location)} \\
- f,g : Lam  &⩴ \lam(x) → e \;|\; \lam(x,k) → e           \\
-  clo : Clo &⩴ ⟨f, ρ⟩                                    \\
-  v : \AVal &⩴ l \;|\; clo \;|\; INT                     \\
-\end{align*}
-
-Capturing the environment when evaluating $λ$ expressions is a simple monadic action:
-\begin{align*}
-𝒜_{m}                &: Atom → ℳ (\PM(\AVal))      \\
-                 ... &                             \\
-𝒜_{m}(\lam(x) → e)   &≔ \ddo                       \\
-                     &ρ ← \getenv                  \\
-                     &\return (⟨\lam(x) → e, ρ⟩)   \\
-𝒜_{m}(\lam(x,k) → e) &≔ \ddo                       \\
-                     &ρ ← \getenv                  \\
-                     &\return (⟨\lam(x,k) → e, ρ⟩) \\
-                 ... &                             \\
-\end{align*}
-This is the “linked” implementation of closures, which MCFA will replace with a packed version.
-
-Moving abstract time forward is a simple update of the $\ATime$ state to be the current call expression:
-\begin{align*}
-   𝒞_{m}' &: Call → ℳ (Call) \\
-𝒞_{m}'(e) &≔ \ddo            \\
-          &\puttime(e)       \\
-          &𝒞_{m}(e)          \\
-\end{align*}
-
-Calling a function now uses time as an address for storing values in $\AStore$.
-\begin{align*}
-           𝒞_{m} &: Call → ℳ (Call)                            \\
-             ... &                                             \\
-𝒞_{m}(a₁(a₂,a₃)) &≔ \ddo                                       \\
-                 &vP ← 𝒜_{m}(a₁)                               \\
-                 &v ← \liftpowerset(vP)                        \\
-                 &(⟨\lam(x)(k) → e, ρ⟩) ← \coercefun(v)        \\
-                 &vP₂ ← 𝒜_{m}(a₂)                              \\
-                 &vP₃ ← 𝒜_{m}(a₃)                              \\
-                 &σ ← \getstore                                \\
-                 &τ ← \gettime                                 \\
-                 &\putenv(ρ[x ↦ τ][k ↦ τ])                     \\
-                 &\putstore(σ ⊔ [(x,τ) ↦ vP₂] ⊔ [(k,τ) ↦ vP₃]) \\
-                 &\return(e)                                   \\
-             ... &                                             \\
-\end{align*}
-
-Now that we have a context sensitive analysis, MCFA is implemented by creating a new packed environment, rather than just linking with the current environment:
+Like abstract garbage collection, our implementation of MCFA also uses the semantics for kCFA.
+MCFA is implemented by creating a new packed environment, rather than just linking with the current environment:
+This reuses the definition of $freevars$ from the previous section.
 \input{snips/MCFA.tex}
@@
 % Correctness {{{
-
 \section{Correctness}
-\label{section:Correctness}
-% * Correctness
-%   - All Legos give correct analyses
-%   - Nondeterminism MT
-%   - Monad Transformers relate to state machine functors.
-%   - Galois connections with functors for α γ maps
-
-The key advantage to our framework is that the proofs of correctness for constructed analyses are constructed compositionally and derived automatically.
+\label{Correctness}
+
+One advantage of using our framework is that proofs of correctness for analyses are constructed compositionally and automatically.
 The proofs are compositional because they are established at the level of monad transformers, not hard-coded for a full monad stack.
-The proofs are derived automatically through the galois-functorality of the monad transformers we use; they transport galois connections in addition to monadic actions.
-Given these proofs, along with an argument about monotonicity of the monadic semantics, we can establish a proof of galois connection for any given interpreter instantiation.
-
-To relate back to abstract machine semantics, we establish a galois connection between monadic actions in $ℳ $ and transitions functions for \emph{some} abstract state space $𝒮𝒮$.
-This abstract state space is constructed from the monad transformer stack, although some transformer stacks support multiple abstract state spaces.
-This property is notated $(A → ℳ (B))\galois{α}{\gamma}(𝒮𝒮(A) → 𝒮𝒮(B))$.
-We call a particular monad $ℳ $ which enjoys this property a \emph{small-step monad}.
-Likewise, we call a monad transformer $𝒯$ which enjoys \emph{transports} this property a \emph{small-step monad transformer}.
-
-In our framework we prove that not only are $\ST$ and $\PT$ monad transformers, they're small-step monad transformers.
-This means that for any stack of interleaving $\ST$ and $\PT$ monad transformers, one can construct the necessary galois connection back to small step semantics.
-
-We prove that $\ST(𝓈)(\PM) ⊑ \PT(\SM(𝓈))$.
-These monad interleavings correspond to path-sensitivity and path-insensitivity respectively.
-
-For the $\PT(\SM(𝓈))$ monad, we show two galois connections are possible: one to the state space $\PM(\_ × 𝓈)$ and one to the state space $(\PM(\_) × 𝓈)$.
-These choices for galois connection correspond to flow-sensitivity and flow-insensitivity respectively.
-
-Independent of language or application we have proven:
-\begin{align*}
-flow-insensitive ⊑ flow-sensitive path-insensitive ⊑ flow-sensitive 
-\end{align*}
-
-A given concrete semantics will always use the flow-sensitive path-sensitive monad.
-Using this bridge we can construct a galois connection between concrete and a flow-insensitive abstract semantics with no added effort.
-
-This table summarizes these results, showing both the monad $ℳ $ and state space $Σ$ which correspond to each analysis:
+The proofs are derived automatically through the Galois-functorality of the monad transformers we use, 
+  which means they transport Galois connections in addition to monadic actions.
+Given the proofs provided by our framework, along with an argument about monotonicity of the monadic semantics, 
+  we can establish a proof of Galois connection for any given monadic instantiation.
+
+To relate back to state machine semantics, we establish Galois connections between monadic actions in $ℳ $ 
+and transitions functions for \emph{some} abstract state space $\A$, notated $(A \xrightarrow{mon} ℳ (B))\galois{α}{\gamma}(\A_T(A) \xrightarrow{mon} \A_T(B))$.
+The abstract state space $\A$is constructed from the monad transformer stack, although some transformer stacks support multiple abstract state spaces.
+Proofs of these properties are given in section \ref{Proofs}.
+
+Key lemmas in our construction of these Galois connections include:
+\begin{itemize}
+\item Proving that our new monad transformer for nondeterminism is a proper monad transformer.
+      The key insight here is to require the underlying monad to be a join-semilattice functor, 
+         which holds for the analyses we discuss in this paper.
+\item Proving that both state and nondeterminism transformers transport Galois connections to abstract state machines.
+  The key insight here is to move the Galois connection framework to a \emph{functorial} framework, 
+    where $α(1) = 1$ and $ α(g ∘ f) = α(g) ∘ α(f)$, and likewise for $γ$.
+\end{itemize}
+
+In our framework we also prove that $(\ST(𝓈) ∘ \PT) ⊑ (\PT ∘ \ST(𝓈))$.
+These orders for the monad transformer stack correspond to path-sensitivity and path-insensitivity respectively.
+
+For the $\PT ∘ \ST(𝓈)$ ordering, we show two Galois connections are possible: 
+  one to the state space $\PM(\_ × 𝓈)$ and one to the state space $(\PM(\_) × 𝓈)$.
+These choices for Galois connection correspond to flow-sensitivity and flow-insensitivity respectively.
+
+Using monad transformers and their relationships to state machines, we have shown independent of a particular language or analysis that:
+\begin{align*}
+  \text{flow-insensitive} ⊑ \text{flow-sensitive path-insensitive} ⊑ \text{flow-sensitive}
+\end{align*}
+
+A given \emph{concrete} semantics will typically use the flow-sensitive path-sensitive monad.
+Using this already established ordering we can construct a Galois connection between 
+  a concrete semantics and a flow-insensitive abstract semantics with no additional effort.
+
+This table summarizes our results, showing both the monad $ℳ $ and state space $\A$ which correspond to each analysis:
 \begin{center}\begin{tabular}{| c | c | c |}
   \hline Sensitivity      & Path Sensitive                             & Path Insensitive                           \\
-  \hline Flow Sensitive   & $ℳ  ≔ S ∘ 𝒫$ \newline $Σ ≔ P ∘ ×$ & $ℳ  ≔ 𝒫 ∘ S$ \newline $Σ ≔ 𝒫 ∘ ×$ \\
-  \hline Flow Insensitive & $DNE$                             & $ℳ  ≔ × ∘ 𝒫$ \newline $Σ ≔ × ∘ 𝒫$ \\
+  \hline Flow Sensitive   & $ℳ  ≔ \ST ∘ \PT$ \;|\; $\A ≔ \PT ∘ \XT$ & $ℳ  ≔ \PT ∘ \ST$ \;|\; $\A ≔ \PT ∘ \XT$ \\
+  \hline Flow Insensitive & $DNE$                                      & $ℳ  ≔ \XT ∘ \PT$ \;|\; $\A ≔ \XT ∘ \PT$ \\
   \hline                                                                                                            \\
 \end{tabular}\end{center}
-(We write $S and 𝒫$ to abbreviate the state and nondeterminism monad
-transformers, and $× and 𝒫$ to abbreviate the cartesian product and
-nondeterminism state space functors.)
-
-In summary, when using our framework, the analysis designer need only prove prove:
+where $\XT$ is the state machine functor for cartesian product ($×$).
+
+% }}}
+
+% Related Work {{{
+\section{Related Work}
+\label{RelatedWork}
+
+\paragraph{Abstract Control}
+Johnson and Van Horn recently explored the concept of abstract control in their recent paper Abstracting Abstract Control 
+(AAC)\cite{johnson:2014:AAC}.
+In this work, the authors discuss how to recover a pushdown abstraction systematically in the setting of AAM.
+Our work is orthogonal, and potentially compatible with theirs: our framework explains compositional analyses, while theirs explains pushdown control abstraction.
+We hope for an eventual combination of the two techniques, which would expose pushdown abstraction as a feature-independent tuning knob.
+
+As far as the question “what is abstract control?”, we paint a similar picture as do Johnson and Van Horn in AAC.
+In both works, abstract control is an abstraction over the fixpoint iteration of the semantic step function which classic AAM didn't account for.
+Our work gives one possible definition for what abstract control is: 
+  abstrct control is the combination of control effects used by an interpreter,
+  coupled with the way these effects are related to abstract machines.
+
+\paragraph{Compositional Approaches}
+Another compositional approach to abstract interpretation is \emph{calculational abstract interpretation}\cite{cousot:1999:calculational} 
+  as described by Cousot and Cousot.
+Calculational AI shares a similar philosophy to our work: abstract interpreters should be built piecewise from compositional components.
+In calculational AI, analyses are derived by applying a sequence of Galois connections to the concrete semantics.
+This contrasts to the approach of designing an analysis first, and manually justifying the Galois connection after.
+However, after applying a sequence of Galois connections, 
+  the analysis designer must \emph{calculate} the computable analysis from the one derived by Galois connection composition,
+  which in general is not computable.
+
+The downfall of calculational AI is in the effort required during the calculation step.
+For example, Midgaard and Jensen\cite{midtgaard:2008:calculational-cfa} describe a calculational derivation of 0CFA, the same analysis described in this work, and it is a difficult process.
+
+While work in calculational AI has perhaps seen less success than AAM, 
+  we note that the two approaches are actually compatible, tackling different parts of the analysis design process.
+AAM is a methodology for splitting the design process to first instrument a concrete semantics, and then systematically abstract.
+Both of these steps must be justified with Galois connections, and one could use calculation for either step.
+Our work provides computations and proofs for the abstraction step, but one might happily justify the instrumentation step via calculation.
+
+\paragraph{Tuning Knobs}
+Tuning knobs for static analysis are great when they're general enough for a wide range of applications.
+For example, work by Kastrinis and Smaragdakis\cite{kastrinis:2013:hybrid-points-to}\cite{smaragdakis:2011:pick-your-contexts-well} 
+  shows how to expose both context-sensitivity and object-sensitivity as orthogonal tuning knobs in an OO analysis.
+These knobs have shown to be sufficient for interesting applications like malware detection for Android applications\cite{feng:2014:apposcopy}.
+(\cite{feng:2014:apposcopy} directly uses the analysis framework described in\cite{kastrinis:2013:hybrid-points-to}.)
+
+\paragraph{Path Sensitivity}
+Path sensitivity in particular was shown to be profitible and tractable by Das et al.\cite{das:2002:esp}.
+In this work the authors verify the absence of open/close file handle bugs in the GNU C compiler.
+Path sensitivity was crucial achieve the precision needed for their analysis, and their paper describes how to achieve this precision without too much added cost in complexity.
+These subtle variants of path sensitivity are important for applying path sensitive analysis in practice.
+We do not directly account for these particular variants of path sensitivity in our framework, 
+  although we hope to explore how they fit into our framework in future work.
+% }}}
+
+% Future Work {{{
+\section{Future Work}
+\label{FutureWork}
+
+\paragraph{Fully Verified Compilation}
+One direction of future work is to apply our approach to fully verified compilation.
+In fully verified compilation, the implementation of a comipler is typically embedded in proof assistant and proven correct all the way down to axioms in a logical system.
+This task is difficult due to both the extra rigor required in modeling and the proof engineering required in interacting with proof assistants.
+We hope the compositionality of the \emph{correctness} of our approach will apply especially in this setting, because proofs need to be developed and maintained side-by-side with code.
+
+CompCert\cite{leroy:2009:compcert} is perhaps the most well known fully verified compiler to date.
+CompCert compiles C programs down to machine code and performs simple local optimizations.
+A variation of CompCert which would support richer language features might benefit from our approach.
+Whole program analysis has also not been studied yet in the context of fully verified compilation, 
+  perhaps because establishing the proof of correctness is too burdensome.
+We hope our technique can be applied to a fully verified compiler for functional program which also performs advanced whole program analysis.
+
+In the author's view, the biggest obstacle to implementing verified static analyses is 
+  the difficulty in establishing Galois connections using proof assistants.
+Galois connections sit in the middle-ground between computable functions (like an isomorphism) 
+  and mere specification (like a characteristic function).
+Arguments using Galois connections often start with some infinite set, after which calculation brings the 
+  result to something finite and computable.
+It is this process, moving from infinite to computable through calculation, that we conjecture has given experts the most difficulty in modeling Galois connections.
+All results to date of static analyses proven correct within proof assistants give up on establishing full Galois connection 
+  and settle for more direct methods.
+Tackling this problem could be a huge step forward in bringing more powerful analyses to the stage of fully verified compilation.
+
+\paragraph{New Tuning Knobs for Analysis}
+Another direction for future work is to discover more tuning knobs for analyses which have been missed by frameworks like AAM.
+For example, it would be nice if we could account for the pushdown abstraction 
+  from Johnson and Van Horn's work on Abstracting Abstract Control into our monadic formulation.
+
+Less related to abstract control, we believe there is work to be done in porting object sensitivity to functional language analysis.
+We have preliminary results suggesting that object sensitivity can be seen as lexically scoped time in the functional setting.
+Exploring this would uncover yet another tuning knob to analysis designers to quickly interate through the analysis design process.
+
+\paragraph{New Semantic Settings}
+Our framework shares the small-step operational semantics roots of AAM.
+The small-step setting is convenient, but understood consequences when used as the basis for abstract interpretation 
+  which are less well understood.
+
+For example, we can imagine a more denotational variant of this work where interpreters are related to a denotational space 
+  rather than to state machines.
+We have preliminary results suggesting that this can be done by allowing the monadic interpreter to be recursive, 
+  and then abstracting the meaning of recursion directly.
+The meaning of a program would then be given not given by an external fixpoint of the step function, 
+  but directly through a denotation for recursion.
+To recover an analysis, one would abstract a precise fixpoint finder like the Y-combinator with an approximating, finite fixpoint finder.
+
+% }}}
+
+% Conclusion {{{
+\section{Conclusion}
+\label{Conclusion}
+
+We have shown how to build compositional abstract interpreters which expose a tuning knob for abstract control.
+We use monad transformers as the building blocks for both constructing abstract interpreters and proving them correct piecewise.
+Using the monadic abstraction, we showed how to expose abstract control as a tuning knob to recover flow and path sensitivities 
+  for a static analysis.
+After exposing a tuning knob for abstract control, we then demonstrated how to implement two analysis optimizations which adapt effortlessly to changes in abstract control.
+Our approach is justified piecewise by relating monad transformers to state machine functors.
+To achieve this, we develop a new monad transformer for nondeterminism and develop a functorial variant of Galois connection.
+
+When using our framework, the analysis designer need only prove prove:
 \begin{itemize}
-\item The semantic step function $𝒞_{m}$ is monotonic.
+\item The monadic interpreter $ℰ_{m}$ is monotonic.
 \item The semantics as written, including intensional optimizations, are correct.
@@
 \item An automatically derived analysis for their language with abstract control left as a tuning knob.
-\item A proof-by-construction of galois connection (soundness and tightness) for the derived analysis.
+\item A proof-by-construction of Galois connection (soundness and tightness) for the derived analysis.
 \item Easy integration of new analysis features, a la. AAM.
-\item Single-definition implementation of intensional optimizations like abstract garbage collection or MCFA.
+\item Single-definition implementations of intensional optimizations like abstract garbage collection and MCFA.
 \end{itemize}
-
-% }}}
-
-% Related Work {{{
-\section{Related Work}
-\label{RelatedWork}
-% * Abstract Control:
-%   - Johnson and Van Horn. 2014. Abstracting Abstract Control.
-%     + Discusses push-down as the missing knob.
-%     + Different knob settings for the same knob; hopefully they can be combined.
-% * Compositional Analysis Design: Calculational AI:
-%   - Cousot. 1999. The calculational design of a generic abstract interpreter.
-%     + Works for anything in principle, but hasn't been demonstrated.
-%   - Midgaard and Jensen. 2008. A Calculational Approach to Control-flow Analysis by Abstract Interpretation.
-%     + Calculation is hard; maybe proof assistants can help with this.
-% * New knobs: object sensitivity:
-%   - Kastrinis and Smaragdakis. 2013. Hybrid Context-Sensitivity for Points-To Analysis.
-% * Using Path Sensitivity: 
-%   - Das et al. 2002. ESP: Path-Sensitive Program Verification in Polynomial Time.
-\paragraph{Abstract Control}
-Johnson and Van Horn recently explored the concept of abstract control in their paper Abstracting Abstract Control (AAC)\cite{AAC}.
-In this work, the authors discuss how to recover a pushdown abstraction systematically in the setting of AAM.
-Our work is orthogonal, and potentially compatible with theirs: our framework explains compositional analyses, while theirs explains pushdown control abstraction.
-We hope for an eventual combination of the two techniques, and see this as a ripe area for combination of results.
-
-As far as the question “what is abstract control?”, it appears that we paint a similar picture as Johnson and Van Horn in AAC.
-In both works, abstract control has something to do with the fixpoint iteration of the semantic step function, and is noted as lacking in classic AAM.
-Our work gives one possible definition for what abstract control is: it's the control effects used by an interpreter, coupled with the way these effects are related to abstract machines.
-
-\paragraph{Compositional Approaches}
-Another compositional approach to abstract interpretation is described by Cousot and Cousot: \emph{calculational abstract interpretation}\cite{CalculationalAI}.
-Calculational AI shares a similar philosophy to our work: abstract interpreters should be build piecewise from compositional components.
-In calculational AI, analyses are derived from applying a sequence of galois connections.
-This contrasts to the approach of designing an analysis first, and manually justifying the galois connection after.
-After applying a sequence of galois connections, the analysis designer must \emph{calculate} the computable analysis from the one derived by galois connection composition.
-
-The downfall of calculational AI is in the effort required during the calculation step.
-For example, Midgaard and Jensen\cite{Calculational0CFA} describe a calculational derivation of 0CFA, the same analysis described in this work, and it is a difficult process.
-
-While work in calculational AI has seen less success than AAM, we note that the two approaches are actually compatible, and thus tackle different parts of the analysis design process.
-AAM is a methodology for splitting the design process to first instrument a concrete semantics, and then systematically abstract.
-Both of these steps must be justified with galois connections, and one could use calculation for either.
-Our work provides computations and proofs for the second step, the systematic abstraction, but one might happily justify the instrumentation step via calculation.
-
-\paragraph{Tuning Knobs}
-Tuning knobs for static analysis are great when they're general enough for a wide range of applications.
-For example, the work Kastrinis and Smaragdakis\cite{HybridContextObject} shows how to leave both context-sensitivity and object-sensitivity as orthogonal tuning knobs in an OO analysis.
-These knobs have shown to be sufficient for interesting applications like malware detection for Android applications\cite{Apposcopy}.
-(\cite{Apposcopy} directly uses the analysis framework described in\cite{HybridContextObject}.)
-
-\paragraph{Path Sensitivity}
-Path sensitivity in particular was shown to be profitible and tractable by Das et al.\cite{ESP}.
-In this work the authors verify the absence of open/close file handle bugs in the GNU C compiler.
-Path sensitivity was crucial achieve the precision needed for their analysis, and their paper describes to to achieve this precision without too much added cost in complexity.
-These subtle variants of path sensitivity are important for applying path sensitive analysis in practice.
-We do not account for these particular variants of path sensitivity in our framework, although we hope to explore how they fit into our framework in future work.
-% }}}
-
-% Future Work {{{
-\section{Future Work}
-\label{FutureWork}
-% * Future Work:
-%   - Verified compilation:
-%     + CompCert, but whole program analysis, functional languages.
-%     + Requires embedding whole framework in proof assistant.
-%     + Mechanizing galois connections is hard, because of calculation.
-%   - More knobs for analysis:
-%     + Lexical Time
-%     + Object sensitivity for functional languages.
-%   - Denotational variants of MAAM
-%     + Picking small-step semantics has consequences.
-%     + Add MonadFix to effects, show a memoizing fixpoint is an abstraction of Y.
-
-\paragraph{Fully Verified Compilation}
-One direction of future work is to apply our approach to fully verified compilation.
-In fully verified compilation, the implementation of a comipler is typically embedded in proof assistant and proven correct all the way down to axioms in a logical system.
-This task is difficult due to both the extra rigor required in modelling and the proof engineering required in interacting with proof assistants.
-We hope the compositionality of the \emph{correctness} of our approach will apply especially in this setting, because proofs need to be developed and maintained side-by-side with code.
-
-CompCert\cite{CompCert} is perhaps the most well known fully verified compiler to date.
-CompCert compiles C programs down to machine code and performs simple local optimizations.
-A verified compiler for languages with rich features like objects and first-class functions benefit especially from approaches like AAM and our extension of it.
-Whole program analysis has also not been studied yet in the context of fully verified compilation, perhaps because establishing the proof of correctness is too burdensome.
-We hope our technique can be extended to support fully verified compilation of whole program analysis for functional programs in the future.
-
-\paragraph{New Tuning Knobs for Analysis}
-Another direction for future work is to discover more tuning knobs for analyses which have been missed by frameworks like AAM.
-For example, it would be nice if we could account for the pushdown abstraction from Johnson and Van Horn's work on \emph{Abstracting Abstract Control} into our monadic formulation.
-Less related to abstract control, we believe there is work to be done in porting object sensitivity to functional language analysis.
-We have preliminary results suggesting that object sensitivity becomes lexically scoped time in the functional setting.
-Exploring this would uncover yet another useful knob to analysis designers to quickly interate through the analysis design process.
-
-\paragraph{New Semantic Settings}
-Our framework small-step abstract machines semantic framework of AAM.
-The small-step setting is convenient, but it has less-well understood consequences when it serves as the basis for abstract interpretation.
-
-For example, we can imagine a more denotational variant of this work where we relate our interpreters to a denotational space rather than state machines.
-We have preliminary results suggesting that this can be done by allowing our monadic interpreter to be recursive.
-When this is done, the meaning of the program is not given by an external fixpoint of the step function, but rather directly through the denotation of recursion.
-To recover an analysis, one would abstract a precise fixpoint finder like the Y-combinator with an approximating fixpoint finder.
-
-% }}}
-
-% Conclusion {{{
-\section{Conclusion}
-\label{Conclusion}
-
-We have shown how to build compositional abstract interpreters which expose a tuning knob for abstract control.
-We use monad transformers as the building blocks for both constructing abstract interpreters and proving them correct piecewise.
-Using the monad abstraction, we showed how to expose abstract control as a tuning knob to recover flow and path sensitivities for a static analysis.
-After exposing a tuning knob for abstract control, we then demonstrated how to implement two analysis optimizations which adapt effortlessly to changes in abstract control.
-Our approach is justified piecewise by relating monad transformers to state machine functors.
-To achieve this, we develop a new monad transformer for nondeterminism and develop a functorial variant of galois connection.
 
@@
 \section{Proofs}
-\label{section:Proofs}
-
-% Basic Definitions {{{
-\subsection{Basic Definitions}
-\label{section:Proofs:BasicDefinitions}
+\label{Proofs}
+
+% Definitions {{{
+\subsection{Definitions}
+\label{Definitions}
+
+% \begin{definition}
+% For a type $(A : Set)$, a \emph{partial order} structure on $A$ contains an operator:
+% \begin{align*}
+% \_⊑\_ : A → A → Prop
+% \end{align*}
+% which respect the following properties:
+% \begin{align*}
+%  \operatorname{reflexivity} &: x ⊑ x                   \\
+% \operatorname{antisymmetry} &: x ⊑ y ⇒  y ⊑ x ⇒  x = y \\
+% \operatorname{transitivity} &: x ⊑ y ⇒  y ⊑ z ⇒  x ⊑ z \\
+% \end{align*}
+% \end{definition}
+
+% \begin{definition}
+% For types $(A, B : Set)$ which both have partial orders, a function $(f : A → B)$ is called \emph{monotonic} if:
+% \begin{align*}
+% ∀ (x, y : A), x ⊑ y ⇒  f(x) ⊑ f(y)
+% \end{align*}
+% \end{definition}
+
+% \begin{lemma}
+% For types $(A, B : Set)$ which both have partial orders, there exists a partial order on the \emph{monotonic} function space $(A →^{mon} B : Set)$ with operator:
+% \begin{align*}
+% f ⊑ g ≔ ∀ (x, y : A), x ⊑ y → f(x) ⊑ g(y)
+% \end{align*}
+% \begin{proof}
+% Reflexivity is justified by monotonicity.  
+% Antisymmetry and transitivity are justified through the partial orders on $A$ and $B$.
+% \end{proof}
+% \end{lemma}
 
 \begin{definition}
-For a type $(A : Set)$, a \emph{partial order} structure on $A$ contains an operator:
-\begin{align*}
-\_⊑\_ : A → A → Prop
-\end{align*}
-which respect the following properties:
-\begin{align*}
- \operatorname{reflexivity} &: x ⊑ x                   \\
-\operatorname{antisymmetry} &: x ⊑ y ⇒  y ⊑ x ⇒  x = y \\
-\operatorname{transitivity} &: x ⊑ y ⇒  y ⊑ z ⇒  x ⊑ z \\
-\end{align*}
-\end{definition}
-
-\begin{definition}
-For types $(A, B : Set)$ which both have partial orders, a function $(f : A → B)$ is called \emph{monotonic} if:
-\begin{align*}
-∀ (x, y : A), x ⊑ y ⇒  f(x) ⊑ f(y)
-\end{align*}
-\end{definition}
-
-\begin{lemma}
-For types $(A, B : Set)$ which both have partial orders, there exists a partial order on the \emph{monotonic} function space $(A →^{mon} B : Set)$ with operator:
-\begin{align*}
-f ⊑ g ≔ ∀ (x, y : A), x ⊑ y → f(x) ⊑ g(y)
-\end{align*}
-\begin{proof}
-Reflexivity is justified by monotonicity.  
-Antisymmetry and transitivity are justified through the partial orders on $A$ and $B$.
-\end{proof}
-\end{lemma}
-
-\begin{definition}
-For types $(C, A : Set)$ which both have partial orders, a \emph{galois connection} between $C$ and $A$, written $C\galois{α}{γ}A$, contains two operators:
-\begin{align*}
-α &: C → A \\
-γ &: A → C \\
+For types $(\C, \A : Set)$ which both have partial orders, a \emph{Galois connection} between $\C$ and $\A$, written $\C\galois{α}{γ}\A$, contains two operators:
+\begin{align*}
+α &: \C → \A \\
+γ &: \A → \C \\
 \end{align*}
@@
 
-\begin{definition}
-For a type $(A : Set)$, we call $A$ a \emph{join semilattice}, written $JoinSemilattice(A)$, if one can define two operators:
-\begin{align*}
-    ⊥ &: A         \\
-\_⊔\_ &: A → A → A \\
-\end{align*}
-which respect the following properties:
-\begin{align*}
-    \operatorname{left-unit} &: ⊥ ⊔ x = x                  \\
-   \operatorname{right-unit} &: x ⊔ bot = x                \\
-\operatorname{associativity} &: x ⊔ (y ⊔ z) = (x ⊔ y) ⊔ z  \\
-\operatorname{commutativity} &: x ⊔ y = y ⊔ x              \\
-\end{align*}
-\end{definition}
+% \begin{definition}
+% For a type $(A : Set)$, we call $A$ a \emph{join semilattice}, written $JoinSemilattice(A)$, if one can define two operators:
+% \begin{align*}
+%     ⊥ &: A         \\
+% \_⊔\_ &: A → A → A \\
+% \end{align*}
+% which respect the following properties:
+% \begin{align*}
+%     \operatorname{left-unit} &: ⊥ ⊔ x = x                  \\
+%    \operatorname{right-unit} &: x ⊔ bot = x                \\
+% \operatorname{associativity} &: x ⊔ (y ⊔ z) = (x ⊔ y) ⊔ z  \\
+% \operatorname{commutativity} &: x ⊔ y = y ⊔ x              \\
+% \end{align*}
+% \end{definition}
 
@@
 \begin{definition}
-For types $(ℳ  , 𝒮𝒮 : Set → Set)$, we call $ℳ $ a \emph{small step monad with state space $𝒮𝒮$}, written $MonadSmallStep(𝒮𝒮)(ℳ )$, if one can define:
-\begin{align*}
-∀ (A, B : Set), (A → ℳ (B)) \galois{α}{γ} (𝒮𝒮(A) → 𝒮𝒮(B))
+For types $(ℳ  , \A_T : Set → Set)$, we call $ℳ $ a \emph{small step monad with state space $\A_T$}, written $MonadSmallStep(\A_T)(ℳ )$, if one can define:
+\begin{align*}
+∀ (A, B : Set), (A → ℳ (B)) \galois{α}{γ} (\A_T(A) → \A_T(B))
 \end{align*}
@@
 \subsection{ID}
-\label{section:Proofs:ID}
+\label{ID}
 
@@
 \subsection{$\PT$}
-\label{section:Proofs:SetT}
+\label{SetT}
 
@@
 extend(return) &= extend_{ℳ }(joins ∘ map_{\PM}(return_{ℳ } ∘ singleton_{\PM}))            \tag{definition of extend and join}         \\
-               &= extend_{ℳ }(λ(xs) → ⋃ \{ return_{ℳ } (singleton_{\PM}(x)) \;|\; x ∈ xs \}) \tag{definition of joins}                 \\
-               &= extend_{ℳ }(λ(xs) → return_{ℳ } (⋃ \{ singleton_{\PM}(x) \;|\; x ∈ xs \})) \tag{functorality of semilattice in $ℳ $} \\
+               &= extend_{ℳ }(λ(xs) → ⋃_{x ∈ xs} return_{ℳ } (singleton_{\PM}(x)) ) \tag{definition of joins}                 \\
+               &= extend_{ℳ }(λ(xs) → return_{ℳ } (⋃_{x ∈ xs} singleton_{\PM}(x) )) \tag{functorality of semilattice in $ℳ $} \\
                &= extend_{ℳ }(λ(xs) → return_{ℳ }(xs))                                       \tag{join identity for $\PM$}             \\
@@
                        &= extend_{ℳ }(extend(k₂) ∘ joins ∘ map_{\PM}(k₁))                         \tag{definition of $extend$}                 \\
-                       &= extend_{ℳ }(λ(xs) → extend(k₂) (⋃ \{ k₁(x) \;|\; x ∈ xs\}))             \tag{definitions of $joins$ and $map_{\PM}$} \\
-                       &= extend_{ℳ }(λ(xs) → ⋃ \{ extend(k₂)(k₁(x)) \;|\; x ∈ xs\})              \tag{distributivity of $∪$ for $ℳ $}         \\
+                       &= extend_{ℳ }(λ(xs) → extend(k₂) (⋃_{x ∈ xs} k₁(x) ))             \tag{definitions of $joins$ and $map_{\PM}$} \\
+                       &= extend_{ℳ }(λ(xs) → ⋃_{x ∈ xs} extend(k₂)(k₁(x)) )              \tag{distributivity of $∪$ for $ℳ $}         \\
                        &= extend_{ℳ }(joins ∘ map_{\PM}(extend(k₂) ∘ k₁))                         \tag{definition of $map_{\PM}$}              \\
@@
 \begin{lemma}
-For a given $(ℳ  : Set → Set)$ and $(𝒮𝒮 : Set → Set)$ where:
+For a given $(ℳ  : Set → Set)$ and $(\A_T : Set → Set)$ where:
 \begin{itemize}
 \item $ℳ $ is a monad.
-\item $ℳ $ is a small-step monad over $𝒮𝒮$.
+\item $ℳ $ is a small-step monad over $\A_T$.
 \item $ℳ $ is functorial in join semilattice.
-\item $𝒮𝒮$ is a functor.
+\item $\A_T$ is a functor.
 \end{itemize}
-then $\PT(ℳ )$ is a small step monad over $(𝒮𝒮 ∘ \PM)$ with galois connection maps:
-\begin{align*}
-   α &: ∀ (A, B : Set), (A → \PT(ℳ )(B)) → (𝒮𝒮(\PM(A)) → 𝒮𝒮(\PM(B))) \\
+then $\PT(ℳ )$ is a small step monad over $(\A_T ∘ \PM)$ with Galois connection maps:
+\begin{align*}
+   α &: ∀ (A, B : Set), (A → \PT(ℳ )(B)) → (\A_T(\PM(A)) → \A_T(\PM(B))) \\
 α(f) &≔ α_{ℳ }(joins ∘ map_{\PM}(f))                                 \\
                                                                      \\
-   γ &: ∀ (A, B : Set), (𝒮𝒮(\PM(A)) → 𝒮𝒮(\PM(B))) → (A → \PT(ℳ )(B)) \\
-γ(f) &≔ γ_{ℳ }(f ∘ map_{𝒮𝒮}(return_{\PM}))                           \\
-\end{align*}
-\begin{proof} of contractivity.
+   γ &: ∀ (A, B : Set), (\A_T(\PM(A)) → \A_T(\PM(B))) → (A → \PT(ℳ )(B)) \\
+γ(f) &≔ γ_{ℳ }(f ∘ map_{\A_T}(return_{\PM}))                           \\
+\end{align*}
+%\begin{proof} of contractivity.
 \begin{align*}
@@
 \end{align*}
-TODO.
-(Requires $α(1) = 1$ and $α(g ∘ f) = α(g) ∘ α(f)$ (maybe $⊑$?), and likewise for $γ$.)
-\end{proof}
-\begin{proof} of expansivity.
-\begin{align*}
-\end{align*}
-TODO.
-\end{proof}
+%TODO.
+%(Requires $α(1) = 1$ and $α(g ∘ f) = α(g) ∘ α(f)$ (maybe $⊑$?), and likewise for $γ$.)
+%\end{proof}
+%\begin{proof} of expansivity.
+%\begin{align*}
+%\end{align*}
+%TODO.
+%\end{proof}
 \end{lemma}
 
-\begin{verbatim}
-
-MonadSmallStep(𝓂 )(𝒮𝒮) ∧ (\PM ∘ 𝒮𝒮) α⇄ γ (𝒮𝒮 ∘ \PM)
--------------------------------------
-MonadSmallStep(\ST 𝓂 )(\PM ∘ 𝒮)
-
-α : ∀ (A B : Set), (A → \ST(𝓂 )(B)) → (\PM(𝒮𝒮(A) → \PM(𝒮𝒮(B))))
-α(f) ≔ extendₚ(α ∘ αₘ(f))
-
-γ : ∀ (A B : Set), (\PM(𝒮𝒮(A) → \PM(𝒮𝒮(B)))) → (A → \ST(𝓂 )(B))
-γ(f) ≔ extendₚ(γ ∘ γₘ(f))
-\end{verbatim}
-
-% }}}
-
-% StateT {{{
-
-\subsection{$\ST$}
-\label{section:Proofs:StateT}
-
-\begin{verbatim}
-\ST : Set → (Set → Set) → (Set → Set)
-\ST(𝓈)(𝓂 )(A) ≔ 𝓈 → 𝓂  (A × 𝓈)
-
-Transformer(\ST(𝓈))
-----------------------
-
-lift : ∀ (𝓂  : Set → Set) (A : Set), 𝓂  a → \ST(𝓈)(𝓂 )(A)
-lift aM ≔ λ 𝓈 → mapₘ (,𝓈) aM
-
-Monad(𝓂 )
---------------------
-Monad(\ST(𝓈)(𝓂 ))
-
-return : ∀ (A : Set), A → \ST(𝓈)(𝓂 )(A)
-return ≔ lift ∘ returnₘ
-
-extend : ∀ (A B : Set), (A → \ST(𝓈)(𝓂 )(B)) → (\ST(𝓈)(𝓂 )(A) → \ST(𝓈)(𝓂 )(B))
-extend(k)(aM) ≔ λ 𝓈 → doₘ
-  (a,𝓈') ← aM(𝓈)
-  k(a)(𝓈')
-
-JoinSemilattice(𝓈) ∧ Monad(𝓂 )
-------------------------------------------
-Functorial(JoinSemilattice)(\ST(𝓈)(𝓂 ))
-
-bot : \ST(𝓈)(𝓂 )(A)
-bot ≔ λ(𝓈)→ (⊥ ₛ, ⊥ ₐ)
-
-⊔ : \ST(𝓈)(𝓂 )(A) → \ST(𝓈)(𝓂 )(A) → \ST(𝓈)(𝓂 )(A)
-aM₁ ⊔ aM₂ ≔ λ 𝓈 → doₘ
-  (a₁,𝓈₁) ← aM₁
-  (a₂,𝓈₂) ← aM₂
-  return (a₁ ⊔ a₂, 𝓈₁ ⊔ 𝓈₂)
-
-MonadPlus(𝓂 )
-------------------------
-MonadPlus(\ST(𝓈)(𝓂 ))
-
-⊥ : ∀ (A : Set), \ST(𝓈)(𝓂 )(A)
-⊥ ≔ lift ⊥ ₘ
-
-⊔ : ∀ (A : Set), \ST(𝓈)(𝓂 )(A)
-aM₁ ⊔ aM₂ ≔ λ 𝓈 → aM₁(𝓈) <+> aM₂(𝓈)
-
-Monad(𝓂 )
-----------------------------
-Monad\ST(𝓈)(\ST(𝓈)(𝓂 ))
-
-get : \ST 𝓈 𝓂  𝓈
-get ≔ λ 𝓈 → return (𝓈,𝓈)
-
-put : 𝓈 → \ST 𝓈 𝓂  1
-put(𝓈) ≔ λ (𝓈') → return (∙,𝓈)
-
-MonadSmallStep(𝓂 )(𝒮𝒮)
---------------------------------------
-MonadSmallStep(\ST(𝓈)(𝓂 ))(𝒮𝒮(\_ × 𝓈))
-
-α : ∀ (A B : Set), (A → \ST(𝓈)(𝓂 )(B)) → (𝒮𝒮(A × 𝓈) → 𝒮𝒮(B × 𝓈))
-α(f) ≔ αₘ (λ (a,𝓈) → f(a)(𝓈))
-
-γ : ∀ (A B : Set), (𝒮𝒮(A × 𝓈) → 𝒮𝒮(B × 𝓈)) → (A → \ST(𝓈)(𝓂 )(B))
-γ(f) ≔ λ(a,𝓈) → γₘ (f)(a,s)
-\end{verbatim}
+% \begin{verbatim}
+% 
+% MonadSmallStep(𝓂 )(\A_T) ∧ (\PM ∘ \A_T) α⇄ γ (\A_T ∘ \PM)
+% -------------------------------------
+% MonadSmallStep(\ST 𝓂 )(\PM ∘ 𝒮)
+% 
+% α : ∀ (A B : Set), (A → \ST(𝓂 )(B)) → (\PM(\A_T(A) → \PM(\A_T(B))))
+% α(f) ≔ extendₚ(α ∘ αₘ(f))
+% 
+% γ : ∀ (A B : Set), (\PM(\A_T(A) → \PM(\A_T(B)))) → (A → \ST(𝓂 )(B))
+% γ(f) ≔ extendₚ(γ ∘ γₘ(f))
+% \end{verbatim}
+% 
+% % }}}
+% 
+% % StateT {{{
+% 
+% \subsection{$\ST$}
+% \label{StateT}
+% 
+% \begin{verbatim}
+% \ST : Set → (Set → Set) → (Set → Set)
+% \ST(𝓈)(𝓂 )(A) ≔ 𝓈 → 𝓂  (A × 𝓈)
+% 
+% Transformer(\ST(𝓈))
+% ----------------------
+% 
+% lift : ∀ (𝓂  : Set → Set) (A : Set), 𝓂  a → \ST(𝓈)(𝓂 )(A)
+% lift aM ≔ λ 𝓈 → mapₘ (,𝓈) aM
+% 
+% Monad(𝓂 )
+% --------------------
+% Monad(\ST(𝓈)(𝓂 ))
+% 
+% return : ∀ (A : Set), A → \ST(𝓈)(𝓂 )(A)
+% return ≔ lift ∘ returnₘ
+% 
+% extend : ∀ (A B : Set), (A → \ST(𝓈)(𝓂 )(B)) → (\ST(𝓈)(𝓂 )(A) → \ST(𝓈)(𝓂 )(B))
+% extend(k)(aM) ≔ λ 𝓈 → doₘ
+%   (a,𝓈') ← aM(𝓈)
+%   k(a)(𝓈')
+% 
+% JoinSemilattice(𝓈) ∧ Monad(𝓂 )
+% ------------------------------------------
+% Functorial(JoinSemilattice)(\ST(𝓈)(𝓂 ))
+% 
+% bot : \ST(𝓈)(𝓂 )(A)
+% bot ≔ λ(𝓈)→ (⊥ ₛ, ⊥ ₐ)
+% 
+% ⊔ : \ST(𝓈)(𝓂 )(A) → \ST(𝓈)(𝓂 )(A) → \ST(𝓈)(𝓂 )(A)
+% aM₁ ⊔ aM₂ ≔ λ 𝓈 → doₘ
+%   (a₁,𝓈₁) ← aM₁
+%   (a₂,𝓈₂) ← aM₂
+%   return (a₁ ⊔ a₂, 𝓈₁ ⊔ 𝓈₂)
+% 
+% MonadPlus(𝓂 )
+% ------------------------
+% MonadPlus(\ST(𝓈)(𝓂 ))
+% 
+% ⊥ : ∀ (A : Set), \ST(𝓈)(𝓂 )(A)
+% ⊥ ≔ lift ⊥ ₘ
+% 
+% ⊔ : ∀ (A : Set), \ST(𝓈)(𝓂 )(A)
+% aM₁ ⊔ aM₂ ≔ λ 𝓈 → aM₁(𝓈) <+> aM₂(𝓈)
+% 
+% Monad(𝓂 )
+% ----------------------------
+% Monad\ST(𝓈)(\ST(𝓈)(𝓂 ))
+% 
+% get : \ST 𝓈 𝓂  𝓈
+% get ≔ λ 𝓈 → return (𝓈,𝓈)
+% 
+% put : 𝓈 → \ST 𝓈 𝓂  1
+% put(𝓈) ≔ λ (𝓈') → return (∙,𝓈)
+% 
+% MonadSmallStep(𝓂 )(\A_T)
+% --------------------------------------
+% MonadSmallStep(\ST(𝓈)(𝓂 ))(\A_T(\_ × 𝓈))
+% 
+% α : ∀ (A B : Set), (A → \ST(𝓈)(𝓂 )(B)) → (\A_T(A × 𝓈) → \A_T(B × 𝓈))
+% α(f) ≔ αₘ (λ (a,𝓈) → f(a)(𝓈))
+% 
+% γ : ∀ (A B : Set), (\A_T(A × 𝓈) → \A_T(B × 𝓈)) → (A → \ST(𝓈)(𝓂 )(B))
+% γ(f) ≔ λ(a,𝓈) → γₘ (f)(a,s)
+% \end{verbatim}
 
@@
 % Bibliography {{{
-\bibliography{davdar}{}
+\bibliography{davdar}
 \bibliographystyle{plain}

### 2014-10-25 writeup-old/sections/03AAMByExample/01Concrete/00SS.hs
rearranging

@@ -1,3 +1,0 @@
-type Env = Map Name Val
-data Val = LitV Lit | Clo [Name] Call Env
-type StateSpace = Maybe (Call, Env)

### 2014-10-25 writeup-old/sections/03AAMByExample/03AbstractSemantics/02BindMany.hs

@@ -1,11 +1,0 @@
-bindMany :: (AAM aam) => aam 
-         -> Time aam -> [Name] -> [Atom] 
-         -> Env aam -> Store d aam 
-         -> Maybe (Env aam, Store d aam)
-bindMany _   _ [] [] e s = Just (e, s)
-bindMany aam t (x:xs) (xa:xas) e s = case bindMany aam t xs xas e s of
-  Nothing -> Nothing
-  Just (e', s') ->
-    let l = alloc aam x t
-    in Just (mapInsert x l e', mapInsertWith (\/) l xv s')
-bindMany _ _ _ _ _ _ = Nothing

### 2014-10-25 writeup-old/sections/03AAMByExample/03AbstractSemantics/03Step.hs

@@ -1,10 +1,0 @@
-step :: (Delta d, AAM aam) => d -> aam 
-     -> StateSpace d aam -> StatSpace d aam
-step d aam ss = eachInSet ss $ \ (c,e,s,t) -> 
-  eachInSet (call d aam t c e s) $ \ (c',e',s') ->
-    setSingleton (c',e',s',tick aam c' t)
-
-exec :: (Delta d, AAM aam) => d -> aam 
-     -> Call -> StateSpace d aam
-exec d aam c0 = iter (collect (step d aam)) $ 
-  setSingleton (c0, mapEmpty, mapEmpty, tzero aam)

### 2014-10-25 writeup-old/sections/03AAMByExample/04RecoveringConcrete/01Delta.hs

@@ -1,16 +1,0 @@
-data C_Delta = C_Delta
-data ConcreteVal aam = LitV Lit | Clo [Name] Call (Env aam)
-instance Delta C_Delta where
-  type Val C_Delta = ConcreteVal
-  lit C_Delta l = Lit l
-  clo C_Delta xs c e = Clo xs c e
-  op C_Delta Add1 (LitV (I i)) = Just $ LitV $ I $ i+1
-  op C_Delta Sub1 (LitV (I i)) = Just $ LitV $ I $ i-1
-  op C_Delta IsNonNeg (LitV (I i)) 
-    | i >= 0 = Just $ LitV $ B $ True
-    | otherwise = Just $ LitV $ B $ False
-  op C_Delta _ _ = Nothing
-  elimBool C_Delta (LitV (B b)) = setSingleton b
-  elimBool C_Delta _ = setEmpty
-  elimClo C_Delta (Clo xs c e) = setSingleton (xs, c, e)
-  elimClo C_Delta _ = setEmpty

### 2014-10-25 writeup-old/sections/03AAMByExample/05Recovering0CFA/00AAM.hs

@@ -1,7 +1,0 @@
-data ZCFA_AAM = ZCFA_AAM
-instance AAM ZCFA_AAM where
-  type Time ZCFA_AAM = ()
-  type Addr ZCFA_AAM = Name
-  tzero ZCFA_AAM = ()
-  tick ZCFA_AAM _ () = ()
-  alloc ZCFA_AAM x () = x

### 2014-10-25 writeup-old/sections/03AAMByExample/05Recovering0CFA/02Exec.hs

@@ -1,2 +1,0 @@
-execSymZCFA :: Call -> StateSpace Sym_Delta ZCFA_AAM
-execSymZCFA = exec Sym_Delta ZCFA_AAM

### 2014-10-25 writeup-old/sections/04MonadicAAM/00MonadicStyle/00Nondeterminism/01Nondet.hs

@@ -1,5 +1,0 @@
-type Nondet = []
-instance Monad Nondet where
-  return x = [x]
-  [] >>= k = []
-  (x:xs) >>= k = k x ++ (xs >>= k)

### 2014-10-25 writeup-old/sections/04MonadicAAM/00MonadicStyle/00Nondeterminism/02NondetMonadPlus.hs

@@ -1,3 +1,0 @@
-instance MonadPlus [] where
-  mzero = []
-  mplus = (++)

### 2014-10-25 writeup-old/sections/04MonadicAAM/00MonadicStyle/01State/03StateMonadPlus.hs

@@ -1,4 +1,0 @@
-instance (MonadPlus m) => MonadPlus (StateT s m) where
-  mzero = StateT $ \ _ -> mzero
-  mplus xM yM = StateT $ \ s -> 
-    runStateT xM s `mplus` runStateT yM s

### 2014-10-25 writeup-old/sections/04MonadicAAM/00MonadicStyle/01State/06MonadStateInterior.hs

@@ -1,8 +1,0 @@
-instance (MonadState s m) => MonadState s (StateT s' m) where
-  get = StateT $ \ s -> do
-    s' <- get
-    return (s',s)
-  put s' = StateT $ \ s -> do
-    put s'
-    return ((),s)
-

### 2014-10-25 writeup-old/sections/04MonadicAAM/00MonadicStyle/01State/08Atom.hs

@@ -1,14 +1,0 @@
-atom :: (Delta d) => d -> aam -> Atom -> M d aam (Val d aam)
-atom d aam (LitA l) = return $ lit d l
-atom d aam (Var x) = do
-  e <- get $ E aam
-  s <- get $ S d aam
-  case mapLookup x e of
-    Nothing -> mzero
-    Just l -> setMSum $ mapLookup l s
-atom d aam (Prim o a) = do
-  v <- atom d aam a
-  mFromMaybe $ op d o v
-atom d aam (Lam xs c) = do
-  e <- get $ E aam
-  return $ clo d xs c e

### 2014-10-25 writeup-old/sections/04MonadicAAM/00MonadicStyle/01State/10StateSpace.hs

@@ -1,1 +1,0 @@
-type StateSpace d aam = Set (Call, Env aam, Store d aam, Time aam)

### 2014-10-25 writeup-old/sections/04MonadicAAM/00MonadicStyle/01State/10Step.hs

@@ -1,6 +1,0 @@
-step :: (Delta d, AAM aam) => d -> aam 
-     -> StateSpace d aam -> StatSpace d aam
-step d aam ss = eachInSet ss $ \ (c,e,s,t) -> 
-  eachInSet (call d aam t c e s) $ \ (c',e',s') ->
-    setSingleton (c',e',s',tick aam c' t)
-

### 2014-10-25 writeup-old/sections/04MonadicAAM/01Generalizing/06StateTMonadStep.hs

@@ -1,6 +1,0 @@
-newtype StateT_SS s m a = StateT_SS { runStateT_SS :: SS m (a, Cell s) }
-instance (Pointed (SS m), JoinLattice (Cell s)) => Pointed (StateT_SS s m) where
-  point a = point (a, bot)
-instance (MonadStep m) => MonadStep (StateT s m) where
-  SS (StateT s m) = StateT_SS s m
-  step f xSS = step (\ (a, s) -> runStateT (f a) s) $ runStateT_SS xSS

### 2014-10-25 writeup-old/sections/04MonadicAAM/02RecoveringConcrete/00Point.hs

@@ -1,1 +1,0 @@
-data Point a = Top | Bot | Point a

### 2014-10-25 writeup-old/sections/04MonadicAAM/02RecoveringConcrete/02PointMonadPlus.hs

@@ -1,7 +1,0 @@
-instance MonadPlus Point where
-  mzero = Bot
-  mplus Bot yM = yM
-  mplus xM Bot = xM
-  mplus Top yM = Top
-  mplus xM Top = Top
-  mplus (Point x) (Point y) = Top

### 2014-10-25 writeup-old/sections/04MonadicAAM/02RecoveringConcrete/03PointMonadStep.hs

@@ -1,5 +1,0 @@
-instance Pointed Point where
-  point = Point
-instance MonadStep Point where
-  type SS Point = Point
-  step f xs = xs >>= f

### 2014-10-25 writeup-old/sections/04MonadicAAM/03RecoveringKCFA/01Exec.hs

@@ -1,2 +1,0 @@
-execSymkCFA :: Int -> Call -> SS (A_M Sym_Delta ZCFA_AAM) Call
-execSymkCFA k = exec Sym_Delta (kCFA_AAM k)

### 2014-10-25 writeup-old/sections/04MonadicAAM/04Optimizations/00Widen/01BeforeM.hs

@@ -1,1 +1,0 @@
-type A_M d aam = StateT (E aam) (StateT (S d aam) (StateT (T aam) Nondet))

### 2014-10-25 writeup-old/sections/04MonadicAAM/04Optimizations/00Widen/04NondetT.hs

@@ -1,1 +1,0 @@
-newtype NondetT m a = NondetT { runNondetT :: m [a] }

### 2014-10-25 writeup-old/sections/04MonadicAAM/04Optimizations/00Widen/05NondetTMonad.hs

@@ -1,6 +1,0 @@
-instance (JoinLatticeF m) => Monad (NonDetT m) where
-  return x = NonDetT $ return [x]
-  xM >>= k = case joinLatticeF :: JoinLatticeW (m [b]) of
-    JoinLatticeW -> do
-      xs <- runNondetT xM
-      listJoin $ map (runNonDetT . k) xs

### 2014-10-25 writeup-old/sections/04MonadicAAM/04Optimizations/00Widen/06NondetTMonadPlus.hs

@@ -1,4 +1,0 @@
-instance (JoinLatticeF m) => MonadPlus (NonDetT m) where
-  mzero = NonDetT mzero
-  mplus xM yM = case joinLatticeF :: JoinLatticeW (m [a]) of
-    JoinLatticeW -> runNondetT xM \/ runNonDetT yM

### 2014-10-25 writeup-old/sections/04MonadicAAM/04Optimizations/00Widen/07NondetTMonadState.hs

@@ -1,7 +1,0 @@
-instance (MonadState s m) => MonadState s (NondetT m) where
-  get = NondetT $ do
-    s <- get
-    return [s]
-  put s = NondetT $ do
-    put s
-    return [()]

### 2014-10-25 writeup-old/sections/04MonadicAAM/04Optimizations/00Widen/08NondetTMonadStep.hs

@@ -1,8 +1,0 @@
-instance (Pointed m) => Pointed (NondetT m) where
-  point x = NondetT $ point [x]
-instance (MonadStep m) => MonadStep (NondetT m) where
-  type SS (NondetT m) = NondetT m
-  step (f :: a -> NondetT m b) xM = 
-    NondetT $ 
-    step (\ xs -> runNondetT $ NondetT (return xs) >>= f) $ 
-    runNondetT xM

### 2014-10-25 writeup/pldi15.tex
first push on paper content.  focusing on ripping out features piecewise and relating making correct-by-construction arguments along the way.

% any author declaration will be ignored  when using 'plid' option (for double blind review)
\documentclass[pldi]{sigplanconf}

% From PLDI Template {-{

% \usepackage{SIunits}            % typset units correctly
\usepackage{courier}            % standard fixed width font
\usepackage[scaled]{helvet} % see www.ctan.org/get/macros/latex/required/psnfss/psnfss2e.pdf
\usepackage{url}                  % format URLs
\usepackage{listings}          % format code
\usepackage{enumitem}      % adjust spacing in enums
\usepackage[colorlinks=true,allcolors=blue,breaklinks,draft=false]{hyperref}   % hyperlinks, including DOIs and URLs in bibliography
% known bug: http://tex.stackexchange.com/questions/1522/pdfendlink-ended-up-in-different-nesting-level-than-pdfstartlink
\newcommand{\doi}[1]{doi:~\href{http://dx.doi.org/#1}{\Hurl{#1}}}   % print a hyperlinked DOI

% }-}

% Setup {-{

\usepackage{fontspec}
\setmainfont{CMU Serif}   % unicode variant of CM

\usepackage{amsmath}      % \begin{align*}
\usepackage{amssymb}      % \leadsto
\usepackage{bm}           % bold maths
\usepackage{unicode-math} % unicode maths
\usepackage{galois}       % \galois

\usepackage{float}        % \begin{figure}
\floatstyle{boxed}
\restylefloat{figure}

% }-}

% Macros {-{

\newcommand{\lamif}{\textbf{λ-IF}\@}

\newcommand{\ifz}[3]{\pmb{\operatorname{if0}} (#1)\{#2\}\{#3\}}
\newcommand{\lam}{\pmb{λ}}
\newcommand{\indor}{\;|\;}
\newcommand{\plus}{\pmb{+}}
\newcommand{\minus}{\pmb{-}}
\newcommand{\op}{\mathbin{op}}
\newcommand{\ap}{\;\;}
\newcommand{\steps}{\leadsto}
\newcommand{\ddo}{\underline{\operatorname{do}}}
\newcommand{\case}[1]{\text{\underline{case} $#1$ \underline{of}}}
\newcommand{\pushkon}{\operatorname{push-Kon}}
\newcommand{\popkon}{\operatorname{pop-Kon}}
\newcommand{\getenv}{\operatorname{get-Env}}
\newcommand{\putenv}{\operatorname{put-Env}}
\newcommand{\ret}{\underline{\operatorname{return}}}
\newcommand{\toclo}{\operatorname{to-clo}}
\newcommand{\mif}[3]{\text{\underline{if} $#1$ \underline{then} $#2$ \underline{else} $#3$}}
\newcommand{\stepf}{\operatorname{step}}
\newcommand{\atomf}{\operatorname{atom}}
\newcommand{\branchOnSet}{\operatorname{branch-on-set}}
\newcommand{\intI}{\operatorname{int-I}}
\newcommand{\ifzE}{\operatorname{if0-E}}
\newcommand{\cloI}{\operatorname{clo-I}}
\newcommand{\cloE}{\operatorname{clo-E}}
\newcommand{\st}{\;|\;}

% }-}

\begin{document}

\title{Modular Static Analysis: Orthogonal Design, Correct by Construction}
\maketitle

% Abstract {-{
% \begin{abstract}
% \end{abstract}
% }-}

% Introduction {-{
% \section{Introduction}

% }-}

% Modular Analysis by Example {{{
\section{Modular Analysis by Example}

To demonstrate our framework we will grow a concrete semantics into an executable analysis in small steps.
Each step will expose some property of the analysis to be tuned independent of other properties.
We use an applied lambda calculus as an example language to demonstrate our approach.

% }}}

% Concrete Semantics {-{
\subsection{Concrete Semantics}

Our example language is \lamif, an applied lambda caluclus with integers and conditional statements.
The syntax for \lamif\ is given in figure~\ref{lamif:syntax}.

\begin{figure}[t]
\caption{\lamif\ syntax.}
\label{lamif:syntax}
\input{snips/Lam-IF.tex}
\end{figure}

We will use a standard small-step semantics for \lamif, which is shown in figure~\ref{lamif:semantics}.

\begin{figure}[t]
\caption{\lamif\ semantics.}
\label{lamif:semantics}
{\scriptsize
\input{snips/Lam-IF-Semantics-SS.tex}
\input{snips/Lam-IF-Semantics-Rules.tex}
}
\end{figure}

The analyses we will consider will be stated as the least fixed point of some abstract collecting semantics.
A collecting semantics for \lamif\ is defined as:
\begin{equation*}
  μ(ς). ς₀ ⊔ ς ⊔ \{ ς' \st ς \steps ς' \}
\end{equation*}
where $ς₀ := ⟨ e₀ , ⊥ , ∙ ⟩$ is the injection of the initial program $e₀$ into the state space $Σ$.

% }-}

% From Semantics to Interpreter {-{
\section{From Semantics to Interpreter}

On our way to an executable analysis for \lamif, we first evolve the small-step relation into an executable small-step interpreter.
We write this interpreter in monadic style for two reasons.
First, it is much easier to add new states and rules to a monadic interpreter.
Second, we will exploit the monadic abstraction to expose flow and path sensitivities in a later interpreter.

The monad for the interpreter will have two types effects: state and partiality.
Two operations carry state effects: get and put; and one operation carries a partiality effect: fail.

We construct a monad $ℳ $ which combines state effects for $Env$ and $Kon$ with partiality effects.
The type $ℳ $ is just a simple type which supports the monad, state, and partiality operations.
The interpreter for \lamif\ which uses $ℳ $ and monadic effects is given in figure~\ref{lamif:minterpreter}.

\begin{figure}[t]
\caption{\lamif\ monadic interpreter.}
\label{lamif:minterpreter}
{\scriptsize
\input{snips/Lam-IF-MInterpreter.tex}
}
\end{figure}

As before, we must construct a collecting semantics in order to define an executable analysis.
To do this, we must relate our step function back to some state machine step \emph{function}.
One we c % HERE

% }-}

% Exposing the Abstract Domain {-{
\section{Exposing the Abstract Domain}

We will now expose the choice of abstract domain for the interpreter.
However, giving a finite abstract domain for integers alone will not yet give a computable analysis.
There will still be an infinite number of possible abstract state spaces due to the recursion between closures and environments.
We will address this issue in a future modification to the interpreter.

To expose the choice of abstract domain we introduce a type $AVal$ behind an abstract interface.
\begin{alignat*}{1}
    AVal &∈ Set                \\
   \intI &: ℤ → AVal           \\
   \ifzE &: AVal → 𝒫 (𝔹)       \\
   \cloI &: Clo → AVal         \\
   \cloE &: AVal → 𝒫 (Clo)     \\
  \delta &: Op → AVal → AVal → AVal
\end{alignat*}

To establish the correctness of our interpreter $\stepf$, we must prove that it is monotonic in $AVal$, $intI$, $if0E$, $cloI$ and $cloE$.
This can be done independent of a specific $AVal$.
Then we need only establish $Val \galois{α}{γ} AVal$ for some $AVal$ in order to justify the correctness of $\stepf$ instantiated with $AVal$.

Exposing the abstract domain forces us to reconsider our monadic effects.
For example, a branch on a possibly unknown integer value must return multiple possible machine states.
Therefore we must trade our partiality monad for a nonteterminism monad.
The monadic abstraction supports nondeterminism nicely in this setting.
The helper function $\branchOnSet$ will return multiple times for each value in the set.
This allows the interpreter to be written in a style that need only consider one state at a time.
The monadic interpreter whith the abstract domain exposd is given in figure~\ref{lamif:minterpreter:adom}.
(The implementation of $\stepf$ is unchanged.)

\begin{figure}[t]
\caption{\lamif\ monadic interpreter with abstract domain exposed.}
\label{lamif:minterpreter:adom}
{\scriptsize
\input{snips/Lam-IF-MInterpreter-ADom.tex}
}
\end{figure}

% }-}

\end{document}

### 2014-10-25 writeup/snips/Lam-IF-Semantics-Rules.tex
first push on paper content.  focusing on ripping out features piecewise and relating making correct-by-construction arguments along the way.

\begin{alignat*}{1}
  ⟨ a , ρ , [ □ \op e ] ∷ \overline{fr} ⟩ &\steps ⟨ e                         , ρ , [ 𝒜_ρ⟦a⟧ \op □ ] ∷ \overline{fr} ⟩         \\
  ⟨ a , ρ , [ v \op □ ] ∷ \overline{fr} ⟩ &\steps ⟨ \delta(\op , v , 𝒜_ρ⟦a⟧ ) , ρ , \overline{fr} ⟩                            \\
%
  ⟨ a , ρ , [ □                   \ap e ] ∷ \overline{fr} ⟩ &\steps ⟨ e , ρ               , [ 𝒜_ρ⟦a⟧ \ap □ ] ∷ \overline{fr} ⟩ \\
  ⟨ a , ρ , [ ⟨ \lam(x) → e, ρ' ⟩ \ap □ ] ∷ \overline{fr} ⟩ &\steps ⟨ e , ρ'[x ↦ 𝒜_ρ⟦a⟧ ] , \overline{fr} ⟩                    \\
%
  ⟨ a , ρ , [ \ifz{□}{e₁}{e₂} ] ∷ \overline{fr} ⟩ &\steps 
  \begin{cases} 
      ⟨ e₁ , ρ , \overline{fr} ⟩ & \text{if $𝒜_ρ⟦a⟧ = 0$}            \\
      ⟨ e₂ , ρ , \overline{fr} ⟩ & \text{otherwise} 
  \end{cases} \\
%
  ⟨ e₁ \op e₂        , ρ , \overline{fr} ⟩ &\steps ⟨ e₁   , ρ , [ □ \op e₂ ]        ∷ \overline{fr} ⟩               \\
  ⟨ e₁ \ap e₂        , ρ , \overline{fr} ⟩ &\steps ⟨ e₁   , ρ , [ □ \ap e₂ ]        ∷ \overline{fr} ⟩               \\
  ⟨ \ifz{e₁}{e₂}{e₃} , ρ , \overline{fr} ⟩ &\steps ⟨ e₁   , ρ , [ \ifz{□}{e₂}{e₃} ] ∷ \overline{fr} ⟩
\end{alignat*}

### 2014-10-27 writeup/pldi15.tex
interpreter and exposing abstract domain content in.  now working on exposing flow sensitivities.

@@
 \newcommand{\case}[1]{\text{\underline{case} $#1$ \underline{of}}}
+\newcommand{\getkon}{\operatorname{get-Kon}}
+\newcommand{\putkon}{\operatorname{put-Kon}}
 \newcommand{\pushkon}{\operatorname{push-Kon}}
@@
 \newcommand{\putenv}{\operatorname{put-Env}}
-\newcommand{\ret}{\underline{\operatorname{return}}}
+\newcommand{\ret}{\operatorname{return}}
 \newcommand{\toclo}{\operatorname{to-clo}}
@@
 \newcommand{\cloE}{\operatorname{clo-E}}
+\newcommand{\aintI}{\widehat{\operatorname{int-I}}}
+\newcommand{\aifzE}{\widehat{\operatorname{if0-E}}}
+\newcommand{\acloI}{\widehat{\operatorname{clo-I}}}
+\newcommand{\acloE}{\widehat{\operatorname{clo-E}}}
+\newcommand{\adelta}{\widehat{\delta}}
 \newcommand{\st}{\;|\;}
+\newcommand{\coerce}[1]{\operatorname{↓_{#1}}}
+\newcommand{\failonbot}{\operatorname{↑_⊥}}
+\newcommand{\liftnondet}{\operatorname{↑_𝒫 }}
+\newcommand{\AVal}{\widehat{Val}}
+\newcommand{\AEnv}{\widehat{Env}}
+\newcommand{\mplus}{\mathbin{⟨+⟩}}
 
@@
 
-% Modular Analysis by Example {{{
+% Modular Analysis by Example {-{
 \section{Modular Analysis by Example}
@@
 To demonstrate our framework we will grow a concrete semantics into an executable analysis in small steps.
+% Should I resist the pun? :P -DD
 Each step will expose some property of the analysis to be tuned independent of other properties.
 We use an applied lambda calculus as an example language to demonstrate our approach.
-
-% }}}
 
@@
 \label{lamif:syntax}
-\input{snips/Lam-IF.tex}
+\input{snips/00_Semantics/00_Syntax.tex}
 \end{figure}
@@
 \label{lamif:semantics}
-{\scriptsize
-\input{snips/Lam-IF-Semantics-SS.tex}
-\input{snips/Lam-IF-Semantics-Rules.tex}
-}
+\input{snips/00_Semantics/01_Semantics.tex}
 \end{figure}
 
-The analyses we will consider will be stated as the least fixed point of some abstract collecting semantics.
+Analyses in our framework are least fixed point computations over a collecting semantics.
 A collecting semantics for \lamif\ is defined as:
@@
 % From Semantics to Interpreter {-{
-\section{From Semantics to Interpreter}
+\subsection{From Semantics to Interpreter}
+\label{interpreter}
 
@@
 First, it is much easier to add new states and rules to a monadic interpreter.
-Second, we will exploit the monadic abstraction to expose flow and path sensitivities in a later interpreter.
+Second, we will exploit the monadic abstraction to expose flow and path sensitivities in a later section.
 
-The monad for the interpreter will have two types effects: state and partiality.
-Two operations carry state effects: get and put; and one operation carries a partiality effect: fail.
+The interpreter will use two types monadic effects: state and partiality.
+The operations for state effects are get and put, and the operation for the partiality effect is $⊥$.
+For reference: monadic, state and partiality operations have the following types:
 
-We construct a monad $ℳ $ which combines state effects for $Env$ and $Kon$ with partiality effects.
-The type $ℳ $ is just a simple type which supports the monad, state, and partiality operations.
-The interpreter for \lamif\ which uses $ℳ $ and monadic effects is given in figure~\ref{lamif:minterpreter}.
+\input{snips/01_Interpreter/00_MonadOps.tex}
+%\input{snips/01_Interpreter/01_StateOps.tex}
+%\input{snips/01_Interpreter/02_PartialityOps.tex}
 
-\begin{figure}[t]
-\caption{\lamif\ monadic interpreter.}
-\label{lamif:minterpreter}
-{\scriptsize
-\input{snips/Lam-IF-MInterpreter.tex}
-}
-\end{figure}
+We construct a monad $ℳ $ for our interpreter which combines state effects for $Env$ and $Kon$ with a partiality effect.
+The type $ℳ $ is just a simple type:
+\input{snips/01_Interpreter/03_MonadType.tex}
+
+We factor the definition of the interpreter into two parts: one for atomic and one for compound expressions.
+Atomic expressions return to the top frame of the stack:
+\input{snips/01_Interpreter/04_Atomic.tex}
+
+Our interpreter uses helper functions $𝒜 ⟦\_⟧$, $\pushkon$, $\popkon$, $\delta$, $\coerce{Clo}$, and $\failonbot$.
+The function $𝒜 ⟦\_⟧$ is the denotation function which uses the monadic state effect for environments:
+\input{snips/01_Interpreter/05_AtomDen.tex}
+
+Functions $\pushkon$ and $\popkon$ interact with effects for both continuation state and partiality:
+\input{snips/01_Interpreter/06_PushPopKon.tex}
+
+The function $\delta$ gives the expected denotation to operators:
+\input{snips/01_Interpreter/07_Delta.tex}
+
+The function $\coerce{Clo}$ is a function from types which include closure values to partial closures.
+\input{snips/01_Interpreter/08_CoerceClo.tex}
+
+The function $\failonbot$ lifts partial values to monadic values, mapping $⊥$ to $⊥$.
+\input{snips/01_Interpreter/09_FailOnBot.tex}
+
+Our small-step interpreter is defined by dispatching to $\atomf$ or pushing continuation frames for compound expressions.
+\input{snips/01_Interpreter/10_Exp.tex}
+
+\paragraph{Correctness and Execution}
 
 As before, we must construct a collecting semantics in order to define an executable analysis.
-To do this, we must relate our step function back to some state machine step \emph{function}.
-One we c % HERE
+To do this, we must relate our small-step interpreter to a state machine transition function.
+That is, we must relate our interpreter of type $Exp → ℳ (Exp)$ to a state transition of type $Σ → Σ$ for \emph{some} $Σ$.
+
+For this interpreter the state space $Σ$ is defined $Σ := {(Exp × Env × Kon)}_⊥$.
+To establish the correctness of $Σ$ we construct an isomorphism between $Exp → ℳ (Exp)$ and $Σ → Σ$.
+We use the standard isomorphisim between Kleisli morphisms $α → ℳ (β)$ and $ℳ (α) → ℳ (β)$ where $ℳ $ is the partiality monad ${(\_)}_⊥$.
+To execute the analysis we transport the interpreter through the isomorphism and find the least fixed point of the collecting semantics.
 
@@
 % Exposing the Abstract Domain {-{
-\section{Exposing the Abstract Domain}
+\subsection{Exposing the Abstract Domain}
 
@@
 
-To expose the choice of abstract domain we introduce a type $AVal$ behind an abstract interface.
-\begin{alignat*}{1}
-    AVal &∈ Set                \\
-   \intI &: ℤ → AVal           \\
-   \ifzE &: AVal → 𝒫 (𝔹)       \\
-   \cloI &: Clo → AVal         \\
-   \cloE &: AVal → 𝒫 (Clo)     \\
-  \delta &: Op → AVal → AVal → AVal
-\end{alignat*}
+To expose the choice of abstract domain we introduce a type $\AVal$ behind an abstract interface:
+\input{snips/02_AbstractDomain/00_AVal.tex}
 
-To establish the correctness of our interpreter $\stepf$, we must prove that it is monotonic in $AVal$, $intI$, $if0E$, $cloI$ and $cloE$.
-This can be done independent of a specific $AVal$.
-Then we need only establish $Val \galois{α}{γ} AVal$ for some $AVal$ in order to justify the correctness of $\stepf$ instantiated with $AVal$.
+This abstraction forces us to use an abstract environment, mapping variables to abstract values:
+\input{snips/02_AbstractDomain/01_AEnv.tex}
+and re-implement $𝒜 ⟦\_⟧$:
+\input{snips/02_AbstractDomain/02_AtomDen.tex}
 
-Exposing the abstract domain forces us to reconsider our monadic effects.
-For example, a branch on a possibly unknown integer value must return multiple possible machine states.
-Therefore we must trade our partiality monad for a nonteterminism monad.
-The monadic abstraction supports nondeterminism nicely in this setting.
-The helper function $\branchOnSet$ will return multiple times for each value in the set.
-This allows the interpreter to be written in a style that need only consider one state at a time.
-The monadic interpreter whith the abstract domain exposd is given in figure~\ref{lamif:minterpreter:adom}.
-(The implementation of $\stepf$ is unchanged.)
+Before implementing $\atomf$ we must reconsider our monadic effects.
+Because we have abstracted values, our interpreter must consider \emph{sets} of possible states resulting from conditionals and function calls.
+We introduce this nondeterminism as an effect in the monad, taking the place of partiality (which was just nondeterminism with zero or one branches).
+Nondeterminism has two operators, $⊥$ as we've seen before, and $\mplus$ which branches to both left and right monadic computation.
+The effects for our monadic type are now:
+\input{snips/02_AbstractDomain/03_Effects.tex}
+and to support these effects we redefine $ℳ $ to be:
+\input{snips/02_AbstractDomain/04_MonadType.tex}
 
-\begin{figure}[t]
-\caption{\lamif\ monadic interpreter with abstract domain exposed.}
-\label{lamif:minterpreter:adom}
-{\scriptsize
-\input{snips/Lam-IF-MInterpreter-ADom.tex}
-}
-\end{figure}
+Our definition for $\atomf$ now uses nondeterminism effects:
+\input{snips/02_AbstractDomain/05_Atomic.tex}
+
+Helper function $\liftnondet$ is the analog of $\failonbot$ for lifting sets into monads with nondeterminism:
+\input{snips/02_AbstractDomain/06_LiftNonDet.tex}
+
+\paragraph{Correctness and Execution}
+
+To construct a collecting semantics we again construct an isomorphism between $Exp → ℳ (Exp)$ and $Σ$ where $Σ := 𝒫 (Exp × \AEnv × Kon)$.
+This isomorphism is again the standard one between Kleisli morphisms $α → ℳ (β)$ and $ℳ (α) → ℳ (β)$ where $ℳ $ is the powerset monad $𝒫 $.
+
+Now that we have a parameter to the analysis, $\AVal$, we can discuss what an analysis is and what it means for it to be correct.
+An analysis is some instantiation of $\AVal$ to $\AVal₁$ where $Val\galois{α}{γ}\AVal₁$.
+Furthermore, each operation in the $\AVal$ interface must be an abstraction of concrete versions, namely $\intI ⊑ γ (\aintI₁)$, $\ifzE ⊑ γ (\aifzE₁)$, etc.
+This local argument about $\AVal₁$, coupled with the monotonicity of $\stepf$ and the isomorphism between $Exp → ℳ (Exp)$ and $Σ → Σ$, 
+  gives us an end to end correctness argument for our executable abstract interpreter.
+At this point our analysis is still not yet decidable; the abstract state space is still infinite.
+
+We will expose two more feature, flow sensitivity and context sensitivity, before we achieve a decidable analysis.
+However, the \emph{correctness} of our analysis will be carried through every example.
+
+% }-}
+
+% Exposing Flow Sensitivity {-{
+\subsection{Exposing Flow Sensitivity}
+
+Novel in this work is the ability to expose flow sensitivity as an orthogonal parameter to a static analysis.
+
+% }-}
+
+% Exposing Context Sensitivity {-{
+\subsection{Exposing Context Sensitivity}
+
+% }-}
+
+% Intentional Optimizations {-{
+\subsection{Intentional Optimizations}
+
+\subsubsection{Abstract Garbage Collection}
+
+\subsubsection{Flat Closures}
+
+% }-}
+
+% }-}
+
+% Summary {-{
+\section{Summary}
+
+Our final interpreter enjoys the following properties:
+\begin{itemize}
+\item Abstract Domain can be altered independent of other features.
+\item Flow Sensitivity can be altered independent of other features.
+\item Context Sensitivity (and its granularity) can be altered indepndend of other features.
+\item Intensional optimizations can be implemented once and integrate seamlessly with any of the above choices.
+\item New state space components and their widening properties can be added seamlessly using compositional building blocks.
+\item Interpreters are automatically related to a collecting semantics, enabling a straight forward least-fixed-point implementation strategy.
+\item All analyses are correct by construction with no additional proof burdon beyond interpreter monotonicity and correctness of orthoganal parameters piecewise.
+\end{itemize}
 

### 2014-11-04 writeup/pldi15.tex
markdown and beginning next big push

--- 2014-10-25 writeup/pldi15.tex
@@
 \usepackage{fontspec}
-\setmainfont{CMU Serif}   % unicode variant of CM
-
-\usepackage{amsmath}      % \begin{align*}
-\usepackage{amssymb}      % \leadsto
-\usepackage{bm}           % bold maths
-\usepackage{unicode-math} % unicode maths
-\usepackage{galois}       % \galois
-
-\usepackage{float}        % \begin{figure}
-\floatstyle{boxed}
-\restylefloat{figure}
+\setmonofont[Scale=0.7]{DejaVu Sans Mono}
 
@@
 
-\newcommand{\lamif}{\textbf{λ-IF}\@}
-
-\newcommand{\ifz}[3]{\pmb{\operatorname{if0}} (#1)\{#2\}\{#3\}}
-\newcommand{\lam}{\pmb{λ}}
-\newcommand{\indor}{\;|\;}
-\newcommand{\plus}{\pmb{+}}
-\newcommand{\minus}{\pmb{-}}
-\newcommand{\op}{\mathbin{op}}
-\newcommand{\ap}{\;\;}
-\newcommand{\steps}{\leadsto}
-\newcommand{\ddo}{\underline{\operatorname{do}}}
-\newcommand{\case}[1]{\text{\underline{case} $#1$ \underline{of}}}
-\newcommand{\pushkon}{\operatorname{push-Kon}}
-\newcommand{\popkon}{\operatorname{pop-Kon}}
-\newcommand{\getenv}{\operatorname{get-Env}}
-\newcommand{\putenv}{\operatorname{put-Env}}
-\newcommand{\ret}{\underline{\operatorname{return}}}
-\newcommand{\toclo}{\operatorname{to-clo}}
-\newcommand{\mif}[3]{\text{\underline{if} $#1$ \underline{then} $#2$ \underline{else} $#3$}}
-\newcommand{\stepf}{\operatorname{step}}
-\newcommand{\atomf}{\operatorname{atom}}
-\newcommand{\branchOnSet}{\operatorname{branch-on-set}}
-\newcommand{\intI}{\operatorname{int-I}}
-\newcommand{\ifzE}{\operatorname{if0-E}}
-\newcommand{\cloI}{\operatorname{clo-I}}
-\newcommand{\cloE}{\operatorname{clo-E}}
-\newcommand{\st}{\;|\;}
-
 % }-}
@@
 
-\title{Modular Static Analysis: Orthogonal Design, Correct by Construction}
+\title{Modular Metatheory for Abstract Interpreters}
 \maketitle
@@
 
-% Introduction {-{
-% \section{Introduction}
-
-% }-}
-
-% Modular Analysis by Example {{{
-\section{Modular Analysis by Example}
-
-To demonstrate our framework we will grow a concrete semantics into an executable analysis in small steps.
-Each step will expose some property of the analysis to be tuned independent of other properties.
-We use an applied lambda calculus as an example language to demonstrate our approach.
-
-% }}}
-
-% Concrete Semantics {-{
-\subsection{Concrete Semantics}
-
-Our example language is \lamif, an applied lambda caluclus with integers and conditional statements.
-The syntax for \lamif\ is given in figure~\ref{lamif:syntax}.
-
-\begin{figure}[t]
-\caption{\lamif\ syntax.}
-\label{lamif:syntax}
-\input{snips/Lam-IF.tex}
-\end{figure}
-
-We will use a standard small-step semantics for \lamif, which is shown in figure~\ref{lamif:semantics}.
-
-\begin{figure}[t]
-\caption{\lamif\ semantics.}
-\label{lamif:semantics}
-{\scriptsize
-\input{snips/Lam-IF-Semantics-SS.tex}
-\input{snips/Lam-IF-Semantics-Rules.tex}
-}
-\end{figure}
-
-The analyses we will consider will be stated as the least fixed point of some abstract collecting semantics.
-A collecting semantics for \lamif\ is defined as:
-\begin{equation*}
-  μ(ς). ς₀ ⊔ ς ⊔ \{ ς' \st ς \steps ς' \}
-\end{equation*}
-where $ς₀ := ⟨ e₀ , ⊥ , ∙ ⟩$ is the injection of the initial program $e₀$ into the state space $Σ$.
-
-% }-}
-
-% From Semantics to Interpreter {-{
-\section{From Semantics to Interpreter}
-
-On our way to an executable analysis for \lamif, we first evolve the small-step relation into an executable small-step interpreter.
-We write this interpreter in monadic style for two reasons.
-First, it is much easier to add new states and rules to a monadic interpreter.
-Second, we will exploit the monadic abstraction to expose flow and path sensitivities in a later interpreter.
-
-The monad for the interpreter will have two types effects: state and partiality.
-Two operations carry state effects: get and put; and one operation carries a partiality effect: fail.
-
-We construct a monad $ℳ $ which combines state effects for $Env$ and $Kon$ with partiality effects.
-The type $ℳ $ is just a simple type which supports the monad, state, and partiality operations.
-The interpreter for \lamif\ which uses $ℳ $ and monadic effects is given in figure~\ref{lamif:minterpreter}.
-
-\begin{figure}[t]
-\caption{\lamif\ monadic interpreter.}
-\label{lamif:minterpreter}
-{\scriptsize
-\input{snips/Lam-IF-MInterpreter.tex}
-}
-\end{figure}
-
-As before, we must construct a collecting semantics in order to define an executable analysis.
-To do this, we must relate our step function back to some state machine step \emph{function}.
-One we c % HERE
-
-% }-}
-
-% Exposing the Abstract Domain {-{
-\section{Exposing the Abstract Domain}
-
-We will now expose the choice of abstract domain for the interpreter.
-However, giving a finite abstract domain for integers alone will not yet give a computable analysis.
-There will still be an infinite number of possible abstract state spaces due to the recursion between closures and environments.
-We will address this issue in a future modification to the interpreter.
-
-To expose the choice of abstract domain we introduce a type $AVal$ behind an abstract interface.
-\begin{alignat*}{1}
-    AVal &∈ Set                \\
-   \intI &: ℤ → AVal           \\
-   \ifzE &: AVal → 𝒫 (𝔹)       \\
-   \cloI &: Clo → AVal         \\
-   \cloE &: AVal → 𝒫 (Clo)     \\
-  \delta &: Op → AVal → AVal → AVal
-\end{alignat*}
-
-To establish the correctness of our interpreter $\stepf$, we must prove that it is monotonic in $AVal$, $intI$, $if0E$, $cloI$ and $cloE$.
-This can be done independent of a specific $AVal$.
-Then we need only establish $Val \galois{α}{γ} AVal$ for some $AVal$ in order to justify the correctness of $\stepf$ instantiated with $AVal$.
-
-Exposing the abstract domain forces us to reconsider our monadic effects.
-For example, a branch on a possibly unknown integer value must return multiple possible machine states.
-Therefore we must trade our partiality monad for a nonteterminism monad.
-The monadic abstraction supports nondeterminism nicely in this setting.
-The helper function $\branchOnSet$ will return multiple times for each value in the set.
-This allows the interpreter to be written in a style that need only consider one state at a time.
-The monadic interpreter whith the abstract domain exposd is given in figure~\ref{lamif:minterpreter:adom}.
-(The implementation of $\stepf$ is unchanged.)
-
-\begin{figure}[t]
-\caption{\lamif\ monadic interpreter with abstract domain exposed.}
-\label{lamif:minterpreter:adom}
-{\scriptsize
-\input{snips/Lam-IF-MInterpreter-ADom.tex}
-}
-\end{figure}
-
-% }-}
+\input{pldi15.autogen.tex}
 

### 2014-11-04 writeup/stash/pldi15-dead-end.tex
markdown and beginning next big push

% any author declaration will be ignored  when using 'plid' option (for double blind review)
\documentclass[pldi]{sigplanconf}

% From PLDI Template {-{

% \usepackage{SIunits}            % typset units correctly
\usepackage{courier}            % standard fixed width font
\usepackage[scaled]{helvet} % see www.ctan.org/get/macros/latex/required/psnfss/psnfss2e.pdf
\usepackage{url}                  % format URLs
\usepackage{listings}          % format code
\usepackage{enumitem}      % adjust spacing in enums
\usepackage[colorlinks=true,allcolors=blue,breaklinks,draft=false]{hyperref}   % hyperlinks, including DOIs and URLs in bibliography
% known bug: http://tex.stackexchange.com/questions/1522/pdfendlink-ended-up-in-different-nesting-level-than-pdfstartlink
\newcommand{\doi}[1]{doi:~\href{http://dx.doi.org/#1}{\Hurl{#1}}}   % print a hyperlinked DOI

% }-}

% Setup {-{

\usepackage{fontspec}
\setmainfont{CMU Serif}   % unicode variant of CM

\usepackage{amsmath}      % \begin{align*}
\usepackage{amssymb}      % \leadsto
\usepackage{bm}           % bold maths
\usepackage{unicode-math} % unicode maths
\usepackage{galois}       % \galois

\usepackage{float}        % \begin{figure}
\floatstyle{boxed}
\restylefloat{figure}

% }-}

% Macros {-{

\newcommand{\lamif}{\textbf{λ-IF}\@}

\newcommand{\ifz}[3]{\pmb{\operatorname{if0}} (#1)\{#2\}\{#3\}}
\newcommand{\lam}{\pmb{λ}}
\newcommand{\indor}{\;|\;}
\newcommand{\plus}{\pmb{+}}
\newcommand{\minus}{\pmb{-}}
\newcommand{\op}{\mathbin{op}}
\newcommand{\ap}{\;\;}
\newcommand{\steps}{\leadsto}
\newcommand{\ddo}{\underline{\operatorname{do}}}
\newcommand{\case}[1]{\text{\underline{case} $#1$ \underline{of}}}
\newcommand{\getkon}{\operatorname{get-Kon}}
\newcommand{\putkon}{\operatorname{put-Kon}}
\newcommand{\pushkon}{\operatorname{push-Kon}}
\newcommand{\popkon}{\operatorname{pop-Kon}}
\newcommand{\getenv}{\operatorname{get-Env}}
\newcommand{\putenv}{\operatorname{put-Env}}
\newcommand{\ret}{\operatorname{return}}
\newcommand{\toclo}{\operatorname{to-clo}}
\newcommand{\mif}[3]{\text{\underline{if} $#1$ \underline{then} $#2$ \underline{else} $#3$}}
\newcommand{\stepf}{\operatorname{step}}
\newcommand{\atomf}{\operatorname{atom}}
\newcommand{\branchOnSet}{\operatorname{branch-on-set}}
\newcommand{\intI}{\operatorname{int-I}}
\newcommand{\ifzE}{\operatorname{if0-E}}
\newcommand{\cloI}{\operatorname{clo-I}}
\newcommand{\cloE}{\operatorname{clo-E}}
\newcommand{\aintI}{\widehat{\operatorname{int-I}}}
\newcommand{\aifzE}{\widehat{\operatorname{if0-E}}}
\newcommand{\acloI}{\widehat{\operatorname{clo-I}}}
\newcommand{\acloE}{\widehat{\operatorname{clo-E}}}
\newcommand{\adelta}{\widehat{\delta}}
\newcommand{\st}{\;|\;}
\newcommand{\coerce}[1]{\operatorname{↓_{#1}}}
\newcommand{\failonbot}{\operatorname{↑_⊥}}
\newcommand{\liftnondet}{\operatorname{↑_𝒫 }}
\newcommand{\AVal}{\widehat{Val}}
\newcommand{\AEnv}{\widehat{Env}}
\newcommand{\mplus}{\mathbin{⟨+⟩}}
\newcommand{\AAddr}{\widehat{Addr}}
\newcommand{\ATime}{\widehat{Time}}
\newcommand{\AStore}{\widehat{Store}}
\newcommand{\aalloc}{\widehat{alloc}}
\newcommand{\atick}{\widehat{tick}}

% }-}

\begin{document}

\title{Modular Static Analysis: Orthogonal Design, Correct by Construction}
\maketitle

% Abstract {-{
% \begin{abstract}
% \end{abstract}
% }-}

% Introduction {-{
% \section{Introduction}

% }-}

% Modular Analysis by Example {-{
\section{Modular Analysis by Example}

To demonstrate our framework we will grow a concrete semantics into an executable analysis in small steps.
% Should I resist the pun? :P -DD
Each step will expose some property of the analysis to be tuned independent of other properties.
We use an applied lambda calculus as an example language to demonstrate our approach.

% Concrete Semantics {-{
\subsection{Concrete Semantics}

Our example language is \lamif, an applied lambda caluclus with integers and conditional statements.
The syntax for \lamif\ is given in figure~\ref{lamif:syntax}.

\begin{figure}[t]
\caption{\lamif\ syntax.}
\label{lamif:syntax}
\input{snips/00_Semantics/00_Syntax.tex}
\end{figure}

We will use a standard small-step semantics for \lamif, which is shown in figure~\ref{lamif:semantics}.

\begin{figure}[t]
\caption{\lamif\ semantics.}
\label{lamif:semantics}
\input{snips/00_Semantics/01_Semantics.tex}
\end{figure}

Analyses in our framework are least fixed point computations over a collecting semantics.
A collecting semantics for \lamif\ is defined as:
\begin{equation*}
  μ(ς). ς₀ ⊔ ς ⊔ \{ ς' \st ς \steps ς' \}
\end{equation*}
where $ς₀ := ⟨ e₀ , ⊥ , ∙ ⟩$ is the injection of the initial program $e₀$ into the state space $Σ$.

% }-}

% From Semantics to Interpreter {-{
\subsection{From Semantics to Interpreter}
\label{interpreter}

On our way to an executable analysis for \lamif, we first evolve the small-step relation into an executable small-step interpreter.
We write this interpreter in monadic style for two reasons.
First, it is much easier to add new states and rules to a monadic interpreter.
Second, we will exploit the monadic abstraction to expose flow and path sensitivities in a later section.

The interpreter will use two types monadic effects: state and partiality.
The operations for state effects are get and put, and the operation for the partiality effect is $⊥$.
For reference: monadic, state and partiality operations have the following types:

\input{snips/01_Interpreter/00_MonadOps.tex}
%\input{snips/01_Interpreter/01_StateOps.tex}
%\input{snips/01_Interpreter/02_PartialityOps.tex}

We construct a monad $ℳ $ for our interpreter which combines state effects for $Env$ and $Kon$ with a partiality effect.
The type $ℳ $ is just a simple type:
\input{snips/01_Interpreter/03_MonadType.tex}

We factor the definition of the interpreter into two parts: one for atomic and one for compound expressions.
Atomic expressions return to the top frame of the stack:
\input{snips/01_Interpreter/04_Atomic.tex}

Our interpreter uses helper functions $𝒜 ⟦\_⟧$, $\pushkon$, $\popkon$, $\delta$, $\coerce{Clo}$, and $\failonbot$.
The function $𝒜 ⟦\_⟧$ is the denotation function which uses the monadic state effect for environments:
\input{snips/01_Interpreter/05_AtomDen.tex}

Functions $\pushkon$ and $\popkon$ interact with effects for both continuation state and partiality:
\input{snips/01_Interpreter/06_PushPopKon.tex}

The function $\delta$ gives the expected denotation to operators:
\input{snips/01_Interpreter/07_Delta.tex}

The function $\coerce{Clo}$ is a function from types which include closure values to partial closures.
\input{snips/01_Interpreter/08_CoerceClo.tex}

The function $\failonbot$ lifts partial values to monadic values, mapping $⊥$ to $⊥$.
\input{snips/01_Interpreter/09_FailOnBot.tex}

Our small-step interpreter is defined by dispatching to $\atomf$ or pushing continuation frames for compound expressions.
\input{snips/01_Interpreter/10_Exp.tex}

\paragraph{Correctness and Execution}

As before, we must construct a collecting semantics in order to define an executable analysis.
To do this, we must relate our small-step interpreter to a state machine transition function.
That is, we must relate our interpreter of type $Exp → ℳ (Exp)$ to a state transition of type $Σ → Σ$ for \emph{some} $Σ$.

For this interpreter the state space $Σ$ is defined $Σ := {(Exp × Env × Kon)}_⊥$.
To establish the correctness of $Σ$ we construct an isomorphism between $Exp → ℳ (Exp)$ and $Σ → Σ$.
We use the standard isomorphisim between Kleisli morphisms $α → ℳ (β)$ and $ℳ (α) → ℳ (β)$ where $ℳ $ is the partiality monad ${(\_)}_⊥$.
To execute the analysis we transport the interpreter through the isomorphism and find the least fixed point of the collecting semantics.

% }-}

% Exposing the Abstract Domain {-{
\subsection{Exposing the Abstract Domain}

We will now expose the choice of abstract domain for the interpreter.
However, giving a finite abstract domain for integers alone will not yet give a computable analysis.
There will still be an infinite number of possible abstract state spaces due to the recursion between closures and environments.
We will address this issue in a future modification to the interpreter.

To expose the choice of abstract domain we introduce a type $\AVal$ behind an abstract interface:
\input{snips/02_AbstractDomain/00_AVal.tex}

This abstraction forces us to use an abstract environment, mapping variables to abstract values:
\input{snips/02_AbstractDomain/01_AEnv.tex}
and re-implement $𝒜 ⟦\_⟧$:
\input{snips/02_AbstractDomain/02_AtomDen.tex}

Before implementing $\atomf$ we must reconsider our monadic effects.
Because we have abstracted values, our interpreter must consider \emph{sets} of possible states resulting from conditionals and function calls.
We introduce this nondeterminism as an effect in the monad, taking the place of partiality (which was just nondeterminism with zero or one branches).
Nondeterminism has two operators, $⊥$ as we've seen before, and $\mplus$ which branches to both left and right monadic computation.
The effects for our monadic type are now:
\input{snips/02_AbstractDomain/03_Effects.tex}
and to support these effects we redefine $ℳ $ to be:
\input{snips/02_AbstractDomain/04_MonadType.tex}

Our definition for $\atomf$ now uses nondeterminism effects:
\input{snips/02_AbstractDomain/05_Atomic.tex}

Helper function $\liftnondet$ is the analog of $\failonbot$ for lifting sets into monads with nondeterminism:
\input{snips/02_AbstractDomain/06_LiftNonDet.tex}

\paragraph{Correctness and Execution}

To construct a collecting semantics we again construct an isomorphism between $Exp → ℳ (Exp)$ and $Σ$ where $Σ := 𝒫 (Exp × \AEnv × Kon)$.
This isomorphism is again the standard one between Kleisli morphisms $α → ℳ (β)$ and $ℳ (α) → ℳ (β)$ where $ℳ $ is the powerset monad $𝒫 $.

Now that we have a parameter to the analysis, $\AVal$, we can discuss what an analysis is and what it means for it to be correct.
An analysis is some instantiation of $\AVal$ to $\AVal₁$ where $Val\galois{α}{γ}\AVal₁$.
Furthermore, each operation in the $\AVal$ interface must be an abstraction of concrete versions, namely $\intI ⊑ γ (\aintI₁)$, $\ifzE ⊑ γ (\aifzE₁)$, etc.
This local argument about $\AVal₁$, coupled with the monotonicity of $\stepf$ and the isomorphism between $Exp → ℳ (Exp)$ and $Σ → Σ$, 
  gives us an end to end correctness argument for our executable abstract interpreter.
At this point our analysis is still not yet decidable; the abstract state space is still infinite.

% }-}

% Exposing Context Sensitivity {-{
\subsection{Exposing Context Sensitivity}

To expose context sensitivity (also called call-site sensitivity) we instantiate the AAM approach of cutting recursion in the state space to our framework.
This conceptually takes place in two steps.
First we introduce indirection between values and environments through a store and a type of addresses.
Second we abstract the choice of address to achieve an analysis.

We introduce abstract time and addresses behind an interface:
\input{snips/03_ContextSensitivity/00_AAM.tex}
which alters the definition of the monad to contain new state components:
\input{snips/03_ContextSensitivity/01_StateSpace.tex}

[LEFT OFF HERE]

% }-}

% Exposing Flow Sensitivity {-{
\subsection{Exposing Flow Sensitivity}

Novel in this work is the ability to expose flow sensitivity as an orthogonal parameter to a static analysis.
To do this we exploit both the monadic abstraction and our correctness framework which relates monadic actions to state machine transitions.
\emph{Our insight is that both computational and correctness components of flow-sensitivity can be obtain by varying the monad for the interpreter alone.}

The anlyses we have constructed thus far are path sensitive, meaning they will follow branching control flow precisely.
We can see that the analysis is path sensitive by inspecting the state space $Σ$ that it induces: $ Σ := 𝒫 (Exp × \AEnv × Kon × \ATime × \AStore)$.
A flow insensitive analysis pulls the store out of the powerset to a single global store: $Σ := (𝒫 (Exp × \AEnv × Kon × \ATime) × \AStore)$.
A flow sensitive path insensitive analysis shares the state space of path sensitivity but employs store widening to keep the state space explosion under control.
In this section we will show how variations in the monad can recover all three analysis variants: path sensitive, flow sensitive, and flow insensitive.

% }-}

% Intentional Optimizations {-{
\subsection{Intentional Optimizations}

\subsubsection{Abstract Garbage Collection}

\subsubsection{Flat Closures}

% }-}

% }-}

% Summary {-{
\section{Summary}

Our final interpreter enjoys the following properties:
\begin{itemize}
\item Abstract Domain can be altered independent of other features.
\item Flow Sensitivity can be altered independent of other features.
\item Context Sensitivity (and its granularity) can be altered indepndend of other features.
\item Intensional optimizations can be implemented once and integrate seamlessly with any of the above choices.
\item New state space components and their widening properties can be added seamlessly using compositional building blocks.
\item Interpreters are automatically related to a collecting semantics, enabling a straight forward least-fixed-point implementation strategy.
\item All analyses are correct by construction with no additional proof burdon beyond interpreter monotonicity and correctness of orthoganal parameters piecewise.
\end{itemize}

% }-}

\end{document}

### 2014-11-07 writeup/pldi15.tex
some sed preprocessing for -- comments.  also cleaned up what make does
monadic interpreter and fixes to makeg
added matt's abstract

@@
 
-% Abstract {-{
-% \begin{abstract}
-% \end{abstract}
-% }-}
+Abstract {-{
+\begin{abstract}
 
-\input{pldi15.autogen.tex}
+The design and implementation of static analyzers have becoming increasingly systematic.
+In fact, design and implementation have remained seemingly on the verge of full mechanization for several years.
+A stumbling block in full mechanization has been the ad hoc nature of soundness proofs accompanying each analyzer.
+While design and implementation is largely systematic, 
+  soundness proofs can change significantly with (apparently) minor changes to the semantics and analyzers themselves.
+We finally reconcile the systematic construction of static analyzers with their proofs of soundness 
+  via a mechanistic Galois-connection-based metatheory for static analyzers.
+
+\end{abstract}
+}-}
+
+\input{tmp/autogen/pldi15.markdown.tex}
 

### 2014-11-11 writeup/Process.hs
on the way to automating markdown -> latex

@@ -1,0 +1,231 @@
+{-# LANGUAGE ScopedTypeVariables #-}
+{-# LANGUAGE OverloadedStrings #-}
+
+module Process where
+
+import Text.Pandoc
+import Text.Pandoc.Walk
+import Data.Monoid
+import Control.Monad
+import System.Process
+import Text.Regex.TDFA
+import Text.Regex.TDFA.Text
+import qualified Data.Text as T
+import Data.Text (Text)
+import qualified Data.Text.IO as T
+import Data.List
+
+varNames :: [Text]
+varNames =
+  [ "Var"
+  , "Atom"
+  , "IOp"
+  , "EOp"
+  , "Exp"
+  , "Env"
+  , "Store"
+  , "Time"
+  , "KAddr"
+  , "OAddr"
+  , "KStore"
+  , "M"
+  , "bind"
+  , "return"
+  , "bind-unit"
+  , "bind-associativity"
+  ]
+
+varNameMacros :: [(Text,Text)]
+varNameMacros = flip map varNames $ \ n ->
+  (n, T.concat ["\\mathtt{", n, "}" ])
+
+-- These get applied in reverse order!
+macroTable :: [(Text, Text)]
+macroTable =
+  varNameMacros
+  ++
+  -- math
+  [ ( "ρ"      , " \\rho "             )
+  , ( "α"      , " \\alpha "           )
+  , ( "β"      , " \\beta "            )
+  , ( "τ"      , " \\tau "             )
+  , ( "σ"      , " \\sigma "           )
+  , ( "ς"      , " \\varsigma "        )
+  , ( "Σ"      , " \\Sigma "           )
+  , ( "θ"      , " \\theta "           )
+  , ( "κ"      , " \\kappa "           )
+  , ( "μ"      , " \\mu "              )
+  , ( "π"      , " \\pi "              )
+  , ( "ℤ"      , " \\Int "             )
+  , ( "×"      , " \\times "           )
+  , ( "⇀"      , " \\rightharpoonup "  ) 
+  , ( "→"      , " \\rightarrow "      ) 
+  , ( "□"      , " \\square "          ) 
+  , ( "~~>"    , " \\rightsquigarrow " )
+  , ( "↦"      , " \\mapsto "          )
+  , ( "∷"      , " \\cons "            )
+  , ( "≠"      , " \\neq "             )
+  , ( "≟"      , " \\stackrel{?}{=}"   )
[…]

### 2014-11-11 writeup/Process.hs
last leg of typesetting

@@ -65,10 +65,12 @@ macroText :: Text -> Text
 macroText = appEndo $ execWriter $ forM_ (reverse allMacros) $ \ (s,r,m) -> do
   let escaped = escapeRegex s
       withMode = case m of
-        Word -> T.concat ["\\<",escaped,"\\>"]
+        Word -> T.concat ["([^-]|^)\\<",escaped,"\\>([^-]|$)"]
         Anywhere -> escaped
       regex = mkRegex $ T.unpack withMode
-      replace = T.unpack $ T.concat [" ", r, " "]
+      replace = T.unpack $ T.concat $ case m of
+        Word -> [" \\1", r, "\\2 "]
+        Anywhere -> [" ", r, " "]
   tell $ Endo $ \ t -> T.pack $ subRegex regex (T.unpack t) replace
 
 main :: IO ()
@@ -95,7 +97,7 @@ stripComments = newlines . map fixEmpties . filter (not . isComment) . T.lines
     fixEmpties s = if T.unpack s =~ ("^\\s*$" :: String) then "" else s
 
 addPars :: Text -> Text
-addPars = newlines . {- addPar . -} T.lines
+addPars = newlines . addPar . T.lines
   where
     addPar :: [Text] -> [Text]
     addPar = intercalate ["\n<!-- -->","\\par","<!-- -->\n"] . splitOn [""]
@@ -189,6 +191,6 @@ mapAllButLast _ [a] = [a]
 mapAllButLast f (x:xs) = f x:mapAllButLast f xs
 
 addAlignEndings :: [Text] -> [Text]
-addAlignEndings = {- mapAllButLast -} map $ \ t -> T.unwords [t, "\\\\"]
+addAlignEndings = mapAllButLast {- map -} $ \ t -> T.unwords [t, "\\\\"]
 
 -- }}}

### 2014-11-12 writeup/Process.hs
added support for raw characters and rearranged macro expansion

@@ -27,44 +27,30 @@ data MatchMode =
   | Anywhere  -- "+" will match in "a+b"
   deriving (Read, Show)
 
-mathttNames :: [Text]
-mathttNames =
-    map (snd . head)
-  $ tableRows 
-  $ unsafePerformIO 
-  $ parseTableIO "Process_mathttNames.tbl"
+extractMacro :: [(Text, Text)] -> (Text, Text, MatchMode)
+extractMacro row = fromJust $ do
+  s <- lookup "Search For" row
+  r <- lookup "Replace With" row
+  let r' = if r == "_" then s else r
+  m <- liftM (read . T.unpack) $ lookup "Match Mode" row
+  return (s, r', m)
 
-mathttMacros :: [(Text,Text,MatchMode)]
-mathttMacros = flip map mathttNames $ \ n ->
-  (n, T.concat ["\\operatorname{\\mathtt{", n, "}}" ], Word)
+getMacros :: FilePath -> [(Text, Text, MatchMode)]
+getMacros = map extractMacro . tableRows . unsafePerformIO . parseTableIO
 
-mathitNames :: [Text]
-mathitNames =
-    map (snd . head)
-  $ tableRows
-  $ unsafePerformIO
-  $ parseTableIO "Process_mathitNames.tbl"
+alterMacro :: (Text -> Text) -> (Text, Text, MatchMode) -> (Text, Text, MatchMode)
+alterMacro f (s,r,m) = (s,f r,m)
 
-mathitMacros :: [(Text,Text,MatchMode)]
-mathitMacros = flip map mathitNames $ \ n ->
-  (n, T.concat ["\\mbox{\\emph{", n, "}}" ], Word)
-
-macros :: [(Text,Text,MatchMode)]
-macros = 
-    map extractMacro
-  $ tableRows
-  $ unsafePerformIO 
-  $ parseTableIO "Process_macros.tbl"
-  where
-    extractMacro :: [(Text, Text)] -> (Text, Text, MatchMode)
-    extractMacro row = fromJust $ do
-      s <- lookup "Search For" row
-      r <- lookup "Replace With" row
-      m <- liftM (read . T.unpack) $ lookup "Match Mode" row
-      return (s, r, m)
+wrapMacro :: Text -> (Text, Text, MatchMode) -> (Text, Text, MatchMode)
+wrapMacro c = alterMacro $ \ r -> T.concat [ "\\" , c ,  "{" , r , "}" ]
 
[…]

### 2014-11-13 writeup/pldi15.tex
reworked semantics. continuations are now allocated. description of effects is done abstractly and concrete instances are given later.
unicode fixes.  abstract interpreter briefly described. most of the content is in. now to make it pretty :)
on the way to automating markdown -> latex
such automation. much amaze.
Typesetting.
last leg of typesetting
No copyright space.
fixed font, prose in progress
Cut typesetting.
fleshed out recovering interpreters section a little more
adding latex commands in markdown is now supported.
progress
...
added support for raw characters and rearranged macro expa

@@
-% any author declaration will be ignored  when using 'plid' option (for double blind review)
-\documentclass[pldi]{sigplanconf}
+%lambdlambdaa any author declaration will be ignored  when using 'plid' option (for double blind review)
+\documentclass[pldi,nocopyrightspace]{sigplanconf}
 
@@
 \usepackage{enumitem}      % adjust spacing in enums
+\usepackage{amsmath}
 \usepackage[colorlinks=true,allcolors=blue,breaklinks,draft=false]{hyperref}   % hyperlinks, including DOIs and URLs in bibliography
@@
 
-\usepackage{fontspec}
-\setmonofont[Scale=0.7]{DejaVu Sans Mono}
+\usepackage{amsmath}
+\usepackage{amsthm}
+\usepackage{amsfonts}
+\usepackage{amssymb}
+\usepackage{mathtools}
+\usepackage{bm}
+\usepackage{stmaryrd}
+\usepackage{galois}
+\usepackage{float}
+\floatstyle{boxed}
+\restylefloat{figure}
+
+\newtheorem{proposition}{Proposition}
+\newtheorem{corollary}{Corollary}
+\newtheorem{definition}{Definition}
+
+% This makes newlines not introduce paragraphs.
+% Paragraphs must be explicitly marked with \par
+% WHAT WIZARDRY IS THIS
+\endlinechar=9\relax
 
@@
 
+\newcommand\litM{M}
+\newcommand\litTime{Time}
+\newcommand\litStore{Store}
+\newcommand\litVal{Val}
+
+\newcommand{\itop}[1]{\operatorname{\mathit{#1}}}
+\newcommand{\ttbfop}[1]{\operatorname{\mathtt{\boldsymbol{\mathbf{#1}}}}}
+\newcommand{\ttbfbin}[1]{\mathbin{\mathtt{\mathbf{\boldsymbol{#1}}}}}
+\newcommand{\ttop}[1]{\operatorname{\mathtt{#1}}}
+\newcommand{\Concrete}[1]{\mathtt{\mathbf{#1}}}
+\newcommand{\Abstract}[1]{\widehat{\mathtt{\mathbf{#1}}}}
+
 % }-}
@@
 
-\title{Modular Metatheory for Abstract Interpreters}
+% These are supposed to make align* not have breaks around it...
+\setlength{\abovedisplayskip}{0em}
+\setlength{\belowdisplayskip}{0em}
+\setlength{\abovedisplayshortskip}{0em}
+\setlength{\belowdisplayshortskip}{0em}
+
+%\title{The Marriage of Monad Transformers and Abstract Interpreters: Modular Metatheory for Program Analysis}
+\title{Monad Transformers and Modular Abstract Interpreters:\\ Reusable Metatheory for Program Analysis}
+\authorinfo{David Darais}{Harvard University, USA}{darais@seas.harvard.edu}
+\authorinfo{Matthew Might}{University of Utah, USA}{might@cs.utah.edu}
+\authorinfo{David Van Horn}{University of Maryland, USA}{dvanhorn@cs.umd.edu}
 \maketitle
 
-Abstract {-{
+% Abstract {-{
 \begin{abstract}
@@
 \end{abstract}
-}-}
+% }-}
 
@@
 
+\bibliographystyle{abbrvnat}
+\bibliography{dvanhorn,davdar}
+
 \end{document}

### 2014-11-13 writeup/pldi15.tex
Reworking abstract.

@@
 \usepackage{float}
+\usepackage{latexsym}
 \floatstyle{boxed}
@@
 %\title{The Marriage of Monad Transformers and Abstract Interpreters: Modular Metatheory for Program Analysis}
-\title{Monad Transformers and Modular Abstract Interpreters:\\ Reusable Metatheory for Program Analysis}
+\title{Galois Transformers and Modular Abstract Interpreters:\\ Reusable Metatheory for Program Analysis}
 \authorinfo{David Darais}{Harvard University, USA}{darais@seas.harvard.edu}
@@
 
-The design and implementation of static analyzers have becoming increasingly systematic.
-In fact, design and implementation have remained seemingly on the verge of full mechanization for several years.
-A stumbling block in full mechanization has been the ad hoc nature of soundness proofs accompanying each analyzer.
-While design and implementation is largely systematic, 
-  soundness proofs can change significantly with (apparently) minor changes to the semantics and analyzers themselves.
-We finally reconcile the systematic construction of static analyzers with their proofs of soundness 
-  via a mechanistic Galois-connection-based metatheory for static analyzers.
+The design and implementation of static analyzers have becoming
+increasingly systematic.  In fact, design and implementation have
+remained seemingly on the verge of full mechanization for several
+years.  A stumbling block in full mechanization has been the ad hoc
+nature of soundness proofs accompanying each analyzer.  While design
+and implementation is largely systematic, soundness proofs can change
+significantly with (apparently) minor changes to the semantics and
+analyzers themselves.
 
+\par We solve the problem of systematically constructing static
+analyzers by introducing the notion of a \emph{Galois transformer}, a
+monad transformer that satisfies the properties of a Galois
+connection.  In concert with a monadic interpreter, we are able to
+define a library of monad transfomers implementing classic context,
+path, and heap (in-)senstive building blocks which can be composed
+together to realize a program analysis monad, \emph{independent of the
+  language being analyzed}.
+
+\par Significantly, a Galois transformer can be proved sound once and
+for all, making it a reusable analysis component.  As new analaysis
+features and abstractions are developed and mixed in, soundness proofs
+need not be reconstructed as the composition of a monad transfomer
+stack is sound by virtue of its consituents.  We believe Galois
+transformers provide a viable foundation for resuable and composable 
+metatheory for program analysis.
 \end{abstract}

### 2014-11-13 writeup/pldi15.tex
comma

@@
 features and abstractions are developed and mixed in, soundness proofs
-need not be reconstructed as the composition of a monad transfomer
+need not be reconstructed, as the composition of a monad transfomer
 stack is sound by virtue of its consituents.  We believe Galois

### 2014-11-13 writeup/pldi15.tex
Spelling in abstract.

--- 2014-11-13 writeup/pldi15.tex
@@
 connection.  In concert with a monadic interpreter, we are able to
-define a library of monad transfomers implementing classic context,
-path, and heap (in-)senstive building blocks which can be composed
+define a library of monad transformers implementing classic context,
+path, and heap (in-)sensitive building blocks which can be composed
 together to realize a program analysis monad, \emph{independent of the
@@
 \par Significantly, a Galois transformer can be proved sound once and
-for all, making it a reusable analysis component.  As new analaysis
+for all, making it a reusable analysis component.  As new analysis
 features and abstractions are developed and mixed in, soundness proofs
-need not be reconstructed as the composition of a monad transfomer
-stack is sound by virtue of its consituents.  We believe Galois
-transformers provide a viable foundation for resuable and composable 
+need not be reconstructed as the composition of a monad transformer
+stack is sound by virtue of its constituents.  We believe Galois
+transformers provide a viable foundation for reusable and composable 
 metatheory for program analysis.

### 2014-11-13 writeup/pldi15.tex

@@
 features and abstractions are developed and mixed in, soundness proofs
-need not be reconstructed as the composition of a monad transformer
-stack is sound by virtue of its constituents.  We believe Galois
-transformers provide a viable foundation for reusable and composable 
+need not be reconstructed, as the composition of a monad transfomer
+stack is sound by virtue of its consituents.  We believe Galois
+transformers provide a viable foundation for resuable and composable 
 metatheory for program analysis.

### 2014-11-13 writeup/pldi15.tex
todos and edits
build was broken; fixed now
Updated abstract.

--- 2014-11-13 writeup/pldi15.tex
@@
 \newcommand\litVal{Val}
+\newcommand\litKAddr{KAddr}
+\newcommand\litKStore{KStore}
+\newcommand\litEnv{Env}
+\newcommand\litClo{Clo}
 
@@
 The design and implementation of static analyzers have becoming
-increasingly systematic.  In fact, design and implementation have
-remained seemingly on the verge of full mechanization for several
-years.  A stumbling block in full mechanization has been the ad hoc
-nature of soundness proofs accompanying each analyzer.  While design
-and implementation is largely systematic, soundness proofs can change
-significantly with (apparently) minor changes to the semantics and
-analyzers themselves.
+increasingly systematic.  In fact, for large classes of analyzers, design and
+implementation have remained seemingly (and now stubbornly) on the verge of full mechanization
+for several years.  A stumbling block in full mechanization has been the
+\emph{ad hoc} nature of soundness proofs accompanying each analyzer.
+While design and implementation is largely systematic, soundness proofs
+can change significantly with what seem like minor changes to the
+semantics and analyzers themselves.  An achievement of this work is to systematize, parmaterize and modularize the proofs of soundness, so as to make them composoble across analytic properties.
 
@@
 connection.  In concert with a monadic interpreter, we are able to
-define a library of monad transfomers implementing classic context,
-path, and heap (in-)senstive building blocks which can be composed
+define a library of monad transformers that implement and represent classic context-,
+path-, and heap-(in-?)sensitive building blocks.  Moreover, these can be composed
 together to realize a program analysis monad, \emph{independent of the
@@
 \par Significantly, a Galois transformer can be proved sound once and
-for all, making it a reusable analysis component.  As new analaysis
+for all, making it a reusable analysis component.  As new analysis
 features and abstractions are developed and mixed in, soundness proofs
 need not be reconstructed, as the composition of a monad transfomer
-stack is sound by virtue of its consituents.  We believe Galois
+stack is sound by virtue of its constituents.  Galois
 transformers provide a viable foundation for resuable and composable 
 metatheory for program analysis.
+
+Finally, these Galois transformers shift the level of abstraction in analysis design and implementation
+to a level where non-specialists have the ability to synthesize sound analyzers
+over a number of parameters.
 \end{abstract}

### 2014-11-13 writeup/pldi15.tex
small edits to abstract

@@
 
-The design and implementation of static analyzers have becoming
-increasingly systematic.  In fact, for large classes of analyzers, design and
-implementation have remained seemingly (and now stubbornly) on the verge of full mechanization
+The design and implementation of static analyzers have becoming increasingly
+systematic.  In fact, for large classes of analyzers, design and implementation
+have remained seemingly (and now stubbornly) on the verge of full mechanization
 for several years.  A stumbling block in full mechanization has been the
-\emph{ad hoc} nature of soundness proofs accompanying each analyzer.
-While design and implementation is largely systematic, soundness proofs
-can change significantly with what seem like minor changes to the
-semantics and analyzers themselves.  An achievement of this work is to systematize, parmaterize and modularize the proofs of soundness, so as to make them composoble across analytic properties.
+\emph{ad hoc} nature of soundness proofs accompanying each analyzer. While
+design and implementation is largely systematic, soundness proofs can change
+significantly with what seem like minor changes to the semantics and analyzers
+themselves.  An achievement of this work is to systematize, parameterize and
+modularize the proofs of soundness, so as to make them composable across
+analytic properties.
 
-\par We solve the problem of systematically constructing static
-analyzers by introducing the notion of a \emph{Galois transformer}, a
-monad transformer that satisfies the properties of a Galois
-connection.  In concert with a monadic interpreter, we are able to
-define a library of monad transformers that implement and represent classic context-,
-path-, and heap-(in-?)sensitive building blocks.  Moreover, these can be composed
-together to realize a program analysis monad, \emph{independent of the
-  language being analyzed}.
+\par
 
-\par Significantly, a Galois transformer can be proved sound once and
-for all, making it a reusable analysis component.  As new analysis
-features and abstractions are developed and mixed in, soundness proofs
-need not be reconstructed, as the composition of a monad transfomer
-stack is sound by virtue of its constituents.  Galois
-transformers provide a viable foundation for resuable and composable 
-metatheory for program analysis.
+We solve the problem of systematically constructing static analyzers by
+introducing the notion of a \emph{Galois transformer}, a monad transformer that
+transports Galois connection properties.  In concert with a monadic
+interpreter, we are able to define a library of monad transformers that
+implement and represent classic context-, path-, and heap-(in-)sensitive
+building blocks.  Moreover, these can be composed together to realize a program
+analysis monad, \emph{independent of the language being analyzed}.
 
-Finally, these Galois transformers shift the level of abstraction in analysis design and implementation
-to a level where non-specialists have the ability to synthesize sound analyzers
-over a number of parameters.
+\par
+
+Significantly, a Galois transformer can be proved sound once and for all,
+making it a reusable analysis component.  As new analysis features and
+abstractions are developed and mixed in, soundness proofs need not be
+reconstructed, as the composition of a monad transformer stack is sound by
+virtue of its constituents.  Galois transformers provide a viable foundation
+for reusable and composable metatheory for program analysis.
+
+\par
+
+Finally, these Galois transformers shift the level of abstraction in analysis
+design and implementation to a level where non-specialists have the ability to
+synthesize sound analyzers over a number of parameters.
+
 \end{abstract}

### 2014-11-13 writeup/pldi15.tex
edits
Balance bib.
hangers

@@
 \usepackage{latexsym}
+\usepackage{balance}
 \floatstyle{boxed}
 \restylefloat{figure}
+\usepackage[font=small,skip=0pt]{caption}
 
@@
 %\title{The Marriage of Monad Transformers and Abstract Interpreters: Modular Metatheory for Program Analysis}
-\title{Galois Transformers and Modular Abstract Interpreters:\\ Reusable Metatheory for Program Analysis}
+\title{Galois Transformers and Modular Abstract Interpreters}
+\subtitle{Reusable Metatheory for Program Analysis}
 \authorinfo{David Darais}{Harvard University, USA}{darais@seas.harvard.edu}
@@
 design and implementation is largely systematic, soundness proofs can change
-significantly with what seem like minor changes to the semantics and analyzers
-themselves.  An achievement of this work is to systematize, parameterize and
-modularize the proofs of soundness, so as to make them composable across
-analytic properties.
+significantly with seemingly minor changes to the semantics or analyzers.  An
+achievement of this work is to systematize, parameterize and modularize the
+proofs of soundness, so as to make them composable across analytic properties.
 
@@
 We solve the problem of systematically constructing static analyzers by
-introducing the notion of a \emph{Galois transformer}, a monad transformer that
+introducing \emph{Galois transformers}: monad transformers that
 transports Galois connection properties.  In concert with a monadic
-interpreter, we are able to define a library of monad transformers that
-implement and represent classic context-, path-, and heap-(in-)sensitive
-building blocks.  Moreover, these can be composed together to realize a program
-analysis monad, \emph{independent of the language being analyzed}.
+interpreter, we define a library of monad transformers that
+implement building blocks for classic analysis parameters like context-, path-,
+and heap-(in-)sensitivity. Moreover, these can be composed together 
+% to realize a program analysis monad, 
+\emph{independent of the language being analyzed}.
 
@@
 
+\balance
 \bibliographystyle{abbrvnat}

### 2014-12-13 cs252r-fa14-writeup/paper.tex
Paper stub

\documentclass[10pt,letter,english]{article}

\usepackage{amsmath}
\usepackage{amscd}
\usepackage{amssymb}

\usepackage{tikz}
\usetikzlibrary{arrows}

\usepackage[sc,osf]{mathpazo}
\linespread{1.05}         % Palatino needs more leading (space between lines)
\usepackage[T1]{fontenc}
\fontencoding{T1}
\fontfamily{ppl}
\fontseries{m}
\fontshape{n}
\fontsize{10}{13}
\selectfont

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% Set Theory Concepts

\newcommand{\buildset}[2]{\{ #1 \;|\; #2 \}}
\newcommand{\compose}[0]{\circ}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% Textual Macros

\newcommand{\aam}[0]{\textsc{aam}}
\newcommand{\maam}[0]{\textsc{maam}}
\newcommand{\js}[0]{JavaScript}
\newcommand{\lambdajs}[0]{$\lambda_{JS}$}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{document}

\title{MJSAI: Designing a Sound, Configurable, Efficient, and Pushdown %
              Monadic Static~Analyzer for JavaScript}
\author{{David Darais and Daniel King}\\
        School of Engineering and Applied Sciences\\
        Harvard University}

\maketitle

\section*{Abstract}

We present the design and implementation of a static analyzer for \js{}. We
leverage the Monadic Abstracting Abstract Machines (\maam{}) technique to create
an analysis which is both configurable and correct by construction. We recast
pushdown static analysis in terms of \aam{}. This recasting permits us to also
create a pushdown \maam{} analysis for \js{}.

\section{Introduction}

This work had three goals:

\begin{itemize}
\item design a configurable static analysis using the Monadic Abstracting
  Abstract Machines (\maam{}) technique
\item implement the analysis, again using the \maam{} technique
\item analyze the performance of many different analysis techniques, including
  pushdown analysis, against the JSAI \cite{jsai} benchmark suite
\end{itemize}

We completed most of the first two goals; however, we have not analyzed the
performance of the implementation. We underestimated the time necessary to
implement the JavaScript semantics in the \maam{} Huskily framework. We are also
currently unable to scale the analysis to real \js{} programs. Finally, we spent
a not insignificant amount of time understanding what pushdown analysis is and
how it can be implemented in the \aam{} and \maam{} frameworks.

The remainder of the paper presents the design and implementation of the
analysis, including subsections on our understanding of pushdown analysis.

\section{Design}
\subsection{Pushdown Analysis in \maam{}}

Pushdown analysis means different things to different people. There is a
significant amount of work on first-order pushdown semantics, but we focus on
understanding the higher-order setting. There are primarily two papers which
describe a pushdown analysis for a higher-order setting: Vardoulakis and
Shivers' \emph{CFA2} \cite{cfa2} and Earl, Might, and Van Horn's \emph{PDCFA}
\cite{pdcfa}. Both papers share the idea of \emph{precise call-return matching}
and some form of \emph{memoization}. The former improves the precision and the
latter improves the time-efficiency of the analysis. The \emph{CFA2} analysis
also precisely tracks so-called \emph{stack} variables. We will not describe
this technique in detail other than to say that it involves identifying
variables which can safely live on the call stack and tracking their values
exactly. Unfortunately, even if one ignores the special \emph{stack} variables,
it is not obvious that the two techniques describe the same style of
analysis. The presentations styles differ dramatically: \emph{CFA2} describes a
Cousot \& Cousot style Abstract Interpretation whereas \emph{PDCFA} describes
the conversion of an abstract machine into a Dyck state graph.

We propose a unification of these two approaches using the \aam{} technique of
Might and Van Horn \cite{aam}. There are two simple ideas that lead to a
pushdown style analysis: a separate, path-sensitive store for continuations and
a memo table for function calls. We claim, without proof, that the
path-sensitive continuation-store provides \emph{precise call-return matching}.
The memo table for function calls clearly provides \emph{memoization}.

We can now describe a AAM-style approach for producing a pushdown abstract
machine from a concrete CEK machine:
\begin{enumerate}
\item Break recursion in the environment, caused by closures, with a store,
  which maps addresses to values, and change the environment to map variables to
  addresses.
\item Break recursion in the continuation stack by creating a
  continuation-store, which maps continuation-addresses to either the halt frame
  or a pair of a frame and a continuation-address.
\item Add a memo table which maps a triplets of a command, an environments, and
  a store to a value.
\end{enumerate}

\begin{figure}
\centering
% \begin{tikzpicture}[->,>=stealth',shorten >=1pt,auto,node distance=3cm,
%   thick,main node/.style={circle,fill=blue!20,draw,font=\sffamily\Large\bfseries}]

%   \node[main node] (1) {\cek{c}{$\rho$}{$\kappa$}};
%   \node[main node] (2) [right of = 1] {\cesk{c}{$\rho$}{$\sigma$}{$\kappa$}};
%   \node[main node] (3) [right of = 2] {\cessk{c}{$\rho$}{$\sigma$}{$\kappa\sigma$}{$\kappa$}};
%   \node[main node] (4) [right of = 3] {\cesskc{c}{$\rho$}{$\sigma$}{$\kappa\sigma$}{$\kappa$}{$\Chi$}};

%   \path[->]
%     (1) edge (2) edge (3) edge (4);

% \end{tikzpicture}
\caption{An AAM-style approach for producing a pushdown abstract machine} \label{fig:M1}
\end{figure}

\subsection{A \js{} Analysis in \maam{}}

\js{} has a large, complex specification and many implementations which
practically define the expected semantics of \js{}. Rather than describe and
implement our own semantics, we implement the \lambdajs{} semantics as described
in \cite{lambdajs}. Additionally, we reused their desugaring program.

We reformulated the \lambdajs{} semantics into a reduction machine
semantics. The \lambdajs{} evaluation contexts became our \emph{Frame} data
type. Additionally, we do not divide the evaluation contexts into two sets and
use syntactic stepping rules to implement exceptional control flow. We choose to
implement exceptional control flow with meta functions that crawl the stack
looking for appropriate return points.

\section{Implementation}

\section{Related Work}

\section{Conclusion}

\bibliography{paper}{}
\bibliographystyle{acm}

\end{document}

### 2014-12-13 cs252r-fa14-writeup/paper.tex
Add a couple macros and fix up section headings
finish picture

@@
 
+\newcommand{\partto}{\rightharpoonup}
+
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 \newcommand{\lambdajs}[0]{$\lambda_{JS}$}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%% Machine Configurations
+
+\newcommand{\cek}[3]{\langle #1, #2, #3 \rangle}
+\newcommand{\cesk}[4]{\langle #1, #2, #3, #4 \rangle}
+\newcommand{\cessk}[5]{\langle #1, #2, #3, #4, #5 \rangle}
+\newcommand{\cesskc}[6]{\langle #1, #2, #3, #4, #5, #6 \rangle}
 
@@
 \section{Design}
-\subsection{Pushdown Analysis in \maam{}}
+\subsection{Pushdown Analysis in MAAM}
 
@@
 
-We can now describe a AAM-style approach for producing a pushdown abstract
+We can now describe a \aam{}-style approach for producing a pushdown abstract
 machine from a concrete CEK machine:
@@
 \centering
-% \begin{tikzpicture}[->,>=stealth',shorten >=1pt,auto,node distance=3cm,
-%   thick,main node/.style={circle,fill=blue!20,draw,font=\sffamily\Large\bfseries}]
 
-%   \node[main node] (1) {\cek{c}{$\rho$}{$\kappa$}};
-%   \node[main node] (2) [right of = 1] {\cesk{c}{$\rho$}{$\sigma$}{$\kappa$}};
-%   \node[main node] (3) [right of = 2] {\cessk{c}{$\rho$}{$\sigma$}{$\kappa\sigma$}{$\kappa$}};
-%   \node[main node] (4) [right of = 3] {\cesskc{c}{$\rho$}{$\sigma$}{$\kappa\sigma$}{$\kappa$}{$\Chi$}};
+\begin{equation*}
+\begin{array}{c c l c l}
+  c            &:& Exp        && \\
+  \rho         &:& Env        &=& Var   \partto Addr \\
+  \sigma       &:& Store      &=& Addr  \partto Value \\
+  \kappa\sigma &:& KStore     &=& KAddr \partto Frame \times KAddr \\
+  \kappa       &:& KAddr      && \\
+  \chi         &:& Cache      &=& Exp \times Store \times Env \partto Value
+\end{array}
+\end{equation*}
 
-%   \path[->]
-%     (1) edge (2) edge (3) edge (4);
+\begin{tikzpicture}[->,>=stealth',shorten >=1pt,auto,node distance=3.5cm,
+  thick,main node/.style={font=\sffamily\bfseries}]
 
-% \end{tikzpicture}
+  \node[main node] (101) at (0, 1)
+    {$\cek{c}{\rho}{\kappa}$};
+  \node[main node] (102) [right of = 101]
+    {$\cesk{c}{\rho}{\sigma}{\kappa}$};
+  \node[main node] (103) [right of = 102]
+    {$\cessk{c}{\rho}{\sigma}{\kappa\sigma}{\kappa}$};
+  \node[main node] (104) [right of = 103]
+    {$\cesskc{c}{\rho}{\sigma}{\kappa\sigma}{\kappa}{\chi}$};
+
+  \path[->]
+    (101) edge node {Step 1} (102)
+    (102) edge node {Step 2} (103)
+    (103) edge node {Step 3} (104);
+
+\end{tikzpicture}
 \caption{An AAM-style approach for producing a pushdown abstract machine} \label{fig:M1}
@@
 
-\subsection{A \js{} Analysis in \maam{}}
+\subsection{A JavaScript Analysis in MAAM}
 

### 2014-12-13 cs252r-fa14-writeup/paper.tex
Some related work on lambda js

@@
 
+\subsection{The Essence of JavaScript: \lambdajs{}}
+
+The \lambdajs{} work provides a complete, practical foundation to perform
+research on real-world JavaScript programs. The \lambdajs{} work identifies a
+simplified language that captures the essential features of \js{}. The
+\lambdajs{} language is the composition of four distinct semantics: function and
+object semantics, mutable references, prototypal inheritance, and control
+operators. In the authors' experience, implementing this semantics is time
+consuming, but mindless. The complexity of \js{} is not apparent in the
+\lambdajs{} model.
+
+The complexity of \js{} is fully captured by a desugaring phase which replaces
+certain forms, such as \texttt{with}, with simpler \js{} code. Additionally, the
+\lambdajs{} team has an implementation of the \js{} standard library which
+abstracts away the details of string and number coercions. These coercions are
+specified by many pages of non-trivial pseudo-code in the ECMAScript 3rd Edition
+standard.
+
 \section{Conclusion}

### 2014-12-13 cs252r-fa14-writeup/paper.tex
Cleanup whitespace
Describe implementation in a picture

@@
 
+\begin{figure}
+\centering
+
+\begin{tikzpicture}[->,>=stealth',shorten >=1pt,auto,node distance=5cm,
+  thick,main node/.style={font=\sffamily\bfseries}]
+
+  \node[main node] (101) at (0, 1)
+    {JS Code};
+  \node[main node] (102) [right of = 101]
+    {\lambdajs{} Code};
+  \node[main node] (103) [right of = 102]
+    {Analysis Results};
+
+  \path[->]
+    (101) edge node {Parse \& Desugar} (102)
+    (102) edge node {\maam{} Analyzer} (103);
+
+\end{tikzpicture}
+\caption{The Architecture of Our Analysis System. The second arrow is original work.}
+\end{figure}
+
 \section{Related Work}
-
 \subsection{The Essence of JavaScript: \lambdajs{}}

### 2014-12-13 cs252r-fa14-writeup/paper.tex
writing stuff for steve

@@
 
+\newtheorem{proposition}[theorem]{Proposition}
+
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 We present the design and implementation of a static analyzer for \js{}. We
-leverage the Monadic Abstracting Abstract Machines (\maam{}) technique to create
-an analysis which is both configurable and correct by construction. We recast
-pushdown static analysis in terms of \aam{}. This recasting permits us to also
-create a pushdown \maam{} analysis for \js{}.
+leverage the Monadic Abstracting Abstract Machines (\maam{}) technique to
+create an analysis which is both configurable and correct by construction. We
+also recast pushdown static analysis in terms of \aam{}. This recasting permits
+us to also easily create a pushdown \maam{} analysis for \js{}.
 
@@
 \begin{itemize}
-\item design a configurable static analysis using the Monadic Abstracting
-  Abstract Machines (\maam{}) technique
-\item implement the analysis, again using the \maam{} technique
-\item analyze the performance of many different analysis techniques, including
-  pushdown analysis, against the JSAI \cite{jsai} benchmark suite
+\item Design a configurable static analysis for javascript using the
+  Monadic Abstracting Abstract Machines (\maam{}, \cite{maam}) design process.
+\item Implement the analysis using the \maam{} framework as implemented in Haskell.
+\item Analyze the performance of many different analysis techniques, including
+  pushdown analysis, against the JSAI \cite{jsai} benchmark suite.
 \end{itemize}
 
-We completed most of the first two goals; however, we have not analyzed the
-performance of the implementation. We underestimated the time necessary to
-implement the JavaScript semantics in the \maam{} Huskily framework. We are also
-currently unable to scale the analysis to real \js{} programs. Finally, we spent
-a not insignificant amount of time understanding what pushdown analysis is and
-how it can be implemented in the \aam{} and \maam{} frameworks.
+We completed the first goal, part of the second, and none of the third.
+%
+We underestimated the time necessary to implement the semantics of \js{} in the
+\maam{} Haskell framework.
+%
+We were unable to complete a full implementation of the \js{} spec and
+therefore unable to run our interpreter on the full JSAI~\cite{jsai} benchmark
+suite.
+%
+However, we were successful in running our analysis on programs which reside in
+the subset of \js{} that we were able to implement.
+%
+We also spent a significant amoutn of time understanding pushdown analysis and
+recasting it in the \aam{} and \maam frameworks.
 
@@
 pushdown style analysis: a separate, path-sensitive store for continuations and
-a memo table for function calls. We claim, without proof, that the
+a memo table for caching analysis results. We claim, without proof, that the
 path-sensitive continuation-store provides \emph{precise call-return matching}.
-The memo table for function calls clearly provides \emph{memoization}.
 
-We can now describe a \aam{}-style approach for producing a pushdown abstract
-machine from a concrete CEK machine:
+We now describe an AAM-style approach to pushdown static analysis for a standard CEK machine.
+%
+Recall that a CEK machine is a state machine with three components:
+%
 \begin{enumerate}
-\item Break recursion in the environment, caused by closures, with a store,
-  which maps addresses to values, and change the environment to map variables to
-  addresses.
-\item Break recursion in the continuation stack by creating a
-  continuation-store, which maps continuation-addresses to either the halt frame
-  or a pair of a frame and a continuation-address.
-\item Add a memo table which maps a triplets of a command, an environments, and
-  a store to a value.
+\item The control expression $(c : Exp)$
+\item The execution environment $(\rho : Var \rightarrow Val)$
+\item The continuation $(k : List Frame)$
 \end{enumerate}
+%
+In AAM-style, we first design an instrument concrete semantics.
+%
+This instrumented concrete semantics will yield a semantically equivalent
+concrete interpreter, but with implementation choices which, after abstraction,
+will yield an analysis with different properties.
+%
+To instrument the semantics for a pushdown analysis we:
+\begin{enumerate}
+\item Break the recursion between environments and values a. la. \citet{aam}
+\item Break the recursion between cons-cells in the list type for continuations.
+%
+This step is also in the discipline of \citet{aam}, although this particular
+choice is not made in their original paper.
+%
+\item Add a memo-table which caches the results of the analysis.
+%
+This memo-table caches results \emph{invariant to the continuation present during evaluation}.
+\end{enumerate}
+
+From steps 1. and 2. we end up with a C-E-S-KS-K* machine with the following components:
+\item The control expression $(c : Exp)$ as before
+\item The environment $(\rho : Var \rightarrow Addr)$
+\item The store $(\sigma : Addr \rightarrow Val)$
+\item The continuation store $(\kappa\sigma : KAddr \rightarrow \mathcal{P}(Frame + \{Null\}))$
+\item The continuation pointer $(\kappa-ptr : KAddr)$
+
+\begin{proposition}
+A step relation
+\begin{equation*}
+\mapsto^{PD} \in CESKSK* \times CESKSK* 
+\end{equation*}
+is definable which agrees with the standard semantics
+\begin{equation*}
+\mapsto \in CEK \times CEK 
+\end{equation*}
+\end{proposition}
+
+Step 3. yields a larger machine, C-E-S-KS-K*-\$ machine, which adds a new component:
+\begin{equation*}
+\$ : C \times E \times S \rightarrow Val + \{SEEN\} + \{UNSEEN\}
+\end{equation*}
+%
+The fact that the cache can memoize invariant to the continuation is
+traditionally understoon in the pushdown literature as by-product of viewing
+the semantics a pushdown automata.
+%
+We propose another intuition: the CEK machine comes from a \emph{reduction}
+semantics--a setting where reduction is defined invariant to context to begin
+with.
 

### 2014-12-13 src/Lang/JS/Main.hs

@@ -21,5 +21,5 @@ main = do
   jsFiles <- filesFromDirM "tinytests"
   traverseOn jsFiles $ \ jsFile -> do
     e <- fromFile jsFile
-    pprint $ execM e
-    --pprint e
+    --pprint $ execM e
+    pprint e

### 2014-12-13 src/Lang/JS/Pretty.hs

@@ -20,10 +20,15 @@ instance (Pretty e) => Pretty (PreExp e) where
   pretty (Func xs e) = pretty $ VarLam xs e
   pretty (ObjE xes) = pretty $ toMap xes
   pretty (Let xes e) = P.atLevel 0 $ P.hvsep
-    [ P.hsep [ P.key "let" 
-             , P.collection "[" "]" "," $ mapOn xes $ \ (x,e') -> P.hsep [ pretty x, P.keyPun ":=", P.botLevel $ P.align $ pretty e' ]
-             , P.key "in" 
-             ]
+    [ P.hsep 
+        [ P.key "let" 
+        , P.collection "[" "]" "," $ mapOn xes $ \ (x,e') -> P.hsep 
+            [ pretty x
+            , P.keyPun ":="
+            , P.botLevel $ P.align $ pretty e' 
+            ]
+        , P.key "in" 
+        ]
     , P.align $ pretty e
     ]
   pretty (App e es) = P.app $ pretty e : map pretty es
@@ -274,9 +279,9 @@ instance Pretty Obj where
 
 instance Pretty AValue where
   pretty (LitA l) = pretty l
-  pretty NumA = P.con "ℝ"
-  pretty StrA = P.con "S"
   pretty (CloA c) = pretty c
   pretty (ObjA o) = pretty o
   pretty (LocA l) = pretty l
-
+  pretty BoolA = P.con "𝔹"
+  pretty StrA  = P.con "𝕊"
+  pretty NumA  = P.con "ℝ"

### 2014-12-13 cs252r-fa14-writeup/paper.tex
bibtex stuff
Some stuff on implementation

@@
 
+\newtheorem{theorem}{Theorem}[section]
 \newtheorem{proposition}[theorem]{Proposition}
@@
 \begin{enumerate}
-\item Break the recursion between environments and values a. la. \citet{aam}
+\item Break the recursion between environments and values a. la. AAM \cite{aam}
 \item Break the recursion between cons-cells in the list type for continuations.
 %
-This step is also in the discipline of \citet{aam}, although this particular
+This step is also in the discipline of AAM \cite{aam}, although this particular
 choice is not made in their original paper.
@@
 From steps 1. and 2. we end up with a C-E-S-KS-K* machine with the following components:
+\begin{enumerate}
 \item The control expression $(c : Exp)$ as before
@@
 \item The continuation pointer $(\kappa-ptr : KAddr)$
+\end{enumerate}
 
@@
   \kappa       &:& KAddr      && \\
-  \chi         &:& Cache      &=& Exp \times Store \times Env \partto Value
+  \$         &:& Cache      &=& Exp \times Store \times Env \partto Value
 \end{array}
@@
   \node[main node] (104) [right of = 103]
-    {$\cesskc{c}{\rho}{\sigma}{\kappa\sigma}{\kappa}{\chi}$};
+    {$\cesskc{c}{\rho}{\sigma}{\kappa\sigma}{\kappa}{\$}$};
 
@@
 
-We reformulated the \lambdajs{} semantics into a reduction machine
-semantics. The \lambdajs{} evaluation contexts became our \emph{Frame} data
+We reformulated the \lambdajs{} semantics into a state machine with explicit
+continuations. The \lambdajs{} evaluation contexts became our \emph{Frame} data
 type. Additionally, we do not divide the evaluation contexts into two sets and
@@
 
+Our static analysis system is the composition of significant existing work. This
+exemplifies the ethos of \maam{}: implement only the semantics and derive
+analyses for free. In particular, we have four primary components:
+
+\begin{itemize}
+\item The \lambdajs{} Parser \& Desugarer
+\item The \lambdajs{} Standard Libarary
+\item A \lambdajs{} State Machine with Explicit Continuations
+\item The \maam{} Framework
+\end{itemize}
+
+The third item is the main contribution of this project.
+
+\subsection{MAAM}
+
+The \maam{} framework separates the semantics of a programming language from the
+control flow. This same implementation of \lambdajs{} can be used with different
+monads to produce path and flow sensitive analysis, as well as other kinds of
+analysis. With an appropriate set of monads, traditional points-to constraints
+could be generated.
+
+\subsection{\lambdajs{} State Machine with Explicit Continuations}
+
+We translated the deterministic, reduction semantics of \lambdajs{} directly
+into a monadic state machine with explicit continuations. The reduction
+semantics maps naturally to a state machine with explicit continuations. The
+evaluation contexts become \emph{frames} and nested reduction contexts become a
+list of frames.
+
+Adapting the deterministic semantics to a monadic semantics consists primarily
+of identifying control branching locations and inserting monadic splits at those
+points. We do not insert monadic splits at value joins as this would result in
+state space explosions. Inserting monadic splits only at control branches
+enables the monad to control path and flow sensitivity.
+
+We replaced substitution in the \lambdajs{} semantics with an
+environment. Finally, we replaced the syntactic exception handling code with a
+meta-function that searches the stack for the appropriate handler.
+
+We, the authors, found the entire translation process completely straightforward
+and mechanical, albeit time consuming due to the size of \lambdajs{} when
+including the primitive operators.
+
 \section{Related Work}

### 2014-12-13 cs252r-fa14-writeup/paper.tex
stuff

@@
 \selectfont
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%% LstListings
+
+\usepackage{listings}
+\lstset{
+  frame=none,
+  xleftmargin=2pt,
+  stepnumber=1,
+  numbers=left,
+  numbersep=5pt,
+  numberstyle=\ttfamily\tiny\color[gray]{0.3},
+  belowcaptionskip=\bigskipamount,
+  captionpos=b,
+  escapeinside={*'}{'*},
+  language=haskell,
+  tabsize=2,
+  emphstyle={\bf},
+  commentstyle=\it,
+  stringstyle=\mdseries\rmfamily,
+  showspaces=false,
+  keywordstyle=\bfseries\rmfamily,
+  columns=flexible,
+  basicstyle=\small\sffamily,
+  showstringspaces=false,
+  morecomment=[l]\%,
+  mathescape=true,
+}
 
@@
 
-\title{MJSAI: Designing a Sound, Configurable, Efficient, and Pushdown %
+\title{MJSLAAM: Designing a Sound, Configurable, Efficient, and Pushdown %
               Monadic Static~Analyzer for JavaScript}
@@
 
-We present the design and implementation of a static analyzer for \js{}. We
+We present the design and implementation of MJSLAAM, a static analyzer for \js{}. We
 leverage the Monadic Abstracting Abstract Machines (\maam{}) technique to
@@
 us to also easily create a pushdown \maam{} analysis for \js{}.
+We call our framework MJSLAAM for Monadic JavaScript Language Abstract Abstract
+Machine.
 
@@
 \item Design a configurable static analysis for javascript using the
-  Monadic Abstracting Abstract Machines (\maam{}, \cite{maam}) design process.
+  Monadic Abstracting Abstract Machines (\maam{}) \cite{maam} design process.
 \item Implement the analysis using the \maam{} framework as implemented in Haskell.
@@
 %
-However, we were successful in running our analysis on programs which reside in
-the subset of \js{} that we were able to implement.
+However, we were successful in running our analysis on small programs which
+reside in the subset of \js{} that we were able to implement.
 %
-We also spent a significant amoutn of time understanding pushdown analysis and
+We also spent a significant amount of time understanding pushdown analysis and
 recasting it in the \aam{} and \maam frameworks.
@@
 \item The execution environment $(\rho : Var \rightarrow Val)$
-\item The continuation $(k : List Frame)$
+\item The continuation $(k : List(Frame))$
 \end{enumerate}
 %
-In AAM-style, we first design an instrument concrete semantics.
+In AAM-style we first design an instrument concrete semantics.
 %
@@
 concrete interpreter, but with implementation choices which, after abstraction,
-will yield an analysis with different properties.
+yield an analysis with pushdown properties.
 %
@@
 \item The continuation store $(\kappa\sigma : KAddr \rightarrow \mathcal{P}(Frame + \{Null\}))$
-\item The continuation pointer $(\kappa-ptr : KAddr)$
+\item The continuation pointer $(\kappa{}ptr : KAddr)$
 \end{enumerate}
@@
 \begin{equation*}
-\mapsto^{PD} \in CESKSK* \times CESKSK* 
+\mapsto^{K} \in CESKSK^* \times CESKSK^* 
 \end{equation*}
@@
 
-\section{Implementation}
+\begin{figure}
+\centering
+
+\begin{lstlisting}
+data Frame =
+    LetK [(Name, Set AValue)] Name [(Name, TExp)] TExp
+  | AppL [TExp]
+  | AppR (Set AValue) [(Set AValue)] [TExp]
+  | ObjK [(String, (Set AValue))] Name [(Name, TExp)]
+    -- Array Dereferencing
+  | FieldRefL TExp
+  | FieldRefR (Set AValue)
+    -- Array Assignment
+  | FieldSetA TExp         TExp
+  | FieldSetN (Set AValue) TExp
+  | FieldSetV (Set AValue) (Set AValue)
+    -- Property Deletion
+  | DeleteL TExp
+  | DeleteR (Set AValue)
+    -- Fig 2. Mutable References
+  | RefSetL TExp
+  | RefSetR (Set AValue)
+  | RefK
+  | DeRefK
+    -- Fig 8. Control Operators
+  | IfK TExp TExp
+  | SeqK TExp
+  | WhileL TExp TExp
+  | WhileR TExp TExp
+  | LabelK Label
+  | BreakK Label
+  | TryCatchK TExp Name
+  | TryFinallyL TExp
+  | TryFinallyR (Set AValue)
+  | ThrowK
+    -- Fig 9. Primitive Operations
+  | PrimOpK Op [(Set AValue)] [TExp]
+  deriving (Eq, Ord)
+instance PartialOrder Frame where pcompare = discreteOrder
+\end{lstlisting}
+\caption{The Frame, i.e., Continuation Datatype}
+\end{figure}
 
@@
 
-\begin{tikzpicture}[->,>=stealth',shorten >=1pt,auto,node distance=5cm,
-  thick,main node/.style={font=\sffamily\bfseries}]
+\begin{lstlisting}
+pushFrame :: (Analysis $\varsigma$ m) => Frame -> m ()
+pushFrame fr = do
+  fp  <- getL konL
+  fp' <- nextFramePtr
+  modifyL kstoreL (mapInsertWith (\/) fp' (singleton (fr, fp))
+  putL konL fp'
 
-  \node[main node] (101) at (0, 1)
-    {JS Code};
-  \node[main node] (102) [right of = 101]
-    {\lambdajs{} Code};
-  \node[main node] (103) [right of = 102]
-    {Analysis Results};
+popFrame :: (Analysis $\varsigma$ m) => m Frame
+popFrame = do
+  fp <- getL konL
+  k$\sigma$ <- getL kstoreL
+  (fr, fp') <- mset (mjoin (liftMaybeSet (k$\sigma$ # fp)))
+  putL konL fp'
+  return fr
+\end{lstlisting}
+\caption{Pushing and Popping Stack Frames}
+\end{figure}
 
-  \path[->]
-    (101) edge node {Parse \& Desugar} (102)
-    (102) edge node {\maam{} Analyzer} (103);
-
-\end{tikzpicture}
-\caption{The Architecture of Our Analysis System. The second arrow is original work.}
-\end{figure}
+\section{Implementation}
 
@@
 
-We, the authors, found the entire translation process completely straightforward
-and mechanical, albeit time consuming due to the size of \lambdajs{} when
-including the primitive operators.
+We found the entire translation process completely straightforward and
+mechanical, albeit time consuming due to the size of \lambdajs{} and number of
+primitive operations.
+
+\begin{figure}
+\centering
+
+\begin{tikzpicture}[->,>=stealth',shorten >=1pt,auto,node distance=5cm,
+  thick,main node/.style={font=\sffamily\bfseries}]
+
+  \node[main node] (101) at (0, 1)
+    {JS Code};
+  \node[main node] (102) [right of = 101]
+    {\lambdajs{} Code};
+  \node[main node] (103) [right of = 102]
+    {Analysis Results};
+
+  \path[->]
+    (101) edge node {Parse \& Desugar} (102)
+    (102) edge node {\maam{} Analyzer} (103);
+
+\end{tikzpicture}
+\caption{The Architecture of Our Analysis System. The second arrow is original work.}
+\end{figure}
 

### 2014-12-13 cs252r-fa14-writeup/paper.tex
move figures apart

--- 2014-12-13 cs252r-fa14-writeup/paper.tex
@@
 
-\begin{figure}
-\centering
-
-\begin{tikzpicture}[->,>=stealth',shorten >=1pt,auto,node distance=5cm,
-  thick,main node/.style={font=\sffamily\bfseries}]
-
-  \node[main node] (101) at (0, 1)
-    {JS Code};
-  \node[main node] (102) [right of = 101]
-    {\lambdajs{} Code};
-  \node[main node] (103) [right of = 102]
-    {Analysis Results};
-
-  \path[->]
-    (101) edge node {Parse \& Desugar} (102)
-    (102) edge node {\maam{} Analyzer} (103);
-
-\end{tikzpicture}
-\caption{The Architecture of Our Analysis System. The second arrow is original work.}
-\end{figure}
-
 Our static analysis system is the composition of significant existing work. This

### 2014-12-13 cs252r-fa14-writeup/paper.tex
Add a lstlisting for data Frame

@@
 \selectfont
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%% LstListings
+
+\usepackage{listings}
+\lstset{
+  frame=none,
+  xleftmargin=2pt,
+  stepnumber=1,
+  numbers=left,
+  numbersep=5pt,
+  numberstyle=\ttfamily\tiny\color[gray]{0.3},
+  belowcaptionskip=\bigskipamount,
+  captionpos=b,
+  escapeinside={*'}{'*},
+  language=haskell,
+  tabsize=2,
+  emphstyle={\bf},
+  commentstyle=\it,
+  stringstyle=\mdseries\rmfamily,
+  showspaces=false,
+  keywordstyle=\bfseries\rmfamily,
+  columns=flexible,
+  basicstyle=\small\sffamily,
+  showstringspaces=false,
+  morecomment=[l]\%,
+}
 
@@
 
+\begin{figure}
+\centering
+
+\begin{lstlisting}
+data Frame =
+    LetK [(Name, Set AValue)] Name [(Name, TExp)] TExp
+  | AppL [TExp]
+  | AppR (Set AValue) [(Set AValue)] [TExp]
+  | ObjK [(String, (Set AValue))] Name [(Name, TExp)]
+    -- Array Dereferencing
+  | FieldRefL TExp
+  | FieldRefR (Set AValue)
+    -- Array Assignment
+  | FieldSetA TExp         TExp
+  | FieldSetN (Set AValue) TExp
+  | FieldSetV (Set AValue) (Set AValue)
+    -- Property Deletion
+  | DeleteL TExp
+  | DeleteR (Set AValue)
+    -- Fig 2. Mutable References
+  | RefSetL TExp
+  | RefSetR (Set AValue)
+  | RefK
+  | DeRefK
+    -- Fig 8. Control Operators
+  | IfK TExp TExp
+  | SeqK TExp
+  | WhileL TExp TExp
+  | WhileR TExp TExp
+  | LabelK Label
+  | BreakK Label
+  | TryCatchK TExp Name
+  | TryFinallyL TExp
+  | TryFinallyR (Set AValue)
+  | ThrowK
+    -- Fig 9. Primitive Operations
+  | PrimOpK Op [(Set AValue)] [TExp]
+  deriving (Eq, Ord)
+instance PartialOrder Frame where pcompare = discreteOrder
+\end{lstlisting}
+
+\end{figure}
+
 \section{Implementation}

### 2014-12-13 cs252r-fa14-writeup/paper.tex
writing

--- 2014-12-13 cs252r-fa14-writeup/paper.tex
@@
 
+The AAM-inspired design process for our final concrete interpreter is
+summarized in Figure 1.
+
 \begin{figure}
@@
 implement our own semantics, we implement the \lambdajs{} semantics as described
-in \cite{lambdajs}. Additionally, we reused their desugaring program.
-
-We reformulated the \lambdajs{} semantics into a state machine with explicit
-continuations. The \lambdajs{} evaluation contexts became our \emph{Frame} data
-type. Additionally, we do not divide the evaluation contexts into two sets and
-use syntactic stepping rules to implement exceptional control flow. We choose to
-implement exceptional control flow with meta functions that crawl the stack
-looking for appropriate return points.
+by Guha et. al. \cite{lambdajs}. Additionally, we reused their Haksell library
+as our desugaring front-end.
 
@@
 
+We reformulated the \lambdajs{} reduction semantics into a state machine
+semantics with explicit continuations. The \lambdajs{} evaluation contexts
+became our \emph{Frame} data type. Additionally, we do not divide the
+evaluation contexts into two sets and use syntactic stepping rules to implement
+exceptional control flow. We choose to implement exceptional control flow with
+meta functions that crawl the stack looking for appropriate return points.
+
 \section{Implementation}
@@
 analysis. With an appropriate set of monads, traditional points-to constraints
-could be generated.
+could be generated. We note that we did not in the end implement each of these
+analyses for \js{}, although we plan to do so in our continuation of this work.
 

### 2014-12-13 cs252r-fa14-writeup/paper.tex
Add some code

--- 2014-12-13 cs252r-fa14-writeup/paper.tex
@@
   morecomment=[l]\%,
+  mathescape=true,
 }
@@
 \end{lstlisting}
+\caption{The Frame, i.e., Continuation Datatype}
+\end{figure}
 
+\begin{figure}
+\centering
+
+\begin{lstlisting}
+pushFrame :: (Analysis $\varsigma$ m) => Frame -> m ()
+pushFrame fr = do
+  fp  <- getL konL
+  fp' <- nextFramePtr
+  modifyL kstoreL (mapInsertWith (\/) fp' (singleton (fr, fp))
+  putL konL fp'
+
+popFrame :: (Analysis $\varsigma$ m) => m Frame
+popFrame = do
+  fp <- getL konL
+  k$\sigma$ <- getL kstoreL
+  (fr, fp') <- mset (mjoin (liftMaybeSet (k$\sigma$ # fp)))
+  putL konL fp'
+  return fr
+\end{lstlisting}
+\caption{Pushing and Popping Stack Frames}
 \end{figure}

### 2014-12-13 cs252r-fa14-writeup/paper.tex
ready to send to Steve

--- 2014-12-13 cs252r-fa14-writeup/paper.tex
@@
 \usepackage{amssymb}
+\usepackage{hyperref}
 
@@
 \begin{itemize}
-\item Design a configurable static analysis for javascript using the
+  \item Design a configurable static analysis for \js{} using the
   Monadic Abstracting Abstract Machines (\maam{}) \cite{maam} design process.
@@
 We also spent a significant amount of time understanding pushdown analysis and
-recasting it in the \aam{} and \maam frameworks.
+recasting it in the \aam{} and \maam{} frameworks.
+%
+Finally, we plan on continuing this work during the coming weeks to realize the
+full potential of our efforts, with the aim of accomplishing all three project
+goals.
 
@@
 The fact that the cache can memoize invariant to the continuation is
-traditionally understoon in the pushdown literature as by-product of viewing
+traditionally understood in the pushdown literature as by-product of viewing
 the semantics a pushdown automata.
@@
 
-\begin{figure}
-\centering
-
-\begin{lstlisting}
-data Frame =
-    LetK [(Name, Set AValue)] Name [(Name, TExp)] TExp
-  | AppL [TExp]
-  | AppR (Set AValue) [(Set AValue)] [TExp]
-  | ObjK [(String, (Set AValue))] Name [(Name, TExp)]
-    -- Array Dereferencing
-  | FieldRefL TExp
-  | FieldRefR (Set AValue)
-    -- Array Assignment
-  | FieldSetA TExp         TExp
-  | FieldSetN (Set AValue) TExp
-  | FieldSetV (Set AValue) (Set AValue)
-    -- Property Deletion
-  | DeleteL TExp
-  | DeleteR (Set AValue)
-    -- Fig 2. Mutable References
-  | RefSetL TExp
-  | RefSetR (Set AValue)
-  | RefK
-  | DeRefK
-    -- Fig 8. Control Operators
-  | IfK TExp TExp
-  | SeqK TExp
-  | WhileL TExp TExp
-  | WhileR TExp TExp
-  | LabelK Label
-  | BreakK Label
-  | TryCatchK TExp Name
-  | TryFinallyL TExp
-  | TryFinallyR (Set AValue)
-  | ThrowK
-    -- Fig 9. Primitive Operations
-  | PrimOpK Op [(Set AValue)] [TExp]
-  deriving (Eq, Ord)
-instance PartialOrder Frame where pcompare = discreteOrder
-\end{lstlisting}
-\caption{The Frame, i.e., Continuation Datatype}
-\end{figure}
-
-\begin{figure}
-\centering
-
-\begin{lstlisting}
-pushFrame :: (Analysis $\varsigma$ m) => Frame -> m ()
-pushFrame fr = do
-  fp  <- getL konL
-  fp' <- nextFramePtr
-  modifyL kstoreL (mapInsertWith (\/) fp' (singleton (fr, fp))
-  putL konL fp'
-
-popFrame :: (Analysis $\varsigma$ m) => m Frame
-popFrame = do
-  fp <- getL konL
-  k$\sigma$ <- getL kstoreL
-  (fr, fp') <- mset (mjoin (liftMaybeSet (k$\sigma$ # fp)))
-  putL konL fp'
-  return fr
-\end{lstlisting}
-\caption{Pushing and Popping Stack Frames}
-\end{figure}
+% \begin{figure}
+% \centering
+% 
+% \begin{lstlisting}
+% data Frame =
+%     LetK [(Name, Set AValue)] Name [(Name, TExp)] TExp
+%   | AppL [TExp]
+%   | AppR (Set AValue) [(Set AValue)] [TExp]
+%   | ObjK [(String, (Set AValue))] Name [(Name, TExp)]
+%     -- Array Dereferencing
+%   | FieldRefL TExp
+%   | FieldRefR (Set AValue)
+%     -- Array Assignment
+%   | FieldSetA TExp         TExp
+%   | FieldSetN (Set AValue) TExp
+%   | FieldSetV (Set AValue) (Set AValue)
+%     -- Property Deletion
+%   | DeleteL TExp
+%   | DeleteR (Set AValue)
+%     -- Fig 2. Mutable References
+%   | RefSetL TExp
+%   | RefSetR (Set AValue)
+%   | RefK
+%   | DeRefK
+%     -- Fig 8. Control Operators
+%   | IfK TExp TExp
+%   | SeqK TExp
+%   | WhileL TExp TExp
+%   | WhileR TExp TExp
+%   | LabelK Label
+%   | BreakK Label
+%   | TryCatchK TExp Name
+%   | TryFinallyL TExp
+%   | TryFinallyR (Set AValue)
+%   | ThrowK
+%     -- Fig 9. Primitive Operations
+%   | PrimOpK Op [(Set AValue)] [TExp]
+%   deriving (Eq, Ord)
+% instance PartialOrder Frame where pcompare = discreteOrder
+% \end{lstlisting}
+% \caption{The Frame, i.e., Continuation Datatype}
+% \end{figure}
 
@@
 The third item is the main contribution of this project.
+The architecture of our executable analysis is shown in Figure 2.
+
+\begin{figure}
+\centering
+
+\begin{tikzpicture}[->,>=stealth',shorten >=1pt,auto,node distance=5cm,
+  thick,main node/.style={font=\sffamily\bfseries}]
+
+  \node[main node] (101) at (0, 1)
+    {JS Code};
+  \node[main node] (102) [right of = 101]
+    {\lambdajs{} Code};
+  \node[main node] (103) [right of = 102]
+    {Analysis Results};
+
+  \path[->]
+    (101) edge node {Parse \& Desugar} (102)
+    (102) edge node {\maam{} Analyzer} (103);
+
+\end{tikzpicture}
+\caption{The Architecture of Our Analysis System. The second arrow is original work.}
+\end{figure}
 
@@
 analyses for \js{}, although we plan to do so in our continuation of this work.
+Our implementation is accessible at \url{https://github.com/davdar/maam}
+under the path "src/Lang/JS/".
 
@@
 
+We show two code snippets from our implementation in Figure 3.
+[pushFrame] shows the monadic action for pushing a continuation frame, which
+consists of: getting the current continuation pointer; getting the 
+continuation pointer to allocate; and adding a new continuation frame to the
+continuation store.
+[popFrame] shows the monadic action for popping continuation frames. The popped
+frame will scrutenized to determine the  control state of the program, and
+as a result, the set of frames returned from the continuation store must be
+merged into the monad. This merging is done with the [mset] operator 
+$(mset : \forall(a). \mathcal{P}(a) \rightarrow m(a))$.
+
 \begin{figure}
@@
 
-\begin{tikzpicture}[->,>=stealth',shorten >=1pt,auto,node distance=5cm,
-  thick,main node/.style={font=\sffamily\bfseries}]
+\begin{lstlisting}
+pushFrame :: (Analysis $\varsigma$ m) => Frame -> m ()
+pushFrame fr = do
+  fp  <- getL konL
+  fp' <- FramePtr
+  modifyL kstoreL (mapInsertWith (\/) fp' (singleton (fr, fp))
+  putL konL fp'
 
-  \node[main node] (101) at (0, 1)
-    {JS Code};
-  \node[main node] (102) [right of = 101]
-    {\lambdajs{} Code};
-  \node[main node] (103) [right of = 102]
-    {Analysis Results};
-
-  \path[->]
-    (101) edge node {Parse \& Desugar} (102)
-    (102) edge node {\maam{} Analyzer} (103);
-
-\end{tikzpicture}
-\caption{The Architecture of Our Analysis System. The second arrow is original work.}
+popFrame :: (Analysis $\varsigma$ m) => m Frame
+popFrame = do
+  fp <- getL konL
+  k$\sigma$ <- getL kstoreL
+  (fr, fp') <- mset (mjoin (liftMaybeSet (k$\sigma$ # fp)))
+  putL konL fp'
+  return fr
+\end{lstlisting}
+\caption{Pushing and Popping Stack Frames}
 \end{figure}

### 2015-03-17 writeup/oopsla2015.tex
small edits and transition to oopsla format

--- 2014-11-13 writeup/pldi15.tex
@@
-%lambdlambdaa any author declaration will be ignored  when using 'plid' option (for double blind review)
-\documentclass[pldi,nocopyrightspace]{sigplanconf}
+\documentclass[10pt,preprint]{sigplanconf}
 
-% From PLDI Template {-{
+% OOPSLA Setup {-{
 
-% \usepackage{SIunits}            % typset units correctly
-\usepackage{courier}            % standard fixed width font
-\usepackage[scaled]{helvet} % see www.ctan.org/get/macros/latex/required/psnfss/psnfss2e.pdf
-\usepackage{url}                  % format URLs
-\usepackage{listings}          % format code
-\usepackage{enumitem}      % adjust spacing in enums
-\usepackage{amsmath}
-\usepackage[colorlinks=true,allcolors=blue,breaklinks,draft=false]{hyperref}   % hyperlinks, including DOIs and URLs in bibliography
-% known bug: http://tex.stackexchange.com/questions/1522/pdfendlink-ended-up-in-different-nesting-level-than-pdfstartlink
-\newcommand{\doi}[1]{doi:~\href{http://dx.doi.org/#1}{\Hurl{#1}}}   % print a hyperlinked DOI
+% % \usepackage{SIunits}            % typset units correctly
+% \usepackage{courier}            % standard fixed width font
+% \usepackage[scaled]{helvet} % see www.ctan.org/get/macros/latex/required/psnfss/psnfss2e.pdf
+% \usepackage{url}                  % format URLs
+% \usepackage{listings}          % format code
+% \usepackage{enumitem}      % adjust spacing in enums
+% \usepackage{amsmath}
+% \usepackage[colorlinks=true,allcolors=blue,breaklinks,draft=false]{hyperref}   % hyperlinks, including DOIs and URLs in bibliography
+% % known bug: http://tex.stackexchange.com/questions/1522/pdfendlink-ended-up-in-different-nesting-level-than-pdfstartlink
+% \newcommand{\doi}[1]{doi:~\href{http://dx.doi.org/#1}{\Hurl{#1}}}   % print a hyperlinked DOI
 
@@
 \usepackage[font=small,skip=0pt]{caption}
+% \usepackage[belowskip=-15pt,aboveskip=0pt]{caption}
 
@@
 
+% OOPSLA Prelude {-{
+
+\special{papersize=8.5in,11in}
+\setlength{\pdfpageheight}{\paperheight}
+\setlength{\pdfpagewidth}{\paperwidth}
+
+\conferenceinfo{CONF 'yy}{Month d--d, 20yy, City, ST, Country} 
+\copyrightyear{20yy} 
+\copyrightdata{978-1-nnnn-nnnn-n/yy/mm} 
+\doi{nnnnnnn.nnnnnnn}
+
+% -- I'm not sure which one of these we use (or none?)... -DCD
+%
+% Uncomment one of the following two, if you are not going for the 
+% traditional copyright transfer agreement.
+
+%\exclusivelicense                % ACM gets exclusive license to publish, 
+                                  % you retain copyright
+
+%\permissiontopublish             % ACM gets nonexclusive license to publish
+                                  % (paid open-access papers, 
+                                  % short abstracts)
+
+\titlebanner{Preprint}             % These are ignored unless
+% \preprintfooter{preprint footer}   % 'preprint' option specified.
+
+% }-}
+
 % These are supposed to make align* not have breaks around it...
-\setlength{\abovedisplayskip}{0em}
-\setlength{\belowdisplayskip}{0em}
-\setlength{\abovedisplayshortskip}{0em}
-\setlength{\belowdisplayshortskip}{0em}
+% \setlength{\abovedisplayskip}{0em}
+% \setlength{\belowdisplayskip}{0em}
+% \setlength{\abovedisplayshortskip}{0em}
+% \setlength{\belowdisplayshortskip}{0em}
 
-%\title{The Marriage of Monad Transformers and Abstract Interpreters: Modular Metatheory for Program Analysis}
 \title{Galois Transformers and Modular Abstract Interpreters}
 \subtitle{Reusable Metatheory for Program Analysis}
-\authorinfo{David Darais}{Harvard University, USA}{darais@seas.harvard.edu}
+\authorinfo{David Darais}{University of Maryland, USA}{darais@cs.umd.edu}
 \authorinfo{Matthew Might}{University of Utah, USA}{might@cs.utah.edu}
@@
 
+\category{CR-number}{subcategory}{third-level}
+\terms term1, term2
+\keywords keyword1, keyword2
+
 \input{tmp/autogen/pldi15.markdown.tex}
+
+\appendix
+\section{Appendix Title}
+
+Appendix body.
+
+\acks
+
+Acknowledgments body.
 

### 2015-03-17 writeup/Process.hs

@@ -83,13 +83,13 @@ ops = def
 
 main :: IO ()
 main = do
-  s <- T.readFile "pldi15.markdown"
+  s <- T.readFile "paper.markdown"
   let pre = preProcess s
       md = readMarkdown ops  $ T.unpack pre
       post = postProcess md
   system "mkdir -p tmp/autogen"
-  T.writeFile "tmp/autogen/pldi15.markdown.pre" pre
-  T.writeFile "tmp/autogen/pldi15.markdown.tex" $ T.pack $ writeLaTeX def post
+  T.writeFile "tmp/autogen/paper.markdown.pre" pre
+  T.writeFile "tmp/autogen/paper.markdown.tex" $ T.pack $ writeLaTeX def post
 
 -- Pre Processing {{{
 

### 2015-03-21 src/FP/Core.hs
Finally figured out flow sens.
So, flow sensitivity was wrong in the initial paper, and then I fixed it
and it was subtly wrong again. I finally fixed it so that flow
sensitivity accurately represents "one set of facts per program point".

The issue now is the examples. Flow sensitivity is only helpful if the
facts at one program point are different in an interesting way than from
another program point. In imperative languages this happens through
strong update. In our little functional language, information (the
store) is growing strictly monotonically. This means that a
flow-insensitive and

@@ -218,7 +218,7 @@ poiterHistory f = loop
   where
     loop x = 
       let x' = f x
-      in if x ⊑ x' 
+      in if x' ⊑ x
         then [x]
         else x : loop x'
 
@@ -294,7 +294,7 @@ diffs = loop bot
   where
     loop :: a -> [a] -> [a]
     loop _ [] = []
-    loop accum (x:xs) = x \\ accum : loop (x \/ accum) xs
+    loop accum (x:xs) = (x \\ accum) : loop (x \/ accum) xs
 
 collectDiffs :: (PartialOrder a, JoinLattice a, Difference a) => (a -> a) -> a -> [a]
 collectDiffs f = diffs . collectHistory f
@@ -303,7 +303,8 @@ collectDiffs f = diffs . collectHistory f
 
 -- Functors {{{
 
-class Commute t u where commute :: t (u a)  -> u (t a)
+class Commute t u where 
+  commute :: t (u a)  -> u (t a)
 class Unit (t :: * -> *) where unit :: a -> t a
 class Functor (t :: * -> *) where map :: (a -> b) -> (t a -> t b)
 
@@ -647,9 +648,10 @@ elemAtN n t = case foldlk ff (Inr zer) t of
     ff (Inr i) x' k = if i == n then Inl x' else k $ Inr $ suc i
     ff (Inl _) _ _ = error "internal error"
 
-toList :: (Iterable a t)             => t -> [a]     ; toList = foldr (:) []
-toSet  :: (Ord a, Iterable a t)      => t -> Set a   ; toSet  = foldr insert empty
-toMap  :: (Ord k, Iterable (k, v) t) => t -> Map k v ; toMap  = foldr (uncurry mapInsert) mapEmpty
+toList    :: (Iterable a t)                     => t -> [a]     ; toList = foldr (:) []
+toSet     :: (Ord a, Iterable a t)              => t -> Set a   ; toSet  = foldr insert empty
+toMap     :: (Ord k, Iterable (k, v) t)         => t -> Map k v ; toMap  = foldr (uncurry mapInsert) mapEmpty
+toMapJoin :: (Ord k, Iterable (k, v) t, Join v) => t -> Map k v ; toMapJoin = foldr (uncurry $ mapInsertWith (\/)) mapEmpty
 
 -- }}}
 
[…]

### 2015-03-21 src/FP/Pretty.hs

@@ -358,7 +358,7 @@ instance (Pretty a) => Pretty (SetWithTop a) where
 instance (Pretty k, Pretty v) => Pretty (Map k v) where
   pretty = collection "{" "}" "," . map prettyMapping . fromMap
     where
-      prettyMapping (k, v) = nest 2 $ hvsep [hsep [pretty k, pun "=>"], pretty v]
+      prettyMapping (k, v) = nest 2 $ hvsep [nest (-2) $ hsep [pretty k, pun "=>"], pretty v]
 instance (Ord a, Pretty a) => Pretty (ListSet a) where pretty = pretty . toSet
 instance (Ord a, Pretty a) => Pretty (ListSetWithTop a) where pretty = pretty . setFromListWithTop
 

### 2015-03-21 src/Lang/LamIf/Analyses.hs

@@ -6,13 +6,14 @@ import Lang.LamIf.Val
 import Lang.LamIf.Semantics
 import Lang.LamIf.Monads
 import Lang.LamIf.CPS
+import Lang.LamIf.StateSpace
 
 -- These instances are defined in MAAM.Time
 timeChoices :: [(String, ExTime)]
 timeChoices =
-  [ ("*" , ExTime (W :: UniTime Cτ)     )
-  , ("1" , ExTime (W :: UniTime (Kτ 1)) )
-  , ("0" , ExTime (W :: UniTime Zτ)     )
+  [ ("*" , ExTime (W :: UniTime (Cτ Ψ))     )
+  , ("1" , ExTime (W :: UniTime (Kτ 1 Ψ)) )
+  , ("0" , ExTime (W :: UniTime (Zτ Ψ))     )
   ]
 
 -- These instances are defined in Lang.CPS.Val
@@ -25,9 +26,10 @@ valChoices =
 -- These instances are defined in MAAM.MonadStep and Lang.CPS.Monads
 monadChoices :: [(String, ExMonad)]
 monadChoices =
-  [ ( "ps" , ExMonad (W :: UniMonad PSΣ PSΣ𝒫 PS) )
-  , ( "fs" , ExMonad (W :: UniMonad FSΣ FSΣ𝒫 FS) )
-  , ( "fi" , ExMonad (W :: UniMonad FIΣ FIΣ𝒫 FI) )
+  [ ( "ps" , ExMonad (W :: UniMonad PSΣ PSΣ𝒫 PS) $ pretty . map (setMap (mapSnd 𝓈σ) . unPSΣ𝒫) )
+  , ( "fs" , ExMonad (W :: UniMonad FSΣ FSΣ𝒫 FS) $ 
+               pretty . map (filter ((/= bot) . snd) . map (\ (a, (_, σ)) -> (a, mapFilter (/= bot) σ)) . toList . unFSΣ𝒫) )
+  , ( "fi" , ExMonad (W :: UniMonad FIΣ FIΣ𝒫 FI) $ pretty . joins . map (snd . unFIΣ𝒫) )
   ]
 
 -- These are defined in Lang.CPS.Semantics

### 2015-03-21 src/Lang/LamIf/CPS.hs

@@ -30,7 +30,7 @@ type Call = StampedFix LocNum (PreCall Name)
 makePrisms ''PreCall
 
 freeVarsLam :: [Name] -> PreCall Name Call -> Set Name
-freeVarsLam xs c = freeVarsCall c \-\ toSet xs
+freeVarsLam xs c = freeVarsCall c \\ toSet xs
 
 freeVarsPico :: Pico -> Set Name
 freeVarsPico (Lit _) = bot
@@ -43,7 +43,7 @@ freeVarsAtom (LamF x kx c) = freeVarsLam [x, kx] $ stampedFix c
 freeVarsAtom (LamK x c) = freeVarsLam [x] $ stampedFix c
 
 freeVarsCall :: PreCall Name Call -> Set Name
-freeVarsCall (Let x a c) = freeVarsAtom (stamped a) \/ (freeVarsCall (stampedFix c) \-\ single x)
+freeVarsCall (Let x a c) = freeVarsAtom (stamped a) \/ (freeVarsCall (stampedFix c) \\ single x)
 freeVarsCall (If ax tc fc) = freeVarsPico ax \/ joins (freeVarsCall . stampedFix ^$ [tc, fc])
 freeVarsCall (AppF fx ax kx) = joins $ freeVarsPico ^$ [fx, ax, kx]
 freeVarsCall (AppK kx ax) = joins $ freeVarsPico ^$ [kx, ax]

### 2015-03-21 src/Lang/LamIf/Examples.hs

@@ -47,33 +47,23 @@ withOptions os e =
     , localSetL P.maxRibbonWidthL 40 $ pretty c
     , P.vsep $ mapOn os $ \ (info, o) -> 
         case runWithOptions o c of
-          ExSigma ς -> P.vsep
+          ExSigma pty ς -> P.vsep
             [ pretty info
-            , pretty ς
+            , pty ς
             ]
     ]
 
 examplesMain :: IO ()
 examplesMain = do
-  e <- parseFile "data/lam-src/simp4.lam"
+  e <- parseFile "data/lam-src/flow-example.lam"
   let os =
-         makeOptions
-         ["*"]
-         ["*"]
-         ["concrete"]
-         ["ps"]
-         ["no"]
-         ["link"]
-         ["*"]
-         ["*"]
-         ++
          makeOptions
          ["0"]
          ["0"]
          ["abstract"]
-         ["fi"]
-         ["no"]
+         ["fi", "fs", "ps"]
+         ["yes"]
          ["link"]
-         ["*"]
-         ["*"]
+         ["app"]
+         ["app"]
   pprint $ withOptions os e

### 2015-03-21 src/Lang/LamIf/Monads.hs

@@ -7,117 +7,107 @@ import Lang.LamIf.StateSpace
 import Lang.LamIf.Pretty ()
 
 -- Path Sensitive
-type PSΣ' val lτ dτ ψ = (ID :.: ListSet) :.: (,) (𝒮 val lτ dτ ψ)
-
-newtype PSΣ val lτ dτ ψ a = PSΣ { runPSΣ :: ListSet (a, 𝒮 val lτ dτ ψ) }
+type PSΣ' val lτ dτ = (ID :.: ListSet) :.: (,) (𝒮 val lτ dτ)
+newtype PSΣ val lτ dτ a = PSΣ { unPSΣ :: ListSet (a, 𝒮 val lτ dτ) }
   deriving (PartialOrder, Bot, Join, JoinLattice, Pretty)
-instance Morphism2 (PSΣ val lτ dτ ψ) (PSΣ' val lτ dτ ψ)  where
-  morph2 = Compose . Compose . ID . map swap . runPSΣ
-instance Morphism2 (PSΣ' val lτ dτ ψ) (PSΣ val lτ dτ ψ) where
+instance Morphism2 (PSΣ val lτ dτ) (PSΣ' val lτ dτ)  where
+  morph2 = Compose . Compose . ID . map swap . unPSΣ
+instance Morphism2 (PSΣ' val lτ dτ) (PSΣ val lτ dτ) where
   morph2 = PSΣ . map swap . unID . unCompose . unCompose
-instance Isomorphism2 (PSΣ val lτ dτ ψ) (PSΣ' val lτ dτ ψ) where
-instance (TimeC lτ, TimeC dτ) => Inject (PSΣ val lτ dτ Ψ) where
+instance Isomorphism2 (PSΣ val lτ dτ) (PSΣ' val lτ dτ) where
+instance (TimeC lτ, TimeC dτ) => Inject (PSΣ val lτ dτ) where
   inj = PSΣ . inj . (,bot)
 
-newtype PSΣ𝒫 val lτ dτ ψ a = PSΣ𝒫 { runPSΣ𝒫 :: Set (a, 𝒮 val lτ dτ ψ) }
-  deriving (PartialOrder, Bot, Join, JoinLattice, Pretty)
[…]

### 2015-03-21 src/Lang/LamIf/Pretty.hs

@@ -83,31 +83,32 @@ instance (Pretty n, Pretty c) => Pretty (PreCall n c) where
 instance (Pretty n) => Functorial Pretty (PreCall n) where
   functorial = W
 
-instance (Pretty (lτ ψ), Pretty (dτ ψ)) => Pretty (Addr lτ dτ ψ) where
+instance (Pretty lτ, Pretty dτ) => Pretty (Addr lτ dτ) where
   pretty (Addr loc lτ dτ) = P.collection "<" ">" "," 
     [ exec [P.pun "x=", P.align $ pretty loc]
     , exec [P.pun "lτ=", P.align $ pretty lτ]
     , exec [P.pun "dτ=", P.align $ pretty dτ]
     ]
 
-instance (Pretty (lτ ψ), Pretty (dτ ψ)) => Pretty (Clo lτ dτ ψ) where
+instance (Pretty lτ, Pretty dτ) => Pretty (Clo lτ dτ) where
   pretty (Clo l _xs _c _ρ lτ) = P.collection "<" ">" "," 
     [ exec [P.pun "λ=", pretty l]
     , exec [P.pun "lτ=", P.align $ pretty lτ]
     ]
 
+makePrettyUnion ''𝒮Cxt
 makePrettyUnion ''𝒮
 
-instance (Pretty (lτ ψ), Pretty (dτ ψ)) => Pretty (CVal lτ dτ ψ) where
+instance (Pretty lτ, Pretty dτ) => Pretty (CVal lτ dτ) where
   pretty (LitC l) = pretty l
   pretty (CloC c) = pretty c
   pretty BotC = P.lit "⊥"
 
-instance (Pretty (lτ ψ), Pretty (dτ ψ)) => Pretty (AVal lτ dτ ψ) where
+instance (Pretty lτ, Pretty dτ) => Pretty (AVal lτ dτ) where
   pretty (LitA l) = pretty l
   pretty IA = P.lit "INT"
   pretty BA = P.lit "BOOL"
   pretty (CloA c) = pretty c
   pretty BotA = P.lit "⊥"
 
-deriving instance (Pretty (val lτ dτ ψ)) => Pretty (Power val lτ dτ ψ)
+deriving instance (Pretty (val lτ dτ)) => Pretty (Power val lτ dτ)

### 2015-03-21 src/Lang/LamIf/Semantics.hs

@@ -14,23 +14,24 @@ type Ψ = LocNum
 -- - the monad m
 
 type TimeC τ =
-  ( Time τ
-  , Bot (τ Ψ)
-  , Ord (τ Ψ)
-  , Pretty (τ Ψ)
+  ( Time Ψ τ
+  , Bot τ
+  , Ord τ
+  , Pretty τ
   )
 
 type ValC lτ dτ val =
-  ( Val lτ dτ Ψ (val lτ dτ Ψ)
-  , Ord (val lτ dτ Ψ)
-  , PartialOrder (val lτ dτ Ψ)
-  , JoinLattice (val lτ dτ Ψ)
-  , Pretty (val lτ dτ Ψ)
+  ( Val lτ dτ val
+  , Ord val
+  , PartialOrder val
+  , JoinLattice val
+  , Difference val
+  , Pretty val
   )
 
 type MonadC val lτ dτ m =
   ( Monad m, MonadBot m, MonadPlus m
-  , MonadState (𝒮 val lτ dτ Ψ) m
+  , MonadState (𝒮 val lτ dτ) m
   )
 
 -- This type class aids type inference. The functional dependencies tell the
@@ -45,67 +46,67 @@ class
 
 -- Some helper types
 type GC m = Call -> m ()
-type CreateClo lτ dτ m = LocNum -> [Name] -> Call -> m (Clo lτ dτ Ψ)
+type CreateClo lτ dτ m = LocNum -> [Name] -> Call -> m (Clo lτ dτ)
 type TimeFilter = Call -> Bool
 
 -- Generate a new address
-new :: (Analysis val lτ dτ m) => Name -> m (Addr lτ dτ Ψ)
+new :: (Analysis val lτ dτ m) => Name -> m (Addr lτ dτ)
 new x = do
   lτ <- getL 𝓈lτL
   dτ <- getL 𝓈dτL
   return $ Addr x lτ dτ
 
 -- bind a name to a value in an environment
-bind :: (Analysis val lτ dτ m) => Name -> val lτ dτ Ψ -> Map Name (Addr lτ dτ Ψ) -> m (Map Name (Addr lτ dτ Ψ))
+bind :: (Analysis val lτ dτ m) => Name -> val -> Map Name (Addr lτ dτ) -> m (Map Name (Addr lτ dτ))
 bind x vD ρ = do
   l <- new x
   modifyL 𝓈σL $ mapInsertWith (\/) l vD
   return $ mapInsert x l ρ
 
 -- bind a name to a value in _the_ environment
[…]

### 2015-03-21 src/Lang/LamIf/StateSpace.hs

@@ -4,36 +4,50 @@ import FP
 import Lang.LamIf.Syntax
 import Lang.LamIf.CPS
 
-data Addr lτ dτ (ψ :: *) = Addr
+data Addr lτ dτ = Addr
   { addrLocation :: Name
-  , addrLexicalTime :: lτ ψ
-  , addrDynamicTime :: dτ ψ
+  , addrLexicalTime :: lτ
+  , addrDynamicTime :: dτ
   } deriving (Eq, Ord)
 
-type Env lτ dτ ψ = Map Name (Addr lτ dτ ψ)
-type Store val lτ dτ ψ = Map (Addr lτ dτ ψ) (val lτ dτ ψ)
+type Env lτ dτ = Map Name (Addr lτ dτ)
+type Store val lτ dτ = Map (Addr lτ dτ) val
 
-data 𝒮 val lτ dτ ψ = 𝒮
-  { 𝓈lτ :: lτ ψ
-  , 𝓈dτ :: dτ ψ
-  , 𝓈ρ :: Env lτ dτ ψ
-  , 𝓈σ :: Store val lτ dτ ψ
+data 𝒮Cxt lτ dτ = 𝒮Cxt
+  { 𝓈Cxtlτ :: lτ
+  , 𝓈Cxtdτ :: dτ
+  , 𝓈Cxtρ :: Env lτ dτ
+  } deriving (Eq, Ord)
+makeLenses ''𝒮Cxt
+instance (Bot lτ, Bot dτ) => Bot (𝒮Cxt lτ dτ) where bot = 𝒮Cxt bot bot bot
+
+data 𝒮 val lτ dτ = 𝒮
+  { 𝓈Cxt :: 𝒮Cxt lτ dτ
+  , 𝓈σ :: Store val lτ dτ
   } deriving (Eq, Ord)
 makeLenses ''𝒮
-instance (Bot (lτ ψ), Bot (dτ ψ)) => Bot (𝒮 val lτ dτ ψ) where
-  bot = 𝒮 bot bot bot bot
+instance (Bot lτ, Bot dτ) => Bot (𝒮 val lτ dτ) where bot = 𝒮 bot bot
 
-data Clo lτ dτ ψ = Clo 
+𝓈lτL :: Lens (𝒮 val lτ dτ) lτ
+𝓈lτL = 𝓈CxtlτL <.> 𝓈CxtL
+
[…]

### 2015-03-21 src/Lang/LamIf/Val.hs

@@ -5,39 +5,39 @@ import Lang.LamIf.StateSpace
 import Lang.LamIf.Syntax
 
 -- Concrete
-data CVal lτ dτ ψ = LitC Lit | CloC (Clo lτ dτ ψ) | BotC
+data CVal lτ dτ = LitC Lit | CloC (Clo lτ dτ) | BotC
   deriving (Eq, Ord)
 makePrisms ''CVal
 
-instance (Eq (lτ ψ), Eq (dτ ψ)) => PartialOrder (CVal lτ dτ ψ) where
+instance (Eq lτ, Eq dτ) => PartialOrder (CVal lτ dτ) where
   pcompare BotC _ = PLT
   pcompare _ BotC = PGT
   pcompare v1 v2 | v1 == v2  = PEQ
                  | otherwise = PUN
 
-instance (Ord (dτ ψ), Ord (lτ ψ)) => Val lτ dτ ψ (CVal lτ dτ ψ) where
-  lit :: Lit -> CVal lτ dτ ψ
+instance (Ord lτ, Ord dτ) => Val lτ dτ (CVal lτ dτ) where
+  lit :: Lit -> CVal lτ dτ
   lit = LitC
-  clo :: Clo lτ dτ ψ -> CVal lτ dτ ψ
+  clo :: Clo lτ dτ -> CVal lτ dτ
   clo = CloC
-  binop :: BinOp -> CVal lτ dτ ψ -> CVal lτ dτ ψ -> CVal lτ dτ ψ
+  binop :: BinOp -> CVal lτ dτ -> CVal lτ dτ -> CVal lτ dτ
   binop Add (LitC (I n1)) (LitC (I n2)) = LitC $ I $ n1 + n2
   binop Sub (LitC (I n1)) (LitC (I n2)) = LitC $ I $ n1 - n2
   binop GTE (LitC (I n1)) (LitC (I n2)) = LitC $ B $ n1 >= n2
   binop _ _ _ = BotC
-  elimBool :: CVal lτ dτ ψ -> Set Bool
+  elimBool :: CVal lτ dτ -> Set Bool
   elimBool (LitC (B b)) = single b
   elimBool _ = empty
-  elimClo :: CVal lτ dτ ψ -> Set (Clo lτ dτ ψ)
+  elimClo :: CVal lτ dτ -> Set (Clo lτ dτ)
   elimClo (CloC c) = single c
   elimClo _ = empty
 
 -- Abstract
[…]

### 2015-03-21 src/MAAM/MonadStep.hs

@@ -22,36 +22,49 @@ deriving instance (MonadStep ς m, Functor m) => MonadStep (ς :.: (,) 𝓈1) (A
 -- Flow Insensitive
 instance (MonadStep ς m, Functorial JoinLattice m) => MonadStep (ς :.: ListSet) (ListSetT m) where
   mstepγ :: forall a b. (a -> ListSetT m b) -> (ς :.: ListSet) a -> (ς :.: ListSet) b
-  mstepγ f = 
-    with (functorial :: W (JoinLattice (m (ListSet b)))) $
-    onComposeIso $ (mstepγ :: forall a' b'. (a' -> m b') -> (ς a' -> ς b')) $ joins . map (unListSetT . f)
+  mstepγ f = onComposeIso $ mstepγ_m ff
+    where
+      mstepγ_m :: forall a' b'. (a' -> m b') -> (ς a' -> ς b')
+      mstepγ_m = mstepγ
+      ff :: ListSet a -> m (ListSet b)
+      ff = with (functorial :: W (JoinLattice (m (ListSet b)))) $
+        joins . map (unListSetT . f)
 
 -- Flow Insensitive with top
-instance (MonadStep ς m, Functorial JoinLattice m, Unit m) => MonadStep (ς :.: ListSetWithTop) (ListSetWithTopT m) where
+instance (MonadStep ς m, Functorial JoinLattice m, Functorial Top m) => MonadStep (ς :.: ListSetWithTop) (ListSetWithTopT m) where
   mstepγ :: forall a b. (a -> ListSetWithTopT m b) -> (ς :.: ListSetWithTop) a -> (ς :.: ListSetWithTop) b
-  mstepγ f = 
-    with (functorial :: W (JoinLattice (m (ListSetWithTop b)))) $
-    onComposeIso $ (mstepγ :: forall a' b'. (a' -> m b') -> (ς a' -> ς b')) $ listSetWithTopElim (unit top) joins . map (unListSetWithTopT . f)
+  mstepγ f = onComposeIso $ mstepγ_m ff
+    where
+      mstepγ_m :: forall a' b'. (a' -> m b') -> (ς a' -> ς b')
+      mstepγ_m = mstepγ
+      ff :: ListSetWithTop a -> m (ListSetWithTop b)
+      ff = 
+        with (functorial :: W (JoinLattice (m (ListSetWithTop b)))) $
+        with (functorial :: W (Top (m (ListSetWithTop b)))) $
[…]

### 2015-03-21 src/MAAM/Time.hs

@@ -4,9 +4,9 @@ import FP
 import qualified FP.Pretty as P
 import GHC.TypeLits
 
-class Time τ where
-  tzero :: τ ψ
-  tick :: ψ -> τ ψ -> τ ψ
+class Time ψ τ | τ -> ψ where
+  tzero :: τ
+  tick :: ψ -> τ -> τ
 
 -- Concrete
 newtype Cτ ψ = Cτ [ψ]
@@ -15,7 +15,7 @@ instance Functorial Eq Cτ where functorial = W
 instance Functorial Ord Cτ where functorial = W
 instance Functorial Pretty Cτ where functorial = W
 instance Bot (Cτ ψ) where bot = nil
-instance Time Cτ where { tzero = bot ; tick = (&) }
+instance Time ψ (Cτ ψ) where { tzero = bot ; tick = (&) }
 
 -- kCFA
 newtype Kτ (k :: Nat) ψ = Kτ [ψ]
@@ -24,7 +24,7 @@ instance Functorial Eq (Kτ k) where functorial = W
 instance Functorial Ord (Kτ k) where functorial = W
 instance Functorial Pretty (Kτ k) where functorial = W
 instance Bot (Kτ k ψ) where bot = nil
-instance (KnownNat k) => Time (Kτ k) where
+instance (KnownNat k) => Time ψ (Kτ k ψ) where
   tzero = bot
   tick x y = fromList $ firstN (natVal (P :: P k)) $ toList $ x & y
 
@@ -37,4 +37,4 @@ instance Functorial Eq Zτ where functorial = W
 instance Functorial Ord Zτ where functorial = W
 instance Functorial Pretty Zτ where functorial = W
 instance Bot (Zτ ψ) where bot = Zτ
-instance Time Zτ where { tzero = bot ; tick = const id }
+instance Time ψ (Zτ ψ) where { tzero = bot ; tick = const id }

### 2015-03-23 src/Lang/LamIf/CPS.hs
flow example worked out

@@ -29,22 +29,26 @@ instance (Eq n, Eq c) => PartialOrder (PreCall n c) where pcompare = discreteOrd
 type Call = StampedFix LocNum (PreCall Name)
 makePrisms ''PreCall
 
-freeVarsLam :: [Name] -> PreCall Name Call -> Set Name
-freeVarsLam xs c = freeVarsCall c \\ toSet xs
+freeVarsLam :: Set Name -> [Name] -> Call -> Set Name
+freeVarsLam ρ xs c = freeVarsCall (ρ \/ toSet xs) c
 
-freeVarsPico :: Pico -> Set Name
-freeVarsPico (Lit _) = bot
-freeVarsPico (Var x) = single x
+freeVarsPico :: Set Name -> Pico -> Set Name
+freeVarsPico _ (Lit _) = bot
+freeVarsPico ρ (Var x) 
+  | ρ ? x = bot
+  | otherwise = single x
 
-freeVarsAtom :: PreAtom Name Call -> Set Name
-freeVarsAtom (Pico p) = freeVarsPico p
-freeVarsAtom (Prim _ a1 a2) = freeVarsPico a1 \/ freeVarsPico a2
-freeVarsAtom (LamF x kx c) = freeVarsLam [x, kx] $ stampedFix c
-freeVarsAtom (LamK x c) = freeVarsLam [x] $ stampedFix c
+freeVarsAtom :: Set Name -> Atom -> Set Name
+freeVarsAtom ρ a = case stamped a of
+  Pico p -> freeVarsPico ρ p
+  Prim _ a1 a2 -> freeVarsPico ρ a1 \/ freeVarsPico ρ a2
+  LamF x kx c -> freeVarsLam ρ [x, kx] c
+  LamK x c -> freeVarsLam ρ [x] c
 
-freeVarsCall :: PreCall Name Call -> Set Name
-freeVarsCall (Let x a c) = freeVarsAtom (stamped a) \/ (freeVarsCall (stampedFix c) \\ single x)
-freeVarsCall (If ax tc fc) = freeVarsPico ax \/ joins (freeVarsCall . stampedFix ^$ [tc, fc])
-freeVarsCall (AppF fx ax kx) = joins $ freeVarsPico ^$ [fx, ax, kx]
-freeVarsCall (AppK kx ax) = joins $ freeVarsPico ^$ [kx, ax]
[…]

### 2015-03-23 writeup/Process.hs

@@ -47,7 +47,7 @@ wrapMacro c = alterMacro $ \ r -> T.concat [ "\\" , c ,  "{" , r , "}" ]
 allMacros :: [(Text,Text,MatchMode)]
 allMacros = concat
   [ getMacros "Process_pre_macros.tbl"
-  , map (wrapMacro "ttop") $ getMacros "Process_tt_op.tbl"
+  , map (wrapMacro "ttbfop") $ getMacros "Process_ttbf_op.tbl"
   , map (wrapMacro "itop") $ getMacros "Process_it_op.tbl"
   , getMacros "Process_post_macros.tbl"
   ]
@@ -119,6 +119,7 @@ postProcess = walkInlineMath . walkBlocksMath . walkInlineRaw . walkBlocksRaw
   where
     walkBlocksRaw = walk $ \ (b :: Block) -> case b of
       CodeBlock (_,[c],_) s
+        | "rawmacro" `isPrefixOf` c -> RawBlock (Format "latex") $ T.unpack $ macroText $ T.pack s
         | "raw" `isPrefixOf` c -> RawBlock (Format "latex") s
       _ -> b
     walkInlineRaw = walk $ \ (i :: Inline) -> case i of

### 2015-03-23 writeup/oopsla2015.tex
Finally figured out flow sens.
So, flow sensitivity was wrong in the initial paper, and then I fixed it
and it was subtly wrong again. I finally fixed it so that flow
sensitivity accurately represents "one set of facts per program point".

The issue now is the examples. Flow sensitivity is only helpful if the
facts at one program point are different in an interesting way than from
another program point. In imperative languages this happens through
strong update. In our little functional language, information (the
store) is growing strictly monotonically. This means that a
flow-insensitive and

@@
-\documentclass[10pt,preprint]{sigplanconf}
+\documentclass[nocopyrightspace,preprint,10pt]{sigplanconf}
 
@@
 \usepackage[font=small,skip=0pt]{caption}
-% \usepackage[belowskip=-15pt,aboveskip=0pt]{caption}
 
@@
 \newcommand\litClo{Clo}
+\newcommand\litelse{else}
 
 \newcommand{\itop}[1]{\operatorname{\mathit{#1}}}
-\newcommand{\ttbfop}[1]{\operatorname{\mathtt{\boldsymbol{\mathbf{#1}}}}}
+\newcommand{\ttop}[1]{\operatorname{\mathtt{#1}}}
+\newcommand{\ttbfop}[1]{\ttop{\boldsymbol{\mathbf{#1}}}}
 \newcommand{\ttbfbin}[1]{\mathbin{\mathtt{\mathbf{\boldsymbol{#1}}}}}
-\newcommand{\ttop}[1]{\operatorname{\mathtt{#1}}}
+\newcommand{\keyword}[1]{\underline{\ttbfop{#1}}}
+\newcommand{\keywordop}[1]{\ttbfbin{#1}}
 \newcommand{\Concrete}[1]{\mathtt{\mathbf{#1}}}
@@
 
-\conferenceinfo{CONF 'yy}{Month d--d, 20yy, City, ST, Country} 
-\copyrightyear{20yy} 
-\copyrightdata{978-1-nnnn-nnnn-n/yy/mm} 
+% \conferenceinfo{CONF 'yy}{Month d--d, 20yy, City, ST, Country} 
+% \copyrightyear{20yy} 
+% \copyrightdata{978-1-nnnn-nnnn-n/yy/mm} 
 \doi{nnnnnnn.nnnnnnn}
@@
 % These are supposed to make align* not have breaks around it...
-% \setlength{\abovedisplayskip}{0em}
-% \setlength{\belowdisplayskip}{0em}
-% \setlength{\abovedisplayshortskip}{0em}
-% \setlength{\belowdisplayshortskip}{0em}
+\setlength{\abovedisplayskip}{0em}
+\setlength{\belowdisplayskip}{0em}
+\setlength{\abovedisplayshortskip}{0em}
+\setlength{\belowdisplayshortskip}{0em}
 
@@
 \subtitle{Reusable Metatheory for Program Analysis}
-\authorinfo{David Darais}{University of Maryland, USA}{darais@cs.umd.edu}
-\authorinfo{Matthew Might}{University of Utah, USA}{might@cs.utah.edu}
-\authorinfo{David Van Horn}{University of Maryland, USA}{dvanhorn@cs.umd.edu}
+% \authorinfo{David Darais}{University of Maryland, USA}{darais@cs.umd.edu}
+% \authorinfo{Matthew Might}{University of Utah, USA}{might@cs.utah.edu}
+% \authorinfo{David Van Horn}{University of Maryland, USA}{dvanhorn@cs.umd.edu}
+\authorinfo{}{}{}
 \maketitle
@@
 
-The design and implementation of static analyzers have becoming increasingly
-systematic.  In fact, for large classes of analyzers, design and implementation
-have remained seemingly (and now stubbornly) on the verge of full mechanization
-for several years.  A stumbling block in full mechanization has been the
-\emph{ad hoc} nature of soundness proofs accompanying each analyzer. While
-design and implementation is largely systematic, soundness proofs can change
-significantly with seemingly minor changes to the semantics or analyzers.  An
-achievement of this work is to systematize, parameterize and modularize the
-proofs of soundness, so as to make them composable across analytic properties.
+The design and implementation of static analyzers have becoming
+increasingly systematic.  Yet for a given language or analysis
+feature, it often requires tedious and error prone work to implement
+an analyzer and prove it sound.
+%
+In short, static analysis features and their proofs of soundness do
+not compose well, causing a dearth of reuse in both implementation and
+metatheory.
 
@@
 
-\category{CR-number}{subcategory}{third-level}
-\terms term1, term2
-\keywords keyword1, keyword2
+% \category{CR-number}{subcategory}{third-level}
+% \terms term1, term2
+% \keywords keyword1, keyword2
 
-\input{tmp/autogen/pldi15.markdown.tex}
+\input{tmp/autogen/paper.markdown.tex}
 
-\appendix
-\section{Appendix Title}
-
-Appendix body.
-
-\acks
-
-Acknowledgments body.
+% \appendix
+% \section{Appendix Title}
+% 
+% Appendix body.
+% 
+% \acks
+% 
+% Acknowledgments body.
 

### 2015-03-23 writeup/oopsla2015.tex
PLDI summary review and discussion in appendix.

@@
 
+\appendix
+\section{PLDI 2015 Summary review}
+
+This paper was previously submitted to PLDI 2015. It advanced through
+the first round with favorable reviews, but received a short negative
+review in the second round stating ``this paper seems like a rather
+straightforward and not very deep extension of the PLDI'13 paper on
+monadic abstract interpreters.''
+
+The summary review of the PC discussion stated:
+
+\begin{quote}
+There was extensive discussion of this paper at the PC meeting, with
+arguments on both sides. From the perspective of the functional
+programming experts, there is only a small increment from the previous
+PLDI'13 paper---moving from monads to monad transformers is a
+seemingly obvious step with little novelty. From the perspective of
+the static analysis experts, the paper presents a very interesting
+view on how to structure abstract interpreters that shows how to apply
+existing work in the functional world to static analysis. Ultimately
+we looked at the section on Galois transfomers (sic) as the point of
+potential novelty from the functional perspective, and found that
+there was too little there to convince people that Galois transformers
+are a non-trivial variation of monad transformers. We encourage the
+authors to continue this work and to try and clarify the novelty from
+the functional perspective (or target the paper more closely to the
+static analysis community). We also encourage the authors to work on
+the exposition in the paper to make it more accessible.
+\end{quote}
+
+The present version of this paper has aimed to clarify the novelty of
+this work and explain how, while the idea of ``going from monads to
+monad transformers'' is in some sense an obvious idea, there is no way
+to get there from the PLDI'13 paper of Sergey et al., without
+significant research.  That research is the subject of this paper.
+
+We have also incorporated all of the feedback toward making the paper
+more accessible.
+
+For completeness, we include our response to the claim ``this paper
+seems like a rather straightforward and not very deep extension of the
+PLDI'13 paper on monadic abstract interpreters.''  Much of this
+content has been incorporated into the paper.
+
+\begin{quote}
+We respectfully disagree with this assessment of the paper. Let us clarify the
+relationship with Sergey et al \cite{dvanhorn:Sergey2013Monadic} (PLDI'13).
+
+The primary contributions of Sergey et al are:
+
+\begin{itemize}
+\item Monads can serve as a good organizational mechanism for constructing abstract
+  interpreters.
+\item Different monads can be used to describe a wide range of abstract
+  interpreters.
+\end{itemize}
+
+In our view, shortcomings of Sergey et al are:
+
+\begin{itemize}
+\item No restrictions are placed on the monads used, allowing for
+  absurd and unsound analyses.
+
+\item There is no soundness framework for the approach. Instantiated
+  analyses must be reasoned about on an ad-hoc basis after inlining
+  the abstractions.
+
+\item The monads are not compositional or reusable across languages as
+  claimed in the paper.
+\end{itemize}
+
+Our paper remedies these shortcomings:
+
+\begin{itemize}
+\item Our interface between interpreters and monads is well-defined
+  and based on language-independent monadic effects, allowing one to
+  reason on both sides of the interface.
+
+\item Our framework constructs end-to-end Galois connections for
+  proofs of soundness.
+
+\item Monads in our framework can be defined and proved correct
+independently of a particular language.  Furthermore, constructing a
+new large monad in our framework is a simple composition of the monad
+transformers described in the paper.
+\end{itemize}
+
+In addition to these improvements, our paper has the following
+desirable properties:
+
+\begin{itemize}
+\item In our framework, intentional analysis optimizations like abstract garbage
+  collection and shape analysis can be defined and proven correct for all
+  concrete and abstract interpreters at once (for a particular language).
+
+\item Our framework fully compartmentalizes the choice of control sensitivity for the
+  analysis.
+\end{itemize}
+
+Alternatively, our paper can be seen as an extension of Liang et al
+\cite{dvanhorn:Liang1995Monad} applied to abstract interpreters rather
+than standard interpreters. We believe this is a deep insight into the
+theory of monad transformers and modular interpreters not present in
+Sergey et al.
+\end{quote}
+
 \end{document}

### 2015-03-25 writeup/oopsla2015.tex
builds now, sorry it was broked
added tikz diagrams for definition of Galois transformer
I think I'm done adding new content

@@
 \usepackage{balance}
+\usepackage{array}
+\usepackage{tikz}
+\usetikzlibrary{matrix}
 \floatstyle{boxed}
@@
 
+\newtheorem{theorem}{Theorem}
 \newtheorem{proposition}{Proposition}
@@
 interpreter, we define a library of monad transformers that
-implement building blocks for classic analysis parameters like context-, path-,
-and heap-(in-)sensitivity. Moreover, these can be composed together 
+implement building blocks for classic analysis parameters like context, path,
+and heap (in)sensitivity. Moreover, these can be composed together 
 % to realize a program analysis monad, 

### 2015-03-26 writeup/oopsla2015.tex
almost down to 14 pages.
almost done
Small typo in abstract.

@@
 \usepackage{tikz}
+\usepackage{etoolbox}
 \usetikzlibrary{matrix}
@@
 
-% These are supposed to make align* not have breaks around it...
-\setlength{\abovedisplayskip}{0em}
-\setlength{\belowdisplayskip}{0em}
-\setlength{\abovedisplayshortskip}{0em}
-\setlength{\belowdisplayshortskip}{0em}
+% These make align* have smaller breaks around it. -DCD
+\newcommand{\zerodisplayskips}{%
+  \setlength{\abovedisplayskip}{4pt}
+  \setlength{\belowdisplayskip}{4pt}
+  \setlength{\abovedisplayshortskip}{0pt}
+  \setlength{\belowdisplayshortskip}{0pt}}
+% \appto{\normalsize}{\zerodisplayskips}
+\appto{\small}{\zerodisplayskips}
+
+% This makes floats have less space between them and text.
+\setlength{\textfloatsep}{8pt}
 
@@
 introducing \emph{Galois transformers}: monad transformers that
-transports Galois connection properties.  In concert with a monadic
+transport Galois connection properties.  In concert with a monadic
 interpreter, we define a library of monad transformers that

### 2015-03-26 writeup/oopsla2015.tex
hit the page limit exactly

@@
 introducing \emph{Galois transformers}: monad transformers that
-transport Galois connection properties.  In concert with a monadic
+transports Galois connection properties.  In concert with a monadic
 interpreter, we define a library of monad transformers that

### 2015-03-26 writeup/oopsla2015.tex
final cut?

--- 2015-03-26 writeup/oopsla2015.tex
@@
 
-The design and implementation of static analyzers have becoming
-increasingly systematic.  Yet for a given language or analysis
-feature, it often requires tedious and error prone work to implement
-an analyzer and prove it sound.
-%
-In short, static analysis features and their proofs of soundness do
-not compose well, causing a dearth of reuse in both implementation and
-metatheory.
+The design and implementation of static analyzers has become increasingly
+systematic.  Yet for a given language or analysis feature, it often requires
+tedious and error prone work to implement an analyzer and prove it sound. In
+short, static analysis features and their proofs of soundness do not compose
+well, causing a dearth of reuse in both implementation and metatheory.
 
@@
 We solve the problem of systematically constructing static analyzers by
-introducing \emph{Galois transformers}: monad transformers that
-transport Galois connection properties.  In concert with a monadic
-interpreter, we define a library of monad transformers that
-implement building blocks for classic analysis parameters like context, path,
-and heap (in)sensitivity. Moreover, these can be composed together 
-% to realize a program analysis monad, 
-\emph{independent of the language being analyzed}.
+introducing \emph{Galois transformers}: monad transformers that transport
+Galois connection properties.  In concert with a monadic interpreter, we define
+a library of monad transformers that implement building blocks for classic
+analysis parameters like context, path, and heap (in)sensitivity. Moreover,
+these can be composed together \emph{independent of the language being
+analyzed}.
 

### 2015-03-26 writeup/oopsla2015.tex
killed a comma

@@
 
-The present version of this paper has aimed to clarify the novelty of
-this work and explain how, while the idea of ``going from monads to
-monad transformers'' is in some sense an obvious idea, there is no way
-to get there from the PLDI'13 paper of Sergey et al., without
-significant research.  That research is the subject of this paper.
+The present version of this paper has aimed to clarify the novelty of this work
+and explain how, while the idea of ``going from monads to monad transformers''
+is in some sense an obvious idea, there is no way to get there from the PLDI'13
+paper of Sergey et al. without significant research. That research is the
+subject of this paper.
 

### 2015-03-30 writeup/arxiv-src/paper.markdown.tex
arxiv source and edited to public version

\section{Introduction}\label{introduction}

\par

Traditional practice in program analysis via abstract interpretation is
to fix a language (as a concrete semantics) and an abstraction (as an
abstraction map, concretization map or Galois connection) before
constructing a static analyzer that is sound with respect to both the
abstraction and the concrete semantics. Thus, each pairing of
abstraction and semantics requires a one-off manual derivation of the
abstract semantics and construction of a proof of soundness.

\par

Work has focused on endowing abstractions with knobs, levers, and dials
to tune precision and compute efficiently. These parameters come with
overloaded meanings such as object, context, path, and heap
sensitivities, or some combination thereof. These efforts develop
families of analyses \emph{for a specific language} and prove the
framework sound.

\par

But this framework approach suffers from many of the same drawbacks as
the one-off analyzers. They are language-specific, preventing reuse of
concepts across languages, and require similar re-implementations and
soundness proofs. This process is still manual, tedious, difficult and
error-prone. And, changes to the structure of the parameter-space
require a completely new proof of soundness. And, it prevents fruitful
insights and results developed in one paradigm from being applied to
others, e.g., functional to object-oriented and \emph{vice versa}.

\par

We propose an automated alternative to structuring and implementing
program analysis. Inspired by Liang, Hudak, and Jones's
\emph{Monad
Transformers and Modular Interpreters} \cite{dvanhorn:Liang1995Monad},
we propose to start with concrete interpreters written in a specific
monadic style. Changing the monad will change the interpreter from a
concrete interpreter into an abstract interpreter. As we show, classical
program abstractions can be embodied as language-independent monads.
Moreover, these abstractions can be written as monad
\emph{transformers}, thereby allowing their composition to achieve new
forms of analysis. We show that these monad transformers obey the
properties of \emph{Galois connections}
\cite{dvanhorn:Cousot1979Systematic} and introduce the concept of a
\emph{Galois transformer}, a monad transformer which transports (1)
Galois connections and (2) mappings to an executable transition system.

\par

Most significantly, Galois transformers can be proved sound once and
used everywhere. Abstract interpreters, which take the form of monad
transformer stacks coupled with a monadic interpreter, inherit the
soundness properties of each element in the stack. This approach enables
reuse of abstractions across languages and lays the foundation for a
modular metatheory of program analysis.

\par

\paragraph{Setup}

We describe a simple language and a garbage-collecting allocating
semantics as the starting point of analysis design (Section
\ref{semantics}). We then briefly discuss three types of path and flow
sensitivity and their corresponding variations in analysis precision
(Section \ref{path-and-flow-sensitivity-in-analysis}).

\par

\paragraph{Monadic Abstract Interpreters}

We develop an abstract interpreter for our example language as a monadic
function with parameters (Section \ref{analysis-parameters}), one of
which is a monadic effect interface combining state and nondeterminism
effects (Section \ref{the-analysis-monad}). These monadic effects, state
and nondeterminism, encode arbitrary relational small-step state-machine
semantics and correspond to state-machine components and relational
nondeterminism respectively.

\par

Interpreters written in this style can be reasoned about using various
laws, including monadic effect laws, and therefore verified correct
independent of any particular choice of parameters. Likewise,
instantiations for these parameters can be reasoned about in isolation
from their instantiation. When instantiated, our generic interpreter is
capable of recovering the concrete semantics and a family of abstract
interpreters, with variations in abstract domain, abstract garbage
collection, mcfa, call-site sensitivity, object sensitivity, and path
and flow sensitivity (Section \ref{recovering-analyses}).

\par

\paragraph{Isolating Path and Flow Sensitivity}

We give specific monads for instantiating the interpreter from Section
\ref{the-interpreter} to path-sensitive, flow-sensitive and
flow-insensitive analyses (Section
\ref{varying-path-and-flow-sensitivity}). This leads to an isolated
understanding of path and flow sensitivity as mere variations in the
monad used for execution. Furthermore, these monads are language
independent, allowing one to reuse the same path and flow sensitivity
machinery for any language of interest, and compose seamlessly with
other analysis parameters.

\par

\paragraph{Galois Transformers}

To ease the construction of monads for building abstract interpreters
and their proofs of correctness, we develop a framework of Galois
transformers (Section \ref{a-compositional-monadic-framework}). Galois
transformers are an extension of monad transformers which transport (1)
Galois connections and (2) mappings to an executable transition system
(Section \ref{galois-transformers-1}). Our Galois transformer framework
allows us to both execute and reason about the correctness of an
abstract interpreter piecewise for each transformer in a stack. Galois
transformers are language independent and they can be proven correct one
and for all in isolation from a particular semantics.

\par

\paragraph{Implementation}

We have implemented our technique as a Haskell library and example
client analysis (Section \ref{implementation-1}). Developers are able to
reuse our language-independent framework for prototyping the design
space of analysis features for their language of choice. Our
implementation is publicly available on Hackage\footnote{
http://hackage.haskell.org/package/maam
}, Haskell's package manager.

\par

\paragraph{Contributions}

We make the following contributions:

\par

\begin{itemize}
\itemsep1pt\parskip0pt\parsep0pt
\item
  A methodology for constructing monadic abstract interpreters based on
  \emph{monadic effects}\footnote{
  This is in contrast to \citet{dvanhorn:Sergey2013Monadic} where monadic
  interpreters are constructed based on \emph{denotation functions}. See our
  Section \ref{related-work} for more details.}.
\item
  A compositional, language-independent library for constructing monads
  with various analysis properties based on \emph{monad transformers}.
\item
  A compositional, language-independent proof framework for constructing
  Galois connections using \emph{Galois transformers}, an extension of
  monad transformers which transport (1) Galois connections and (2)
  mappings to an executable transition system.
\item
  Two new general purpose monad transformers for nondeterminism which
  are not present in any previous work on monad transformers (even
  outside static analysis literature). Although applicable to settings
  other than static analysis, these two transformers give rise naturally
  to variations in path and flow sensitivity when applied to abstract
  interpreters.
\item
  An isolated understanding of path and flow sensitivity in analysis as
  properties of the interpreter monad, which we develop independently of
  other analysis features.
\end{itemize}

\par

\section{Semantics}\label{semantics}

\par

To demonstrate our framework we design an abstract interpreter for
$ \ttbfop{\lambda{}IF} $, a simple applied lambda calculus shown in
Figure~\ref{SS}. $ \ttbfop{\lambda{}IF} $ extends traditional lambda
calculus with integers, addition, subtraction and conditionals. We write
$ \keywordop{ \mathbin{\bm{@}} } $ as explicit abstract syntax for
function application. The state-space $ \Sigma $ for
$ \ttbfop{\lambda{}IF} $ makes allocation explicit using two separate
stores for values ($ \itop{Store} $) and for the stack
($ \itop{KStore} $).

\par

\begin{figure}

\small\begin{alignat*}{3}
i  \in  &  \mathbb{Z}  \\
x  \in  &  \itop{Var}  \\
a  \in  &  \itop{Atom}  &&  \Coloneqq  i  \;|\;  x  \;|\;   \keyword{\lambda} (x).e \\
 \oplus   \in  &  \itop{IOp}  &&  \Coloneqq   \keywordop{+}   \;|\;   \keywordop{-}  \\
 \odot   \in  &  \itop{Op}  &&  \Coloneqq   \oplus   \;|\;   \keywordop{ \mathbin{\bm{@}} }  \\
e  \in  &  \itop{Exp}  &&  \Coloneqq  a  \;|\;  e  \odot  e  \;|\;   \keyword{if0} (e) \{ e \}  \{ e \}  \\
 \  \\
 \tau   \in  &  \itop{Time}  &&  \coloneqq   \mathbb{Z}  \\
l  \in  &  \itop{Addr}  &&  \coloneqq   \itop{Var}   \times   \itop{Time}  \\
 \rho   \in  &  \itop{Env}  &&  \coloneqq   \itop{Var}   \rightharpoonup   \itop{Addr}  \\
 \sigma   \in  &  \itop{Store}  &&  \coloneqq   \itop{Addr}   \rightharpoonup   \itop{Val}  \\
c  \in  &  \itop{Clo}  &&  \Coloneqq   \langle  \keyword{\lambda} (x).e, \rho  \rangle  \\
v  \in  &  \itop{Val}  &&  \Coloneqq  i  \;|\;  c \\
 \kappa l  \in  &  \itop{KAddr}  &&  \coloneqq   \itop{Time}  \\
 \kappa  \sigma   \in  &  \itop{KStore}  &&  \coloneqq   \itop{KAddr}   \rightharpoonup   \itop{Frame}   \times   \itop{KAddr}  \\
fr  \in  &  \itop{Frame}  &&  \Coloneqq   \langle  \square   \odot  e \rangle   \;|\;   \langle v  \odot   \square  \rangle   \;|\;   \langle  \keyword{if0} ( \square ) \{ e \}  \{ e \}  \rangle  \\
 \varsigma   \in  &  \Sigma  &&  \Coloneqq   \itop{Exp}   \times   \itop{Env}   \times   \itop{Store}   \times   \itop{KAddr}   \times   \itop{KStore} 
\end{alignat*}\normalsize

\caption{ $ \ttbfop{\lambda{}IF} $ Syntax and Concrete State Space }
\label{SS} \end{figure}

\par

Guided by the syntax and semantics of $ \ttbfop{\lambda{}IF} $ defined
in this section we develop interpretation parameters in Section
\ref{analysis-parameters}, a monadic interpreter in Section
\ref{the-interpreter}, and both concrete and abstract instantiations for
the interpretation parameters in Section \ref{recovering-analyses}. The
variations in path and flow sensitivity developed in sections
\ref{varying-path-and-flow-sensitivity} and
\ref{a-compositional-monadic-framework} are independent of this (or any
other) semantics.

\par

We give semantics to atomic expressions and primitive operators
denotationally through
$A \llbracket  \underline{\hspace{0.5em}}  \rrbracket $ and
$ \delta  \llbracket  \underline{\hspace{0.5em}}  \rrbracket $, and to
compound expressions relationally through $ \rightsquigarrow $, as shown
in Figure \ref{ConcreteSemantics}. We will recover these semantics from
a concrete instantiation of our generic abstract interpreter in Section
\ref{recovering-analyses}.

\par

\begin{figure}

\small\begin{align*}
&\hspace{0em} A \llbracket  \underline{\hspace{0.5em}}  \rrbracket   \in   \itop{Atom}   \rightarrow   (\itop{Env}   \times   \itop{Store}   \rightharpoonup   \itop{Val})  \\
&\hspace{0em} A \llbracket i \rrbracket ( \rho , \sigma )  \coloneqq  i \\
&\hspace{0em} A \llbracket x \rrbracket ( \rho , \sigma )  \coloneqq   \sigma ( \rho (x)) \\
&\hspace{0em} A \llbracket  \keyword{\lambda} (x).e \rrbracket ( \rho , \sigma )  \coloneqq   \langle  \keyword{\lambda} (x).e, \rho  \rangle   \\
&\hspace{0em}  \delta  \llbracket  \underline{\hspace{0.5em}}  \rrbracket   \in   \itop{IOp}   \rightarrow  ( \mathbb{Z}   \times   \mathbb{Z}   \rightarrow   \mathbb{Z} ) \\
&\hspace{0em}  \delta  \llbracket  \keywordop{+}  \rrbracket (i _1 ,i _2 )  \coloneqq  i _1  + i _2  \\
&\hspace{0em}  \delta  \llbracket  \keywordop{-}  \rrbracket (i _1 ,i _2 )  \coloneqq  i _1  - i _2  \\
&\hspace{0em}  \underline{\hspace{0.5em}}{  \rightsquigarrow  }\underline{\hspace{0.5em}}   \in   \mathcal{P} ( \Sigma   \times   \Sigma ) \\
&\hspace{0em}  \langle e _1   \odot  e _2 , \rho , \sigma , \kappa l, \kappa  \sigma , \tau  \rangle   \rightsquigarrow   \langle e _1 , \rho , \sigma , \tau , \kappa  \sigma ', \tau +1 \rangle   \;\;\text{where}\;\;   \\
&\hspace{2em}  \kappa  \sigma '  \coloneqq   \kappa  \sigma  \lbrack  \tau   \mapsto  ( \langle  \square   \odot  e _2  \rangle , \kappa l) \rbrack  \\
&\hspace{0em}  \langle a, \rho , \sigma , \kappa l, \kappa  \sigma , \tau  \rangle   \rightsquigarrow   \langle e, \rho , \sigma , \tau , \kappa  \sigma ', \tau +1 \rangle   \;\;\text{where}\;\;  \\
&\hspace{2em}  \begin{aligned}[l]  & ( \langle  \square   \odot  e \rangle , \kappa l')  \coloneqq   \kappa  \sigma ( \kappa l)  \\  &  \kappa  \sigma '  \coloneqq   \kappa  \sigma  \lbrack  \tau   \mapsto  ( \langle A \llbracket a \rrbracket ( \rho , \sigma )  \odot   \square  \rangle , \kappa l') \rbrack   \end{aligned}  \\
&\hspace{0em}  \langle a, \rho , \sigma , \kappa l, \kappa  \sigma , \tau  \rangle   \rightsquigarrow   \langle e, \rho '', \sigma ', \kappa l', \kappa  \sigma , \tau +1 \rangle   \;\;\text{where}\;\;   \\
&\hspace{2em}  \begin{aligned}[l]  & ( \langle  \langle  \keyword{\lambda} (x).e, \rho ' \rangle   \keywordop{ \mathbin{\bm{@}} }   \square  \rangle , \kappa l')  \coloneqq   \kappa  \sigma ( \kappa l)  \\  &  \rho ''  \coloneqq   \rho ' \lbrack x  \mapsto  (x, \tau ) \rbrack   \\  &  \sigma '  \coloneqq   \sigma  \lbrack (x, \tau )  \mapsto  A \llbracket a \rrbracket ( \rho , \sigma ) \rbrack   \end{aligned}  \\
&\hspace{0em}  \langle i _2 , \rho , \sigma , \kappa l, \kappa  \sigma , \tau  \rangle   \rightsquigarrow   \langle i, \rho , \sigma , \kappa l', \kappa  \sigma , \tau +1 \rangle   \;\;\text{where}\;\;   \\
&\hspace{2em}  \begin{aligned}[l]  & ( \langle i _1   \oplus   \square  \rangle , \kappa l')  \coloneqq   \kappa  \sigma ( \kappa l)  \\  & i  \coloneqq   \delta  \llbracket  \oplus  \rrbracket (i _1 ,i _2 )  \end{aligned}  \\
&\hspace{0em}  \langle i, \rho , \sigma , \kappa l, \kappa  \sigma , \tau  \rangle   \rightsquigarrow   \langle e, \rho , \sigma , \kappa l', \kappa  \sigma , \tau +1 \rangle   \;\;\text{where}\;\;   \\
&\hspace{2em}  \begin{aligned}[l]  & ( \langle  \keyword{if0} ( \square ) \{ e _1  \}  \{ e _2  \}  \rangle , \kappa l')  \coloneqq   \kappa  \sigma ( \kappa l)  \\  & e  \coloneqq  e _1   \;\;\text{when}\;\;  i = 0  \;;\;  e _2   \;\;\text{when}\;\;  i  \neq  0  \end{aligned} 
\end{align*}\normalsize

\caption{Concrete Semantics}
\label{ConcreteSemantics}

\end{figure}

\par

Our abstract interpreter will support abstract garbage
collection~\cite{dvanhorn:Might:2006:GammaCFA}, the concrete analogue of
which is just standard garbage collection. We include abstract garbage
collection for two reasons. First, it is one of the few techniques that
results in both performance \emph{and} precision improvements for
abstract interpreters. Second, later we will systematically recover both
concrete and abstract garbage collectors through a single monadic
garbage collector.

\par

Garbage collection is defined using a reachability function $ \itop{R} $
which computes transitively reachable address in $ \sigma $:

\small\begin{align*}
&\hspace{0em}  \itop{R}   \in   \itop{Store}   \times   \itop{Env}   \times   \itop{Exp}   \rightarrow   \mathcal{P}  (\itop{Addr})  \\
&\hspace{0em}  \itop{R}(  \sigma , \rho ,e)  \coloneqq   \mu (X).  \\
&\hspace{2em} X  \cup   \itop{R} _0  ( \rho ,e)  \cup   \{ l'  \;|\;  l'  \in   \itop{R-Val}(  \sigma (l))  \;;\;  l  \in  X \} 
\end{align*}\normalsize

We write $ \mu (X). f(X)$ as the least-fixed-point of a function $f$.
This definition uses two helper functions: $ \itop{R} _0  $ for
computing the initial reachable set and $ \itop{R-Val} $ for computing
addresses reachable from values.

\small\begin{align*}
&\hspace{0em}  \itop{R} _0    \in   \itop{Env}   \times   \itop{Exp}   \rightarrow   \mathcal{P}  (\itop{Addr})  \\
&\hspace{0em}  \itop{R} _0  ( \rho ,e)  \coloneqq   \{  \rho (x)  \;|\;  x  \in   \itop{FV}( e) \}  \\
&\hspace{0em}  \itop{R-Val}   \in   \itop{Val}   \rightarrow   \mathcal{P}  (\itop{Addr})  \\
&\hspace{0em}  \itop{R-Val}( i)  \coloneqq   \{  \}  \\
&\hspace{0em}  \itop{R-Val}(  \langle  \keyword{\lambda} (x).e, \rho  \rangle )  \coloneqq   \{  \rho (y)  \;|\;  y  \in   \itop{FV}(  \keyword{\lambda} (x).e) \} 
\end{align*}\normalsize

We omit the definition of $ \itop{FV} $, which is the standard recursive
definition for computing free variables of an expression. Analogously,
$ \itop{KR} $ is the set of transitively reachable stack-frame addresses
in $ \kappa  \sigma $:

\small\begin{align*}
&\hspace{0em}  \itop{KR}   \in   \itop{KStore}   \times   \itop{KAddr}   \rightarrow   \mathcal{P}  (\itop{KAddr})  \\
&\hspace{0em}  \itop{KR}(  \kappa  \sigma , \kappa l _0 )  \coloneqq   \mu (X). X  \cup   \{  \kappa l _0  \}   \cup   \{  \pi  _2 ( \kappa  \sigma ( \kappa l))  \;|\;   \kappa l  \in  X \} 
\end{align*}\normalsize

\par

Our final semantics is given via the step relation
$ \underline{\hspace{0.5em}}{  \rightsquigarrow   ^{ \itop{gc} }   }\underline{\hspace{0.5em}} $
which nondeterministically either takes a semantic step or performs
garbage collection.

\small\begin{align*}
&\hspace{0em}  \underline{\hspace{0.5em}}{  \rightsquigarrow   ^{ \itop{gc} }   }\underline{\hspace{0.5em}}   \in   \mathcal{P} ( \Sigma   \times   \Sigma ) \\
&\hspace{0em}  \varsigma   \rightsquigarrow   ^{ \itop{gc} }    \varsigma '  \;\;\text{where}\;\;   \varsigma   \rightsquigarrow   \varsigma ' \\
&\hspace{0em}  \langle e, \rho , \sigma , \kappa l, \kappa  \sigma , \tau  \rangle   \rightsquigarrow   ^{ \itop{gc} }    \langle e, \rho , \sigma ', \kappa l, \kappa  \sigma ', \tau  \rangle   \;\;\text{where}\;\;   \\
&\hspace{2em}  \begin{aligned}[l]  &  \sigma '  \coloneqq   \{ l  \mapsto   \sigma (l)  \;|\;  l  \in   \itop{R}(  \sigma , \rho ,e) \}   \\  &  \kappa  \sigma '  \coloneqq   \{  \kappa l  \mapsto   \kappa  \sigma ( \kappa l)  \;|\;   \kappa l  \in   \itop{KR}(  \kappa  \sigma , \kappa l) \}   \end{aligned} 
\end{align*}\normalsize

An execution of the semantics is the least-fixed-point of a collecting
semantics:

\small\begin{align*}
&\hspace{0em}  \itop{Analysis}   \coloneqq   \mu (X).X  \cup   \{  \varsigma  _0  \}   \cup   \{   \varsigma '  \;|\;   \varsigma   \rightsquigarrow   ^{ \itop{gc} }    \varsigma '  \;;\;   \varsigma   \in  X  \} 
\end{align*}\normalsize

where $ \varsigma  _0 $ is the injection of the initial program $e _0 $:

\small\begin{align*}
&\hspace{0em}  \varsigma  _0   \coloneqq   \langle e _0 , \bot , \bot ,0, \bot ,1 \rangle 
\end{align*}\normalsize

The analyses we present in this paper will be proven correct in Section
\ref{recovering-analyses} by establishing a Galois connection with this
concrete collecting semantics.

\par

\section{Path and Flow Sensitivity in
Analysis}\label{path-and-flow-sensitivity-in-analysis}

\par

In this paper we identify three specific variants of path and flow
sensitivity in analysis: path-sensitive, flow-sensitive and
flow-insensitive. Our framework exposes the essence of path and flow
sensitivity through a monadic effect interface in Section
\ref{analysis-parameters}, and we recover each of these variations
through specific monad instances in Sections
\ref{varying-path-and-flow-sensitivity} and
\ref{a-compositional-monadic-framework}.

\par

Consider a combination of if-statements in our example language
$ \ttbfop{\lambda{}IF} $ (extended with let-bindings) where an analysis
cannot determine the value of $N$: \small\begin{alignat*}{3}

& 1 \mathbin{\colon}   \keyword{let}\;\;  x  \coloneqq            &&  \hspace{1em}  \hspace{1em}  \keyword{in}                  \\
&  \hspace{1em}  \hspace{1em} 2 \mathbin{\colon}   \keyword{if0} (N) \{           &&  \hspace{1em}  \hspace{1em} 5 \mathbin{\colon}   \keyword{let}\;\;  y  \coloneqq         \\
&  \hspace{1em}  \hspace{1em}  \hspace{1em}  \hspace{1em} 3 \mathbin{\colon}   \keyword{if0} (N) \{ 1 \}  \{ 2 \}    &&  \hspace{1em}  \hspace{1em}  \hspace{1em}  \hspace{1em} 6 \mathbin{\colon}   \keyword{if0} (N) \{ 5 \}  \{ 6 \}   \\
&  \hspace{1em}  \hspace{1em}  \}   \;\;\keyword{\litelse}\;\;   \{             &&  \hspace{1em}  \hspace{1em}  \keyword{in}                  \\
&  \hspace{1em}  \hspace{1em}  \hspace{1em}  \hspace{1em} 4 \mathbin{\colon}   \keyword{if0} (N) \{ 3 \}  \{ 4 \}   \}  &&  \hspace{1em}  \hspace{1em} 7 \mathbin{\colon}   \keyword{exit} (x, y)

\end{alignat*}\normalsize \paragraph{Path-Sensitive} A path-sensitive
analysis tracks both data and control flow precisely. At program points
3 and 4 the analysis considers separate worlds:

\small\begin{alignat*}{1}
3 \mathbin{\colon}   \{ N=0 \}  \quad 4 \mathbin{\colon}   \{ N \neq 0 \} 
\end{alignat*}\normalsize

At program points 5 and 6 the analysis continues in two separate,
precise worlds:

\small\begin{alignat*}{1}
5,6 \mathbin{\colon}   \{ N=0 ,\;\;  x=1 \}   \{ N \neq 0 ,\;\;  x=4 \} 
\end{alignat*}\normalsize

At program point 7 the analysis correctly correlates the values for $x$
and $y$:

\small\begin{alignat*}{1}
7 \mathbin{\colon}   \{ N=0 ,\;\;  x=1 ,\;\;  y=5 \}   \{ N \neq 0 ,\;\;  x=4 ,\;\;  y=6 \} 
\end{alignat*}\normalsize

\par

\paragraph{Flow-Sensitive}

A flow-sensitive analysis collects a \emph{single} set of facts for each
variable \emph{at each program point}. At program points 3 and 4, the
analysis considers separate worlds:

\small\begin{alignat*}{1}
3 \mathbin{\colon}   \{ N=0 \}  \quad 4 \mathbin{\colon}   \{ N \neq 0 \} 
\end{alignat*}\normalsize

Each nested if-statement then evaluates only one side of the branch,
resulting in values $1$ and $4$. At program points 5 and 6 the analysis
is only allowed one set of facts, so it must merge the possible values
that $x$ and $N$ could take:

\small\begin{alignat*}{1}
5,6 \mathbin{\colon}   \{ N \in  \mathbb{Z}  ,\;\;  x \in  \{ 1,4 \}  \} 
\end{alignat*}\normalsize

The analysis then explores both branches at program point 6 resulting in
no correlation between values for $x$ and $y$:

\small\begin{alignat*}{1}
7 \mathbin{\colon}   \{ N \in  \mathbb{Z}  ,\;\;  x \in  \{ 1,4 \}  ,\;\;  y \in  \{ 5,6 \}  \} 
\end{alignat*}\normalsize

\par

\paragraph{Flow-Insensitive}

A flow-insensitive analysis collects a \emph{single} set of facts about
each variable which must hold true \emph{for the entire program}.
Because the value of $N$ is unknown at \emph{some} point in the program,
the value of $x$ must consider both branches of the nested if-statement.
This results in the global set of facts giving four values to $x$.

\small\begin{alignat*}{1}
 \{ N \in  \mathbb{Z}  ,\;\;  x \in  \{ 1,2,3,4 \}  ,\;\;  y \in  \{ 5,6 \}  \} 
\end{alignat*}\normalsize

\par

In our framework we capture the variation of path and flow sensitivity
as a purely orthogonal parameter to the abstract interpreter. Path and
flow sensitivity will then compose seamlessly with choices of call-site
sensitivity, object sensitivity, abstract garbage collection, mcfa a la
\citet{dvanhorn:Might2010Resolving}, shape analysis, abstract domain,
etc. Most importantly, we empower the analysis designer to
compartmentalize the path and flow sensitivity of each component in the
abstract state space independently. Constructing an analysis which is
flow-sensitive in the data-store and path-sensitive in the stack-store
is just as easy as constructing a single path or flow sensitivity across
the board, and one can alternate between them for free.

\par

\section{Analysis Parameters}\label{analysis-parameters}

\par

Before writing an abstract interpreter we first design its parameters.
The interpreter, which we develop in Section \ref{the-interpreter}, will
be designed such that variations in these parameters will recover both
concrete and a family of abstract interpreters, which we show in Section
\ref{recovering-analyses}. To do this we extend the ideas developed in
\citet{davdar:van-horn:2010:aam} with a new parameter for path and flow
sensitivity.

\par

There will be three parameters to our abstract interpreter, one of which
is novel in this work:

\par

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\itemsep1pt\parskip0pt\parsep0pt
\item
  The monad, novel in this work, captures the state and control effects
  of the interpreter and gives rise to path and flow sensitivity.
\item
  The abstract domain, which for this language is merely an abstraction
  for integers.
\item
  Abstract Time, capturing call-site and object sensitivities.
\end{enumerate}

\par

We place each of these parameters behind an abstract interface and leave
their implementations opaque when defining the monadic interpreter in
Section \ref{the-interpreter}. Each of these parameters come with
reasoning principles. These principles allow us to reason about the
correctness of the generic interpreter independent of a particular
instantiation of the parameters. The goal is to factor as much of the
proof effort into what we can say about the generic interpreter
independent of parameter instantiation. Then, an instantiation of the
interpreter need only justify that each parameter meets its local
interface.

\par

\subsection{The Analysis Monad}\label{the-analysis-monad}

\par

The monad for the interpreter captures the \emph{effects} of
interpretation. There are two effects in the interpreter: state and
nondeterminism. The state effect will mediate how the interpreter
interacts with state cells in the state space: $ \itop{Env} $,
$ \itop{Store} $, $ \itop{KAddr} $ and $ \itop{KStore} $. The
nondeterminism effect will mediate branching in the execution of the
interpreter. One of our results is that path and flow sensitivity can be
recovered by altering how these effects interact in the monad.

\par

We use monadic state and nondeterminism effects to abstract over
arbitrary \emph{relational small-step state-machine semantics}. State
effects correspond to the components of the state-machine and
nondeterminism effects correspond to potential nondeterminism in the
relation's definition.

\par

We briefly review monad, state and nondeterminism operators and their
laws. For more details about monad transformers and monad laws we refer
the reader to \cite{dvanhorn:Liang1995Monad}.

\par

\paragraph{Monadic Sequencing}

A type operator $ \itop{M} $ is a monad if it supports $ \itop{bind} $,
a sequencing operator, and its unit $ \itop{return} $.

\small\begin{alignat*}{2}
 \itop{M}  &  \mathbin{\colon}   \itop{Type}   \rightarrow   \itop{Type}  \\
 \itop{bind}  &  \mathbin{\colon}   \forall   \alpha   \beta ,  \itop{M}(  \alpha )  \rightarrow  ( \alpha   \rightarrow   \itop{M}(  \beta ))  \rightarrow   \itop{M}(  \beta ) \\
 \itop{return}  &  \mathbin{\colon}   \forall   \alpha ,  \alpha   \rightarrow   \itop{M}(  \alpha )
\end{alignat*}\normalsize

We use monad laws (left and right units, and associativity) to reason
about the interpreter in the absence of a particular implementation of
$ \itop{bind} $ and $ \itop{return} $. We use semicolon notation as
syntactic sugar for $ \itop{bind} $; for example,
$a  \leftarrow  m  \;;\;  k(a)$ is sugar for $ \itop{bind}( m)(k)$. We
replace semicolons with line breaks headed by a $ \ttbfop{do} $ command
for multiline monadic definitions.

\par

\paragraph{State Effect}

A type operator $ \itop{M} $ supports the monadic state effect for a
type $s$ if it supports $ \itop{get} $ and $ \itop{put} $ actions over
$s$. \small\begin{alignat*}{4}

   \itop{M}  &  \mathbin{\colon}   \itop{Type}   \rightarrow   \itop{Type}  &  \hspace{1em}   \hspace{1em} \itop{get}  &  \mathbin{\colon}   \itop{M}( s)       \\
  s &  \mathbin{\colon}   \itop{Type}         &  \hspace{1em}   \hspace{1em} \itop{put}  &  \mathbin{\colon}  s  \rightarrow   \itop{M}( 1)

\end{alignat*}\normalsize We use the state monad laws (get-get, get-put,
put-get, put-put) to reason about state effects.

\par

\paragraph{Nondeterminism Effect}

A type operator $ \itop{M} $ supports the monadic nondeterminism effect
if it supports an alternation operator $ \mathbin{\langle + \rangle} $
and its unit $ \itop{mzero} $.

\small\begin{alignat*}{2}
 \itop{M}  &  \mathbin{\colon}   \itop{Type}   \rightarrow   \itop{Type}  \\
 \underline{\hspace{0.5em}}{  \mathbin{\langle + \rangle}  }\underline{\hspace{0.5em}}  &  \mathbin{\colon}   \forall   \alpha ,  \itop{M}(  \alpha )  \times   \itop{M}(  \alpha )  \rightarrow   \itop{M}(  \alpha ) \\
 \itop{mzero}  &  \mathbin{\colon}   \forall   \alpha ,  \itop{M}(  \alpha )
\end{alignat*}\normalsize

Nondeterminism laws state that $ \itop{M}(  \alpha )$ must have a
join-semilattice structure, that $ \itop{mzero} $ be a zero for
$ \itop{bind} $, and that $ \itop{bind} $ distributes through
$ \mathbin{\langle + \rangle} $.

\par

\paragraph{Monad Examples}

The state monad over $s$ is defined with
$s  \rightarrow  ( \alpha   \times  s)$ and supports monadic sequencing
($ \itop{bind} $ and $ \itop{return} $) and state effects
($ \itop{get} $ and $ \itop{put} $). The nondeterminism monad is defined
with $ \mathcal{P} ( \alpha )$ and supports monadic sequencing
($ \itop{bind} $ and $ \itop{return} $) and nondeterminism effects
($ \underline{\hspace{0.5em}}{  \mathbin{\langle + \rangle}  }\underline{\hspace{0.5em}} $
and $ \itop{mzero} $).

\par

Our interpreter will be defined up to this effect interface and will
avoid referencing an explicit configuration $ \varsigma $ or collections
of results. This level of indirection will then be exploited in Section
\ref{varying-path-and-flow-sensitivity}, where different monads will
meet the same effect interface but yield different analysis properties.

\par

\subsection{The Abstract Domain}\label{the-abstract-domain}

\par

The abstract domain is the $ \itop{Val} $ type in the semantics. To
parameterize over the abstract domain we make $ \itop{Val} $ opaque, but
require that it support various operations.

\par

$ \itop{Val} $ must be a join-semilattice with $ \bot $ and $ \sqcup $
respecting the usual laws. We require $ \itop{Val} $ to be a
join-semilattice so it can be merged in updates to $ \itop{Store} $ to
preserve soundness.

\small\begin{alignat*}{2}
 \bot   \mathbin{\colon}   \itop{Val}  &  \hspace{1em}  \hspace{1em}  \underline{\hspace{0.5em}}{  \sqcup  }\underline{\hspace{0.5em}}   \mathbin{\colon}   \itop{Val}   \times   \itop{Val}   \rightarrow   \itop{Val} 
\end{alignat*}\normalsize

\par

$ \itop{Val} $ must also support conversions to and from concrete
values. These conversions take the form of introduction and elimination
rules. Introduction rules inject concrete values into abstract values.
Elimination rules project abstract values into a \emph{finite} set of
concrete observations. For example, we do not require that abstract
values support elimination to integers, only the finite observation of
comparison with zero.

\small\begin{alignat*}{4}
 \itop{int-I}  &  \mathbin{\colon}   \mathbb{Z}   \rightarrow   \itop{Val}  &&  \hspace{1em}   \hspace{1em} \itop{int-if0-E}  &&  \mathbin{\colon}   \itop{Val}   \rightarrow   \mathcal{P} (Bool) \\
 \itop{clo-I}  &  \mathbin{\colon}   \itop{Clo}   \rightarrow   \itop{Val}  &&  \hspace{1em}   \hspace{1em} \itop{clo-E}  &&  \mathbin{\colon}   \itop{Val}   \rightarrow   \mathcal{P}  (\itop{Clo}) 
\end{alignat*}\normalsize

\par

The laws for the introduction and elimination rules induce a Galois
connection between $ \mathcal{P} ( \mathbb{Z} )$ and $ \itop{Val} $:

\small\begin{alignat*}{2}
 \{  \ttbfop{true}  \}  &  \sqsubseteq   \itop{int-if0-E}  (\itop{int-I}( i))  \ttbfop{if}  i = 0 \\
 \{  \ttbfop{false}  \}  &  \sqsubseteq   \itop{int-if0-E}  (\itop{int-I}( i))  \ttbfop{if}  i  \neq  0 \\
 \bigsqcup  _{\substack{ b  \in   \itop{int-if0-E}( v)  \\  i  \in   \theta (b) }}   \itop{int-I}( i) &  \sqsubseteq  v \\
 \;\;\text{where}\;\;   \begin{aligned}[l]   \theta  (\ttbfop{true})   \\   \theta  (\ttbfop{false})   \end{aligned}  &  \begin{aligned}[l]  &  \coloneqq   \{ 0 \}   \\  &  \coloneqq   \{ i  \;|\;  i  \in   \mathbb{Z}   \;;\;  i  \neq  0 \}   \end{aligned} 
\end{alignat*}\normalsize

Closures must follow similar laws, inducing a Galois connection between
$ \mathcal{P}  (\itop{Clo}) $ and $ \itop{Val} $:

\small\begin{alignat*}{2}
 \{ c \}  &  \sqsubseteq   \itop{clo-E}( cloI(c)) \\
 \bigsqcup  _{\substack{ c  \in   \itop{clo-E}( v) }}   \itop{clo-I}( c) &  \sqsubseteq  v
\end{alignat*}\normalsize

Finally, $ \delta $ must be sound and complete w.r.t. the Galois
connection between concrete values and $ \itop{Val} $:

\small\begin{alignat*}{2}
 \itop{int-I}( i _1  + i _2 ) &  \sqsubseteq   \delta  \llbracket  \keywordop{+}  \rrbracket  (\itop{int-I}( i _1 ) ,\itop{int-I}( i _2 )) \\
 \itop{int-I}( i _1  - i _2 ) &  \sqsubseteq   \delta  \llbracket  \keywordop{-}  \rrbracket  (\itop{int-I}( i _1 ) ,\itop{int-I}( i _2 )) \\
 \bigsqcup  _{\substack{ b _1   \in   \itop{int-if0-E}( v _1 )  \\  b _2   \in   \itop{int-if0-E}( v _2 )  \\  i  \in   \theta (b _1 ,b _2 ) }}   \itop{int-I}( i) &  \sqsubseteq   \delta  \llbracket  \odot  \rrbracket (v _1 ,v _2 ) \\
 \;\;\text{where}\;\;   \begin{aligned}[l]   \theta (  \ttbfop{true}  ,  \ttbfop{true}  )  \\   \theta (  \ttbfop{true}  ,  \ttbfop{false}  )  \\   \theta (  \ttbfop{false}  ,  \ttbfop{true}  )  \\   \theta (  \ttbfop{false}  ,  \ttbfop{false}  )  \end{aligned}  &  \begin{aligned}[l]  &  \coloneqq   \{ 0 \}   \\  &  \coloneqq   \{ i  \;|\;  i  \in   \mathbb{Z}   \;;\;  i  \neq  0  \}   \\  &  \coloneqq   \{ i  \;|\;  i  \in   \mathbb{Z}   \;;\;  i  \neq  0 \}   \\  &  \coloneqq   \mathbb{Z}   \end{aligned} 
\end{alignat*}\normalsize

\par

Supporting additional primitive types like booleans, lists, or arbitrary
inductive datatypes is analogous. Introduction functions inject the type
into $ \itop{Val} $ and elimination functions project a finite set of
discrete observations. Introduction, elimination and $ \delta $
operators must all be sound and complete following a Galois connection
discipline.

\par

\subsection{Abstract Time}\label{abstract-time}

\par

The interface for abstract time is familiar from
\citet{davdar:van-horn:2010:aam} (AAM), which introduces abstract time
as a single parameter to control various forms of context sensitivity,
and \citet{dvanhorn:Smaragdakis2011Pick}, which instantiates the
parameter to achieve both call-site and object sensitivity. We only
demonstrate call-site sensitivity in this presentation, but our
language-independent library supports object sensitivity, the
implementation of which is a straightforward application of
\citet{dvanhorn:Smaragdakis2011Pick}.

\small\begin{alignat*}{1}
 \itop{Time}   \mathbin{\colon}   \itop{Type}   \hspace{1em}  \hspace{1em}   \itop{tick}   \mathbin{\colon}   \itop{Exp}   \times   \itop{KAddr}   \times   \itop{Time}   \rightarrow   \itop{Time} 
\end{alignat*}\normalsize

Remarkably, we need not state laws for $ \itop{tick} $. The interpreter
will merge values which reside at the same address to preserve
soundness. Therefore, any supplied implementations of $ \itop{tick} $ is
valid from a soundness perspective. However, different choices in
$ \itop{tick} $ will yield different trade-offs in precision and
performance of the abstract interpreter.

\par

\section{The Interpreter}\label{the-interpreter}

\par

We now present a monadic interpreter for $ \ttbfop{\lambda{}IF} $
parameterized over $ \itop{M} $, $ \itop{Val} $ and $ \itop{Time} $ from
Section \ref{analysis-parameters}. We instantiate these parameters to
obtain an analysis in Section \ref{recovering-analyses}.

\par

First we implement
$A \llbracket  \underline{\hspace{0.5em}}  \rrbracket $ as a
\emph{monadic} denotation for atomic expressions, shown in Figure
\ref{InterpreterStep}. The monadic
$A \llbracket  \underline{\hspace{0.5em}}  \rrbracket $ is a
straightforward translation of the
$A \llbracket  \underline{\hspace{0.5em}}  \rrbracket $ shown in Figure
\ref{ConcreteSemantics} from a pure function to a monadic function with
state effects. $ \itop{get-Env} $ and $ \itop{get-Store} $ are primitive
operations for monadic state. $ \itop{clo-I} $ comes from the interface
for $ \itop{Val} $.

\par

Next we implement $ \itop{step} $, a \emph{monadic} small-step
\emph{function} for compound expressions, also shown in Figure
\ref{InterpreterStep}. The monadic $ \itop{step} $ is a straightforward
translation of the $ \itop{step} $ shown in Figure
\ref{ConcreteSemantics} from a relation to a monadic function with state
and nondeterminism effects.

\par

\begin{figure}

\small\begin{align*}
&\hspace{0em} A \llbracket  \underline{\hspace{0.5em}}  \rrbracket   \in   \itop{Atom}   \rightarrow   \itop{M}(  \itop{Val})  \\
&\hspace{0em} A \llbracket i \rrbracket   \coloneqq   \itop{return}  (\itop{int-I}( i)) \\
&\hspace{0em} A \llbracket x \rrbracket   \coloneqq   \ttbfop{do}  \\
&\hspace{2em}  \rho   \leftarrow   \itop{get-Env}  \\
&\hspace{2em}  \sigma   \leftarrow   \itop{get-Store}  \\
&\hspace{2em}  \ttbfop{if}  x  \in   \rho   \;\ttbfop{then}\;   \itop{return}(  \sigma ( \rho (x)))  \;\ttbfop{\litelse}\;   \itop{return}(  \bot ) \\
&\hspace{0em} A \llbracket  \keyword{\lambda} (x).e \rrbracket   \coloneqq   \ttbfop{do}  \\
&\hspace{2em}  \rho   \leftarrow   \itop{get-Env}  \\
&\hspace{2em}  \itop{return}  (\itop{clo-I}(  \langle  \keyword{\lambda} (x).e, \rho  \rangle )) \\
&\hspace{0em}  \itop{step}   \mathbin{\colon}   \itop{Exp}   \rightarrow   \itop{M}  (\itop{Exp})  \\
&\hspace{0em}  \itop{step}( e)  \coloneqq   \ttbfop{do}  \\
&\hspace{2em}  \itop{tickM}( e) \\
&\hspace{2em} e'  \leftarrow   \ttbfop{case}  e  \ttbfop{of}  \\
&\hspace{3em} e _1   \odot  e _2   \rightarrow   \ttbfop{do}  \\
&\hspace{4em}  \itop{push}(  \langle  \square   \odot  e _2  \rangle ) \\
&\hspace{4em}  \itop{return}( e _1 ) \\
&\hspace{3em} a  \rightarrow   \ttbfop{do}  \\
&\hspace{4em} fr  \leftarrow   \itop{pop}  \\
&\hspace{4em} v  \leftarrow  A \llbracket a \rrbracket  \\
&\hspace{4em} e  \leftarrow   \ttbfop{case}  fr  \ttbfop{of}  \\
&\hspace{5em}  \langle  \square   \odot  e \rangle   \rightarrow   \ttbfop{do}  \\
&\hspace{6em}  \itop{push}(  \langle v  \odot   \square  \rangle ) \\
&\hspace{6em}  \itop{return}( e) \\
&\hspace{5em}  \langle v'  \keywordop{ \mathbin{\bm{@}} }   \square  \rangle   \rightarrow   \ttbfop{do}  \\
&\hspace{6em}  \langle  \keyword{\lambda} (x).e, \rho ' \rangle   \leftarrow   \itop{\uparrow_p}  (\itop{clo-E}( v')) \\
&\hspace{6em}  \tau   \leftarrow   \itop{get-Time}  \\
&\hspace{6em}  \sigma   \leftarrow   \itop{get-Store}  \\
&\hspace{6em}  \itop{put-Env}(  \rho ' \lbrack x  \mapsto  (x, \tau ) \rbrack ) \\
&\hspace{6em}  \itop{put-Store}(  \sigma   \sqcup   \lbrack (x, \tau )  \mapsto   \{ v \}  \rbrack ) \\
&\hspace{6em}  \itop{return}( e) \\
&\hspace{5em}  \langle v'  \oplus   \square  \rangle   \rightarrow   \ttbfop{do}  \\
&\hspace{6em}  \itop{return}(  \delta  \llbracket  \oplus  \rrbracket (v',v)) \\
&\hspace{5em}  \langle  \keyword{if0} ( \square ) \{ e _1  \}  \{ e _2  \}  \rangle   \rightarrow   \ttbfop{do}  \\
&\hspace{6em} b  \leftarrow   \itop{\uparrow_p}  (\itop{int-if0-E}( v)) \\
&\hspace{6em}  \ttbfop{if}( b)  \;\ttbfop{then}\;   \itop{return}( e _1 )  \;\ttbfop{\litelse}\;   \itop{return}( e _2 ) \\
&\hspace{2em}  \itop{gc}( e') \\
&\hspace{2em}  \itop{return}( e') \\
&\hspace{0em}  \itop{gc}   \mathbin{\colon}   \itop{Exp}   \rightarrow   \itop{M}( 1) \\
&\hspace{0em}  \itop{gc}( e)  \coloneqq   \ttbfop{do}  \\
&\hspace{2em}  \rho   \leftarrow   \itop{get-Env}  \\
&\hspace{2em}  \sigma   \leftarrow   \itop{get-Store}  \\
&\hspace{2em}  \kappa  \sigma   \leftarrow   \itop{get-KStore}  \\
&\hspace{2em}  \itop{put-Store}(  \{ l  \mapsto   \sigma (l)  \;|\;  l  \in   \itop{R}(  \sigma , \rho ,e)) \\
&\hspace{2em}  \itop{put-KStore}(  \{  \kappa l  \mapsto   \kappa  \sigma ( \kappa l)  \;|\;   \kappa l  \in   \itop{KR}(  \kappa  \sigma , \kappa l) \} )
\end{align*}\normalsize

\caption{Monadic Semantics}
\label{InterpreterStep}

\end{figure}

\par

$ \itop{step} $ uses helper functions $ \itop{push} $ and $ \itop{pop} $
for manipulating stack frames, $ \itop{\uparrow_p} $ for lifting values
from $ \mathcal{P} $ into $ \itop{M} $, and a monadic version of
$ \itop{tick} $ called $ \itop{tickM} $, each of which are shown in
Figure \ref{InterpreterHelpers}. The interpreter looks deterministic,
however the nondeterminism is abstracted away behind
$ \itop{\uparrow_p} $ and monadic bind
$x  \leftarrow  e _1   \;;\;  e _2 $.

\par

We implement abstract garbage collection in a general away using the
monadic effect interface, also shown in Figure \ref{InterpreterStep}.
$ \itop{R} $ and $ \itop{KR} $ are as defined in Section
\ref{semantics}. Again, this is a straightforward translation from a
pure function to a monadic function with state effects.

\par

\paragraph{Preserving Soundness}

In generalizing the semantics to account for nondeterminism, updates to
both the data-store and stack-store must merge values rather than
performing a strong update. This is because we place no restriction on
the semantics for $ \itop{Time} $ and therefore must preserve soundness
in the presence of reused addresses.

\par

To support $ \sqcup $ for stores (in observation of soundness) we modify
our definitions of $ \itop{Store} $ and $ \itop{KStore} $.

\small\begin{align*}
&\hspace{0em}  \sigma    \in   \itop{Store}    \mathbin{\colon}   \itop{Addr}   \rightarrow   \itop{Val}  \\
&\hspace{0em}  \kappa  \sigma   \in   \itop{KStore}   \mathbin{\colon}   \itop{KAddr}   \rightarrow   \mathcal{P}  (\itop{Frame}   \times   \itop{KAddr}) 
\end{align*}\normalsize

\par

\begin{figure}

\small\begin{align*}
&\hspace{0em}  \itop{\uparrow_p}   \mathbin{\colon}   \forall   \alpha ,  \mathcal{P} ( \alpha )  \rightarrow   \itop{M}(  \alpha ) \\
&\hspace{0em}  \itop{\uparrow_p} ( \{ a _1  .. a _n  \} )  \coloneqq   \itop{return}( a _1 )  \mathbin{\langle + \rangle}  ..  \mathbin{\langle + \rangle}   \itop{return}( a _n ) \\
&\hspace{0em}  \itop{push}   \mathbin{\colon}   \itop{Frame}   \rightarrow   \itop{M}( 1) \\
&\hspace{0em}  \itop{push}( fr)  \coloneqq   \ttbfop{do}  \\
&\hspace{2em}  \kappa l  \leftarrow   \itop{get-KAddr}  \\
&\hspace{2em}  \kappa  \sigma   \leftarrow   \itop{get-KStore}  \\
&\hspace{2em}  \kappa l'  \leftarrow   \itop{get-Time}  \\
&\hspace{2em}  \itop{put-KStore}(  \kappa  \sigma   \sqcup   \lbrack  \kappa l'  \mapsto   \{ fr \mathbin{::}  \kappa l \}  \rbrack ) \\
&\hspace{2em}  \itop{put-KAddr}(  \kappa l') \\
&\hspace{0em}  \itop{pop}   \mathbin{\colon}   \itop{M}  (\itop{Frame})  \\
&\hspace{0em}  \itop{pop}   \coloneqq   \ttbfop{do}  \\
&\hspace{2em}  \kappa l  \leftarrow   \itop{get-KAddr}  \\
&\hspace{2em}  \kappa  \sigma   \leftarrow   \itop{get-KStore}  \\
&\hspace{2em} fr \mathbin{::}  \kappa l'  \leftarrow   \itop{\uparrow_p} ( \kappa  \sigma ( \kappa l)) \\
&\hspace{2em}  \itop{put-KAddr}(  \kappa l') \\
&\hspace{2em}  \itop{return}( fr) \\
&\hspace{0em}  \itop{tickM}   \mathbin{\colon}   \itop{Exp}   \rightarrow   \itop{M}( 1) \\
&\hspace{0em}  \itop{tickM}( e) =  \ttbfop{do}  \\
&\hspace{2em}  \tau   \leftarrow   \itop{get-Time}  \\
&\hspace{2em}  \kappa l  \leftarrow   \itop{get-KAddr}  \\
&\hspace{2em}  \itop{put-Time}(  \itop{tick}( e, \kappa l, \tau ))
\end{align*}\normalsize

\caption{Monadic helper functions}
\label{InterpreterHelpers}

\end{figure}

\par

\paragraph{Execution}

In the concrete semantics, execution takes the form of a
least-fixed-point computation over a collecting semantics. This in
general requires a join-semilattice structure for some $ \Sigma $ and a
transition function $ \Sigma   \rightarrow   \Sigma $. However, we no
longer have a function $ \Sigma   \rightarrow   \Sigma $; we have a
monadic function $ \itop{Exp}   \rightarrow   \itop{M}  (\itop{Exp}) $
which does not immediately admit a least-fixed-point iteration to
execute the analysis.

\par

To solve this we require that monadic actions
$ \itop{Exp}   \rightarrow   \itop{M}  (\itop{Exp}) $ form a Galois
connection with a transition system $ \Sigma   \rightarrow   \Sigma $.
This Galois connection serves two purposes. First, it allows us to
implement the analysis by converting our interpreter to the transition
system $ \Sigma   \rightarrow   \Sigma $ through $ \gamma $. Second,
this Galois connection serves to \emph{transport other Galois
connections} as part of our correctness framework. This will allow us to
construct Galois connections between monads
$m _1   \galois{\alpha}{\gamma}  m _2 $ and derive Galois connections
between transition systems
$ \Sigma  _1   \galois{\alpha}{\gamma}   \Sigma  _2 $ for free.

\par

A collecting-semantics execution of our interpreter is then defined as
the least-fixed-point iteration of $ \itop{step} $ transported through
the Galois connection:

\small\begin{align*}
&\hspace{0em}  \itop{Analysis}   \coloneqq   \mu (X). X  \sqcup   \varsigma  _0   \sqcup   \gamma  (\itop{step}) (X)
\end{align*}\normalsize

where $ \varsigma  _0 $ is the injection of the initial program $e _0 $
into $ \Sigma $ and $ \gamma $ has type
$ (\itop{Exp}   \rightarrow   \itop{M}  (\itop{Exp}) )  \rightarrow  ( \Sigma   \rightarrow   \Sigma )$.

\par

\section{Recovering Analyses}\label{recovering-analyses}

\par

In Section \ref{the-interpreter} we defined a monadic interpreter with
the uninstantiated parameters from Section \ref{analysis-parameters}:
$ \itop{M} $, $ \itop{Val} $ and $ \itop{Time} $. To recover a concrete
interpreter we instantiate these parameters to concrete components
$ \Concrete{\litM} $, $ \Concrete{\litVal} $ and
$ \Concrete{\litTime} $, and to recover an abstract interpreter we
instantiate them to abstract components $ \Abstract{\litM} $,
$ \Abstract{\litVal} $ and $ \Abstract{\litTime} $. The soundness of the
final implementation is thus factored into two steps: \vbox{

\par

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\itemsep1pt\parskip0pt\parsep0pt
\item
  Proving the parameterized monadic interpreter correct for any
  instantiation of $ \itop{M} $, $ \itop{Val} $ and $ \itop{Time} $.
\item
  Constructing Galois connections
  $ \Concrete{\litM}   \galois{\alpha}{\gamma}   \Abstract{\litM} $,
  $ \Concrete{\litVal}   \galois{\alpha}{\gamma}   \Abstract{\litVal} $
  and
  $ \Concrete{\litTime}   \galois{\alpha}{\gamma}      \Abstract{\litTime} $
  piecewise.
\end{enumerate}

\par

} The key benefit of our approach is that (1) can be proved once against
\emph{all} instantiations of $ \itop{M} $, $ \itop{Val} $ and
$ \itop{Time} $ using the reasoning principles established in Section
\ref{analysis-parameters}, greatly simplifying the proof burden when
choosing different abstract components in (2).

\par

\subsection{Recovering a Concrete
Interpreter}\label{recovering-a-concrete-interpreter}

\par

For the concrete value space we instantiate $ \itop{Val} $ to
$ \Concrete{\litVal} $:

\small\begin{align*}
&\hspace{0em} v  \in   \Concrete{\litVal}   \coloneqq   \mathcal{P}  (\Concrete{\litClo}  +  \mathbb{Z} )
\end{align*}\normalsize

The concrete value space $ \Concrete{\litVal} $ has straightforward
introduction and elimination rules:

\small\begin{align*}
&\hspace{0em}  \itop{int-I}   \mathbin{\colon}   \mathbb{Z}   \rightarrow   \Concrete{\litVal}  \\
&\hspace{0em}  \itop{int-I}( i)  \coloneqq   \{ i \}  \\
&\hspace{0em}  \itop{int-if0-E}   \mathbin{\colon}   \Concrete{\litVal}   \rightarrow   \mathcal{P} (Bool) \\
&\hspace{0em}  \itop{int-if0-E}( v)  \coloneqq   \{   \ttbfop{true}   \;|\;  0  \in  v  \}   \cup   \{   \ttbfop{false}   \;|\;   \exists  i  \in  v s.t. i  \neq  0  \} 
\end{align*}\normalsize

and a straightforward concrete $ \delta $:

\small\begin{align*}
&\hspace{0em}  \delta  \llbracket  \underline{\hspace{0.5em}}  \rrbracket   \mathbin{\colon}   \itop{IOp}   \rightarrow   \Concrete{\litVal}   \times   \Concrete{\litVal}   \rightarrow   \Concrete{\litVal}  \\
&\hspace{0em}  \delta  \llbracket  \keywordop{+}  \rrbracket (v _1 ,v _2 )  \coloneqq   \{  i _1  + i _2   \;|\;  i _1   \in  v _1   \;;\;  i _2   \in  v _2   \}  \\
&\hspace{0em}  \delta  \llbracket  \keywordop{-}  \rrbracket (v _1 ,v _2 )  \coloneqq   \{  i _1  - i _2   \;|\;  i _1   \in  v _1   \;;\;  i _2   \in  v _2   \} 
\end{align*}\normalsize

\begin{proposition} $ \Concrete{\litVal} $ satisfies the abstract domain
laws shown in Section \ref{the-abstract-domain}. \end{proposition}

\par

Concrete time $ \Concrete{\litTime} $ captures program contours as a
sequence of products of $ \itop{Exp} $ and $ \Concrete{\litKAddr} $:

\small\begin{align*}
&\hspace{0em}  \tau   \in   \Concrete{\litTime}   \coloneqq   (\itop{Exp}   \times   \itop{KAddr})  ^{ * } 
\end{align*}\normalsize

and $ \itop{tick} $ is just a cons operator:

\small\begin{align*}
&\hspace{0em}  \itop{tick}   \mathbin{\colon}   \itop{Exp}   \times   \Concrete{\litKAddr}   \times   \Concrete{\litTime}   \rightarrow   \Concrete{\litTime}  \\
&\hspace{0em}  \itop{tick}  (e, \kappa l, \tau )  \coloneqq  (e, \kappa l) \mathbin{::}  \tau 
\end{align*}\normalsize

\par

For the concrete monad we instantiate $ \itop{M} $ to a path-sensitive
$ \Concrete{\litM} $ which contains a powerset of concrete state space
components.

\small\begin{align*}
&\hspace{0em}  \psi   \in   \Psi   \coloneqq   \Concrete{\litEnv}   \times   \Concrete{\litStore}   \times   \Concrete{\litKAddr}   \times   \Concrete{\litKStore}   \times   \Concrete{\litTime}  \\
&\hspace{0em} m  \in   \Concrete{\litM}(  \alpha )  \coloneqq   \Psi   \rightarrow   \mathcal{P} ( \alpha   \times   \Psi )
\end{align*}\normalsize

Monadic operators $ \itop{bind} $ and $ \itop{return} $ encapsulate both
state-passing and set-flattening:

\small\begin{align*}
&\hspace{0em}  \itop{bind}   \mathbin{\colon}   \forall   \alpha ,  \Concrete{\litM}(  \alpha )  \rightarrow  ( \alpha   \rightarrow   \Concrete{\litM}(  \beta ))  \rightarrow   \Concrete{\litM}(  \beta ) \\
&\hspace{0em}  \itop{bind}( m)(f)( \psi )  \coloneqq   \\
&\hspace{2em}  \{ (y, \psi '')  \;|\;  (y, \psi '')  \in  f(a)( \psi ')  \;;\;  (a, \psi ')  \in  m( \psi ) \}  \\
&\hspace{0em}  \itop{return}   \mathbin{\colon}   \forall   \alpha ,  \alpha   \rightarrow   \Concrete{\litM}(  \alpha ) \\
&\hspace{0em}  \itop{return}( a)( \psi )  \coloneqq   \{ (a, \psi ) \} 
\end{align*}\normalsize

State effects return singleton sets:

\small\begin{align*}
&\hspace{0em}  \itop{get-Env}   \mathbin{\colon}   \Concrete{\litM}(  \Concrete{\litEnv})  \\
&\hspace{0em}  \itop{get-Env}(  \langle  \rho , \sigma , \kappa , \tau  \rangle )  \coloneqq   \{ ( \rho , \langle  \rho , \sigma , \kappa , \tau  \rangle ) \}  \\
&\hspace{0em}  \itop{put-Env}   \mathbin{\colon}   \Concrete{\litEnv}   \rightarrow   \mathcal{P} (1) \\
&\hspace{0em}  \itop{put-Env}(  \rho ')( \langle  \rho , \sigma , \kappa , \tau  \rangle )  \coloneqq   \{ (1, \langle  \rho ', \sigma , \kappa , \tau  \rangle ) \} 
\end{align*}\normalsize

Nondeterminism effects are implemented with set union:

\small\begin{align*}
&\hspace{0em}  \itop{mzero}   \mathbin{\colon}   \forall   \alpha ,  \Concrete{\litM}(  \alpha ) \\
&\hspace{0em}  \itop{mzero}(  \psi )  \coloneqq   \{  \}  \\
&\hspace{0em}  \underline{\hspace{0.5em}}{  \mathbin{\langle + \rangle}  }\underline{\hspace{0.5em}}   \mathbin{\colon}   \forall   \alpha ,  \Concrete{\litM}(  \alpha )  \times   \Concrete{\litM}(  \alpha )  \rightarrow   \Concrete{\litM}(  \alpha ) \\
&\hspace{0em} (m _1   \mathbin{\langle + \rangle}  m _2 )( \psi )  \coloneqq  m _1 ( \psi )  \cup  m _2 ( \psi )
\end{align*}\normalsize

\begin{proposition} $ \Concrete{\litM} $ satisfies monad, state, and
nondeterminism laws shown in Section \ref{the-analysis-monad}.
\end{proposition}

\par

Finally, we must establish a Galois connection between
$ \itop{Exp}   \rightarrow   \Concrete{\litM}(  \itop{Exp}) $ and
$ \Concrete{\Sigma}   \rightarrow   \Concrete{\Sigma} $ for some choice
of $ \Concrete{\Sigma} $. For the path-sensitive monad
$ \Concrete{\litM} $, $ \Concrete{\Sigma} $ is defined:

\small\begin{align*}
&\hspace{0em}  \Concrete{\Sigma}   \coloneqq   \mathcal{P}  (\itop{Exp}   \times   \Psi )
\end{align*}\normalsize

The Galois connection between $ \Concrete{\litM} $ and
$ \Concrete{\Sigma} $ is similar to the definition of $ \itop{bind} $:

\small\begin{align*}
&\hspace{0em}  \gamma   \mathbin{\colon}   (\itop{Exp}   \rightarrow   \Concrete{\litM}(  \itop{Exp}) )  \rightarrow  ( \Concrete{\Sigma}   \rightarrow   \Concrete{\Sigma} ) \\
&\hspace{0em}  \gamma (f)(e \psi  ^{ * } )  \coloneqq   \{ (e, \psi ')  \;|\;  (e, \psi ')  \in  f(e)( \psi )  \;;\;  (e, \psi )  \in  e \psi  ^{ * }  \}  \\
&\hspace{0em}  \alpha   \mathbin{\colon}  ( \Concrete{\Sigma}   \rightarrow   \Concrete{\Sigma} )  \rightarrow   (\itop{Exp}   \rightarrow   \Concrete{\litM}(  \itop{Exp}) ) \\
&\hspace{0em}  \alpha (f)(e)( \psi )  \coloneqq  f( \{ (e, \psi ) \} )
\end{align*}\normalsize

The injection $ \varsigma  _0 $ for a program $e _0 $ is
$ \{  \langle e _0 , \bot , \bot , \bullet , \bot , \bullet  \rangle  \} $.
\begin{proposition} $ \gamma $ and $ \alpha $ form a Galois connection.
\end{proposition}

\par

\subsection{Recovering an Abstract
Interpreter}\label{recovering-an-abstract-interpreter}

\par

We pick a simple abstraction for integers,
$ \{  \keywordop{-} ,0, \keywordop{+}  \} $, although our technique
scales seamlessly to other abstract value domains.

\small\begin{align*}
&\hspace{0em}  \Abstract{\litVal}   \coloneqq   \mathcal{P}  (\Abstract{\litClo}  +  \{  \keywordop{-} ,0, \keywordop{+}  \} )
\end{align*}\normalsize

Introduction and elimination for $ \Abstract{\litVal} $ are defined:

\small\begin{align*}
&\hspace{0em}  \itop{int-I}   \mathbin{\colon}   \mathbb{Z}   \rightarrow   \Abstract{\litVal}  \\
&\hspace{0em}  \itop{int-I}( i)  \coloneqq   \{  \keywordop{-}  \}   \ttbfop{if}  i < 0 \\
&\hspace{0em}  \itop{int-I}( i)  \coloneqq   \{ 0 \}     \ttbfop{if}  i = 0 \\
&\hspace{0em}  \itop{int-I}( i)  \coloneqq   \{  \keywordop{+}  \}   \ttbfop{if}  i > 0 \\
&\hspace{0em}  \itop{int-if0-E}   \mathbin{\colon}   \Abstract{\litVal}   \rightarrow   \mathcal{P} (Bool) \\
&\hspace{0em}  \itop{int-if0-E}( v)  \coloneqq   \{   \ttbfop{true}   \;|\;  0  \in  v  \}   \cup   \{   \ttbfop{false}   \;|\;   \keywordop{-}   \in  v  \vee   \keywordop{+}   \in  v  \} 
\end{align*}\normalsize

Introduction and elimination for $ \Abstract{\litClo} $ is identical to
the concrete domain.

\par

The abstract $ \delta $ operator is defined:

\small\begin{align*}
&\hspace{0em}  \delta   \mathbin{\colon}   \itop{IOp}   \rightarrow   \Abstract{\litVal}   \times   \Abstract{\litVal}   \rightarrow   \Abstract{\litVal}   \\
&\hspace{0em}  \delta  \llbracket  \keywordop{+}  \rrbracket (v _1 ,v _2 )  \coloneqq   \\
&\hspace{3em}  \{  i  \;|\;  0  \in  v _1   \wedge  i  \in  v _2   \}   \cup   \{  i  \;|\;  i  \in  v _1   \wedge  0  \in  v _2   \}  \\
&\hspace{2em}  \cup   \{   \keywordop{+}   \;|\;   \keywordop{+}   \in  v _1   \wedge   \keywordop{+}   \in  v _2   \}   \cup   \{   \keywordop{-}   \;|\;   \keywordop{-}   \in  v _1   \wedge   \keywordop{-}   \in  v _2   \}   \\
&\hspace{2em}  \cup   \{   \keywordop{-} ,0, \keywordop{+}   \;|\;   \keywordop{+}   \in  v _1   \wedge   \keywordop{-}   \in  v _2   \}   \\
&\hspace{2em}  \cup   \{   \keywordop{-} ,0, \keywordop{+}   \;|\;   \keywordop{-}   \in  v _1   \wedge   \keywordop{+}   \in  v _2   \} 
\end{align*}\normalsize

The definition for
$ \delta  \llbracket  \keywordop{-}  \rrbracket (v _1 ,v _2 )$ is
analogous. \begin{proposition} $ \Abstract{\litVal} $ satisfies the
abstract domain laws shown in Section~\ref{the-abstract-domain}.
\end{proposition} \begin{proposition}
$ \Concrete{\litVal}   \galois{\alpha}{\gamma}   \Abstract{\litVal} $,
and both concrete and abstract definitions for $ \itop{int-I} $,
$ \itop{int-if0-E} $ and $ \delta $ are ordered $ \sqsubseteq $
respectively through the Galois connection. \end{proposition}

\par

Next we abstract $ \itop{Time} $ to $ \Abstract{\litTime} $ as the
finite domain of k-truncated lists of execution contexts:

\small\begin{align*}
&\hspace{0em}  \Abstract{\litTime}   \coloneqq   (\itop{Exp}   \times   \Abstract{\litKAddr})  ^{ * _k  } 
\end{align*}\normalsize

The $ \itop{tick} $ operator becomes cons followed by k-truncation,
which restricts the list to the first-k elements:

\small\begin{align*}
&\hspace{0em}  \itop{tick}   \mathbin{\colon}   \itop{Exp}   \times   \Abstract{\litKAddr}   \times   \Abstract{\litTime}   \rightarrow   \Abstract{\litTime}  \\
&\hspace{0em}  \itop{tick}( e, \kappa l, \tau ) =  \lfloor (e, \kappa l) \mathbin{::}  \tau  \rfloor  _k 
\end{align*}\normalsize

This abstraction for time yields k-call-site sensitivity, or a kCFA
analysis. \begin{proposition}
$ \Concrete{\litTime}   \galois{\alpha}{\gamma}   \Abstract{\litTime} $,
and both concrete and abstract definitions for $ \itop{tick} $ are
ordered $ \sqsubseteq $ through the Galois connection. \end{proposition}

\par

The monad $ \Abstract{\litM} $ need not change in implementation from
$ \Concrete{\litM} $; they are identical up the choice of $ \Psi $.

\small\begin{align*}
&\hspace{0em}  \psi   \in   \Psi   \coloneqq   \Abstract{\litEnv}   \times   \Abstract{\litStore}   \times   \Abstract{\litKAddr}   \times   \Abstract{\litKStore}   \times   \Abstract{\litTime} 
\end{align*}\normalsize

The resulting state space $ \Abstract{\Sigma} $ is finite and its
least-fixed-point iteration will give a sound and computable analysis.

\par

\section{Varying Path and Flow
Sensitivity}\label{varying-path-and-flow-sensitivity}

\par

Sections \ref{the-interpreter} and \ref{recovering-analyses} describe
the construction of a path-sensitive analysis using our framework. In
this section we show an alternate definition for $ \Abstract{\litM} $
which yields a flow-insensitive analysis. Section
\ref{a-compositional-monadic-framework} will generalize the definitions
from this section into compositional components (monad transformers) in
addition to introducing another definition for $ \Abstract{\litM} $
which yields a flow-sensitive analysis.

\par

Before going into the details of the flow-insensitive monad, we wish to
build intuition regarding what one would expect from such a development.

\par

Recall the path-sensitive monad $ \Abstract{\litM} $ and its state space
$ \Abstract{\Sigma} $ from section \ref{recovering-analyses}:

\small\begin{align*}
&\hspace{0em}  \Abstract{\litM}(  \itop{Exp})   \coloneqq   \Psi   \times   \Abstract{\litStore}   \rightarrow   \mathcal{P}  (\itop{Exp}   \times   \Psi   \times   \Abstract{\litStore})  \\
&\hspace{0em}  \Abstract{\Sigma}  (\itop{Exp})   \coloneqq   \mathcal{P}  (\itop{Exp}   \times   \Psi   \times   \Abstract{\litStore}) 
\end{align*}\normalsize

where
$ \Psi   \coloneqq   \Abstract{\litEnv}   \times   \Abstract{\litKAddr}   \times   \Abstract{\litKStore}   \times   \Abstract{\litTime} $.
This is path-sensitive because $ \Abstract{\Sigma}  (\itop{Exp}) $ can
represent arbitrary \emph{relations} between
$ (\itop{Exp}   \times   \Psi )$ and $ \Abstract{\litStore} $.

\par

As discussed in Section \ref{path-and-flow-sensitivity-in-analysis}, a
flow-sensitive analysis will give a single set of facts per program
point. This results in the following monad $ \Abstract{\litM} ^{  fs } $
and state space $ \Abstract{\Sigma}  ^{ fs } $ which encode \emph{finite
maps} to $ \Abstract{\litStore} $ rather than relations:

\small\begin{align*}
&\hspace{0em}  \Abstract{\litM} ^{  fs }  (\itop{Exp})   \coloneqq   \Psi   \times   \Abstract{\litStore}   \rightarrow   \lbrack  (\itop{Exp}   \times   \Psi )  \mapsto   \Abstract{\litStore} \rbrack   \\
&\hspace{0em}  \Abstract{\Sigma}  ^{ fs }  (\itop{Exp})   \coloneqq   \lbrack  (\itop{Exp}   \times   \Psi )  \mapsto   \Abstract{\litStore} \rbrack  
\end{align*}\normalsize

where $ \lbrack  \alpha   \mapsto  s \rbrack $ is notation for a finite
map from $ \alpha $ to $s$.

\par

Finally, a flow-insensitive analysis contains a single global set of
facts, which is represented by pulling $ \Abstract{\litStore} $ out of
the powerset:

\small\begin{align*}
&\hspace{0em}  \Abstract{\litM} ^{  fi }  (\itop{Exp})   \coloneqq   \Psi   \times   \Abstract{\litStore}   \rightarrow   \mathcal{P}  (\itop{Exp}   \times   \Psi )  \times   \Abstract{\litStore}  \\
&\hspace{0em}  \Abstract{\Sigma}  ^{ fi }  (\itop{Exp})   \coloneqq   \mathcal{P}  (\itop{Exp}   \times   \Psi )  \times   \Abstract{\litStore} 
\end{align*}\normalsize

\par

These three resulting structures, $ \Abstract{\Sigma} $,
$ \Abstract{\Sigma}  ^{ fs } $ and $ \Abstract{\Sigma}  ^{ fi } $,
capture the essence of path-sensitive, flow-sensitive and
flow-insensitive iteration, and arise naturally from
$ \Abstract{\litM} $, $ \Abstract{\litM} ^{  fs } $ and
$ \Abstract{\litM} ^{  fi } $, which each have monadic structure. We
only describe $ \Abstract{\litM} ^{  fi } $ directly in this section;
more compositional versions of each of these monads are described in
Section \ref{a-compositional-monadic-framework}, from which
$ \Abstract{\Sigma}  ^{ fs } $ and $ \Abstract{\Sigma}  ^{ fi } $ can be
recovered.

\par

\subsection{Flow Insensitive Monad}\label{flow-insensitive-monad}

\par

For $ \Abstract{\litM} ^{  fi } $ the monad operator $ \itop{bind} $
performs the store merging needed to capture a flow-insensitive
analysis.

\small\begin{align*}
&\hspace{0em}  \itop{bind}   \mathbin{\colon}   \forall   \alpha   \beta ,  \Abstract{\litM} ^{  fi } ( \alpha )  \rightarrow  ( \alpha   \rightarrow   \Abstract{\litM} ^{  fi } ( \beta ))  \rightarrow   \Abstract{\litM} ^{  fi } ( \beta ) \\
&\hspace{0em}  \itop{bind}( m)(f)( \psi , \sigma )  \coloneqq   \\
&\hspace{2em} ( \{ bs _{ 11 }  .. bs _{ 1m _1  }  .. bs _{ n1 }  .. bs _{ nm _n  }  \} , \sigma  _1   \sqcup  ..  \sqcup   \sigma  _n )  \;\;\text{where}\;\;  \\
&\hspace{4em} ( \{ (a _1 , \psi  _1 ) .. (a _n , \psi  _n ) \} , \sigma ')  \coloneqq  m( \psi , \sigma ) \\
&\hspace{4em} ( \{ b \psi  _{ i1 }  .. b \psi  _{ im _i  }  \} , \sigma  _i )  \coloneqq  f(a _i )( \psi  _i , \sigma ')
\end{align*}\normalsize

The unit for $ \itop{bind} $ returns one nondeterminism branch and a
single global store:

\small\begin{align*}
&\hspace{0em}  \itop{return}   \mathbin{\colon}   \forall   \alpha ,  \alpha   \rightarrow   \Abstract{\litM} ^{  fi } ( \alpha ) \\
&\hspace{0em}  \itop{return}( a)( \psi , \sigma )  \coloneqq  ( \{ a, \psi  \} , \sigma )
\end{align*}\normalsize

\par

State effects $ \itop{get-Env} $ and $ \itop{put-Env} $ are also
straightforward, returning one branch of nondeterminism:

\small\begin{align*}
&\hspace{0em}  \itop{get-Env}   \mathbin{\colon}   \Abstract{\litM} ^{  fi }  (\Abstract{\litEnv})  \\
&\hspace{0em}  \itop{get-Env}(  \langle  \rho , \kappa , \tau  \rangle , \sigma )  \coloneqq  ( \{ ( \rho , \langle  \rho , \kappa , \tau  \rangle ) \} , \sigma ) \\
&\hspace{0em}  \itop{put-Env}   \mathbin{\colon}   \Abstract{\litEnv}   \rightarrow   \Abstract{\litM} ^{  fi } (1) \\
&\hspace{0em}  \itop{put-Env}(  \rho ')( \langle  \rho , \kappa , \tau  \rangle , \sigma )  \coloneqq  ( \{ (1, \langle  \rho ', \kappa , \tau  \rangle ) \} , \sigma )
\end{align*}\normalsize

State effects $ \itop{get-Store} $ and $ \itop{put-Store} $ are
analogous to $ \itop{get-Env} $ and $ \itop{put-Env} $.

\par

Nondeterminism operations will union the powerset and join the store
pairwise:

\small\begin{align*}
&\hspace{0em}  \itop{mzero}   \mathbin{\colon}   \forall   \alpha ,  \itop{M}(  \alpha ) \\
&\hspace{0em}  \itop{mzero}(  \psi , \sigma )  \coloneqq  ( \{  \} ,  \bot ) \\
&\hspace{0em}  \underline{\hspace{0.5em}}{  \mathbin{\langle + \rangle}  }\underline{\hspace{0.5em}}   \mathbin{\colon}   \forall   \alpha ,  \itop{M}(  \alpha )  \times   \itop{M}(  \alpha )  \rightarrow   \itop{M}   \alpha   \\
&\hspace{0em} (m _1   \mathbin{\langle + \rangle}  m _2 )( \psi , \sigma )  \coloneqq  (a \psi  ^{ * }  _1   \cup  a \psi  ^{ * }  _2 , \sigma  _1   \sqcup   \sigma  _2 )  \;\;\text{where}\;\;   \\
&\hspace{2em} (a \psi  ^{ * }  _i , \sigma  _i )  \coloneqq  m _i ( \psi , \sigma )
\end{align*}\normalsize

\par

Finally, the Galois connection relating $ \Abstract{\litM} ^{  fi } $ to
a state space transition over $ \Abstract{\Sigma}  ^{ fi } $ must also
compute set unions and store joins pairwise:

\small\begin{align*}
&\hspace{0em}  \Abstract{\Sigma}  ^{ fi }   \coloneqq   \mathcal{P}  (\itop{Exp}   \times   \Psi )  \times   \Abstract{\litStore}  \\
&\hspace{0em}  \gamma   \mathbin{\colon}   (\itop{Exp}   \rightarrow   \Abstract{\litM} ^{  fi }  (\itop{Exp}) )  \rightarrow  ( \Abstract{\Sigma}  ^{ fi }   \rightarrow   \Abstract{\Sigma}  ^{ fi } ) \\
&\hspace{0em}  \gamma (f)(e \psi  ^{ * } , \sigma )  \coloneqq   \\
&\hspace{2em} ( \{ e \psi  _{ 11 }  .. e \psi  _{ n1 }  .. e \psi  _{ nm _n  }  \} ,  \sigma  _1   \sqcup  ..  \sqcup   \sigma  _n )  \;\;\text{where}\;\;   \\
&\hspace{4em}  \{ (e _1 , \psi  _1 ) .. (e _n , \psi  _n ) \}   \coloneqq  e \psi  ^{ * }  \\
&\hspace{4em} ( \{ e \psi  _{ i1 }  .. e \psi  _{ im _i  }  \} , \sigma  _i )  \coloneqq  f(e _i )( \psi  _i , \sigma ) \\
&\hspace{0em}  \alpha    \mathbin{\colon}  ( \Abstract{\Sigma}  ^{ fi }   \rightarrow   \Abstract{\Sigma}  ^{ fi } )  \rightarrow   (\itop{Exp}   \rightarrow   \Abstract{\litM} ^{  fi }  (\itop{Exp}) ) \\
&\hspace{0em}  \alpha (f)(e)( \psi , \sigma )  \coloneqq  f( \{ (e, \psi ) \} , \sigma )
\end{align*}\normalsize

\begin{proposition} $ \gamma $ and $ \alpha $ form a Galois connection
and there exists Galois connections:

\small\begin{alignat*}{1}
 \Concrete{\litM}   \galois{\alpha_1}{\gamma_1}   \Abstract{\litM}   \galois{\alpha_2}{\gamma_2}   \Abstract{\litM} ^{  fi } 
\end{alignat*}\normalsize

\end{proposition} The first Galois connection
$ \Concrete{\litM}   \galois{\alpha_1}{\gamma_1}   \Abstract{\litM} $ is
justified piecewise by the Galois connections between
$ \Concrete{\litVal}   \galois{\alpha}{\gamma}   \Abstract{\litVal} $
and
$ \Concrete{\litTime}   \galois{\alpha}{\gamma}   \Abstract{\litTime} $.
The second Galois connection
$ \Abstract{\litM}   \galois{\alpha_2}{\gamma_2}   \Abstract{\litM} ^{  fi } $
is justified by calculation over their definitions. We aim to recover
this proof more easily through compositional components in Section
\ref{a-compositional-monadic-framework}. \begin{corollary}

\small\begin{alignat*}{1}
 \Concrete{\Sigma}   \galois{\alpha_1}{\gamma_1}   \Abstract{\Sigma}   \galois{\alpha_2}{\gamma_2}   \Abstract{\Sigma}  ^{ fi } 
\end{alignat*}\normalsize

\end{corollary} This property is derived by transporting each Galois
connection between monads through their respective Galois connections to
$ \Sigma $. This is also a direct corollary of Theorem
\ref{galois-theorem} in Section \ref{galois-transformers-1} when cast in
the compositional framework of Galois transformers. \begin{proposition}
The following orderings hold between the three induced transition
relations:

\small\begin{alignat*}{1}
 \alpha  _1   \circ   \Concrete{\gamma}  (\itop{step})   \circ   \gamma  _1   \sqsubseteq   \Abstract{\gamma}  (\itop{step})   \sqsubseteq   \gamma  _2   \circ   \Abstract{\gamma}  ^{ fi }  (\itop{step})   \circ   \alpha  _2 
\end{alignat*}\normalsize

\end{proposition} This is an application of the monotonicity of
$ \itop{step} $ and the Galois connections between monads, each
transported through the Galois connection to their corresponding
transition systems.

\par

We note that the implementation for our interpreter and abstract garbage
collector remain the same for each instantiation; they scale seamlessly
to variations in path and flow sensitivity when instantiated with the
appropriate monad.

\par

\section{A Compositional Monadic
Framework}\label{a-compositional-monadic-framework}

\par

In our development thus far, any modification to the interpreter
requires redesigning the monad $ \Abstract{\litM} $ and constructing new
proofs relating $ \Abstract{\litM} $ to $ \Concrete{\litM} $. We want to
avoid reconstructing complicated monads for our interpreters, especially
as languages and analyses grow and change. Even more, we want to avoid
reconstructing complicated \emph{proofs} that such changes require.
Toward this goal we introduce a compositional framework for constructing
monads which are correct-by-construction by extending the well-known
structure of monad transformer to that of \emph{Galois transformer}.

\par

There are two monadic effects used in our monadic interpreter: state and
nondeterminism. For state we will review the state monad transformer
$S _t  \lbrack s \rbrack $, which is standard (See
\citet{dvanhorn:Liang1995Monad} for more details). For nondeterminism we
develop two monad transformers, $ \mathcal{P}  _t $ and
$FS _t  \lbrack s \rbrack $. These transformers are fully general
purpose, even outside the context of program analysis, and are novel in
this work.

\par

To create a monad with various state and nondeterminism effects, one
must merely summon some composition of these three monads.
\emph{Implementations and proofs for monadic sequencing, state effects,
nondeterminism effects, and mappings to an executable transition system
will come entirely for free.} This means that for a language which has a
different state space than the example in this paper, no added effort is
required to construct a monad stack for that language.

\par

Path and flow sensitivity properties arise from the \emph{order of
composition} of monad transformers. Placing state after nondeterminism
($S _t  \lbrack s \rbrack   \circ   \mathcal{P}  _t $ or
$S _t  \lbrack s \rbrack   \circ  FS _t  \lbrack s' \rbrack $) will
result in $s$ being path-sensitive. Placing state before nondeterminism
($ \mathcal{P}  _t   \circ  S _t  \lbrack s \rbrack $ or
$FS _t  \lbrack s' \rbrack   \circ  S _t  \lbrack s \rbrack $) will
result in $s$ being flow-insensitive. Finally, when
$FS _t  \lbrack s \rbrack $ is used in place of
$S _t  \lbrack s \rbrack   \circ   \mathcal{P}  _t $ or
$ \mathcal{P}  _t   \circ  S _t  \lbrack s \rbrack $, $s$ will be
flow-sensitive. The combination of all three sensitivities looks like
($ \itop{M}   \coloneqq  S _t  \lbrack s _1  \rbrack   \circ  FS _t  \lbrack s _2  \rbrack   \circ  S _t  \lbrack s _3  \rbrack $),
which will induce state space transition system
$ \Sigma  (\itop{Exp})   \coloneqq   \lbrack  (\itop{Exp}   \times  s _1 )  \mapsto  s _2  \rbrack   \times  s _3 $,
which is path-sensitive in $s _1 $, flow-sensitive in $s _2 $ and
flow-insensitive in $s _3 $. Using $S _t  \lbrack s \rbrack $,
$ \mathcal{P}  _t $ and $FS _t  \lbrack s \rbrack $, one can easily
choose which components of the analysis are path-sensitive,
flow-sensitive or flow-insensitive.

\par

In the following definitions we must necessarily use $ \itop{bind} $,
$ \itop{return} $ and other operations from the underlying monad, and we
notate these $ \itop{bind} _m  $, $ \itop{return} _m  $,
$ \ttbfop{do} _m  $, $ \leftarrow  _m $, etc.

\par

\subsection{State Monad Transformer}\label{state-monad-transformer}

\par

Briefly we review the state monad transformer,
$S _t  \lbrack s \rbrack $:

\small\begin{align*}
&\hspace{0em} S _t  \lbrack  \underline{\hspace{0.5em}}  \rbrack   \mathbin{\colon}   (\itop{Type}   \rightarrow   \itop{Type})   \rightarrow   (\itop{Type}   \rightarrow   \itop{Type})  \\
&\hspace{0em} S _t  \lbrack s \rbrack (m)( \alpha )  \coloneqq  s  \rightarrow  m( \alpha   \times  s)
\end{align*}\normalsize

$S _t  \lbrack s \rbrack $ transports monad operations from $m$ to
$S _t  \lbrack s \rbrack (m)$:

\small\begin{align*}
&\hspace{0em}  \itop{bind}   \mathbin{\colon}   \forall   \alpha   \beta ,  \\
&\hspace{2em} S _t  \lbrack s \rbrack (m)( \alpha )  \rightarrow  ( \alpha   \rightarrow  S _t  \lbrack s \rbrack (m)( \beta ))  \rightarrow  S _t  \lbrack s \rbrack (m)( \beta ) \\
&\hspace{0em}  \itop{bind}( m)(f)(s)  \coloneqq  (x,s')  \leftarrow  _m  m(s)  \;;\;  f(x)(s') \\
&\hspace{0em}  \itop{return}   \mathbin{\colon}   \forall   \alpha ,  \alpha   \rightarrow  S _t  \lbrack s \rbrack (m)( \alpha ) \\
&\hspace{0em}  \itop{return}( x)(s)  \coloneqq   \itop{return} _m  (x,s)
\end{align*}\normalsize

$S _t  \lbrack s \rbrack $ supports state effects:
\small\begin{alignat*}{4}

 \itop{get}  &  \mathbin{\colon}  S _t  \lbrack s \rbrack (m)(s)     &  \hspace{1em}   \hspace{1em} \itop{get}( s)     &  \coloneqq   \itop{return} _m  (s,s)  \\
 \itop{put}  &  \mathbin{\colon}  s  \rightarrow  S _t  \lbrack s \rbrack (m)(1) &  \hspace{1em}   \hspace{1em} \itop{put}( s')(s) &  \coloneqq   \itop{return} _m  (1,s')

\end{alignat*}\normalsize Finally, $S _t  \lbrack s \rbrack $ transports
nondeterminism effects from $m$ to $S _t  \lbrack s \rbrack (m)$:

\small\begin{align*}
&\hspace{0em}  \itop{mzero}   \mathbin{\colon}   \forall   \alpha , S _t  \lbrack s \rbrack (m)( \alpha ) \\
&\hspace{0em}  \itop{mzero}( s)  \coloneqq   \itop{mzero} _m    \\
&\hspace{0em}  \underline{\hspace{0.5em}}{  \mathbin{\langle + \rangle}  }\underline{\hspace{0.5em}}   \mathbin{\colon}   \forall   \alpha , S _t  \lbrack s \rbrack (m)( \alpha )  \times  S _t  \lbrack s \rbrack (m)( \alpha )  \rightarrow  S _t  \lbrack s \rbrack (m)( \alpha ) \\
&\hspace{0em} (m _1   \mathbin{\langle + \rangle}  m _2 )(s)  \coloneqq  m _1 (s)  \mathbin{\langle + \rangle}  _m  m _2 (s) 
\end{align*}\normalsize

\par

\subsection{Nondeterminism Monad
Transformer}\label{nondeterminism-monad-transformer}

\par

We have developed a new monad transformer for nondeterminism which
composes with state in both directions. Previous attempts to define a
monad transformer for nondeterminism resulted in operations which do not
respect monad laws or nondeterminism effect laws--ours respects both.

\par

The nondeterminism monad transformer is defined with the expected type,
embedding $ \mathcal{P} $ inside $m$:

\small\begin{align*}
&\hspace{0em}  \mathcal{P}  _t   \mathbin{\colon}   (\itop{Type}   \rightarrow   \itop{Type})   \rightarrow   (\itop{Type}   \rightarrow   \itop{Type})  \\
&\hspace{0em}  \mathcal{P}  _t (m)( \alpha )  \coloneqq  m( \mathcal{P} ( \alpha ))
\end{align*}\normalsize

$ \mathcal{P}  _t $ transports monad operations from $m$ to
$ \mathcal{P}  _t (m)$ \emph{provided that $m$ is a join-semilattice
functor}. The join-lattice functorality of $m$ will be instantiated with
$ \mathcal{P} ( \beta )$.

\small\begin{align*}
&\hspace{0em}  \itop{bind}   \mathbin{\colon}   \forall   \alpha   \beta ,  \mathcal{P}  _t (m)( \alpha )  \rightarrow  ( \alpha   \rightarrow   \mathcal{P}  _t (m)( \beta ))  \rightarrow   \mathcal{P}  _t (m)( \beta ) \\
&\hspace{0em}  \itop{bind}( m)(f)  \coloneqq   \\
&\hspace{2em}  \{ x _1  .. x _n  \}   \leftarrow  _m  m  \;;\;  f(x _1 )  \sqcup  _m  ..  \sqcup  _m  f(x _n )
\end{align*}\normalsize

\small\begin{align*}
&\hspace{0em}  \itop{return}   \mathbin{\colon}   \forall   \alpha ,  \alpha   \rightarrow   \mathcal{P}  _t (m)( \alpha ) \\
&\hspace{0em}  \itop{return}( x)  \coloneqq   \itop{return} _m  ( \{ x \} )
\end{align*}\normalsize

$ \mathcal{P}  _t $ transports state effects from $m$ to
$ \mathcal{P}  _t (m)$:

\small\begin{align*}
&\hspace{0em}  \itop{get}   \mathbin{\colon}   \mathcal{P}  _t (m)(s) \\
&\hspace{0em}  \itop{get}  = map _m ( \lambda (s). \{ s \} ) (\itop{get} _m  ) \\
&\hspace{0em}  \itop{put}   \mathbin{\colon}  s  \rightarrow   \mathcal{P}  _t (m)(1) \\
&\hspace{0em}  \itop{put}( s) = map _m ( \lambda (1). \{ 1 \} ) (\itop{put} _m  (s))
\end{align*}\normalsize

Finally, $ \mathcal{P}  _t $ supports nondeterminism effects through a
straightforward application of the underlying monad's join-semilattice
functorality:

\small\begin{align*}
&\hspace{0em}  \itop{mzero}   \mathbin{\colon}   \forall   \alpha ,  \mathcal{P}  _t (m)( \alpha ) \\
&\hspace{0em}  \itop{mzero}   \coloneqq   \bot  _m  \\
&\hspace{0em}  \underline{\hspace{0.5em}}{  \mathbin{\langle + \rangle}  }\underline{\hspace{0.5em}}   \mathbin{\colon}   \forall   \alpha ,  \mathcal{P}  _t (m)( \alpha ) x  \mathcal{P}  _t (m)( \alpha )  \rightarrow   \mathcal{P}  _t (m)( \alpha ) \\
&\hspace{0em} m _1   \mathbin{\langle + \rangle}  m _2   \coloneqq  m _1   \sqcup  _m  m _2 
\end{align*}\normalsize

\par

\begin{proposition} (1) $ \itop{bind} $ and $ \itop{return} $ satisfy
monad laws, (2) $ \itop{get} $ and $ \itop{put} $ satisfy the state
monad laws, and (3) $ \itop{mzero} $ and $ \mathbin{\langle + \rangle} $
satisfy the nondeterminism monad laws. \end{proposition} The key lemma
in (1) is the functorality of $m$, namely that:

\small\begin{alignat*}{1}
 \itop{return} _m  (x  \sqcup  y) =  \itop{return} _m  (x)  \sqcup  _m   \itop{return} _m  (y)
\end{alignat*}\normalsize

\begin{enumerate}
\def\labelenumi{(\arabic{enumi})}
\setcounter{enumi}{1}
\itemsep1pt\parskip0pt\parsep0pt
\item
  is by simple calculation and (3) is trivial as a consequence of the
  underlying monad being a join-semilattice functor.
\end{enumerate}

\par

\subsection{Flow Sensitivity Monad
Transformer}\label{flow-sensitivity-monad-transformer}

\par

The flow sensitivity monad transformer is a unique monad transformer
that combines state and nondeterminism effects, and does not arise
naturally from composing vanilla nondeterminism and state transformers.
The flow sensitivity monad transformer is defined:

\small\begin{align*}
&\hspace{0em} FS _t  \lbrack  \underline{\hspace{0.5em}}  \rbrack   \mathbin{\colon}   (\itop{Type}   \rightarrow   \itop{Type})   \rightarrow   (\itop{Type}   \rightarrow   \itop{Type})  \\
&\hspace{0em} FS _t  \lbrack s \rbrack (m)( \alpha )  \coloneqq  s  \rightarrow  m( \lbrack  \alpha   \mapsto  s \rbrack )
\end{align*}\normalsize

where $ \lbrack  \alpha   \mapsto  s \rbrack $ is notation for a finite
map from $ \alpha $ to $s$.

\par

$FS _t  \lbrack s \rbrack $ transports monad operations provided that
$s$ is a join-semilattice and $m$ is a join-semilattice functor. The
functorality of $m$ will be instantiated with
$ \lbrack  \beta   \rightarrow  s \rbrack $, which forms a lattice when
$s$ does also.

\small\begin{align*}
&\hspace{0em}  \itop{bind}   \mathbin{\colon}   \forall   \alpha   \beta ,  \\
&\hspace{1em} FS _t  \lbrack s \rbrack (m)( \alpha )  \rightarrow  ( \alpha   \rightarrow  FS _t  \lbrack s \rbrack (m)( \beta ))  \rightarrow  FS _t  \lbrack s \rbrack (m)( \beta ) \\
&\hspace{0em}  \itop{bind}( m)(f)(s)  \coloneqq   \ttbfop{do} _m   \\
&\hspace{2em}  \{ x _1   \mapsto  s _1 ,..,x _n   \mapsto  s _n  \}   \leftarrow  _m  m(s) \\
&\hspace{2em} f(x _1 )(s _1 )  \sqcup  _m  ..  \sqcup  _m  f(x _n )(s _n ) \\
&\hspace{0em}  \itop{return}   \mathbin{\colon}   \forall   \alpha ,  \alpha   \rightarrow  FS _t  \lbrack s \rbrack (m)( \alpha ) \\
&\hspace{0em}  \itop{return}( x)(s)  \coloneqq   \itop{return} _m    \{ x  \mapsto  s \} 
\end{align*}\normalsize

$FS _t  \lbrack s \rbrack $ transports state effects from $m$ to
$FS _t  \lbrack s \rbrack (m)$:

\small\begin{align*}
&\hspace{0em}  \itop{get}   \mathbin{\colon}  FS _t  \lbrack s \rbrack (m)(s) \\
&\hspace{0em}  \itop{get}( s)  \coloneqq   \itop{return} _m    \{ s  \mapsto  s \} 
\end{align*}\normalsize

\small\begin{align*}
&\hspace{0em}  \itop{put}   \mathbin{\colon}  s  \rightarrow  FS _t  \lbrack s \rbrack (m)(1) \\
&\hspace{0em}  \itop{put}( s')(s)  \coloneqq   \itop{return} _m    \{ 1  \mapsto  s' \} 
\end{align*}\normalsize

Finally, $FS _t  \lbrack s \rbrack $ supports nondeterminism effects
provided $s$ is a join-semilattice and $m$ is a join-semilattice
functor:

\small\begin{align*}
&\hspace{0em}  \itop{mzero}   \mathbin{\colon}   \forall   \alpha , FS _t  \lbrack s \rbrack (m)( \alpha ) \\
&\hspace{0em}  \itop{mzero}( s)  \coloneqq   \bot  _m  \\
&\hspace{0em}  \underline{\hspace{0.5em}}{  \mathbin{\langle + \rangle}  }\underline{\hspace{0.5em}}   \mathbin{\colon}   \forall   \alpha , FS _t  \lbrack s \rbrack (m)( \alpha ) x FS _t  \lbrack s \rbrack (m)( \alpha )  \rightarrow  FS _t  \lbrack s \rbrack (m)( \alpha ) \\
&\hspace{0em} (m _1   \mathbin{\langle + \rangle}  m _2 )(s)  \coloneqq  m _1 (s)  \sqcup  _m  m _2 (s)
\end{align*}\normalsize

\begin{proposition} $ \itop{get} $ and $ \itop{put} $ satisfy the state
monad laws, $ \itop{mzero} $ and $ \mathbin{\langle + \rangle} $ satisfy
the nondeterminism monad laws, and
$S _t  \lbrack s \rbrack   \circ   \mathcal{P}  _t   \galois{\alpha_1}{\gamma_1}  FS _t  \lbrack s \rbrack   \galois{\alpha_2}{\gamma_2}   \mathcal{P}  _t   \circ  S _t  \lbrack s \rbrack $.
\end{proposition} These proofs are analogous to those for state and
nondeterminism monad transformers.

\par

\subsection{Mapping to State Spaces}\label{mapping-to-state-spaces}

\par

Both our execution and correctness frameworks require that monadic
actions in $m$ map to state space transitions in $ \Sigma $. We extend
the earlier statement of Galois connection to the transformer setting,
mapping monad \emph{transformer} actions in $T$ to state space
\emph{functor} transitions in $ \Pi $.

\small\begin{align*}
&\hspace{0em} T  \mathbin{\colon}   (\itop{Type}   \rightarrow   \itop{Type})   \rightarrow   (\itop{Type}   \rightarrow   \itop{Type})  \\
&\hspace{0em}  \Pi   \mathbin{\colon}   (\itop{Type}   \rightarrow   \itop{Type})   \rightarrow   (\itop{Type}   \rightarrow   \itop{Type})  \\
&\hspace{0em}  \itop{mstep}   \mathbin{\colon}   \forall   \alpha   \beta ,  \\
&\hspace{2em} ( \alpha   \rightarrow  T(m)( \beta ))  \galois{\alpha}{\gamma}  ( \Pi ( \Sigma  _m )( \alpha )  \rightarrow   \Pi ( \Sigma  _m )( \beta ))
\end{align*}\normalsize

In the type of $ \itop{mstep} $, $m$ is an arbitrary monad whose monadic
actions map to state space $ \Sigma  _m $. The monad transformer $T$
must induce a state space transformer $ \Pi $ for which $ \itop{mstep} $
can be defined. We only show the $ \gamma $ sides of the mappings in
this section, which allow one to execute the analyses.

\par

For the state monad transformer $S _t  \lbrack s \rbrack $ mstep is
defined:

\small\begin{align*}
&\hspace{0em}  \itop{mstep-\gamma}   \mathbin{\colon}   \forall   \alpha   \beta ,  \\
&\hspace{2em} ( \alpha   \rightarrow  S _t  \lbrack s \rbrack (m)( \beta ))  \rightarrow  ( \Sigma  _m ( \alpha   \times  s)  \rightarrow   \Sigma  _m ( \beta   \times  s)) \\
&\hspace{0em}  \itop{mstep-\gamma} (f)  \coloneqq   \itop{mstep_m-\gamma} ( \lambda (a,s). f(a)(s))
\end{align*}\normalsize

For the nondeterminism transformer $ \mathcal{P}  _t $ mstep is defined:

\small\begin{align*}
&\hspace{0em}  \itop{mstep-\gamma}   \mathbin{\colon}   \forall   \alpha   \beta ,  \\
&\hspace{2em} ( \alpha   \rightarrow   \mathcal{P}  _t (m)( \beta ))  \rightarrow  ( \Sigma  _m ( \mathcal{P} ( \alpha ))  \rightarrow   \Sigma  _m ( \mathcal{P} ( \beta ))) \\
&\hspace{0em}  \itop{mstep-\gamma} (f)  \coloneqq   \itop{mstep_m-\gamma} (F)  \;\;\text{where}\;\;   \\
&\hspace{2em} F( \{ x _1  .. x _n  \} ) = f(x _1 )  \sqcup  _m  ..  \sqcup  _m  f(x _n ))
\end{align*}\normalsize

For the flow sensitivity monad transformer $FS _t  \lbrack s \rbrack $
mstep is defined:

\small\begin{align*}
&\hspace{0em}  \itop{mstep-\gamma}   \mathbin{\colon}   \forall   \alpha   \beta ,  \\
&\hspace{2em} ( \alpha   \rightarrow  FS _t  \lbrack s \rbrack (m)( \beta ))  \rightarrow  ( \Sigma  _m ( \lbrack  \alpha   \mapsto  s \rbrack )  \rightarrow   \Sigma  _m ( \lbrack  \beta   \times  s \rbrack )) \\
&\hspace{0em}  \itop{mstep-\gamma} (f)  \coloneqq   \itop{mstep_m-\gamma} (F)  \;\;\text{where}\;\;   \\
&\hspace{2em} F( \{ x _1   \mapsto  s _1  \} ,.., \{ x _n   \mapsto  s _n  \} )  \coloneqq   \\
&\hspace{4em} f(x _1 )(s _1 )  \sqcup  _m  ..  \sqcup  _m  f(x _n )(s _n )
\end{align*}\normalsize

The Galois connections for $ \itop{mstep} $ for
$S _t  \lbrack s \rbrack $, $P _t $ and $FS _t  \lbrack s \rbrack $
crucially require that $ \itop{mstep_m-\gamma} $ and
$ \itop{mstep_m-\alpha} $ be homomorphic, i.e.~that:

\small\begin{align*}
&\hspace{0em}  \alpha (id)  \sqsubseteq   \itop{return}   \;\;\text{and}\;\;   \alpha (f  \circ  g)  \sqsubseteq   \alpha (f)  \mathbin{\langle \circ \rangle}   \alpha (g)
\end{align*}\normalsize

and likewise for $ \gamma $, where $ \mathbin{\langle \circ \rangle} $
is composition in the Kleisli category for the monad $ \itop{M} $.

\par

\subsection{Galois Transformers}\label{galois-transformers-1}

\par

The capstone of our framework is the fact that monad transformers
$S _t  \lbrack s \rbrack $, $FS _t  \lbrack s \rbrack $ and
$ \mathcal{P}  _t $ are also \emph{Galois transformers}.

\par

\begin{definition} A monad transformer $T$ is a Galois transformer if:
\begin{enumerate} \item It transport Galois connections, that is for all
monads $m _1 $ and $m _2 $, $m _1   \galois{\alpha}{\gamma}  m _2 $
implies $T(m _1 )  \galois{\alpha}{\gamma}  T(m _2 )$.

\begin{center}
\begin{tikzpicture}
  \matrix (m) [matrix of math nodes,row sep=3em,column sep=4em,minimum width=2em]
  {
     m_1 & T(m_1) \\
     m_2 & T(m_2) \\
  };
  \path[-stealth]
    (m-1-1) edge [bend left=40] node [right] {$\alpha$}   (m-2-1)
            edge                node [below] {$T$}        (m-1-2)
    (m-2-1) edge [bend left=40] node [left]  {$\gamma$}   (m-1-1)
            edge                node [below] {$T$}        (m-2-2)
    (m-1-2) edge [bend left=40] node [right] {$\alpha_T$} (m-2-2)
    (m-2-2) edge [bend left=40] node [left]  {$\gamma_T$} (m-1-2)
  ;
\end{tikzpicture}
\end{center}

\item

It transports mappings to an executable transition system, that is there
exists $ \Pi $ s.t. for all monads $m$ and functors $ \Sigma $,
$( \alpha   \rightarrow  m( \beta ))  \galois{\alpha}{\gamma}  ( \Sigma ( \alpha )  \rightarrow         \Sigma ( \beta ))$
implies
$( \alpha   \rightarrow  T(m)( \beta ))  \galois{\alpha}{\gamma}  ( \Pi ( \Sigma )( \alpha )  \rightarrow   \Pi ( \Sigma )( \beta ))$.

\begin{center}
\begin{tikzpicture}
  \matrix (m) [matrix of math nodes,row sep=3em,column sep=4em,minimum width=2em]
  {
             \alpha \rightarrow m(\beta)      & \alpha              \rightarrow T(m)(\beta)        \\
     \Sigma(\alpha) \rightarrow \Sigma(\beta) & \Pi(\Sigma)(\alpha) \rightarrow \Pi(\Sigma)(\beta) \\
  };
  \path[-stealth]
    (m-1-1) edge [bend left=40] node [right] {$\alpha$}   (m-2-1)
            edge                node [below] {$T$}        (m-1-2)
    (m-2-1) edge [bend left=40] node [left]  {$\gamma$}   (m-1-1)
            edge                node [below] {$\Pi$}      (m-2-2)
    (m-1-2) edge [bend left=40] node [right] {$\alpha_T$} (m-2-2)
    (m-2-2) edge [bend left=40] node [left]  {$\gamma_T$} (m-1-2)
  ;
\end{tikzpicture}
\end{center}

\end{enumerate} \end{definition} Property (1) transports Galois
connections between monads, and property (2) transports Galois
connections between transition systems. By composing the 2-dimensional
diagrams (1) and (2) into a 3-dimensional diagram (which we do not show)
we establish the following theorem: \begin{theorem} If $T$ is a Galois
transformer, then it is sufficient to prove that underlying monads
$m _1 $ and $m _2 $ form a Galois connection
$m _1   \galois{\alpha}{\gamma}  m _2 $ in order to establish
$ \Pi ( \Sigma  _1 )  \galois{\alpha}{\gamma}   \Pi ( \Sigma  _2 )$.
\label{galois-theorem} \end{theorem} This is the workhorse of our entire
proof framework, allowing us to reason about monadic actions, like the
monadic interpreter $ \itop{step} $ from section \ref{the-interpreter},
and derive properties about the induced transition system, which is how
the analysis is executed, for free. \begin{proposition}
$S _t  \lbrack s \rbrack $, $FS _t  \lbrack s \rbrack $ and
$ \mathcal{P}  _t $ are Galois transformers. \end{proposition} The
proofs are sketched earlier in Section
\ref{a-compositional-monadic-framework}.

\par

\subsection{Building Transformer
Stacks}\label{building-transformer-stacks}

\par

We can now build monad transformer stacks from combinations of
$S _t  \lbrack s \rbrack $, $FS \lbrack s \rbrack  _t $ and
$ \mathcal{P}  _t $ which automatically construct the following
properties:

\par

\begin{itemize}
\itemsep1pt\parskip0pt\parsep0pt
\item
  The resulting monad has the combined effects of all pieces of the
  transformer stack.
\item
  Actions in the resulting monad map to a state space transition system
  $ \Sigma   \rightarrow   \Sigma $ for some $ \Sigma $, allowing one to
  execute the analysis.
\item
  Galois connections between $ \Concrete{\Sigma} $ and
  $ \Abstract{\Sigma} $ are established piecewise from monad transformer
  components.
\end{itemize}

\par

We instantiate our interpreter to the following monad stacks in
decreasing order of precision:

\par

\vspace{4pt}

\small\begin{tabular}{ >{$}l<{$} | >{$}l<{$} | >{$}l<{$} }

S _t   \lbrack \Abstract{\litEnv} \rbrack       & S _t   \lbrack \Abstract{\litEnv} \rbrack        & S _t   \lbrack \Abstract{\litEnv} \rbrack      \\
S _t   \lbrack \Abstract{\litKAddr} \rbrack     & S _t   \lbrack \Abstract{\litKAddr} \rbrack      & S _t   \lbrack \Abstract{\litKAddr} \rbrack    \\
S _t   \lbrack \Abstract{\litKStore} \rbrack    & S _t   \lbrack \Abstract{\litKStore} \rbrack     & S _t   \lbrack \Abstract{\litKStore} \rbrack   \\
S _t   \lbrack \Abstract{\litTime} \rbrack      & S _t   \lbrack \Abstract{\litTime} \rbrack       & S _t   \lbrack \Abstract{\litTime} \rbrack     \\
S _t   \lbrack \Abstract{\litStore} \rbrack     &               &  \mathcal{P}  _t           \\
 \mathcal{P}  _t            & FS _t   \lbrack \Abstract{\litStore} \rbrack     & S _t   \lbrack \Abstract{\litStore} \rbrack    \\

\end{tabular}\normalsize \vspace{4pt}

\par

\noindent
From left to right these give path-sensitive, flow-sensitive and
flow-insensitive analyses for the data-store.

\par

Another benefit of our approach is that we can easily select different
path and flow sensitivity properties for the data-store and stack-store
independent of each other, merely by rearranging the order of
composition.

\par

\vspace{8pt}

\small\begin{tabular}{ >{$}l<{$} | >{$}l<{$} | >{$}l<{$} }

S _t   \lbrack \Abstract{\litEnv} \rbrack       & S _t   \lbrack \Abstract{\litEnv} \rbrack        & S _t   \lbrack \Abstract{\litEnv} \rbrack      \\
S _t   \lbrack \Abstract{\litKAddr} \rbrack     & S _t   \lbrack \Abstract{\litKAddr} \rbrack      & S _t   \lbrack \Abstract{\litKAddr} \rbrack    \\
S _t   \lbrack \Abstract{\litTime} \rbrack      & S _t   \lbrack \Abstract{\litTime} \rbrack       & S _t   \lbrack \Abstract{\litTime} \rbrack     \\
S _t   \lbrack \Abstract{\litKStore} \rbrack    &               &  \mathcal{P}  _t           \\
 \mathcal{P}  _t            & FS _t   \lbrack \Abstract{\litKStore} \rbrack    & S _t   \lbrack \Abstract{\litKStore} \rbrack   \\
S _t   \lbrack \Abstract{\litStore} \rbrack     & S _t   \lbrack \Abstract{\litStore} \rbrack      & S _t   \lbrack \Abstract{\litStore} \rbrack    \\

\end{tabular}\normalsize \vspace{8pt}

\par

\noindent
From left to right these give analysis which are all flow-insensitive in
the data-store, but path-sensitive, flow-sensitive and flow-insensitive
in the stack-store.

\par

\section{Implementation}\label{implementation-1}

\par

We have implemented our framework in Haskell and applied it to compute
analyses for $ \ttbfop{\lambda{}IF} $. Our implementation provides path
sensitivity, flow sensitivity, and flow insensitivity as a
semantics-independent monad library. The code shares a striking
resemblance with the math.

\par

Our implementation is suitable for prototyping and exploring the design
space of static analyzers. Our analyzer supports exponentially more
compositions of analysis features than any current analyzer. For
example, our implementation is the first which can combine arbitrary
choices in call-site, object, path and flow sensitivities. Furthermore,
the user can choose different path and flow sensitivities for each
component of the state space.

\par

Our implementation {\tt maam} supports command-line flags for garbage
collection, mCFA, call-site sensitivity, object sensitivity, and path
and flow sensitivity.

{\small\tt
\par \noindent ./maam prog.lam --gc --mcfa --kcfa=1 --ocfa=2
\par \noindent \hspace{2em} --data-store=flow-sen --stack-store=path-sen
}
\par \noindent

These flags are implemented completely independently of one another and
their combination is applied to a single parameterized monadic
interpreter. Furthermore, using Galois transformers allows us to prove
each combination correct in one fell swoop.

\par

A developer wishing to use our library to develop analyzers for their
language of choice inherits as much of the analysis infrastructure as
possible. We provide call-site, object, path and flow sensitivities as
language-independent libraries. To support analysis for a new language a
developer need only implement:

\par

\begin{itemize}
\itemsep1pt\parskip0pt\parsep0pt
\item
  A monadic semantics for their language, using state and nondeterminism
  effects.
\item
  The abstract value domain, and optionally the concrete value domain if
  they wish to recover concrete execution.
\item
  Intentional optimizations for their semantics like garbage collection
  and mcfa.
\end{itemize}

\par

The developer then receives the following for free through our analysis
library:

\par

\begin{itemize}
\itemsep1pt\parskip0pt\parsep0pt
\item
  A family of monads which implement their effect interface and give
  different path and flow sensitivities.
\item
  An execution engine for each monad to drive the analysis.
\item
  Mechanisms for call-site and object sensitivities.
\end{itemize}

\par

Not only is a developer able to reuse our implementation of call-site,
object, path and flow sensitivities, they need not understand the
execution machinery or soundness proofs for them either. They need only
verify that their monadic semantics is monotonic w.r.t. the analysis
parameters, and that their abstract value domain forms a Galois
connection. The execution and correctness of the final analyzer is
constructed automatically given these two properties.

\par

Our implementation is publicly available and can be installed as a cabal
package: {\small\tt cabal install maam}.

\par

\section{Related Work}\label{related-work}

\par

\paragraph{Overview}

\par

Program analysis comes in many forms such as points-to
\cite{dvanhorn:Andersen1994Program}, flow
\cite{dvanhorn:Jones:1981:LambdaFlow}, or shape analysis
\cite{dvanhorn:Chase1990Analysis}, and the literature is vast. (See
\citet{dvanhorn:hind-paste01,dvanhorn:Midtgaard2012Controlflow} for
surveys.) Much of the research has focused on developing families or
frameworks of analyses that endow the abstraction with a number of
knobs, levers, and dials to tune precision and compute efficiently (some
examples include
\citet{dvanhorn:Shivers:1991:CFA, dvanhorn:nielson-nielson-popl97,
dvanhorn:Milanova2005Parameterized, davdar:van-horn:2010:aam}; there are
many more). These parameters come in various forms with overloaded
meanings such as object \cite{dvanhorn:Milanova2005Parameterized,
dvanhorn:Smaragdakis2011Pick}, context
\cite{dvanhorn:Sharir:Interprocedural,
dvanhorn:Shivers:1991:CFA}, path \cite{davdar:das:2002:esp}, and heap
\cite{davdar:van-horn:2010:aam} sensitivities, or some combination
thereof \cite{dvanhorn:Kastrinis2013Hybrid}.

\par

These various forms can all be cast in the theory of abstraction
interpretation of \citet{dvanhorn:Cousot:1977:AI,
dvanhorn:Cousot1979Systematic} and understood as computable
approximations of an underlying concrete interpreter. Our work
demonstrates that if this underlying concrete interpreter is written in
monadic style, monad transformers are a useful way to organize and
compose these various kinds of program abstractions in a modular and
language-independent way.

\par

This work is inspired by the trifecta combination of
\citeauthor{dvanhorn:Cousot:1977:AI}'s theory of abstract interpretation
based on Galois connections \cite{dvanhorn:Cousot:1977:AI,
dvanhorn:Cousot1979Systematic, dvanhorn:Cousot98-5},
\citeauthor{davdar:Moggi:1989:Monads}'s original monad transformers
\cite{davdar:Moggi:1989:Monads} which were later popularized in
\citeauthor{dvanhorn:Liang1995Monad}'s \emph{Monad Transformers and
Modular Interpreters} \cite{dvanhorn:Liang1995Monad}, and
\citeauthor{dvanhorn:Sergey2013Monadic}'s \emph{Monadic Abstract
Interpreters} \cite{dvanhorn:Sergey2013Monadic}.

\par

\par

\citet{dvanhorn:Liang1995Monad} first demonstrated how monad
transformers could be used to define building blocks for constructing
(concrete) interpreters. Their interpreter monad
\mbox{\(\mathit{InterpM}\)} bears a strong resemblance to ours. We show
this ``building blocks'' approach to interpreter construction also
extends to \emph{abstract} interpreter construction using Galois
transformers. Moreover, we show that these monad transformers can be
proved sound via a Galois connection to their concrete counterparts,
ensuring the soundness of any stack built from sound blocks of Galois
transformers. Soundness proofs of various forms of analysis are
notoriously brittle with respect to language and analysis features. A
reusable framework of Galois transformers offers a potential way forward
for a modular metatheory of program analysis.

\par

\citet{dvanhorn:Cousot98-5} develops a ``calculational approach'' to
analysis design whereby analyses are not designed and then verified
\emph{post facto}, but rather derived by positing an abstraction and
calculating it from the concrete interpreter using Galois connections.
These calculations are done by hand. Our approach offers a limited
ability to automate the calculation process by relying on monad
transformers to combine different abstractions.

\par

We build directly on the work of Abstracting Abstract Machines (AAM) by
\citet{davdar:van-horn:2010:aam} and
\citet{dvanhorn:Smaragdakis2011Pick} in our parameterization of abstract
time to achieve call-site and object sensitivity. More notably, we
follow the AAM philosophy of instrumenting a concrete semantics
\emph{first} and performing a systematic abstraction \emph{second}. This
greatly simplifies the Galois connection arguments during systematic
abstraction. However, this is at the added cost of proving that the
instrumented semantics simulate the original concrete semantics.

\par

\paragraph{Monadic Abstract Interpreters}

\par

\citeauthor{dvanhorn:Sergey2013Monadic} first introduced the concept of
writing abstract interpreters in monadic style in \emph{Monadic Abstract
Interpreters} (MAI)\cite{dvanhorn:Sergey2013Monadic}, in which
variations in analysis are also recovered through new monad
implementations. However, our approach is considerably different from
MAI.

\par

In MAI, the framework's interface is based on \emph{denotation
functions} for every syntactic form of the language (See
``CPSInterface'', Figure 2 in MAI). This design decision has far
reaching consequences for the entire approach. The denotation functions
in MAI are language-specific and specialized to their example language.
MAI uses a single monad stack fixed to the denotation function
interface: state on top of list (Section 5.3.1 in MAI). New analyses are
achieved through multiple denotation functions into this single monad.
Analyses in MAI are all fixed to be path-sensitive, and the methodology
for incorporating other path or flow properties is to surgically
instrument the execution of the analysis with a custom Galois connection
(Section 6.5 in MAI). Lastly, the framework provides no reasoning
principles or proofs of soundness for the denotation function interface.
\emph{A user of MAI must inline the definitions of each analysis and
prove their implementation correct from scratch each time.}

\par

By contrast, our framework's interface is based on state and
nondeterminism \emph{monadic effects} (Section
\ref{the-analysis-monad}). This interface comes equipped with reasoning
principles, allowing one to verify the correctness of a monadic
interpreter \emph{independent of a particular monad}, which is not
possible in MAI. State and nondeterminism monadic effects capture the
essence of \emph{small-step relational semantics}, and are therefore
truly language independent. Our tools are reusable for any semantics
described as a small-step state machine relation. Because we place the
monadic interpreter behind an interface of effects rather than
denotation functions, we are able to introduce language-independent
monads which capture flow-sensitivity and flow-insensitivity (Sections
\ref{varying-path-and-flow-sensitivity} and
\ref{a-compositional-monadic-framework}), and we show how to compose
these features with other analysis design choices (Sections
\ref{analysis-parameters} and \ref{a-compositional-monadic-framework}).
The monadic effect interface also allows us to separate the monad from
the abstract domain, both of which are tightly coupled in MAI. Finally,
our framework is compositional through the use of monad transformers
(Section \ref{a-compositional-monadic-framework}) which construct
execution engines and proofs of soundness for free.

\par

We do not achieve correctness and compositionality \emph{in addition} to
our transition from denotation functions to monadic effects; rather we
achieve correctness and compositionality \emph{through it}; such a
transition is essential, primary and novel to our work.

\par

\paragraph{Widening for Control-Flow}

\par

\citeauthor{dvanhorn:Hardekopf2014Widening} also introduce a unifying
account of control flow properties in \emph{Widening for Control-Flow}
(WCF)\cite{dvanhorn:Hardekopf2014Widening}, accounting for path, flow
and call-site sensitivities . WCF achieves this through an
instrumentation of the abstract machine's state space which is allowed
to track arbitrary contextual information, up to the path-history of the
entire execution. WCF also develops a modular proof framework, proving
the bulk of soundness proofs for each instantiation of the
instrumentation at once.

\par

Our work achieves similar goals, although isolating path and flow
sensitivity is not our primary objective. While WCF is based on a
language-dependent instrumentation of the semantics, we achieve
variations in path and flow sensitivity by modifying
language-independent control properties of the interpreter through a
monad.

\par

Particular strengths of WCF are the wide range of choices for
control-flow sensitivity which are shown to be implementable within the
design, and the modular proof framework. For example, WCF is able to
also account for call-site sensitivity through their design; we must
account for call-site sensitivity through a different mechanism.

\par

Particular strengths of our work is the understanding of path and flow
sensitivity not through instrumentation but through control properties
of the interpreter, and also a modular proof framework, although modular
in a different sense from WCF. We also show how to compose different
path and flow sensitivity choices for independent components of the
state space, like a flow-sensitive data-store and path-sensitive
stack-store, for example.

\par

\section{Conclusion}\label{conclusion}

\par

We have shown that \emph{Galois transformers}, monad transformers that
transport (1) Galois connections and (2) mappings to an executable
transition system, are effective, language-independent building blocks
for constructing program analyzers, and form the basis of a modular,
reusable and composable metatheory for program analysis.

\par

In the end, we hope language independent characterizations of analysis
ingredients will both facilitate the systematic construction of program
analyses and bridge the gap between various communities which often work
in isolation.
